---
title: "Cox Proportional Hazards model: conc. 25(OH)D & DBP"
author: "Debojyoti Das"
date: "2025-07-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

set output directory
```{r output_dir, message=FALSE, warning=FALSE}
output_dir <- "/Users/debda22ö/Projects/T1D_vitamin_D_results/cox_analysis/concentrations/age20"

if(!dir.exists(output_dir)){
  dir.create(output_dir, recursive = TRUE)
} else {
  cat(paste0(output_dir," is already present!!\n"))
}
```

load libraries
```{r load_libraries, message=FALSE, warning=FALSE}
pacman::p_load(
  haven,        
  zoo,
  sjlabelled,
  janitor,       
  data.table,   
  ggplot2,
  comprehenr,  
  stringr,     
  rstatix,
  dplyr,
  tidyr,
  forcats,
  officer,
  cowplot,
  ggExtra,
  ggpubr,
  GGally,
  janitor,
  ggrepel,
  ggtext,
  parameters,
  lme4,
  broom,
  pscl,
  tibble,
  knitr,
  ggeffects,
  cobalt,
  WeightIt,
  marginaleffects,
  survival,
  survminer,
  splines,
  purrr,
  nnet,
  magick,
  patchwork,
  flextable
  )
```


```{r word_sect_properties}
sect_properties <- prop_section(
  page_size = page_size(
    orient = "landscape",
    height = 11.7,
    width = 8.3
  ),
  page_margins = page_mar(
    top = 0.5,
    bottom = 0.5,
    left = 0.5,
    right = 0.5,
    header = 0.3,
    footer = 0.3,
    gutter = 0
  ),
  type = "continuous"
)
```

define paths
```{r define_paths, eval=TRUE}
data_root <- "/Users/debda22ö/Projects/"
data_root <- paste0(data_root,"Linkoping_projects/ABIS_RAW_DATA_SPSS/")
```

load parameters and helper functions
```{r}
source(paste0(data_root,"../../Rmd_and_R_files/T1D_projects/all_parameters_case_control_T1D_vitamin_D.R"))
source(paste0(data_root,"../../Rmd_and_R_files/T1D_projects/cox_T1D_helper.R"))
```


read/load vitamin_d_case_control_plus_abis for analysis
```{r}
vitamin_d_case_control_plus_abis <- read.table(paste0(data_root, "/vitamin_d/vitamin_d_case_control_plus_abis.csv"))
```

reorder factors
```{r}
vitamin_d_case_control_plus_abis$First_degree_T1D_heredity_cat <- factor(
  vitamin_d_case_control_plus_abis$First_degree_T1D_heredity_cat,
  levels = c("No", "Yes")
)

vitamin_d_case_control_plus_abis$Smoking_during_pregnancy <- factor(
  vitamin_d_case_control_plus_abis$Smoking_during_pregnancy,
  levels = c("No", "Yes")
)

vitamin_d_case_control_plus_abis$Parity <- factor(
  vitamin_d_case_control_plus_abis$Parity,
  levels = c("First parity", "Previous parity")
)

vitamin_d_case_control_plus_abis$Preterm_dichotomous <- factor(
  vitamin_d_case_control_plus_abis$Preterm_dichotomous,
  levels = c("Normal", "Preterm")
)

vitamin_d_case_control_plus_abis$Season <- factor(
  vitamin_d_case_control_plus_abis$Season,
  levels = c("Autumn", "Spring", "Summer", "Winter")
)

vitamin_d_case_control_plus_abis$Sex_cat <- factor(
  vitamin_d_case_control_plus_abis$Sex_cat,
  levels = c("Boy", "Girl")
)

vitamin_d_case_control_plus_abis$Fish_any_pregnancy <- factor(
  vitamin_d_case_control_plus_abis$Fish_any_pregnancy,
  levels = c("Seldom", "1–2 times week or more"))

vitamin_d_case_control_plus_abis$Delivery_mode <- factor(
  vitamin_d_case_control_plus_abis$Delivery_mode,
  levels = c("Normal","caesarean", "other problem"),
  labels = c("Normal","Caesarean", "Vaginal delivery with problem"))
```


```{r}
qs_birth_weight <- round(
  quantile(vitamin_d_case_control_plus_abis$Birth_weight, 
           probs = seq(0, 1, 0.25), 
           na.rm = TRUE),2)

vitamin_d_case_control_plus_abis <- vitamin_d_case_control_plus_abis %>%
    dplyr::mutate(
        Birth_weight_cat = cut(
            Birth_weight,
            breaks = qs_birth_weight,
            include.lowest = TRUE,
            labels = c(
                paste0("≤", qs_birth_weight[2]),                                    
                paste0(qs_birth_weight[2], "–", qs_birth_weight[3]),                        
                paste0(qs_birth_weight[3], "–", qs_birth_weight[4]),                           
                paste0(">", qs_birth_weight[4])                                     
            )
        )
    )


```


Cox analysis

conducting a time-to-event Cox regression with:

 - Time scale: Age (from birth)

 - Exposure: Fixed (cord blood 25(OH)D)

 - Follow-up ends: December 31, 2020

Outcome: T1D diagnosis
```{r event_time, warning=FALSE}
vitamin_d_case_control_plus_abis$start_age <- 0  # since follow-up starts at birth

vitamin_d_case_control_plus_abis$INDATUMA <- as.Date(vitamin_d_case_control_plus_abis$INDATUMA, format = "%Y-%m-%d")
```


```{r event_time_formulation, warning=FALSE}
# Set exit date based on whether T1D occurred before 2020-12-31
vitamin_d_case_control_plus_abis$exit_date <- as.Date(ifelse(
  !is.na(vitamin_d_case_control_plus_abis$T1D_age_diagnosis) &
    vitamin_d_case_control_plus_abis$INDATUMA<= as.Date("2020-12-31"),
  vitamin_d_case_control_plus_abis$INDATUMA,
  as.Date("2020-12-31")
))
```


```{r event_time_calculation, warning=FALSE}
# Calculate age at exit in years
vitamin_d_case_control_plus_abis$exit_age <- as.numeric(
  difftime(
    vitamin_d_case_control_plus_abis$exit_date,
    vitamin_d_case_control_plus_abis$child_dob,
    units = "days"
  )
) / 365.25

vitamin_d_case_control_plus_abis$event <- ifelse(
  !is.na(vitamin_d_case_control_plus_abis$INDATUMA) &
    vitamin_d_case_control_plus_abis$INDATUMA <= as.Date("2020-12-31"),
  1,
  0
)
```


```{r cox_data, warning=FALSE}
df_cox_full <- vitamin_d_case_control_plus_abis %>%
  dplyr::select(all_of(
    c(
      "ABISnr",
      "DBP",
      "conc_25OHD",
      "diagnosis",
      "child_dob",
      "First_degree_T1D_heredity_cat",
      "Preterm_group",
      "Preterm_dichotomous",
      "Delivery_mode",
      "Gestational_age",
      "Birth_weight",
      "Birth_weight_cat",
      "Birth_height",
      "Season",
      "Parity",
      "Smoking_during_pregnancy",
      "Vitamin_mineral_supplement_pregnancy",
      "Fish_lake_pregnancy",
      "Fish_Baltic_pregnancy",
      "Fish_other_pregnancy",
      "Fish_any_pregnancy",
      "Sex_cat",
      # "HDIA",
      "DIAGNOS",
      # "INDATUMA",
      # "UTDATUMA",
      # "SOURCE",
      "T1D_age_diagnosis",
      "onset_age_group",
      "Birth_weight_cat",
      "day_of_year",
      # "Ai",
      # "adjusted_25OHD",
      "start_age",
      "exit_date",
      "exit_age",
      "event"
    )
  ))
```

```{r}
df_cox_full <- df_cox_full %>%
  dplyr::mutate(
    Vitamin_mineral_supplement_pregnancy = ifelse(
      is.na(Vitamin_mineral_supplement_pregnancy) | Vitamin_mineral_supplement_pregnancy == "Do not know",
      NA_character_,
      Vitamin_mineral_supplement_pregnancy)
  )
```


```{r}
child_covariates <- c("Preterm_dichotomous","Birth_weight", "Season", "Sex_cat")

maternal_covariates <- c(
  "First_degree_T1D_heredity_cat",
  "Smoking_during_pregnancy",
  "Fish_any_pregnancy",
  "Parity",
  "Delivery_mode",
  "Vitamin_mineral_supplement_pregnancy"
)
```

```{r remove_missing_rows}
df_cox_full <- df_cox_full %>%
  dplyr::select(all_of(
    c(
      "ABISnr",
      "diagnosis",
      "conc_25OHD",
      "DBP",
      "exit_age",
      "event",
      "Birth_weight_cat",
      child_covariates,
      maternal_covariates,
      "onset_age_group"
    )
  )) 
```

save data in long format
```{r}
write.table(df_cox_full, paste0(output_dir,"/long_format_cox_age_full.csv"), row.names = FALSE)
```

Check Proportinal Hazard Assumption
a. conc 25(OH)D
```{r}
fit_lin_conc_25OHD   <- coxph(
  Surv(exit_age, event) ~ conc_25OHD + Preterm_dichotomous + Birth_weight_cat +
    Season + Sex_cat + First_degree_T1D_heredity_cat + Smoking_during_pregnancy +
    Fish_any_pregnancy + Parity + Delivery_mode + Vitamin_mineral_supplement_pregnancy,
  data = df_cox_full
)

fit_spline_conc_25OHD <- coxph(
  Surv(exit_age, event) ~ splines::ns(conc_25OHD, df = 3) +
    Preterm_dichotomous + Birth_weight_cat + Season + Sex_cat +
    First_degree_T1D_heredity_cat + Smoking_during_pregnancy +
    Fish_any_pregnancy + Parity + Delivery_mode + Vitamin_mineral_supplement_pregnancy,
  data = df_cox_full
)

anova(fit_lin_conc_25OHD, fit_spline_conc_25OHD, test="LRT")
```
Schoenfeld residuals against the transformed time

```{r}
zph_lin_conc_25OHD <- cox.zph(fit_lin_conc_25OHD)
zph_lin_conc_25OHD_table <- zph_lin_conc_25OHD$table %>% 
  as.data.frame %>%
  tibble::rownames_to_column(var = "Term") %>%
  dplyr::mutate(Term = sapply(Term,clean_terms))

dimnames(zph_lin_conc_25OHD$y)[[2]] <- sapply(dimnames(zph_lin_conc_25OHD$y)[[2]], clean_terms)

print(zph_lin_conc_25OHD)           # identify offenders
          
pdf(
  paste0(output_dir,"/zph_linear_conc_25OHD_schoenfeld_residuals_vs_transformed_time_PH_check.pdf"),
  # units = "in", 
  width = 11.67, 
  height = 8.67, 
  font = "Times" # type = "cairo",
#    res = 300
  )

## page 1
par(mfrow = c(2,2))
plot(zph_lin_conc_25OHD, var = 1:4, plot = TRUE)

## page 2
par(mfrow = c(2,2))
plot(zph_lin_conc_25OHD, var = 5:8, plot = TRUE)

## page 3
par(mfrow = c(2,2))
plot(zph_lin_conc_25OHD, var = 9:11, plot = TRUE)

dev.off()

img <- image_read_pdf(paste0(output_dir, "/zph_linear_conc_25OHD_schoenfeld_residuals_vs_transformed_time_PH_check.pdf"), density = 300)
image_write(img, paste0(output_dir, "/zph_linear_conc_25OHD_schoenfeld_residuals_vs_transformed_time_PH_check_300dpi.pdf"), format = "pdf")
```

b. DBP
```{r}
fit_lin_DBP  <- coxph(
  Surv(exit_age, event) ~ DBP + Preterm_dichotomous + Birth_weight_cat +
    Season + Sex_cat + First_degree_T1D_heredity_cat + Smoking_during_pregnancy +
    Fish_any_pregnancy + Parity + Delivery_mode + Vitamin_mineral_supplement_pregnancy,
  data = df_cox_full
)

fit_spline_DBP <- coxph(
  Surv(exit_age, event) ~ splines::ns(DBP, df = 3) +
    Preterm_dichotomous + Birth_weight_cat + Season + Sex_cat +
    First_degree_T1D_heredity_cat + Smoking_during_pregnancy +
    Fish_any_pregnancy + Parity + Delivery_mode + Vitamin_mineral_supplement_pregnancy,
  data = df_cox_full
)

anova(fit_lin_DBP, fit_spline_DBP, test="LRT")
```

Schoenfeld residuals against the transformed time

```{r}
zph_lin_DBP <- cox.zph(fit_lin_DBP)
zph_lin_DBP_table <- zph_lin_DBP$table %>% 
  as.data.frame %>%
  tibble::rownames_to_column(var = "Term") %>%
  dplyr::mutate(Term = sapply(Term,clean_terms))

dimnames(zph_lin_DBP$y)[[2]] <- sapply(dimnames(zph_lin_DBP$y)[[2]], clean_terms)

print(zph_lin_DBP_table %>%
        dplyr::arrange(desc(chisq),p))           
          
pdf(
  paste0(output_dir,"/zph_linear_DBP_schoenfeld_residuals_vs_transformed_time_PH_check.pdf"),
  # units = "in", 
  width = 11.67, 
  height = 8.67, 
  font = "Times" # type = "cairo",
#    res = 300
  )

## page 1
par(mfrow = c(2,2))
plot(zph_lin_DBP, var = 1:4, plot = TRUE)

## page 2
par(mfrow = c(2,2))
plot(zph_lin_DBP, var = 5:8, plot = TRUE)

## page 3
par(mfrow = c(2,2))
plot(zph_lin_DBP, var = 9:11, plot = TRUE)

dev.off()

img <- image_read_pdf(paste0(output_dir, "/zph_linear_DBP_schoenfeld_residuals_vs_transformed_time_PH_check.pdf"), density = 300)
image_write(img, paste0(output_dir, "/zph_linear_DBP_schoenfeld_residuals_vs_transformed_time_PH_check_300dpi.pdf"), format = "pdf")
# image_write(img, paste0(output_dir, "/zph_linear_DBP_schoenfeld_residuals_vs_transformed_time_PH_check.tiff"), format = "tiff")
```

```{r}
# Rename columns with suffixes
zph_25OHD_named <- zph_lin_conc_25OHD_table %>%
  rename_with(~ paste0(.x, "_25OHD"))

zph_DBP_named <- zph_lin_DBP_table %>%
  rename_with(~ paste0(.x, "_DBP"))

ft <- flextable::flextable(cbind(zph_25OHD_named, zph_DBP_named)) %>%
  flextable::autofit()
```


```{r}
flextable::save_as_docx(
  `scaled Schoenfeld residuals with time` = ft,
  path = paste0(output_dir,"/scaled_schoenfeld_residuals_with_time.docx"),
  pr_section = sect_properties
)
```


Cox regression	Hazard ratio: relative rate of developing T1D at any given age. 

```{r cox_model_base}
cox_model_base_conc_25OHD <- survival::coxph(
  Surv(exit_age, event = event) ~ conc_25OHD,
  data = df_cox_full %>% dplyr::select(exit_age, event,conc_25OHD)
)

cox_model_base_DBP <- survival::coxph(
  Surv(exit_age, event = event) ~ DBP,
  data = df_cox_full %>% dplyr::select(exit_age, event, DBP)
)

firth_base_conc_25OHD <- coxphf::coxphf(
  Surv(exit_age, event) ~ conc_25OHD,
  data = df_cox_full %>% 
    dplyr::select(exit_age, event, conc_25OHD) %>%
    tidyr::drop_na()
  )

firth_base_DBP <- coxphf::coxphf(
  Surv(exit_age, event) ~ DBP,
  data = df_cox_full %>% 
    dplyr::select(exit_age, event, DBP)  %>%
    tidyr::drop_na()
  )
```


```{r cox_model_child_confounders}
cox_model_child_confounders_conc_25OHD <- survival::coxph(
  Surv(exit_age, event = event) ~ conc_25OHD + Season + Sex_cat,
  data = df_cox_full %>%
    dplyr::select(
      exit_age,
      event,
      conc_25OHD,
      Season,
      Sex_cat
    )
)

cox_model_child_confounders_DBP <- survival::coxph(
  Surv(exit_age, event = event) ~ DBP + Season + Sex_cat,
  data = df_cox_full %>%     dplyr::select(
      exit_age,
      event,
      DBP,
      Season,
      Sex_cat
    )
)
```


```{r firth_model_child_confounders}
firth_child_confounders_conc_25OHD <- coxphf::coxphf(
  Surv(exit_age, event) ~ conc_25OHD + Season + Sex_cat,
  data = df_cox_full %>%
    dplyr::select(
      exit_age,
      event,
      conc_25OHD,
      Season,
      Sex_cat
    ) %>% tidyr::drop_na(),
  maxit = 1000,
  maxstep = 0.01
)

firth_child_confounders_DBP <- coxphf::coxphf(
  Surv(exit_age, event) ~ DBP + Season + Sex_cat,
  data = df_cox_full %>%
    dplyr::select(
      exit_age,
      event,
      DBP,
      Preterm_dichotomous,
      Season,
      Sex_cat,
      Birth_weight_cat
    ) %>% tidyr::drop_na(),
  maxit = 500,
  maxstep = 0.10
)
```



```{r cox_model_maternal_confounders}
cox_model_maternal_confounders_conc_25OHD <- survival::coxph(
  Surv(exit_age, event = event) ~ conc_25OHD + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode +
    Vitamin_mineral_supplement_pregnancy,
  data = df_cox_full %>%
    dplyr::select(
      exit_age,
      event,
      conc_25OHD,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Vitamin_mineral_supplement_pregnancy
    )
)

cox_model_maternal_confounders_DBP <- survival::coxph(
  Surv(exit_age, event = event) ~ DBP + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode +
    Vitamin_mineral_supplement_pregnancy,
  data = df_cox_full %>% 
    dplyr::select(
      exit_age,
      event,
      DBP,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Vitamin_mineral_supplement_pregnancy
    )
)
```


```{r firth_model_maternal_confounders}
firth_maternal_confounders_conc_25OHD <- coxphf::coxphf(
  Surv(exit_age, event = event) ~ conc_25OHD + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode +
    Vitamin_mineral_supplement_pregnancy,
  data = df_cox_full %>% 
    dplyr::select(
      exit_age,
      event,
      conc_25OHD,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Vitamin_mineral_supplement_pregnancy
    ) %>% 
    tidyr::drop_na()
)

firth_maternal_confounders_DBP <- coxphf::coxphf(
  Surv(exit_age, event = event) ~ DBP + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode +
    Vitamin_mineral_supplement_pregnancy,
  data = df_cox_full %>% 
    dplyr::select(
      exit_age,
      event,
      DBP,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Vitamin_mineral_supplement_pregnancy
    ) %>% 
    tidyr::drop_na()
)
```

```{r cox_model_maternal_and_child_confounders}
cox_model_maternal_and_child_confounders_conc_25OHD <- survival::coxph(
  Surv(exit_age, event = event) ~ conc_25OHD + First_degree_T1D_heredity_cat + 
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode + 
    Vitamin_mineral_supplement_pregnancy +
    Season + Sex_cat,
  data = df_cox_full %>%
    dplyr::select(
      exit_age,
      event,
      conc_25OHD,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Vitamin_mineral_supplement_pregnancy,
      Season,
      Sex_cat
    )
)

cox_model_maternal_and_child_confounders_DBP <- survival::coxph(
  Surv(exit_age, event = event) ~ DBP + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode +
    Vitamin_mineral_supplement_pregnancy + 
    Season + Sex_cat,
  data = df_cox_full %>% 
    dplyr::select(
      exit_age,
      event,
      DBP,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Vitamin_mineral_supplement_pregnancy,
      Season,
      Sex_cat
    )
)
```

```{r firth_model_maternal_and_child_confounders}
firth_maternal_and_child_confounders_conc_25OHD <- coxphf::coxphf(
  Surv(exit_age, event = event) ~ conc_25OHD + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode + 
    Vitamin_mineral_supplement_pregnancy +
    Season + Sex_cat,
  data = df_cox_full %>% 
    dplyr::select(
      exit_age,
      event,
      conc_25OHD,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Vitamin_mineral_supplement_pregnancy,
      Season,
      Sex_cat
    ) %>% 
    tidyr::drop_na()
)

firth_maternal_and_child_confounders_DBP <- coxphf::coxphf(
  Surv(exit_age, event = event) ~ DBP + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode + 
    Vitamin_mineral_supplement_pregnancy +
    Season + Sex_cat,
  data = df_cox_full %>% 
    dplyr::select(
      exit_age,
      event,
      DBP,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Vitamin_mineral_supplement_pregnancy,
      Season,
      Sex_cat
    ) %>% 
    tidyr::drop_na()
)
```


```{r cox_model_full}
cox_model_full_conc_25OHD <- survival::coxph(
  Surv(exit_age, event = event) ~ conc_25OHD + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode + 
    Season + Sex_cat + Vitamin_mineral_supplement_pregnancy + 
    Preterm_dichotomous + Birth_weight_cat,
  data = df_cox_full %>%
    dplyr::select(
      exit_age,
      event,
      conc_25OHD,
      Preterm_dichotomous,
      Season,
      Sex_cat,
      Birth_weight_cat,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Vitamin_mineral_supplement_pregnancy
    )
)

cox_model_full_DBP <- survival::coxph(
  Surv(exit_age, event = event) ~ DBP + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode +
    Season + Sex_cat + Vitamin_mineral_supplement_pregnancy + 
    Preterm_dichotomous + Birth_weight_cat,
  data = df_cox_full %>%
    dplyr::select(
      exit_age,
      event,
      DBP,
      Preterm_dichotomous,
      Season,
      Sex_cat,
      Birth_weight_cat,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Vitamin_mineral_supplement_pregnancy
    )
)
```


```{r firth_model_full}
firth_full_conc_25OHD <- coxphf::coxphf(
  Surv(exit_age, event = event) ~ conc_25OHD + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode +
    Season + Sex_cat + Vitamin_mineral_supplement_pregnancy + 
    Preterm_dichotomous + Birth_weight_cat,
  data = df_cox_full %>%
    dplyr::select(
      exit_age,
      event,
      conc_25OHD,
      Preterm_dichotomous,
      Season,
      Sex_cat,
      Birth_weight_cat,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Vitamin_mineral_supplement_pregnancy
    ) %>%
    tidyr::drop_na(),
  maxit = 1000,
  maxstep = 0.01
)

firth_full_DBP <- coxphf::coxphf(
  Surv(exit_age, event = event) ~ DBP + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode +
    Season + Sex_cat + Vitamin_mineral_supplement_pregnancy + 
    Preterm_dichotomous + Birth_weight_cat,
  data = df_cox_full %>%
    dplyr::select(
      exit_age,
      event,
      DBP,
      Preterm_dichotomous,
      Season,
      Sex_cat,
      Birth_weight_cat,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Vitamin_mineral_supplement_pregnancy
    ) %>%
    tidyr::drop_na(),
  maxit = 5000,
  maxstep = 0.01
)
```


create pattern strings to clean-tidy result tables 
```{r}
child_levels <- paste0(
  paste0(unlist(lapply(c(child_covariates[-2], "Birth_weight_cat"), function(x)
  levels(factor(df_cox_full[[x]]))))[-2], collapse = "|"), "|^Preterm$")

child_pat_replace <- paste0(c(child_covariates[-2],"Birth_weight_cat"),collapse = "|")
```

```{r}
maternal_levels <- paste0(
  unique(
    unlist(
      lapply(
        maternal_covariates, 
        function(x) levels(factor(df_cox_full[[x]]))
        )
      )
    ), 
  collapse = "|")

maternal_pat_replace <- paste0(maternal_covariates, collapse = "|")
```

```{r}
all_levels <- (paste0(paste0(
  unique(
    unlist(
      lapply(
        c(child_covariates[-2],"Birth_weight_cat",maternal_covariates), 
        function(x) levels(factor(df_cox_full[[x]]))
        )
      )
    )[-2], 
  collapse = "|"), "|^Preterm$"))

all_pat_replace <- paste0(
  c(child_covariates[-2],"Birth_weight_cat",maternal_covariates), collapse = "|")
```

```{r tidy_model_base}
tidy_cox_model_base_conc_25OHD <- tidy_base_model(cox_model_base_conc_25OHD, "Base", "conc_25OHD")
tidy_cox_model_base_DBP <- tidy_base_model(cox_model_base_DBP, "Base","DBP")

tidy_firth_base_conc_25OHD <- tidy_base_model(firth_base_conc_25OHD, "Base", "conc_25OHD", firth_model = TRUE)
tidy_firth_base_DBP <- tidy_base_model(firth_base_DBP, "Base","DBP", firth_model = TRUE)
```



```{r tidy_model_child_confounders}
tidy_cox_model_child_confounders_conc_25OHD <- tidy_model_specific_covariates(
  model = cox_model_child_confounders_conc_25OHD,
  predictor = "conc_25OHD",
  model_name = "Base + Child confounders", 
  covariates_as_pattern = child_pat_replace, 
  covariate_levels_as_pattern = child_levels
  )

tidy_cox_model_child_confounders_DBP <- tidy_model_specific_covariates(
  model = cox_model_child_confounders_DBP,
  predictor = "DBP",
  model_name = "Base + Child confounders", 
  covariates_as_pattern = child_pat_replace, 
  covariate_levels_as_pattern = child_levels
  )

tidy_firth_child_confounders_conc_25OHD <- tidy_model_specific_covariates(
  model = firth_child_confounders_conc_25OHD,
  predictor = "conc_25OHD",
  model_name = "Base + Child confounders", 
  covariates_as_pattern = child_pat_replace, 
  covariate_levels_as_pattern = child_levels, 
  firth_model = TRUE
  )

tidy_firth_child_confounders_DBP <- tidy_model_specific_covariates(
  model = firth_child_confounders_DBP,
  predictor = "DBP",
  model_name = "Base + Child confounders", 
  covariates_as_pattern = child_pat_replace, 
  covariate_levels_as_pattern = child_levels, 
  firth_model = TRUE
  )
```

```{r tidy_model_maternal_confounders}
tidy_cox_model_maternal_confounders_conc_25OHD <- tidy_model_specific_covariates(
  model = cox_model_maternal_confounders_conc_25OHD,
  predictor = "conc_25OHD",
  model_name = "Base + Maternal confounders", 
  covariates_as_pattern = maternal_pat_replace, 
  covariate_levels_as_pattern = maternal_levels
  )

tidy_cox_model_maternal_confounders_DBP <- tidy_model_specific_covariates(
  model = cox_model_maternal_confounders_DBP,
  predictor = "DBP",
  model_name = "Base + Maternal confounders", 
  covariates_as_pattern = maternal_pat_replace, 
  covariate_levels_as_pattern = maternal_levels
  )

tidy_firth_maternal_confounders_conc_25OHD <- tidy_model_specific_covariates(
  model = firth_maternal_confounders_conc_25OHD,
  predictor = "conc_25OHD",
  model_name = "Base + Maternal confounders", 
  covariates_as_pattern = maternal_pat_replace, 
  covariate_levels_as_pattern = maternal_levels, 
  firth_model = TRUE
  )

tidy_firth_maternal_confounders_DBP <- tidy_model_specific_covariates(
  model = firth_maternal_confounders_DBP,
  predictor = "DBP",
  model_name = "Base + Maternal confounders", 
  covariates_as_pattern = maternal_pat_replace, 
  covariate_levels_as_pattern = maternal_levels, 
  firth_model = TRUE
  )
```

```{r tidy_model_maternal_and_child_confounders}
tidy_cox_model_maternal_and_child_confounders_conc_25OHD <- tidy_model_specific_covariates(
  model = cox_model_maternal_and_child_confounders_conc_25OHD,
  predictor = "conc_25OHD",
  model_name = "Base + Maternal\n + Child confounders", 
  covariates_as_pattern = all_pat_replace, 
  covariate_levels_as_pattern = all_levels
  )

tidy_cox_model_maternal_and_child_confounders_DBP <- tidy_model_specific_covariates(
  model = cox_model_maternal_and_child_confounders_DBP,
  predictor = "DBP",
  model_name = "Base + Maternal\n + Child confounders", 
  covariates_as_pattern = all_pat_replace, 
  covariate_levels_as_pattern = all_levels
  )

tidy_firth_maternal_and_child_confounders_conc_25OHD <- tidy_model_specific_covariates(
  model = firth_maternal_and_child_confounders_conc_25OHD,
  predictor = "conc_25OHD",
  model_name = "Base + Maternal\n + Child confounders", 
  covariates_as_pattern = all_pat_replace, 
  covariate_levels_as_pattern = all_levels, 
  firth_model = TRUE
  )

tidy_firth_maternal_and_child_confounders_DBP <- tidy_model_specific_covariates(
  model = firth_maternal_and_child_confounders_DBP,
  predictor = "DBP",
  model_name = "Base + Maternal\n + Child confounders", 
  covariates_as_pattern = all_pat_replace, 
  covariate_levels_as_pattern = all_levels, 
  firth_model = TRUE
  )
```

```{r tidy_cox_model_full}
tidy_cox_model_full_conc_25OHD <- tidy_model_specific_covariates(
  model = cox_model_full_conc_25OHD,
  predictor = "conc_25OHD",
  model_name = "Base + All confounders\n + Child mediators", 
  covariates_as_pattern = all_pat_replace, 
  covariate_levels_as_pattern = all_levels
  ) 

tidy_cox_model_full_DBP <- tidy_model_specific_covariates(
  model = cox_model_full_DBP,
  predictor = "DBP",
  model_name = "Base + All confounders\n + Child mediators", 
  covariates_as_pattern = all_pat_replace, 
  covariate_levels_as_pattern = all_levels
  ) 

tidy_firth_full_conc_25OHD <- tidy_model_specific_covariates(
  model = firth_full_conc_25OHD,
  predictor = "conc_25OHD",
  model_name = "Base + All confounders\n + Child mediators", 
  covariates_as_pattern = all_pat_replace, 
  covariate_levels_as_pattern = all_levels, 
  firth_model = TRUE
  )

tidy_firth_full_DBP <- tidy_model_specific_covariates(
  model = firth_full_DBP,
  predictor = "DBP",
  model_name = "Base + All confounders\n + Child mediators", 
  covariates_as_pattern = all_pat_replace, 
  covariate_levels_as_pattern = all_levels, 
  firth_model = TRUE
  )
```




```{r merged_all_model_results_conc_25OHD }
merged_all_cox_model_results_conc_25OHD <-  rbind(
  tidy_cox_model_base_conc_25OHD,
  tidy_cox_model_child_confounders_conc_25OHD,
  tidy_cox_model_maternal_confounders_conc_25OHD,
  tidy_cox_model_maternal_and_child_confounders_conc_25OHD,
  tidy_cox_model_full_conc_25OHD
) %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::set_table_properties(width = 1, layout = "autofit") %>%
  flextable::color(
    i = ~ as.numeric(`P-value`) < 0.05,
    j = c("Term", "Level", "P-value"),
    color = "red"
  )

merged_all_firth_model_results_conc_25OHD <-  rbind(
  tidy_firth_base_conc_25OHD,
  tidy_firth_child_confounders_conc_25OHD,
  tidy_firth_maternal_confounders_conc_25OHD,
  tidy_firth_maternal_and_child_confounders_conc_25OHD,
  tidy_firth_full_conc_25OHD
) %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::set_table_properties(width = 1, layout = "autofit") %>%
  flextable::color(
    i = ~ as.numeric(`P-value`) < 0.05,
    j = c("Term", "Level", "P-value"),
    color = "red"
  )
```

```{r merged_all_model_results_DBP}
merged_all_cox_model_results_DBP <-  rbind(
  tidy_cox_model_base_DBP,
  tidy_cox_model_child_confounders_DBP,
  tidy_cox_model_maternal_confounders_DBP, 
  tidy_cox_model_maternal_and_child_confounders_DBP,
  tidy_cox_model_full_DBP
) %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::set_table_properties(width = 1, layout = "autofit") %>%
  flextable::color(
    i = ~ as.numeric(`P-value`) < 0.05,
    j = c("Term", "Level", "P-value"),
    color = "red"
  )

merged_all_firth_model_results_DBP <-  rbind(
  tidy_firth_base_DBP,
  tidy_firth_child_confounders_DBP,
  tidy_firth_maternal_confounders_DBP,
  tidy_firth_maternal_and_child_confounders_DBP,
  tidy_firth_full_DBP
) %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::set_table_properties(width = 1, layout = "autofit") %>%
  flextable::color(
    i = ~ as.numeric(`P-value`) < 0.05,
    j = c("Term", "Level", "P-value"),
    color = "red"
  )
```


```{r save_merged_all_model_results}
flextable::save_as_docx(
  `Cox proportional hazards models: conc. 25(OH)D` = merged_all_cox_model_results_conc_25OHD,
  path = paste0(output_dir,"/flex_merged_all_cox_model_results_conc_25OHD.docx"),
  pr_section = sect_properties
)

flextable::save_as_docx(
  `Cox proportional hazards models: DBP` = merged_all_cox_model_results_DBP,
  path = paste0(output_dir,"/flex_merged_all_cox_model_results_DBP.docx"),
  pr_section = sect_properties
)

flextable::save_as_docx(
  `Cox proportional hazards models with Firth's Penalized Likelihood: conc. 25(OH)D` = merged_all_firth_model_results_conc_25OHD,
  path = paste0(output_dir,"/flex_merged_all_firth_model_results_conc_25OHD.docx"),
  pr_section = sect_properties
)

flextable::save_as_docx(
  `Cox proportional hazards models with Firth's Penalized Likelihood: DBP` = merged_all_firth_model_results_DBP,
  path = paste0(output_dir,"/flex_merged_all_firth_model_results_DBP.docx"),
  pr_section = sect_properties
)
```

Cox models to test effect modification by heredity

1. Interaction model (primary test)

```{r}
cox_model_conc_25OHD_interaction_heredity <- coxph(
  Surv(exit_age, event = event) ~ conc_25OHD * First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode + 
    Vitamin_mineral_supplement_pregnancy + Season + Sex_cat,
  data = df_cox_full %>% 
    dplyr::select(
      exit_age,
      event,
      conc_25OHD,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Vitamin_mineral_supplement_pregnancy,
      Season,
      Sex_cat
    ) %>% 
    tidyr::drop_na()
)

cox_model_conc_25OHD_interaction_season <- coxph(
  Surv(exit_age, event = event) ~ conc_25OHD * Season + Smoking_during_pregnancy + 
    Fish_any_pregnancy + Parity + Delivery_mode + 
    Vitamin_mineral_supplement_pregnancy + First_degree_T1D_heredity_cat + Sex_cat,
  data = df_cox_full %>% 
    dplyr::select(
      exit_age,
      event,
      conc_25OHD,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Vitamin_mineral_supplement_pregnancy,
      Season,
      Sex_cat
    ) %>% 
    tidyr::drop_na()
)

cox_model_conc_25OHD_interaction_vitamin_supplement <- coxph(
  Surv(exit_age, event = event) ~ conc_25OHD * Vitamin_mineral_supplement_pregnancy + 
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode + 
    First_degree_T1D_heredity_cat +  Season + Sex_cat,
  data = df_cox_full %>% 
    dplyr::select(
      exit_age,
      event,
      conc_25OHD,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Vitamin_mineral_supplement_pregnancy,
      Season,
      Sex_cat
    ) %>% 
    tidyr::drop_na()
)
```

```{r}
cox_model_DBP_interaction_heredity <- coxph(
  Surv(exit_age, event = event) ~ DBP * First_degree_T1D_heredity_cat + 
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode +
    Vitamin_mineral_supplement_pregnancy + Season + Sex_cat,
  data = df_cox_full %>% 
    dplyr::select(
      exit_age,
      event,
      DBP,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Vitamin_mineral_supplement_pregnancy,
      Season,
      Sex_cat
    ) %>% 
    tidyr::drop_na()
)

cox_model_DBP_interaction_season <- coxph(
  Surv(exit_age, event = event) ~ DBP * Season + 
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode + 
    Vitamin_mineral_supplement_pregnancy + First_degree_T1D_heredity_cat + Sex_cat,
  data = df_cox_full %>% 
    dplyr::select(
      exit_age,
      event,
      DBP,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Vitamin_mineral_supplement_pregnancy,
      Season,
      Sex_cat
    ) %>% 
    tidyr::drop_na()
)

cox_model_DBP_interaction_vitamin_supplement <- coxph(
  Surv(exit_age, event = event) ~ DBP * Vitamin_mineral_supplement_pregnancy + 
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode + 
    First_degree_T1D_heredity_cat +  Season + Sex_cat,
  data = df_cox_full %>% 
    dplyr::select(
      exit_age,
      event,
      DBP,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Vitamin_mineral_supplement_pregnancy,
      Season,
      Sex_cat
    ) %>% 
    tidyr::drop_na()
)
```


```{r}
tab_conc_25OHD_hered <- make_comp_table(
  cox_model_maternal_and_child_confounders_conc_25OHD,
  cox_model_conc_25OHD_interaction_heredity,
  "conc 25(OH)D × First-degree T1D Heredity"
)

tab_conc_25OHD_season <- make_comp_table(
  cox_model_maternal_and_child_confounders_conc_25OHD,
  cox_model_conc_25OHD_interaction_season,
  "conc 25(OH)D × Season"
)

tab_conc_25OHD_vitamin_supplement <- make_comp_table(
  cox_model_maternal_and_child_confounders_conc_25OHD,
  cox_model_conc_25OHD_interaction_vitamin_supplement,
  "conc 25(OH)D × Vitamin Supplement"
)

tab_DBP_hered <- make_comp_table(
  cox_model_maternal_and_child_confounders_DBP,
  cox_model_DBP_interaction_heredity,
  "DBP × First-degree T1D Heredity"
)

tab_DBP_season <- make_comp_table(
  cox_model_maternal_and_child_confounders_DBP,
  cox_model_DBP_interaction_season,
  "DBP × Season"
)

tab_DBP_vitamin_supplement <- make_comp_table(
  cox_model_maternal_and_child_confounders_DBP,
  cox_model_DBP_interaction_vitamin_supplement,
  "DBP × Vitamin Supplement "
)

anova_interaction <- rbind(
  tab_conc_25OHD_hered,
  tab_conc_25OHD_season,
  tab_conc_25OHD_vitamin_supplement,
  tab_DBP_hered,
  tab_DBP_season,
  tab_DBP_vitamin_supplement
  ) %>%
  flextable::flextable() %>%
  flextable::autofit()

flextable::save_as_docx(
  `Analysis of Deviance Table: Cox model: response is  Surv(time, event)` = anova_interaction,
  path = paste0(output_dir,"/anova_interaction.docx"),
  pr_section = sect_properties
)
```


survival plots for cox model with maternal and child confounders    
```{r}
plots <- lapply(names(cox_model_maternal_and_child_confounders_conc_25OHD$assign), function(v) {
  if(v %in% c("conc_25OHD", "DBP")){
        suppressWarnings(p <- survminer::ggadjustedcurves(
        fit = cox_model_maternal_and_child_confounders_conc_25OHD,
        variable = v,
        method = "single"
    ))
  p + ggtitle(label = "", subtitle = clean_terms(v))
  
  } else {
        suppressWarnings(p <- survminer::ggadjustedcurves(
        fit = cox_model_maternal_and_child_confounders_conc_25OHD,
        variable = v
    ))
    p + labs(color = clean_terms(v)) 

  }
})

names(plots) <- names(cox_model_maternal_and_child_confounders_conc_25OHD$assign)
```

```{r}
plots_per_page <- 4
n_pages <- ceiling(length(plots) / plots_per_page)

pdf(paste0(output_dir,"/adjusted_curves_cox_model_maternal_and_child_confounders_conc_25OHD.pdf"), width = 14, height = 10)  # 2x2 layout

for(i in seq_len(n_pages)) {
  idx <- ((i-1)*plots_per_page + 1):min(i*plots_per_page, length(plots))
  # Combine up to 4 plots
  p <- wrap_plots(plots[idx], ncol = 2, nrow = 2)
  print(p)
}

dev.off()

img <- image_read_pdf(paste0(output_dir, "/adjusted_curves_cox_model_maternal_and_child_confounders_conc_25OHD.pdf"), density = 300)
image_write(img, paste0(output_dir, "/adjusted_curves_cox_model_maternal_and_child_confounders_conc_25OHD_300dpi.pdf"), format = "pdf")
```

<!-- Stratifed Cox -->
```{r strata_wise_cox_base}
strata_wise_cox_base_conc_25OHD <- run_stratified_cox_models(
  data = df_cox_full %>% dplyr::select(-DBP),
  predictor = "conc_25OHD",
  strata_vars = c(child_covariates[-2], "Birth_weight_cat", maternal_covariates)
  )

strata_wise_cox_base_DBP <- run_stratified_cox_models(
  data = df_cox_full %>% dplyr::select(-conc_25OHD),
  predictor = "DBP",
  strata_vars = c(child_covariates[-2], "Birth_weight_cat", maternal_covariates)
  )
```

Diagnostics cox_model_maternal_and_child_confounders_conc_25OHD
```{r}
p_conc_25OHD <- survminer::ggcoxdiagnostics(cox_model_maternal_and_child_confounders_conc_25OHD, type = "schoenfeld")
p_conc_25OHD$data$covariate <- sapply(p_conc_25OHD$data$covariate, add_colon)


p_DBP <- survminer::ggcoxdiagnostics(cox_model_maternal_and_child_confounders_DBP, type = "schoenfeld")
p_DBP$data$covariate <- sapply(p_DBP$data$covariate, add_colon)
```

```{r}
fit_surv_season <- survfit(Surv(exit_age, event) ~ Season, data = df_cox_full)
fit_surv_T1D_heredity <- survfit(Surv(exit_age, event) ~ First_degree_T1D_heredity_cat, data = df_cox_full)
fit_surv_vitamin_supplement <- survfit(Surv(exit_age, event) ~ Vitamin_mineral_supplement_pregnancy, data = df_cox_full)      
```

Season and T1D Heredity survival plots
```{r}
plot_fit_surv_season <- ggsurvplot(
    fit_surv_season, 
    data = df_cox_full, 
    size = 1,                 
    palette = c("#E7B800","#00B800","#E70000", "#2E9FDF"),# custom color palettes
    conf.int = TRUE,          
    pval = TRUE,              
    risk.table = TRUE,        
    risk.table.col = "strata",
    risk.table.height = 0.25,
    ggtheme = theme_light()     
)

plot_fit_surv_heredity <- ggsurvplot(
    fit_surv_T1D_heredity, 
    data = df_cox_full, 
    size = 1,                 
    palette = c("#00B800","#E70000"),# custom color palettes
    conf.int = TRUE,          
    pval = TRUE,              
    risk.table = TRUE,        
    risk.table.col = "strata",
    risk.table.height = 0.25,
    ggtheme = theme_light()     
)

plot_fit_surv_vitamin_supplement <- ggsurvplot(
    fit_surv_vitamin_supplement, 
    data = df_cox_full, 
    size = 1,                 
    palette = c("#00B800","#E70000"),# custom color palettes
    conf.int = TRUE,          
    pval = TRUE,              
    risk.table = TRUE,        
    risk.table.col = "strata",
    risk.table.height = 0.25,
    ggtheme = theme_light()     
)
```

```{r}
fit_surv_T1D_heredity_season_vitamin_supplement <- survival::survfit(
  survival::Surv(exit_age, event) ~ First_degree_T1D_heredity + Season + Vitamin_supplement,
  data = df_cox_full %>% 
    dplyr::rename(
      First_degree_T1D_heredity = First_degree_T1D_heredity_cat,
      Vitamin_supplement = Vitamin_mineral_supplement_pregnancy
      )
  )
```

```{r}
plot_fit_surv_heredity_season_vitamin_supplement <- ggsurvplot(
    fit_surv_T1D_heredity_season_vitamin_supplement, 
    data = df_cox_full  %>% 
    dplyr::rename(
      First_degree_T1D_heredity = First_degree_T1D_heredity_cat,
      Vitamin_supplement = Vitamin_mineral_supplement_pregnancy
      ), 
    size = 1,                 
    conf.int = TRUE,          
    # pval = TRUE,              
    risk.table = TRUE,        
    risk.table.col = "strata",
    risk.table.height = 0.25,
    ggtheme = theme_light()     
) 

plot_fit_surv_heredity_season_vitamin_supplement$plot <- plot_fit_surv_heredity_season_vitamin_supplement$plot + 
  facet_grid(Vitamin_supplement ~ Season)
  # facet_grid(First_degree_T1D_heredity ~ Season)

plot_fit_surv_heredity_season_vitamin_supplement$table$data$strata <- sapply(plot_fit_surv_heredity_season_vitamin_supplement$table$data$strata, clean_terms)
plot_fit_surv_heredity_season_vitamin_supplement$table <- plot_fit_surv_heredity_season_vitamin_supplement$table + 
  theme(legend.position = "none")
plot_fit_surv_heredity_season_vitamin_supplement$plot <- plot_fit_surv_heredity_season_vitamin_supplement$plot +
  theme(legend.position = "bottom")
plot_fit_surv_heredity_season_vitamin_supplement$table <- plot_fit_surv_heredity_season_vitamin_supplement$table + theme( axis.title.y = element_blank())
```


Leave child Mediators "Preterm_dichotomous" and "Birth_weight_cat" 
```{r}
season_stratied_cox_model_maternal_and_child_confounders_conc_25OHD <- run_stratified_cox_models(
  data = df_cox_full %>% dplyr::select(-DBP),
  predictor = "conc_25OHD",
  strata_vars = c("Season"),
  covariate_list = c(
    "Sex_cat",
    "First_degree_T1D_heredity_cat",
    "Smoking_during_pregnancy",
    "Fish_any_pregnancy",
    "Parity",
    "Delivery_mode",
    "Vitamin_mineral_supplement_pregnancy"
  ),
  covariates_as_pattern = all_pat_replace,
  covariate_levels_as_pattern = all_levels
)

T1D_heredity_stratied_cox_model_maternal_and_child_confounders_conc_25OHD <- run_stratified_cox_models(
  data = df_cox_full %>% dplyr::select(-DBP),
  predictor = "conc_25OHD",
  strata_vars = c("First_degree_T1D_heredity_cat"),
  covariate_list = c(
    "Season", 
    "Sex_cat",
    "Smoking_during_pregnancy",
    "Fish_any_pregnancy",
    "Parity",
    "Delivery_mode",
    "Vitamin_mineral_supplement_pregnancy"
  ),
  covariates_as_pattern = all_pat_replace,
  covariate_levels_as_pattern = all_levels
)

vitamin_supplement_stratied_cox_model_maternal_and_child_confounders_conc_25OHD <- run_stratified_cox_models(
  data = df_cox_full %>% dplyr::select(-DBP),
  predictor = "conc_25OHD",
  strata_vars = c("Vitamin_mineral_supplement_pregnancy"),
  covariate_list = c(
    "Season", 
    "Sex_cat",
    "Smoking_during_pregnancy",
    "Fish_any_pregnancy",
    "Parity",
    "Delivery_mode",
    "First_degree_T1D_heredity_cat"
  ),
  covariates_as_pattern = all_pat_replace,
  covariate_levels_as_pattern = all_levels
)
```

```{r}
stratified_conc_25OHD <- rbind(
    season_stratied_cox_model_maternal_and_child_confounders_conc_25OHD[[1]],
    T1D_heredity_stratied_cox_model_maternal_and_child_confounders_conc_25OHD[[1]],
    vitamin_supplement_stratied_cox_model_maternal_and_child_confounders_conc_25OHD[[1]]
    ) %>%
    table2flex()
```

```{r}
flextable::save_as_docx(
  `Sensitivity Analysis: Cox models stratified by Season, First-degree T1D heredity, and Vitamin supplement` = stratified_conc_25OHD,
  path = paste0(output_dir,"/stratified_conc_25OHD.docx"),
  pr_section = sect_properties
)
```


<!-- ```{r strata_wise_cox_child_covariates} -->
<!-- strata_wise_cox_child_covariates_conc_25OHD <- run_stratified_cox_models( -->
<!--   data = df_cox_full %>% dplyr::select(-DBP),  -->
<!--   predictor ="conc_25OHD", -->
<!--   strata_vars = maternal_covariates, -->
<!--   covariate_list = c(child_covariates[-2], "Birth_weight_cat"), -->
<!--   covariates_as_pattern = child_pat_replace, -->
<!--   covariate_levels_as_pattern = child_levels -->
<!--   ) -->

<!-- strata_wise_cox_child_covariates_DBP <- run_stratified_cox_models( -->
<!--   data = df_cox_full %>% dplyr::select(-conc_25OHD),  -->
<!--   predictor = "DBP", -->
<!--   strata_vars = maternal_covariates, -->
<!--   covariate_list = c(child_covariates[-2], "Birth_weight_cat"), -->
<!--   covariates_as_pattern = child_pat_replace, -->
<!--   covariate_levels_as_pattern = child_levels -->
<!--   ) -->
<!-- ``` -->

<!-- ```{r strata_wise_cox_maternal_covariates} -->
<!-- strata_wise_cox_maternal_covariates_conc_25OHD <- run_stratified_cox_models( -->
<!--   data = df_cox_full %>% dplyr::select(-DBP),  -->
<!--   predictor ="conc_25OHD", -->
<!--   strata_vars = c(child_covariates[-2], "Birth_weight_cat"), -->
<!--   covariate_list =  maternal_covariates, -->
<!--   covariates_as_pattern = maternal_pat_replace, -->
<!--   covariate_levels_as_pattern = maternal_levels -->
<!--   ) -->

<!-- strata_wise_cox_maternal_covariates_DBP <- run_stratified_cox_models( -->
<!--   data = df_cox_full %>% dplyr::select(-conc_25OHD),  -->
<!--   predictor = "DBP", -->
<!--   strata_vars = c(child_covariates[-2], "Birth_weight_cat"), -->
<!--   covariate_list =  maternal_covariates, -->
<!--   covariates_as_pattern = maternal_pat_replace, -->
<!--   covariate_levels_as_pattern = maternal_levels -->
<!--   ) -->
<!-- ``` -->


<!-- ```{r strata_flex_tables} -->
<!-- flex_strata_wise_cox_base_conc_25OHD <- table2flex(strata_wise_cox_base_conc_25OHD) -->
<!-- flex_strata_wise_cox_base_DBP <- table2flex(strata_wise_cox_base_DBP) -->

<!-- flex_strata_wise_cox_child_covariates_conc_25OHD <- table2flex(strata_wise_cox_child_covariates_conc_25OHD) -->
<!-- flex_strata_wise_cox_child_covariates_DBP <- table2flex(strata_wise_cox_child_covariates_DBP) -->

<!-- flex_strata_wise_cox_maternal_covariates_conc_25OHD <- table2flex(strata_wise_cox_maternal_covariates_conc_25OHD) -->
<!-- flex_strata_wise_cox_maternal_covariates_DBP <- table2flex(strata_wise_cox_maternal_covariates_DBP) -->
<!-- ``` -->




<!-- ```{r save_flex_strata_wise_cox_models} -->
<!-- flextable::save_as_docx( -->
<!--   `Strata wise cox (base) model: conc. 25(OH)D` = flex_strata_wise_cox_base_conc_25OHD, -->
<!--   path = paste0(output_dir,"/flex_strata_wise_cox_base_conc_25OHD.docx"), -->
<!--   pr_section = sect_properties -->
<!-- ) -->

<!-- flextable::save_as_docx( -->
<!--   `Strata wise cox  (base) model: DBP` = flex_strata_wise_cox_base_DBP, -->
<!--   path = paste0(output_dir,"/flex_strata_wise_cox_base_DBP.docx"), -->
<!--   pr_section = sect_properties -->
<!-- ) -->

<!-- flextable::save_as_docx( -->
<!--   `Strata wise cox model with child covariates: conc. 25(OH)D` = flex_strata_wise_cox_child_covariates_conc_25OHD, -->
<!--   path = paste0(output_dir,"/flex_strata_wise_cox_child_covariates_conc_25OHD.docx"), -->
<!--   pr_section = sect_properties -->
<!-- ) -->

<!-- flextable::save_as_docx( -->
<!--   `Strata wise cox model with child covariates: DBP` = flex_strata_wise_cox_child_covariates_DBP, -->
<!--   path = paste0(output_dir,"/flex_strata_wise_cox_child_covariates_DBP.docx"), -->
<!--   pr_section = sect_properties -->
<!-- ) -->

<!-- flextable::save_as_docx( -->
<!--   `Strata wise cox model with maternal covariates: conc. 25(OH)D` = flex_strata_wise_cox_maternal_covariates_conc_25OHD, -->
<!--   path = paste0(output_dir,"/flex_strata_wise_cox_maternal_covariates_conc_25OHD.docx"), -->
<!--   pr_section = sect_properties -->
<!-- ) -->

<!-- flextable::save_as_docx( -->
<!--   `Strata wise cox model with maternal covariates: DBP` = flex_strata_wise_cox_maternal_covariates_DBP, -->
<!--   path = paste0(output_dir,"/flex_strata_wise_cox_maternal_covariates_DBP.docx"), -->
<!--   pr_section = sect_properties -->
<!-- ) -->
<!-- ``` -->
