---
title: "Landmark analysis - Weighted Cox model: 25(OH)D and DBP"
author: "Debojyoti Das"
date: "2026-05-08"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
```

set output directory
```{r output_dir, message=FALSE, warning=FALSE}
output_dir <- "/Users/debda22/Projects/abis-vitamin-d-type1-diabetes/results/revision/case_and_control_weighted_cox/landmark"

if(!dir.exists(output_dir)){
  dir.create(output_dir, recursive = TRUE)
} else {
  cat(paste0(output_dir," is already present!!\n"))
}
```

load libraries
```{r load_libraries, message=FALSE, warning=FALSE}
pacman::p_load(
  haven,        
  zoo,
  sjlabelled,
  janitor,       
  data.table,   
  ggplot2,
  comprehenr,  
  stringr,     
  rstatix,
  dplyr,
  tidyr,
  forcats,
  officer,
  cowplot,
  ggExtra,
  ggpubr,
  GGally,
  janitor,
  ggrepel,
  ggtext,
  parameters,
  lme4,
  broom,
  pscl,
  tibble,
  knitr,
  ggeffects,
  cobalt,
  WeightIt,
  marginaleffects,
  survival,
  survminer,
  splines,
  purrr,
  nnet,
  magick,
  patchwork,
  flextable
  )
```


```{r word_sect_properties}
sect_properties <- prop_section(
  page_size = page_size(
    orient = "landscape",
    height = 11.7,
    width = 8.3
  ),
  page_margins = page_mar(
    top = 0.5,
    bottom = 0.5,
    left = 0.5,
    right = 0.5,
    header = 0.3,
    footer = 0.3,
    gutter = 0
  ),
  type = "continuous"
)

set_flextable_defaults(
    padding.top = 0, 
    padding.bottom = 0,
    line_spacing = 1
)
```

define paths
```{r define_paths, eval=TRUE}
data_root <- "/Users/debda22/Projects/abis-data"
proj_root <- "/Users/debda22/Projects/abis-vitamin-d-type1-diabetes"
```

load parameters and helper functions
```{r}
source(paste0(proj_root,"/rscripts/all_parameters_case_control_T1D_vitamin_D.R"))
source(paste0(proj_root,"/rscripts/weighted_cox_T1D_helper.R"))
```

```{r}
abis <- readRDS(file.path(data_root,"vitamin_d","covariates.abis.full.cohort.RDS"))
```

```{r}
abis <- abis %>%
  dplyr::filter(!is.na(child_dob))

print("number of participants with DOB:")
print(abis %>% nrow())
print("number of participants in birth questionnaire:")
print(abis %>% count(par_b))
```

window of interest
```{r}
event_time_formulation <- function(data2use, age_cutoff = 5, age_cutoff_lower_limit = 0, age_cutoff_date = as.Date(NA)){

  # Define the age cutoff in years (e.g., 5 for "before age 5")
  # age_cutoff <- 5
  # age_cutoff_lower_limit <- 0
  
  if(is.na(age_cutoff_date)){
  data2use <- data2use %>%
    dplyr::mutate(
    # Each child's age-based censoring date
    age_cutoff_date = lubridate::as_date(child_dob) + lubridate::years(age_cutoff)
  )
  } else {
    data2use <- data2use %>%
        dplyr::mutate(
          # Each child's age-based censoring date
          age_cutoff_date = as.Date(age_cutoff_date)
          )
  }
  
  data2use <- data2use %>%
  dplyr::mutate(
    # Each child's age-based censoring date
    # age_cutoff_date = lubridate::as_date(child_dob) + lubridate::years(age_cutoff),

    # Exit date = minimum of event date and age cutoff date
    exit_date = lubridate::as_date(
      pmin(
        ifelse(!is.na(INDATUMA), INDATUMA, age_cutoff_date),
        age_cutoff_date,
        na.rm = TRUE
      )
    ),

    # Age at exit in years
    exit_age = lubridate::time_length(exit_date - child_dob, "years"),
    # Event indicator: T1D occurred before age cutoff
    event = ifelse(!is.na(INDATUMA) & INDATUMA <= age_cutoff_date, 1, 0)
  )

  data2use <- data2use %>%
  dplyr::mutate(exit_age = exit_age - age_cutoff_lower_limit) %>%
  dplyr::filter(exit_age > 0)

  return(data2use)

}
```


read/load vitamin_d_case_control_plus_abis for analysis
```{r}
vitamin_d_case_control_plus_abis <- read.table(paste0(data_root, "/vitamin_d/vitamin_d_case_control_plus_abis.csv"))
```


```{r}
vitamin_d_case_control_plus_abis <- vitamin_d_case_control_plus_abis %>%
  dplyr::mutate(
    child_dob = lubridate::as_date(child_dob),
    INDATUMA = lubridate::as_date(INDATUMA),
  )

for(c in names(vitamin_d_case_control_plus_abis)){
  curr_class <- class(vitamin_d_case_control_plus_abis[[c]])
  if(curr_class != "Date" & curr_class == "character"){
    vitamin_d_case_control_plus_abis[[c]] <- factor(vitamin_d_case_control_plus_abis[[c]])
  }
}
```

reorder factors
```{r}
vitamin_d_case_control_plus_abis$T1D_heredity_cat <- factor(
  vitamin_d_case_control_plus_abis$First_degree_T1D_heredity_cat,
  levels = c("No", "Yes")
)

vitamin_d_case_control_plus_abis$Smoking_during_pregnancy <- factor(
  vitamin_d_case_control_plus_abis$Smoking_during_pregnancy,
  levels = c("No", "Yes")
)

vitamin_d_case_control_plus_abis$Parity <- factor(
  vitamin_d_case_control_plus_abis$Parity,
  levels = c("First parity", "Previous parity")
)

vitamin_d_case_control_plus_abis$Preterm_dichotomous <- factor(
  vitamin_d_case_control_plus_abis$Preterm_dichotomous,
  levels = c("Normal", "Preterm")
)

vitamin_d_case_control_plus_abis$Season <- factor(
  vitamin_d_case_control_plus_abis$Season,
  levels = c("Autumn", "Spring", "Summer", "Winter")
)

vitamin_d_case_control_plus_abis$Sex_cat <- factor(
  vitamin_d_case_control_plus_abis$Sex_cat,
  levels = c("Boy", "Girl")
)

vitamin_d_case_control_plus_abis$Fish_any_pregnancy <- factor(
  vitamin_d_case_control_plus_abis$Fish_any_pregnancy,
  levels = c("Seldom", "1–2 times week or more"))

vitamin_d_case_control_plus_abis <- vitamin_d_case_control_plus_abis %>%
  dplyr::mutate(
    Delivery_mode = ifelse(
      is.na(Delivery_mode),
      NA_character_,
      ifelse(
        grepl("caesar", Delivery_mode, ignore.case = TRUE),
        "Caesarean",
        "Vaginal"
      )
    ),
    Delivery_mode = factor(Delivery_mode, levels = c("Vaginal", "Caesarean"))
  )
```


```{r}
qs_birth_weight <- round(
  quantile(
    vitamin_d_case_control_plus_abis$Birth_weight, 
    probs = seq(0, 1, 0.25), 
    na.rm = TRUE)
  ,
  2)

vitamin_d_case_control_plus_abis <- vitamin_d_case_control_plus_abis %>%
    dplyr::mutate(
        Birth_weight_cat = cut(
            Birth_weight,
            breaks = qs_birth_weight,
            include.lowest = TRUE,
            labels = c(
                paste0("≤", qs_birth_weight[2]),                                    
                paste0(qs_birth_weight[2], "–", qs_birth_weight[3]),                        
                paste0(qs_birth_weight[3], "–", qs_birth_weight[4]),                           
                paste0(">", qs_birth_weight[4])                                     
            )
        )
    )


```


Weighted Cox analysis

conducting a time-to-event Cox regression with:

 - Time scale: Age (from birth)

 - Exposure: Fixed (cord blood ferritin)

 - Follow-up ends: December 31, 2020

Outcome: T1D diagnosis
```{r event_time, warning=FALSE}
vitamin_d_case_control_plus_abis$start_age <- 0  # since follow-up starts at birth

vitamin_d_case_control_plus_abis$INDATUMA <- as.Date(vitamin_d_case_control_plus_abis$INDATUMA, format = "%Y-%m-%d")
```

```{r cox_data, warning=FALSE}
df_cox_full <- vitamin_d_case_control_plus_abis
```

```{r}
child_covariates <- c("Preterm_dichotomous","Birth_weight_cat", "Season", "Sex_cat")

maternal_covariates <- c(
  "T1D_heredity_cat",
  "Smoking_during_pregnancy",
  "Fish_any_pregnancy",
  "Parity",
  "Delivery_mode",
  "Vitamin_mineral_supplement_pregnancy"
)
```

```{r remove_missing_rows}
df_cox_full <- df_cox_full %>%
  dplyr::select(all_of(
    c(
      "ABISnr",
      "diagnosis",
      "conc_25OHD",
      "DBP",
      child_covariates,
      maternal_covariates,
      "INDATUMA"
    )
  )) %>%
  dplyr::mutate(ABISnr = as.character(ABISnr)) %>%
  dplyr::left_join(
    abis %>% dplyr::select(ABISnr,child_dob),
    by = "ABISnr"
  ) %>%
  dplyr::mutate(conc_25OHD = conc_25OHD/10.0)
```



# =========================================================
# DEFINE WINDOWS
# =========================================================
```{r}
window_definitions <- tibble::tribble(
  ~window_name, ~lower, ~upper, ~admin_end,

  "  0-5 years",   0,  5,  NA,
  " 5-10 years",  5, 10,  NA,
  "10-15 years", 10, 15,  NA,
  "  15+ years",  15, NA, "2020-12-31",
  " 0-12 years", 0, 12,  NA,
  "  12+ years",  12, NA, "2020-12-31"
  )
```

```{r}
# =========================================================
# MAIN FIRTH PENALIZED COX FUNCTION
# (Weights dropped - see methods for justification)
# =========================================================
# run_window_analysis <- function(
#     full_df,
#     subset_df,
#     exposure,
#     covars,
#     window_name,
#     lower,
#     upper = NULL,
#     admin_end = NULL
# ){
# 
#   # ---------------------------------------------
#   # CREATE WINDOWED DATASETS
#   # ---------------------------------------------
#   full_window <- event_time_formulation(
#     data2use = full_df,
#     age_cutoff = upper,
#     age_cutoff_lower_limit = lower,
#     age_cutoff_date = admin_end
#   )
#   subset_window <- event_time_formulation(
#     data2use = subset_df,
#     age_cutoff = upper,
#     age_cutoff_lower_limit = lower,
#     age_cutoff_date = admin_end
#   )
#   # ---------------------------------------------
#   # WINDOW-SPECIFIC SAMPLING FRACTIONS
#   # (Retained for descriptive reporting)
#   # ---------------------------------------------
#   n_cases <-
#     sum(subset_window$event == 1, na.rm = TRUE)
# 
#   n_censored <-
#     sum(
#       subset_window$event == 0 &
#         subset_window$exit_date < subset_window$age_cutoff_date,
#       na.rm = TRUE
#     )
# 
#   n_full_cases    <- sum(full_window$event == 1, na.rm = TRUE)
#   n_full_controls <- sum(full_window$event == 0, na.rm = TRUE)
#   n_subset_cases    <- sum(subset_window$event == 1, na.rm = TRUE)
#   n_subset_controls <- sum(subset_window$event == 0, na.rm = TRUE)
# 
#   f_cases    <- n_subset_cases    / n_full_cases
#   f_controls <- n_subset_controls / n_full_controls
# 
#   # ---------------------------------------------
#   # MODEL FORMULA
#   # ---------------------------------------------
#   formula_txt <- paste0(
#     "Surv(exit_age, event) ~ ",
#     exposure,
#     " + ",
#     paste(covars, collapse = " + ")
#   )
#   formula_obj <- as.formula(formula_txt)
# 
#   # ---------------------------------------------
#   # FIT FIRTH PENALIZED COX MODEL
#   # coxphf does not support weights;
#   # weighting scheme dropped - see methods
#   # ---------------------------------------------
#   model_vars <- c("exit_age", "event", exposure, covars)
# 
#   subset_window_complete <- subset_window %>%
#     tidyr::drop_na(dplyr::all_of(model_vars))
# 
#   n_dropped <- nrow(subset_window) - nrow(subset_window_complete)
#   if (n_dropped > 0) {
#     warning(paste0(
#       "[", window_name, "] ", n_dropped,
#       " row(s) dropped due to missing values before coxphf fit."
#     ))
#   }
# 
#   fit <- coxphf::coxphf(
#     formula = formula_obj,
#     data    = subset_window_complete
#   )
# 
#   subset_window <- subset_window_complete
#   # ---------------------------------------------
#   # EXTRACT RESULTS VIA summary()
#   # broom::tidy() does not correctly parse
#   # coxphf objects - use summary() directly
#   # ---------------------------------------------
#   s <- summary(fit)
#   # Locate the row corresponding to the exposure term
#   coef_names <- names(s$coefficients)           # named numeric vector
#   exposure_idx <- which(coef_names == exposure)
# 
#   if (length(exposure_idx) == 0) {
#     stop(paste("Exposure term '", exposure, "' not found in model output.", sep = ""))
#   }
# 
#   res <- tibble::tibble(
#     term      = exposure,
#     estimate  = exp(s$coefficients[exposure_idx]),  # already exponentiated
#     conf.low  = s$ci.lower[exposure_idx],               # already on HR scale
#     conf.high = s$ci.upper[exposure_idx],               # already on HR scale
#     p.value   = s$prob[exposure_idx],                   # Firth p-value
#     window    = window_name,
#     N                = nrow(subset_window),
#     Cases            = n_cases,
#     Censored         = n_censored,
#     PY               = sum(subset_window$exit_age),
#     Full_cases       = n_full_cases,
#     Full_controls    = n_full_controls,
#     Sampled_cases    = n_subset_cases,
#     Sampled_controls = n_subset_controls,
#     f_cases          = f_cases,
#     f_controls       = f_controls
#   )
# 
#   return(res)
# }
```

# =========================================================
# MAIN COX FUNCTION
# =========================================================
```{r}
run_window_analysis <- function(
    full_df,
    subset_df,
    exposure,
    covars,
    window_name,
    lower,
    upper = NULL,
    admin_end = NULL
){


  # ---------------------------------------------
  # CREATE WINDOWED DATASETS
  # ---------------------------------------------

  full_window <- event_time_formulation(
    data2use = full_df,
    age_cutoff = upper,
    age_cutoff_lower_limit = lower,
    age_cutoff_date = admin_end
  )

  subset_window <- event_time_formulation(
    data2use = subset_df,
    age_cutoff = upper,
    age_cutoff_lower_limit = lower,
    age_cutoff_date = admin_end
  )

  # ---------------------------------------------
  # WINDOW-SPECIFIC SAMPLING FRACTIONS
  # IMPORTANT
  # ---------------------------------------------
  n_cases <-
    sum(subset_window$event == 1, na.rm = TRUE)

  n_censored <-
    sum(
      subset_window$event == 0 &
      subset_window$exit_date < subset_window$age_cutoff_date,
      na.rm = TRUE
    )


  n_full_cases <-
    sum(full_window$event == 1, na.rm = TRUE)

  n_full_controls <-
    sum(full_window$event == 0, na.rm = TRUE)

  n_subset_cases <-
    sum(subset_window$event == 1, na.rm = TRUE)

  n_subset_controls <-
    sum(subset_window$event == 0, na.rm = TRUE)


  f_cases <- n_subset_cases / n_full_cases
  f_controls <- n_subset_controls / n_full_controls

  # ---------------------------------------------
  # DEFINE WEIGHTS
  # ---------------------------------------------

  subset_window <- subset_window %>%
    dplyr::mutate(
      ipw =
        ifelse(
          event == 1,
          1 / f_cases,
          1 / f_controls
        )
    )

  # ---------------------------------------------
  # MODEL FORMULA
  # ---------------------------------------------

  formula_txt <- paste0(
    "Surv(exit_age, event) ~ ",
    exposure,
    " + ",
    paste(covars, collapse = " + ")
  )

  formula_obj <- as.formula(formula_txt)

  # print(format(formula_obj))
  # ---------------------------------------------
  # FIT MODEL
  # ---------------------------------------------

  # fit <- coxph(
  #   formula_obj,
  #   data = subset_window,
  #   weights = ipw,
  #   robust = TRUE
  # )

    # 1. Create a variable to store the note
  convergence_note <- NA_character_
  
  # 2. Run the model inside withCallingHandlers
  fit <- withCallingHandlers(
          coxph(
          formula_obj,
          data = subset_window,
          weights = ipw,
          robust = TRUE
        ),
    
        warning = function(w) {
          if (grepl("Loglik converged before variable", w$message)) {
            convergence_note <<- "convergence problem"
          }
          invokeRestart("muffleWarning") 
        }
  )
  

  # ---------------------------------------------
  # EXTRACT RESULTS
  # ---------------------------------------------

  res <- broom::tidy(
    fit,
    exponentiate = TRUE,
    conf.int = TRUE
  ) %>%
    dplyr::filter(term == exposure) %>%
    dplyr::mutate(

      window = window_name,

      N = nrow(subset_window),

      Cases = n_cases,
      Censored = n_censored,
      PY = sum(subset_window$exit_age),


      Full_cases = n_full_cases,
      Full_controls = n_full_controls,

      Sampled_cases = n_subset_cases,
      Sampled_controls = n_subset_controls,

      f_cases = f_cases,
      f_controls = f_controls,
      
      Note = convergence_note
    )

  return(res)
}
```


# =========================================================
# DEFINE EXPOSURES
# =========================================================
```{r}
exposures <- c(
      "conc_25OHD",
      "DBP"
)
```


# =========================================================
# DEFINE COVARIATES
# =========================================================
```{r}
base_covars <- c(
  "T1D_heredity_cat",
  "Smoking_during_pregnancy",
  "Fish_any_pregnancy",
  "Parity",
  "Delivery_mode",
  "Vitamin_mineral_supplement_pregnancy",
  "Preterm_dichotomous",
  "Birth_weight_cat", 
  "Season", 
  "Sex_cat"
)
```


# =========================================================
# RUN ALL WINDOW ANALYSES
# =========================================================
```{r}
all_results <- purrr::cross2(
  exposures,
  seq_len(nrow(window_definitions))
) %>%

  purrr::map_df(function(x){

    exposure_i <- x[[1]]

    window_i <- window_definitions[x[[2]], ]

    run_window_analysis(

      full_df = abis,

      subset_df = df_cox_full,

      exposure = exposure_i,

      covars = base_covars,

      window_name = window_i$window_name,

      lower = window_i$lower,

      upper = window_i$upper,

      admin_end = window_i$admin_end
    )

  })


```

# =========================================================
# FINAL TABLE
# =========================================================
```{r}
results_table <- all_results %>%

  dplyr::mutate(

    HR_CI =
      sprintf(
        "%.2f (%.2f–%.2f)",
        estimate,
        conf.low,
        conf.high
      ),
    f_cases = round(f_cases,3),
    f_controls = round(f_controls,3),
    
    p.value =
      format.pval(
        p.value,
        digits = 2,
        eps = 0.001
      )
  ) %>%

  dplyr::select(
    Exposure = term,
    Window = window,
    N,
    Cases,
    HR_CI,
    p.value,
    f_cases,
    f_controls,
    Censored,
    PY,
    Note
  )


```

# =========================================================
# OPTIONAL WIDE FORMAT TABLE
# =========================================================
```{r}
results_wide <- results_table %>%
  dplyr::select(
    Exposure,
    Window,
    HR_CI
  ) %>%
  tidyr::pivot_wider(
    names_from = Window,
    values_from = HR_CI
  )

```


# =========================================================
# VIEW RESULTS
# =========================================================
```{r}
# print(results_table)
# print(results_wide)
```
rearrange results remove duplicate row
```{r}
results_table_reformat <- results_table %>%
  dplyr::select(-Censored)%>%
  dplyr::arrange(Exposure,Window) %>%
  dplyr::mutate(
    Exposure = case_when(
      Exposure == "conc_25OHD" ~ "conc 25OHD",
      TRUE ~ Exposure)
  )

results_table_reformat$Exposure <- factor(
  results_table_reformat$Exposure,
  levels = c(
      "conc 25OHD",
      "DBP"
    )
)

results_table_reformat$Window <- factor(
  results_table_reformat$Window,
  levels = c(
    "  0-5 years",
    " 5-10 years",
    "10-15 years",
    "  15+ years",
    " 0-12 years",
    "  12+ years"
    )
)

results_table_reformat <- results_table_reformat %>% 
  dplyr::arrange(Exposure,Window)

df_clean <- results_table_reformat %>%
  dplyr::filter(row_number() == 1 | if_any(everything(), ~ .x != lag(.x)))
```

```{r}
saveRDS(df_clean, file.path(output_dir,"landmark_results_case_control_sampling_frac_cox.RDS"))
```

```{r}
chronological_df <- df_clean %>% 
  dplyr::filter(!(Window %in% c(" 0-12 years", "  12+ years")))

chronological_df_header <- data.frame(
    Exposure = "Chronological",
    Window = as.factor(NA_character_),
    N = NA_integer_,
    Cases = NA_integer_,
    HR_CI = NA_character_,
    p.value =NA_character_,
    f_cases = NA_real_,
    f_controls = NA_real_,
    PY = NA_real_,
    Note = NA_character_
    )

puberty_split_df <- df_clean %>% 
  dplyr::filter(Window %in% c(" 0-12 years", "  12+ years"))

puberty_split_df_header <- data.frame(
    Exposure = "Puberty-Category",
    Window = as.factor(NA_character_),
    N = NA_integer_,
    Cases = NA_integer_,
    HR_CI = NA_character_,
    p.value =NA_character_,
    f_cases = NA_real_,
    f_controls = NA_real_,
    PY = NA_real_,
    Note = NA_character_
    )
```


flextable
```{r}
results_table_flex <- bind_rows(
  chronological_df_header,
  chronological_df,
  puberty_split_df_header,
  puberty_split_df
) %>%
  dplyr::mutate(
    Exposure = if_else(
      row_number() != 1 &
        (!is.na(lag(Exposure, 1)) &
           Exposure == lag(Exposure, 1)),
      NA_character_,
      Exposure
    ),
    Window = if_else(
      row_number() != 1 &
        (!is.na(lag(Window, 1)) &
           Window == lag(Window, 1)),
      NA_character_,
      Window
    )
  ) %>%
  dplyr::rename(
    `HR (95% CI)` = HR_CI,
    `P-value` = p.value,
    `f (cases)` = f_cases,
    `f (controls)` = f_controls
    ) %>%
  dplyr::mutate(
    Window = dplyr::case_when(
      grepl("problem", Note, ignore.case = TRUE) ~ paste0(Window,"\U2021"),
      TRUE ~ Window
    )
  ) %>%
  dplyr::select(-Note) %>%
  flextable() %>%
  flextable::add_footer_lines(
    values = c(
      "\U2021 Converge Problem"
    )
  ) %>%
  flextable::autofit() %>%
  flextable::align(j = 2, align = "right", part = "body")
```


```{r save_merged_all_model_results}
flextable::save_as_docx(
  `Weighted Cox model: landmark analysis` = results_table_flex,
  path = paste0(output_dir,"/flex_merged_all_weighted_cox_results_landmark_analysis.docx"),
  pr_section = sect_properties
)


saveRDS(
  object = results_table_flex,
  file = paste0(output_dir,"/flex_merged_all_weighted_cox_results_landmark_analysis.docx.RDS")
)
```



