---
title: "Subgroup Analysis - Weighted Cox model: conc_25OHD and DBP"
author: "Debojyoti Das"
date: "2026-06-01"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
```

set output directory
```{r output_dir, message=FALSE, warning=FALSE}
output_dir <- "/Users/debda22/Projects/abis-vitamin-d-type1-diabetes/results/revision/case_and_control_weighted_cox/subgroup_analysis"

if(!dir.exists(output_dir)){
  dir.create(output_dir, recursive = TRUE)
} else {
  cat(paste0(output_dir," is already present!!\n"))
}
```

load libraries
```{r load_libraries, message=FALSE, warning=FALSE}
pacman::p_load(
  multipleNCC,
  haven,        
  zoo,
  sjlabelled,
  janitor,       
  data.table,   
  ggplot2,
  comprehenr,  
  stringr,     
  rstatix,
  dplyr,
  tidyr,
  forcats,
  officer,
  cowplot,
  ggExtra,
  ggpubr,
  GGally,
  janitor,
  ggrepel,
  ggtext,
  parameters,
  lme4,
  broom,
  pscl,
  tibble,
  knitr,
  ggeffects,
  cobalt,
  WeightIt,
  marginaleffects,
  survival,
  survminer,
  splines,
  purrr,
  nnet,
  magick,
  patchwork,
  flextable,
  aod
  )
```


```{r word_sect_properties}
sect_properties <- prop_section(
  page_size = page_size(
    orient = "landscape",
    height = 11.7,
    width = 8.3
  ),
  page_margins = page_mar(
    top = 0.5,
    bottom = 0.5,
    left = 0.5,
    right = 0.5,
    header = 0.3,
    footer = 0.3,
    gutter = 0
  ),
  type = "continuous"
)

set_flextable_defaults(
    padding.top = 0, 
    padding.bottom = 0,
    line_spacing = 1
)
```


define paths
```{r define_paths, eval=TRUE}
data_root <- "/Users/debda22/Projects/abis-data"
proj_root <- "/Users/debda22/Projects/abis-vitamin-d-type1-diabetes"
```

load parameters and helper functions
```{r}
source(paste0(proj_root,"/rscripts/all_parameters_case_control_T1D_vitamin_D.R"))
source(paste0(proj_root,"/rscripts/weighted_cox_T1D_helper.R"))
```

```{r}
abis <- readRDS(file.path(data_root,"vitamin_d","covariates.abis.full.cohort.RDS"))
```

```{r}
print("ABIS count:")
print(abis %>% count(par_b))
```

```{r}
abis <- abis %>%
  dplyr::filter(!is.na(child_dob))

print("number of participants with DOB:")
print(abis %>% nrow())
print("number of participants in birth questionnaire:")
print(abis %>% count(par_b))
```


```{r}
# Set exit date based on whether T1D occurred before 2020-12-31
abis$exit_date <- as.Date(ifelse(
  !is.na(abis$INDATUMA) & abis$INDATUMA <= as.Date("2020-12-31"),
  abis$INDATUMA,
  as.Date("2020-12-31")
))

# Calculate age at exit in years
abis$exit_age <- as.numeric(
  difftime(
    abis$exit_date,
    abis$child_dob,
    units = "days"
  )
) / 365.25

abis$event <- ifelse(!is.na(abis$INDATUMA) & abis$INDATUMA <= as.Date("2020-12-31"),1,0)
```

```{r}
print("ABIS - T1D count with birth questionnaire:")
print(abis %>% count(par_b, event))
```


read/load vitamin_d_case_control_plus_abis for analysis
```{r}
vitamin_d_case_control_plus_abis <- read.table(paste0(data_root, "/vitamin_d/vitamin_d_case_control_plus_abis.csv"))
```


```{r}
vitamin_d_case_control_plus_abis <- vitamin_d_case_control_plus_abis %>%
  dplyr::mutate(
    child_dob = lubridate::as_date(child_dob),
    INDATUMA = lubridate::as_date(INDATUMA),
  )

for(c in names(vitamin_d_case_control_plus_abis)){
  curr_class <- class(vitamin_d_case_control_plus_abis[[c]])
  if(curr_class != "Date" & curr_class == "character"){
    vitamin_d_case_control_plus_abis[[c]] <- factor(vitamin_d_case_control_plus_abis[[c]])
  }
}
```

reorder factors
```{r}
vitamin_d_case_control_plus_abis$T1D_heredity_cat <- factor(
  vitamin_d_case_control_plus_abis$First_degree_T1D_heredity_cat,
  levels = c("No", "Yes")
)

vitamin_d_case_control_plus_abis$Smoking_during_pregnancy <- factor(
  vitamin_d_case_control_plus_abis$Smoking_during_pregnancy,
  levels = c("No", "Yes")
)

vitamin_d_case_control_plus_abis$Parity <- factor(
  vitamin_d_case_control_plus_abis$Parity,
  levels = c("First parity", "Previous parity")
)

vitamin_d_case_control_plus_abis$Preterm_dichotomous <- factor(
  vitamin_d_case_control_plus_abis$Preterm_dichotomous,
  levels = c("Normal", "Preterm")
)

vitamin_d_case_control_plus_abis$Season <- factor(
  vitamin_d_case_control_plus_abis$Season,
  levels = c("Autumn", "Spring", "Summer", "Winter")
)

vitamin_d_case_control_plus_abis$Sex_cat <- factor(
  vitamin_d_case_control_plus_abis$Sex_cat,
  levels = c("Boy", "Girl")
)

vitamin_d_case_control_plus_abis$Fish_any_pregnancy <- factor(
  vitamin_d_case_control_plus_abis$Fish_any_pregnancy,
  levels = c("Seldom", "1–2 times week or more"))

vitamin_d_case_control_plus_abis <- vitamin_d_case_control_plus_abis %>%
  dplyr::mutate(
    Delivery_mode = ifelse(
      is.na(Delivery_mode),
      NA_character_,
      ifelse(
        grepl("caesar", Delivery_mode, ignore.case = TRUE),
        "Caesarean",
        "Vaginal"
      )
    ),
    Delivery_mode = factor(Delivery_mode, levels = c("Vaginal", "Caesarean"))
  )
```

```{r}
qs_birth_weight <- round(
  quantile(
    vitamin_d_case_control_plus_abis$Birth_weight, 
    probs = seq(0, 1, 0.25), 
    na.rm = TRUE)
  ,
  2)

vitamin_d_case_control_plus_abis <- vitamin_d_case_control_plus_abis %>%
    dplyr::mutate(
        Birth_weight_cat = cut(
            Birth_weight,
            breaks = qs_birth_weight,
            include.lowest = TRUE,
            labels = c(
                paste0("≤", qs_birth_weight[2]),                                    
                paste0(qs_birth_weight[2], "–", qs_birth_weight[3]),                        
                paste0(qs_birth_weight[3], "–", qs_birth_weight[4]),                           
                paste0(">", qs_birth_weight[4])                                     
            )
        )
    )


```



Weighted Cox analysis

conducting a time-to-event Cox regression with:

 - Time scale: Age (from birth)

 - Exposure: Fixed (cord blood ferritin)

 - Follow-up ends: December 31, 2020

Outcome: T1D diagnosis
```{r event_time, warning=FALSE}
vitamin_d_case_control_plus_abis$start_age <- 0  # since follow-up starts at birth

vitamin_d_case_control_plus_abis$INDATUMA <- as.Date(vitamin_d_case_control_plus_abis$INDATUMA, format = "%Y-%m-%d")
```

```{r}
event_time_formulation <- function(data2use, age_cutoff = 5, age_cutoff_lower_limit = 0, age_cutoff_date = as.Date(NA)){

  # Define the age cutoff in years (e.g., 5 for "before age 5")
  # age_cutoff <- 5
  # age_cutoff_lower_limit <- 0
  
  if(is.na(age_cutoff_date)){
  data2use <- data2use %>%
    dplyr::mutate(
    # Each child's age-based censoring date
    age_cutoff_date = lubridate::as_date(child_dob) + lubridate::years(age_cutoff)
  )
  } else {
    data2use <- data2use %>%
        dplyr::mutate(
          # Each child's age-based censoring date
          age_cutoff_date = as.Date(age_cutoff_date)
          )
  }
  
  data2use <- data2use %>%
  dplyr::mutate(
    # Each child's age-based censoring date
    # age_cutoff_date = lubridate::as_date(child_dob) + lubridate::years(age_cutoff),

    # Exit date = minimum of event date and age cutoff date
    exit_date = lubridate::as_date(
      pmin(
        ifelse(!is.na(INDATUMA), INDATUMA, age_cutoff_date),
        age_cutoff_date,
        na.rm = TRUE
      )
    ),

    # Age at exit in years
    exit_age = lubridate::time_length(exit_date - child_dob, "years"),
    # Event indicator: T1D occurred before age cutoff
    event = ifelse(!is.na(INDATUMA) & INDATUMA <= age_cutoff_date, 1, 0)
  )

  data2use <- data2use %>%
  dplyr::mutate(exit_age = exit_age - age_cutoff_lower_limit) %>%
  dplyr::filter(exit_age > 0)

  return(data2use)

}
```


```{r cox_data, warning=FALSE}
df_cox_full <- vitamin_d_case_control_plus_abis
```


```{r cox_data, warning=FALSE}
df_cox_full <- event_time_formulation(
df_cox_full, 
age_cutoff_lower_limit = 0, 
age_cutoff_date = as.Date("2020-12-31")
)

print(df_cox_full %>% dplyr::count(event))
```

Case–control weighted Cox (using time-to-event):

f <- n_subcohort_control / n_full_cohort_control
f_cases <- n_subcohort_cases / n_full_cohort_cases

Due to financial and practical constraints, cord blood samples were analyzed in 
a randomly selected subset of cohort participants, including 80 of 160 incident
cases and 300 of 16,000 non-cases. Because biomarker measurements were available
only for a stratified random subset defined by outcome status, analyses were 
performed using inverse probability weighted Cox proportional hazards regression. 
Sampling probabilities were estimated separately for cases and non-cases, and 
inverse sampling fractions were used as weights to obtain unbiased hazard ratio 
estimates representative of the full cohort, following methods for two-phase 
outcome-dependent sampling described by Samuelsen (1997) and Borgan et al. (2000).
```{r}
f <- nrow(df_cox_full %>% dplyr::filter(event==0))/ nrow(abis %>% dplyr::filter(event==0))
f_cases <- nrow(df_cox_full %>% dplyr::filter(event==1))/ nrow(abis %>% dplyr::filter(event==1))
```

```{r}
print("Sampling fractions")

print("controls")
print(paste0(nrow(df_cox_full %>% dplyr::filter(event==0)),"/",  nrow(abis %>% dplyr::filter(event==0)), " = ",f))

print("cases")
print(paste0(nrow(df_cox_full %>% dplyr::filter(event==1)),"/",  nrow(abis %>% dplyr::filter(event==1)), " = ",f_cases))

```

```{r}
child_covariates <- c("Preterm_dichotomous","Birth_weight_cat", "Season", "Sex_cat")

maternal_covariates <- c(
  "T1D_heredity_cat",
  "Smoking_during_pregnancy",
  "Fish_any_pregnancy",
  "Parity",
  "Delivery_mode",
  "Vitamin_mineral_supplement_pregnancy"
)
```

```{r remove_missing_rows}
df_cox_full <- df_cox_full %>%
  dplyr::select(all_of(
    c(
      "ABISnr",
      "diagnosis",
      "conc_25OHD",
      "DBP",
      child_covariates,
      maternal_covariates,
      "INDATUMA",
      "event",
      "exit_age"
    )
  )) %>%
  dplyr::mutate(ABISnr = as.character(ABISnr)) %>%
  dplyr::left_join(
    abis %>% dplyr::select(ABISnr,child_dob),
    by = "ABISnr"
  ) %>%
  dplyr::mutate(conc_25OHD = conc_25OHD/10.0)
```


Weighted Cox regression	Hazard ratio: relative rate of developing T1D at any given age. 

```{r}
df_cox_full <- df_cox_full %>%
  dplyr::mutate(
    ipw = if_else(event == 1, 1 / f_cases, 1 / f)
  )
```

```{r}
base_covars <- c(
  "T1D_heredity_cat",
  "Smoking_during_pregnancy",
  "Fish_any_pregnancy",
  "Parity",
  "Delivery_mode",
  "Vitamin_mineral_supplement_pregnancy",
  "Preterm_dichotomous",
  "Birth_weight_cat", 
  "Season", 
  "Sex_cat"
)
```

```{r}
run_interaction_analysis <- function(
    data,
    exposure,
    subgroup,
    covars,
    time = "exit_age",
    event = "event"
) {
    # -----------------------------
  # 1. Helper Function to Tidy coxphf Objects
  # -----------------------------
  # tidy_coxphf <- function(model, exp_term) {
  #   # Extract names of variables
  #   terms <- names(model$coefficients)
  #   idx <- which(terms == exp_term)
  #   
  #   if (length(idx) == 0) {
  #     return(data.frame(
  #       term = exp_term, estimate = NA_real_, std.error = NA_real_,
  #       statistic = NA_real_, p.value = NA_real_, conf.low = NA_real_, conf.high = NA_real_
  #     ))
  #   }
  #   
  #   data.frame(
  #     term      = exp_term,
  #     estimate  = exp(model$coefficients[idx]),
  #     std.error = sqrt(model$var[idx, idx]),
  #     statistic = model$prob[idx], # coxphf provides profile likelihood test p-values directly
  #     p.value   = model$prob[idx],
  #     conf.low  = model$ci.lower[idx],
  #     conf.high = model$ci.upper[idx]
  #   )
  # }
    # -----------------------------
  # 2. Unadjusted Overall Model (New)
  # -----------------------------
  formula_unadj <- as.formula(paste0("Surv(", time, ", ", event, ") ~ ", exposure))
  
  # Note: Weights are removed here for a purely crude/unadjusted estimate. 
  # Add weights = ipw back if your study design requires weighted unadj models.
  model_unadj <- coxph(
    formula_unadj,
    data = data,
    weights = ipw,
    robust = TRUE
    )

  unadj_result <- broom::tidy(model_unadj, exponentiate = TRUE, conf.int = TRUE) %>%
    dplyr::filter(term == exposure) %>%
    dplyr::mutate(
      subgroup = "Overall",
      level = "Unadjusted",
      N = nrow(data),
      Cases = sum(data[[event]] == 1),
      p_interaction = NA_real_,
      Note = NA_character_
    )

  all_vars <- c(time, event, exposure, subgroup, covars)
  
  # -----------------------------
  # 1. Enforce Complete-Case Analysis Globally
  # -----------------------------
  # Correct syntax using if_all() to ensure no missing data across any analytical variables
  data <- data %>%
    dplyr::filter(if_all(all_of(all_vars), ~ !is.na(.))) %>%
    dplyr::mutate(!!subgroup := droplevels(as.factor(.data[[subgroup]])))

  # remove subgroup from covariates
  covars_sub <- setdiff(covars, subgroup)

  # -----------------------------
  # 1. Main adjusted model
  # -----------------------------

  formula_main <- as.formula(
    paste0(
      "Surv(", time, ", ", event, ") ~ ",
      exposure, " + ",
      paste(covars, collapse = " + ")
    )
  )

  
  model_main <- coxph(
    formula_main,
    data = data,
    weights = ipw,
    robust = TRUE
    )

  main_result <- broom::tidy(
    model_main,
    exponentiate = TRUE,
    conf.int = TRUE
  ) %>%
    dplyr::filter(term == exposure) %>%
    dplyr::mutate(
      subgroup = "Overall",
      level = "Adjusted",
      N = nrow(data),
      Cases = sum(data[[event]] == 1),
      p_interaction = NA_real_,
      Note = NA_character_
    )
  
  # -----------------------------
  # 2. Interaction model
  # -----------------------------
  # remove NA subgroup rows (important)
  data <- data %>%
    dplyr::filter(!is.na(.data[[subgroup]]))
  
  # 2. drop unused factor levels (important)
  data[[subgroup]] <- droplevels(data[[subgroup]])
  
  subgroup_levels <- unique(data[[subgroup]])

  # -----------------------------
  # Interaction model
  # -----------------------------
  formula_main_no_subgroup <- as.formula(
    paste0(
      "Surv(", time, ", ", event, ") ~ ",
      exposure, " + ",
      paste(covars_sub, collapse = " + ")
    )
  )
  
  formula_interaction <- as.formula(
      paste0(
        "Surv(", time, ", ", event, ") ~ ",
        exposure, "*", subgroup, " + ",
        paste(covars_sub, collapse = " + ")
      )
    )
  
  model_main_no_subgroup <- coxph(
    formula_main_no_subgroup,
    data = data,
    weights = ipw,
    robust = TRUE
  )

  convergence_note_interaction1 <- NA_character_
  convergence_note_interaction2 <- NA_character_
  
  model_interaction <- withCallingHandlers(
    coxph(
      formula_interaction,
      data = data,
      weights = ipw,
      robust = TRUE
    ),
    warning = function(w) {
      convergence_note_interaction1 <<- w$message 
      invokeRestart("muffleWarning")
    }
    )

  
  # Omnibus LRT p-value for interaction
  beta <- coef(model_interaction)
  interaction_indices <- grep(":", names(beta))
  interaction_test <- withCallingHandlers(
    wald.test(
      b = coef(model_interaction), 
      Sigma = vcov(model_interaction), 
      Terms = interaction_indices
      ),
    warning = function(w) {
      convergence_note_interaction2 <<- w$message 
      invokeRestart("muffleWarning")
    }
    )
  
  p_interaction <- interaction_test$result$chi2["P"]

  convergence_note_interaction <- ifelse((
    is.na(convergence_note_interaction1) &
      is.na(convergence_note_interaction2)
  ),
  NA_character_,
  ifelse(
    !is.na(convergence_note_interaction1) &
      !is.na(convergence_note_interaction2),
    paste0(
      convergence_note_interaction1,
      convergence_note_interaction2
    ),
    ifelse(
      !is.na(convergence_note_interaction1),
      convergence_note_interaction1,
      convergence_note_interaction2
    )
  ))
  # -----------------------------
  # Stratified subgroup models
  # -----------------------------

  subgroup_results <- map_df(
    subgroup_levels,
    function(level) {

      d <- data %>%
        dplyr::filter(.data[[subgroup]] == level)

      formula_sub <- as.formula(
        paste0(
          "Surv(", time, ", ", event, ") ~ ",
          exposure, " + ",
          paste(covars_sub, collapse = " + ")
        )
      )
  
    convergence_note <- NA_character_

    fit <- tryCatch({
        withCallingHandlers(
          coxph(
            formula_sub, 
            data = d, 
            weights = ipw, 
            robust = TRUE, 
            control = coxph.control(iter.max = 100)
            ),
          warning = function(w) {
            convergence_note <<- w$message  
            invokeRestart("muffleWarning")
          }
        )
      }, error = function(e) {
        convergence_note <<- paste0("Math Error: ", e$message)
        NULL
      })
    
          # If the math exploded, return an empty row populated with the error note
      if (is.null(fit)) {
        return(data.frame(
          term = exposure, 
          estimate = NA_real_, 
          std.error = NA_real_,
          robust.se = NA_real_,
          statistic = NA_real_, 
          p.value = NA_real_, 
          conf.low = NA_real_, 
          conf.high = NA_real_,
          subgroup = subgroup, 
          level = as.character(level), 
          N = nrow(d), 
          Cases = sum(d[[event]] == 1),
          p_interaction = NA_real_, 
          Note = convergence_note
        ))
      }

      broom::tidy(
        fit,
        exponentiate = TRUE,
        conf.int = TRUE
      ) %>%
        dplyr::filter(term == exposure) %>%
        dplyr::mutate(
          subgroup = subgroup,
          level = as.character(level),
          N = nrow(d),
          Cases = sum(d[[event]] == 1),
          p_interaction = NA_real_,
          Note = convergence_note
        )
    }
  )

    subgroup_interaction_results <- data.frame(
      term = exposure,
      estimate = NA_real_,
      std.error = NA_real_,
      robust.se = NA_real_,
      statistic = NA_real_,
      p.value = NA_real_,
      conf.low = NA_real_,
      conf.high = NA_real_,
      subgroup = subgroup,
      p_interaction = p_interaction,
      Note = convergence_note_interaction
    )

  dplyr::bind_rows(
    unadj_result,
    main_result,
    subgroup_interaction_results,
    subgroup_results
  ) %>%
    dplyr::select(
      exposure = term,
      subgroup,
      level,
      N,
      Cases,
      estimate,
      conf.low,
      conf.high,
      p.value,
      p_interaction,
      Note
    )
}
```

```{r}
exposures <- c(
      "conc_25OHD",
      "DBP"
)

subgroups <- c(
  "Vitamin_mineral_supplement_pregnancy",
  "Season",
  "Sex_cat",
  "T1D_heredity_cat",
  "Preterm_dichotomous"
)
```


```{r}
results <- tidyr::expand_grid(exposure = exposures, subgroup = subgroups) %>%
  purrr::pmap(
    ~run_interaction_analysis(
      data = df_cox_full,
      exposure = ..1,
      subgroup = ..2,
      covars = base_covars
    )
  )

```


```{r}
results_table <- results %>%
  dplyr::bind_rows() %>%
  dplyr::mutate(
    `HR (95% CI)` = ifelse(
      is.na(estimate),
      NA_character_,
      sprintf(
      "%.2f (%.2f–%.2f)",
      estimate,
      conf.low,
      conf.high
      )
    ),
    .before = 4 ) %>%
  dplyr::mutate(
    p.value = format.pval(p.value, digits = 2, eps = 0.001),
    `p-value interaction` =   ifelse(
      is.na(p_interaction), 
      NA_character_, 
      ifelse(
        p_interaction < 0.001,
        formatC(p_interaction, format = "e", digits = 2),
        formatC(p_interaction, format = "f", digits = 2)
      )
    )
  ) %>%
  dplyr::select(
    -all_of(c("estimate", "conf.low", "conf.high", "p.value", "p_interaction"))
    )
```

rearrange results remove duplicate row
```{r}
results_table_reformat <- results_table %>% 
  dplyr::arrange(exposure,subgroup,level) %>%
  dplyr::mutate(
    level = ifelse(is.na(level),"Interaction",level)
    ) %>%
  dplyr::mutate(
    level = case_when(
      level == "Boy" ~ "Male",
      level == "Girl" ~ "Female",
      TRUE ~ level
      ),
    exposure = case_when(
      exposure == "conc_25OHD" ~ "conc 25OHD",
      TRUE ~ exposure)
  )

results_table_reformat$exposure <- factor(
  results_table_reformat$exposure, 
  levels = c(
      "conc 25OHD",
      "DBP"
    )
  )

results_table_reformat$subgroup <- factor(
  results_table_reformat$subgroup, 
  levels = c(
    "Overall", 
    "Vitamin_mineral_supplement_pregnancy",
    "Season",
    "Sex_cat",
    "T1D_heredity_cat",
    "Preterm_dichotomous"
    )
  )

results_table_reformat$level <- factor(
  results_table_reformat$level, 
  levels = c(
    "Unadjusted", 
    "Adjusted", 
    "Interaction",
    "Autumn", 
    "Spring", 
    "Summer", 
    "Winter",
    "No", 
    "Yes",
    "Male", 
    "Female",
    "Normal",
    "Preterm"
    )
  )

results_table_reformat <- results_table_reformat %>% 
  dplyr::arrange(exposure,subgroup,level)

df_clean <- results_table_reformat %>%
  dplyr::filter(row_number() == 1 | if_any(everything(), ~ .x != lag(.x)))


```

flextable
```{r}
results_table_flex <- df_clean %>% 
   dplyr::mutate(
        subgroup = gsub("\\_"," ",gsub("\\_cat","",subgroup))
        ) %>%
    dplyr::mutate(
        exposure = if_else(row_number() != 1 &
                               (!is.na(lag(exposure,1)) & exposure == lag(exposure, 1)),
                           NA_character_,
                           exposure),
        subgroup = if_else(row_number() != 1 &
                               (!is.na(lag(subgroup,1)) & subgroup == lag(subgroup, 1)),
                           NA_character_,
                           subgroup)
        ) %>% 
  dplyr::mutate(
    level = dplyr::case_when(
      grepl("infinite argument in coxph.wtest", Note, ignore.case = TRUE) |
        (!(level %in% c(
          "Interaction", "Overall"
        )) & is.na(`HR (95% CI)`)) ~ paste0(level, "\U2021"),
      TRUE ~ level
    )
  ) %>%
  dplyr::rename(
    Exposure = exposure,
    Subgroup = subgroup,
    `level/Model` = level,
    ) %>%
  dplyr::select(-Note) %>%
  flextable::flextable() %>% 
  flextable::add_footer_lines(
    values = c(
      "\U2021 Did not converge"
    )
  ) %>%
  flextable::autofit()
```

```{r word_sect_properties}
sect_properties <- prop_section(
  page_size = page_size(
    orient = "landscape",
    height = 11.7,
    width = 8.3
  ),
  page_margins = page_mar(
    top = 0.5,
    bottom = 0.5,
    left = 0.5,
    right = 0.5,
    header = 0.3,
    footer = 0.3,
    gutter = 0
  ),
  type = "continuous"
)

set_flextable_defaults(
    padding.top = 0, 
    padding.bottom = 0,
    line_spacing = 1
)
```

```{r save_merged_all_model_results}
flextable::save_as_docx(
  `Weighted Cox model: subgroup analysis` = results_table_flex,
  path = paste0(output_dir,"/flex_merged_all_weighted_cox_results_subgroup_analysis.docx"),
  pr_section = sect_properties
)


saveRDS(
  object = results_table_flex,
  file = paste0(output_dir,"/flex_merged_all_weighted_cox_results_subgroup_analysis.docx.RDS")
)
```


