---
title: "Weighted Cox model: vitD"
author: "Debojyoti Das"
date: "2026-06-01"
output: html_document
---

```{r setup, include=FALSE}
rm(list = ls())
knitr::opts_chunk$set(echo = TRUE)
```

set output directory
```{r output_dir, message=FALSE, warning=FALSE}
output_dir <- "/Users/debda22/Projects/abis-vitamin-d-type1-diabetes/results/revision/case_and_control_weighted_cox/full"

if(!dir.exists(output_dir)){
  dir.create(output_dir, recursive = TRUE)
} else {
  cat(paste0(output_dir," is already present!!\n"))
}
```

load libraries
```{r load_libraries, message=FALSE, warning=FALSE}
pacman::p_load(
  multipleNCC,
  haven,        
  zoo,
  sjlabelled,
  janitor,       
  data.table,   
  ggplot2,
  comprehenr,  
  stringr,     
  rstatix,
  dplyr,
  tidyr,
  forcats,
  officer,
  cowplot,
  ggExtra,
  ggpubr,
  GGally,
  janitor,
  ggrepel,
  ggtext,
  parameters,
  lme4,
  broom,
  pscl,
  tibble,
  knitr,
  ggeffects,
  cobalt,
  WeightIt,
  marginaleffects,
  survival,
  survminer,
  splines,
  purrr,
  nnet,
  magick,
  patchwork,
  flextable
  )
```


```{r word_sect_properties}
sect_properties <- prop_section(
  page_size = page_size(
    orient = "landscape",
    height = 11.7,
    width = 8.3
  ),
  page_margins = page_mar(
    top = 0.5,
    bottom = 0.5,
    left = 0.5,
    right = 0.5,
    header = 0.3,
    footer = 0.3,
    gutter = 0
  ),
  type = "continuous"
)

set_flextable_defaults(
    padding.top = 0, 
    padding.bottom = 0,
    line_spacing = 1
)
```


define paths
```{r define_paths, eval=TRUE}
data_root <- "/Users/debda22/Projects/abis-data"
proj_root <- "/Users/debda22/Projects/abis-vitamin-d-type1-diabetes/"
```

load parameters and helper functions
```{r}
source(paste0(proj_root,"/rscripts/all_parameters_case_control_T1D_vitamin_D.R"))
source(paste0(proj_root,"/rscripts/weighted_cox_T1D_helper.R"))
```

```{r}
abis <- readRDS(file.path(data_root,"vitamin_d","covariates.abis.full.cohort.RDS"))
```

```{r}
print("ABIS count:")
print(abis %>% count(par_b))
```


```{r}
print("abis cohort")
print(abis %>% dplyr::count(event))

abis <- abis %>%
  dplyr::mutate(has_dob = ifelse(!is.na(child_dob),1,0)) %>%
  dplyr::filter(has_dob == 1)

print("abis cohort with birth questionnaire")
print(abis %>% dplyr::filter(!is.na(child_dob)) %>% dplyr::count(event))
```



```{r}
print("ABIS - T1D count with birth questionnaire:")
print(abis %>% count(par_b, event))
```


read/load vitamin_d_case_control_plus_abis for analysis
```{r}
vitamin_d_case_control_plus_abis <- read.table(paste0(data_root, "/vitamin_d/vitamin_d_case_control_plus_abis.csv"))
```

reorder factors
```{r}
vitamin_d_case_control_plus_abis$T1D_heredity_cat <- factor(
  vitamin_d_case_control_plus_abis$First_degree_T1D_heredity_cat,
  levels = c("No", "Yes")
)

vitamin_d_case_control_plus_abis$Smoking_during_pregnancy <- factor(
  vitamin_d_case_control_plus_abis$Smoking_during_pregnancy,
  levels = c("No", "Yes")
)

vitamin_d_case_control_plus_abis$Parity <- factor(
  vitamin_d_case_control_plus_abis$Parity,
  levels = c("First parity", "Previous parity")
)

vitamin_d_case_control_plus_abis$Preterm_dichotomous <- factor(
  vitamin_d_case_control_plus_abis$Preterm_dichotomous,
  levels = c("Normal", "Preterm")
)

vitamin_d_case_control_plus_abis$Season <- factor(
  vitamin_d_case_control_plus_abis$Season,
  levels = c("Autumn", "Spring", "Summer", "Winter")
)

vitamin_d_case_control_plus_abis$Sex_cat <- factor(
  vitamin_d_case_control_plus_abis$Sex_cat,
  levels = c("Boy", "Girl")
)

vitamin_d_case_control_plus_abis$Fish_any_pregnancy <- factor(
  vitamin_d_case_control_plus_abis$Fish_any_pregnancy,
  levels = c("Seldom", "1–2 times week or more"))


vitamin_d_case_control_plus_abis <- vitamin_d_case_control_plus_abis %>%
  dplyr::mutate(
    Delivery_mode = ifelse(
      is.na(Delivery_mode),
      NA_character_,
      ifelse(
        grepl("caesar", Delivery_mode, ignore.case = TRUE),
        "Caesarean",
        "Vaginal"
      )
    ),
    Delivery_mode = factor(Delivery_mode, levels = c("Vaginal", "Caesarean"))
  )


```


```{r}
qs_birth_weight <- round(
  quantile(
    vitamin_d_case_control_plus_abis$Birth_weight, 
    probs = seq(0, 1, 0.25), 
    na.rm = TRUE)
  ,
  2)

vitamin_d_case_control_plus_abis <- vitamin_d_case_control_plus_abis %>%
    dplyr::mutate(
        Birth_weight_cat = cut(
            Birth_weight,
            breaks = qs_birth_weight,
            include.lowest = TRUE,
            labels = c(
                paste0("≤", qs_birth_weight[2]),                                    
                paste0(qs_birth_weight[2], "–", qs_birth_weight[3]),                        
                paste0(qs_birth_weight[3], "–", qs_birth_weight[4]),                           
                paste0(">", qs_birth_weight[4])                                     
            )
        )
    )


```

Outcome: T1D diagnosis
```{r event_time, warning=FALSE}
vitamin_d_case_control_plus_abis$start_age <- 0  # since follow-up starts at birth

vitamin_d_case_control_plus_abis$INDATUMA <- as.Date(vitamin_d_case_control_plus_abis$INDATUMA, format = "%Y-%m-%d")

# Set exit date based on whether T1D occurred before 2020-12-31
vitamin_d_case_control_plus_abis$exit_date <- as.Date(ifelse(
  !is.na(vitamin_d_case_control_plus_abis$T1D_age_diagnosis) &
    vitamin_d_case_control_plus_abis$INDATUMA<= as.Date("2020-12-31"),
  vitamin_d_case_control_plus_abis$INDATUMA,
  as.Date("2020-12-31")
))
```


```{r event_time_calculation, warning=FALSE}
# Calculate age at exit in years
vitamin_d_case_control_plus_abis$exit_age <- as.numeric(
  difftime(
    vitamin_d_case_control_plus_abis$exit_date,
    vitamin_d_case_control_plus_abis$child_dob,
    units = "days"
  )
) / 365.25

vitamin_d_case_control_plus_abis$event <- ifelse(
  !is.na(vitamin_d_case_control_plus_abis$INDATUMA) &
    vitamin_d_case_control_plus_abis$INDATUMA <= as.Date("2020-12-31"),
  1,
  0
)
```

```{r}
df_cox_full <- vitamin_d_case_control_plus_abis %>%
  dplyr::select(all_of(
    c(
      "ABISnr",
      "diagnosis",
      "DBP",
      "conc_25OHD",
      "child_dob",
      "T1D_heredity_cat",
      "Smoking_during_pregnancy",
      "Fish_any_pregnancy",
      "Parity",
      "Delivery_mode",
      "Season",
      "Sex_cat",
      "Vitamin_mineral_supplement_pregnancy",
      "Preterm_dichotomous",
      "Birth_weight",
      "Birth_weight_cat",
      "INDATUMA",
      "start_age",
      "exit_date",
      "exit_age",
      "event"
    )
  ))
```


```{r}
df_cox_full <- df_cox_full %>%
  dplyr::mutate(
    Vitamin_mineral_supplement_pregnancy = ifelse(
      is.na(Vitamin_mineral_supplement_pregnancy) | Vitamin_mineral_supplement_pregnancy == "Do not know",
      NA_character_,
      Vitamin_mineral_supplement_pregnancy)
  )
```

Case–control weighted Cox (using time-to-event):

f <- n_subcohort_control / n_full_cohort_control
f_cases <- n_subcohort_cases / n_full_cohort_cases

Due to financial and practical constraints, cord blood samples were analyzed in 
a randomly selected subset of cohort participants, including 80 of 160 incident
cases and 296 of 16,336 non-cases. Because biomarker measurements were available
only for a stratified random subset defined by outcome status, analyses were 
performed using inverse probability weighted Cox proportional hazards regression. 
Sampling probabilities were estimated separately for cases and non-cases, and 
inverse sampling fractions were used as weights to obtain unbiased hazard ratio 
estimates representative of the full cohort, following methods for two-phase 
outcome-dependent sampling described by Samuelsen (1997) and Borgan et al. (2000).
```{r}
f <- nrow(df_cox_full %>% dplyr::filter(event==0))/ nrow(abis %>% dplyr::filter(event==0))
f_cases <- nrow(df_cox_full %>% dplyr::filter(event==1))/ nrow(abis %>% dplyr::filter(event==1))
```

```{r}
print("Sampling fractions")

print("controls")
print(paste0(nrow(df_cox_full %>% dplyr::filter(event==0)),"/",  nrow(abis %>% dplyr::filter(event==0)), " = ",f))

print("cases")
print(paste0(nrow(df_cox_full %>% dplyr::filter(event==1)),"/",  nrow(abis %>% dplyr::filter(event==1)), " = ",f_cases))

```
<!-- [1] "Sampling fractions" -->
<!-- [1] "controls" -->
<!-- [1] "296/16176 = 0.0182987141444115" -->
<!-- [1] "cases" -->
<!-- [1] "83/160 = 0.51" -->
```{r}
child_covariates <- c("Preterm_dichotomous","Birth_weight_cat", "Season", "Sex_cat")

maternal_covariates <- c(
  "T1D_heredity_cat",
  "Smoking_during_pregnancy",
  "Fish_any_pregnancy",
  "Parity",
  "Delivery_mode",
  "Vitamin_mineral_supplement_pregnancy"
)
```

```{r remove_missing_rows}
df_cox_full <- df_cox_full %>%
  dplyr::select(all_of(
    c(
      "ABISnr",
      "diagnosis",
      "conc_25OHD",
      "DBP",
      "exit_age",
      "event",
      child_covariates,
      maternal_covariates
    )
  )) %>%
  # re-scale conc_25OHD to estimate HR per per 10 μg/L
  dplyr::mutate(conc_25OHD = conc_25OHD/10.0)

```


Check Proportinal Hazard Assumption
a. conc_25OHD
```{r}
fit_lin_conc_25OHD <- coxph(
  Surv(exit_age, event = event) ~ conc_25OHD +
    Preterm_dichotomous +
    Smoking_during_pregnancy +
    Delivery_mode +
    Birth_weight_cat + 
    Season +
    T1D_heredity_cat +
    Fish_any_pregnancy +
    Parity +
    Vitamin_mineral_supplement_pregnancy,
  data = df_cox_full,
  weights = if_else(event == 1, 1 / f_cases, 1 / f),
  robust = TRUE
)

# Schoenfeld residuals against the transformed time

zph_lin_conc_25OHD <- cox.zph(fit_lin_conc_25OHD)
zph_lin_conc_25OHD_table <- zph_lin_conc_25OHD$table %>% 
  as.data.frame %>%
  tibble::rownames_to_column(var = "Term") %>%
  dplyr::mutate(Term = sapply(Term,clean_terms))%>%
  dplyr::mutate(across(where(is.numeric), format_numeric))

dimnames(zph_lin_conc_25OHD$y)[[2]] <- sapply(dimnames(zph_lin_conc_25OHD$y)[[2]], clean_terms)

print(zph_lin_conc_25OHD)           # identify offenders
          
pdf(
  paste0(output_dir,"/zph_linear_conc_25OHD_schoenfeld_residuals_vs_transformed_time_PH_check.pdf"),
  width = 11.67, 
  height = 8.67, 
  font = "Times" 
  )

## page 1
par(mfrow = c(2,2))
plot(zph_lin_conc_25OHD, var = 1:4, plot = TRUE)

## page 2
par(mfrow = c(2,2))
plot(zph_lin_conc_25OHD, var = 5:8, plot = TRUE)

## page 3
par(mfrow = c(2,2))
plot(zph_lin_conc_25OHD, var = 9:10, plot = TRUE)


dev.off()
```

b. DBP
```{r}
fit_lin_DBP   <- coxph(
    Surv(exit_age, event = event) ~ DBP + 
    Preterm_dichotomous +
    Smoking_during_pregnancy +
    Delivery_mode +
    Birth_weight_cat + 
    Season +
    T1D_heredity_cat +
    Fish_any_pregnancy +
    Parity +
    Vitamin_mineral_supplement_pregnancy,
  data = df_cox_full
)

# Schoenfeld residuals against the transformed time

zph_lin_DBP <- cox.zph(fit_lin_DBP)
zph_lin_DBP_table <- zph_lin_DBP$table %>% 
  as.data.frame %>%
  tibble::rownames_to_column(var = "Term") %>%
  dplyr::mutate(Term = sapply(Term,clean_terms)) %>%
  dplyr::mutate(across(where(is.numeric), format_numeric))

dimnames(zph_lin_DBP$y)[[2]] <- sapply(dimnames(zph_lin_DBP$y)[[2]], clean_terms)

print(zph_lin_DBP)           # identify offenders
          
pdf(
  paste0(output_dir,"/zph_linear_DBP_schoenfeld_residuals_vs_transformed_time_PH_check.pdf"),
  width = 11.67, 
  height = 8.67, 
  font = "Times" 
  )

## page 1
par(mfrow = c(2,2))
plot(zph_lin_DBP, var = 1:4, plot = TRUE)

## page 2
par(mfrow = c(2,2))
plot(zph_lin_DBP, var = 5:8, plot = TRUE)

## page 2
par(mfrow = c(2,2))
plot(zph_lin_DBP, var = 9:10, plot = TRUE)


dev.off()
```



```{r}
ft_conc_25OHD <- flextable::flextable(zph_lin_conc_25OHD_table) %>%
  flextable::autofit()

ft_DBP <- flextable::flextable(zph_lin_DBP_table) %>%
  flextable::autofit()
```


```{r}
flextable::save_as_docx(
  `scaled Schoenfeld residuals with time: conc_25OHD` = ft_conc_25OHD,
  `scaled Schoenfeld residuals with time: DBP` = ft_DBP,
  path = paste0(output_dir,"/scaled_schoenfeld_residuals_with_time.docx"),
  pr_section = sect_properties
)
```


Weighted Cox regression	Hazard ratio: relative rate of developing T1D at any given age. 

```{r cox_model_base}
unadjusted_conc_25OHD <- coxph(
    Surv(exit_age, event = event) ~ conc_25OHD, 
  data = df_cox_full, 
  weights = if_else(event == 1, 1/f_cases, 1 / f),
  robust = TRUE 
  
)

unadjusted_DBP <- coxph(
    Surv(exit_age, event = event) ~ DBP, 
  data = df_cox_full, 
  weights = if_else(event == 1, 1/f_cases, 1 / f),
  robust = TRUE 
  
)
```



```{r weighted_cox_maternal_and_child_confounders}
adjusted_conc_25OHD <-  coxph(
    Surv(exit_age, event = event) ~ conc_25OHD + Preterm_dichotomous + 
      Birth_weight_cat + Season + T1D_heredity_cat + Smoking_during_pregnancy +
    Fish_any_pregnancy + Parity + Delivery_mode + Vitamin_mineral_supplement_pregnancy, 
  data = df_cox_full, 
  weights = if_else(event == 1, 1/f_cases, 1 / f),
  robust = TRUE 
)

adjusted_DBP <-  coxph(
    Surv(exit_age, event = event) ~ DBP + Preterm_dichotomous + 
      Birth_weight_cat + Season + T1D_heredity_cat + Smoking_during_pregnancy +
    Fish_any_pregnancy + Parity + Delivery_mode + Vitamin_mineral_supplement_pregnancy, 
  data = df_cox_full, 
  weights = if_else(event == 1, 1/f_cases, 1 / f),
  robust = TRUE 
)
```


create pattern strings to clean-tidy result tables 

```{r}
all_levels <- (paste0(paste0(
  unique(
    unlist(
      lapply(
        c(child_covariates,maternal_covariates), 
        function(x) levels(factor(df_cox_full[[x]]))
        )
      )
    )[-2], 
  collapse = "|"), "|^Preterm$"))

all_pat_replace <- paste0(
  c(child_covariates,maternal_covariates), collapse = "|")
```


```{r tidy_model_base}
tidy_unadjusted_conc_25OHD <- tidy_base_model(
  unadjusted_conc_25OHD,
  "Unadjusted",
  "conc_25OHD",
  df_cox_full = df_cox_full,
  predictor_increment = "+ 10 μg/L"
)

tidy_unadjusted_DBP <- tidy_base_model(
  unadjusted_DBP,
  "Unadjusted",
  "DBP",
  df_cox_full = df_cox_full ,
  predictor_increment = "+ 1 mg/L"
)
```



```{r tidy_model_maternal_and_child_confounders}
tidy_adjusted_conc_25OHD <- tidy_model_specific_covariates(
  model = adjusted_conc_25OHD,
  predictor = "conc_25OHD",
  model_name = "Adjusted", 
  covariates_as_pattern = all_pat_replace, 
  covariate_levels_as_pattern = all_levels, 
  df_cox_full = df_cox_full,
  predictor_increment = "+ 10 μg/L" 
  )

tidy_adjusted_DBP <- tidy_model_specific_covariates(
  model = adjusted_DBP,
  predictor = "DBP",
  model_name = "Adjusted", 
  covariates_as_pattern = all_pat_replace, 
  covariate_levels_as_pattern = all_levels, 
  df_cox_full = df_cox_full ,
  predictor_increment = "+1 mg/L"
  )
```



```{r merged_all_model_results}
merged_all_weighted_cox_results <-  rbind(
  tidy_unadjusted_conc_25OHD,
  tidy_unadjusted_DBP,
  tidy_adjusted_conc_25OHD,
  tidy_adjusted_DBP
) %>%
  dplyr::mutate(
    Model = if_else(row_number() != 1 &
                      lag(Model, 1) == Model &
                      !(Term %in% c(
                        "conc_25OHD", "DBP"
                      )),
                    NA_character_,
                    Model)
  ) %>%
      dplyr::mutate(
        Model = if_else(row_number() != 1 &
                      (!is.na(lag(Model,1)) & Model == lag(Model, 1)),
                    NA_character_,
                    Model)
      )  %>%
  dplyr::mutate(
    Level = str_replace(
      Level, 
      "(\\d+)", 
      function(x) str_pad(x, 2, side = "left")
      )
    )
```


```{r merged_all_model_results_short_flex}
merged_all_weighted_cox_results_short_flex <- merged_all_weighted_cox_results %>%
  dplyr::filter(Term %in% c("conc 25OHD","DBP"))  %>%
      dplyr::mutate(
        Model = if_else(row_number() != 1 &
                      (!is.na(lag(Model,1)) & Model == lag(Model, 1)),
                    NA_character_,
                    Model)
      ) %>%
  flextable::flextable()%>%
  flextable::autofit() %>%
  flextable::set_table_properties(width = 1, layout = "autofit") %>%
  flextable::color(
    i = ~ as.numeric(`P-value`) < 0.05,
    j = c("Term", "Level", "P-value"),
    color = "red"
  ) 
```


```{r merged_all_model_results_long_flex}
merged_all_weighted_cox_results_long_flex <- merged_all_weighted_cox_results %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::set_table_properties(width = 1, layout = "autofit") %>%
  flextable::color(
    i = ~ as.numeric(`P-value`) < 0.05,
    j = c("Term", "Level", "P-value"),
    color = "red"
  )  %>%
  flextable::align(j = "Level", align = "right")
```



```{r save_merged_all_model_results}
flextable::save_as_docx(
  `Weighted Cox model` = merged_all_weighted_cox_results_short_flex,
  path = paste0(output_dir,"/flex_merged_all_weighted_cox_results_short.docx"),
  pr_section = sect_properties
)

flextable::save_as_docx(
  `Weighted Cox model` = merged_all_weighted_cox_results_long_flex,
  path = paste0(output_dir,"/flex_merged_all_weighted_cox_results_long.docx"),
  pr_section = sect_properties
)

saveRDS(
  object = merged_all_weighted_cox_results_short_flex,
  file = paste0(
    output_dir,
    "/flex_merged_all_weighted_cox_results_short.RDS"
  )
)

saveRDS(
  object = merged_all_weighted_cox_results_long_flex,
  file = paste0(
    output_dir,
    "/flex_merged_all_weighted_cox_results_long.RDS"
  )
)
```