---
title: "Cox Proportional Hazards model: Total 25(OH)D || DBP  || Free 25(OH)D"

author: "Debojyoti Das"
date: "2025-07-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

set output directory
```{r output_dir, message=FALSE, warning=FALSE}
output_dir <- "/Users/debda22/OneDrive - Linköpings universitet/T1D_vitamin_D_results/cox_analysis_revision/concentrations_total_free_25OHD_DBP"

if(!dir.exists(output_dir)){
  dir.create(output_dir, recursive = TRUE)
} else {
  cat(paste0(output_dir," is already present!!\n"))
}
```

load libraries
```{r load_libraries, message=FALSE, warning=FALSE}
pacman::p_load(
  haven,        
  zoo,
  sjlabelled,
  janitor,       
  data.table,   
  ggplot2,
  comprehenr,  
  stringr,     
  rstatix,
  dplyr,
  tidyr,
  forcats,
  officer,
  cowplot,
  ggExtra,
  ggpubr,
  GGally,
  janitor,
  ggrepel,
  ggtext,
  parameters,
  lme4,
  broom,
  pscl,
  tibble,
  knitr,
  ggeffects,
  cobalt,
  WeightIt,
  marginaleffects,
  survival,
  survminer,
  splines,
  purrr,
  nnet,
  magick,
  patchwork,
  flextable,
  ranger,
  iml
  )
```


```{r word_sect_properties}
sect_properties <- prop_section(
  page_size = page_size(
    orient = "landscape",
    height = 11.7,
    width = 8.3
  ),
  page_margins = page_mar(
    top = 0.5,
    bottom = 0.5,
    left = 0.5,
    right = 0.5,
    header = 0.3,
    footer = 0.3,
    gutter = 0
  ),
  type = "continuous"
)
```

define paths
```{r define_paths, eval=TRUE}
data_root <- "/Users/debda22/OneDrive - Linköpings universitet/"
data_root <- paste0(data_root,"Linkoping_projects/ABIS_RAW_DATA_SPSS/")
```

load parameters and helper functions
```{r}
source(paste0(data_root,"../../Rmd_and_R_files/T1D_projects/all_parameters_case_control_T1D_vitamin_D.R"))
source(paste0(data_root,"../../Rmd_and_R_files/T1D_projects/cox_T1D_helper.R"))
```


read/load vitamin_d_case_control_plus_abis for analysis
```{r}
vitamin_d_case_control_plus_abis <- read.table(paste0(data_root, "/vitamin_d/vitamin_d_case_control_plus_abis.csv"))
```

reorder factors
```{r}
vitamin_d_case_control_plus_abis$First_degree_T1D_heredity_cat <- factor(
  vitamin_d_case_control_plus_abis$First_degree_T1D_heredity_cat,
  levels = c("No", "Yes")
)

vitamin_d_case_control_plus_abis$Smoking_during_pregnancy <- factor(
  vitamin_d_case_control_plus_abis$Smoking_during_pregnancy,
  levels = c("No", "Yes")
)

vitamin_d_case_control_plus_abis$Parity <- factor(
  vitamin_d_case_control_plus_abis$Parity,
  levels = c("First parity", "Previous parity")
)

vitamin_d_case_control_plus_abis$Preterm_dichotomous <- factor(
  vitamin_d_case_control_plus_abis$Preterm_dichotomous,
  levels = c("Normal", "Preterm")
)

vitamin_d_case_control_plus_abis$Season <- factor(
  vitamin_d_case_control_plus_abis$Season,
  levels = c("Autumn", "Spring", "Summer", "Winter")
)

vitamin_d_case_control_plus_abis$Sex_cat <- factor(
  vitamin_d_case_control_plus_abis$Sex_cat,
  levels = c("Boy", "Girl")
)

vitamin_d_case_control_plus_abis$Fish_any_pregnancy <- factor(
  vitamin_d_case_control_plus_abis$Fish_any_pregnancy,
  levels = c("Seldom", "1–2 times week or more"))

vitamin_d_case_control_plus_abis$Delivery_mode <- factor(
  vitamin_d_case_control_plus_abis$Delivery_mode,
  levels = c("Normal","caesarean", "other problem"),
  labels = c("Normal","Caesarean", "Unknown"))
```


```{r}
qs_birth_weight <- round(quantile(vitamin_d_case_control_plus_abis$Birth_weight, probs = seq(0, 1, 0.25), na.rm = TRUE),2)

vitamin_d_case_control_plus_abis <- vitamin_d_case_control_plus_abis %>%
    dplyr::mutate(
        Birth_weight_cat = cut(
            Birth_weight,
            breaks = qs_birth_weight,
            include.lowest = TRUE,
            labels = c(
                paste0("≤", qs_birth_weight[2]),                                    
                paste0(qs_birth_weight[2], "–", qs_birth_weight[3]),                        
                paste0(qs_birth_weight[3], "–", qs_birth_weight[4]),                           
                paste0(">", qs_birth_weight[4])                                     
            )
        )
    )


```


Cox analysis

conducting a time-to-event Cox regression with:

 - Time scale: Age (from birth)

 - Exposure: Fixed (cord blood 25(OH)D)

 - Follow-up ends: December 31, 2020

Outcome: T1D diagnosis
```{r event_time, warning=FALSE}
vitamin_d_case_control_plus_abis$start_age <- 0  # since follow-up starts at birth

vitamin_d_case_control_plus_abis$INDATUMA <- as.Date(vitamin_d_case_control_plus_abis$INDATUMA, format = "%Y-%m-%d")
```


```{r event_time_formulation, warning=FALSE}
# Set exit date based on whether T1D occurred before 2020-12-31
vitamin_d_case_control_plus_abis$exit_date <- as.Date(ifelse(
  !is.na(vitamin_d_case_control_plus_abis$T1D_age_diagnosis) &
    vitamin_d_case_control_plus_abis$INDATUMA<= as.Date("2020-12-31"),
  vitamin_d_case_control_plus_abis$INDATUMA,
  as.Date("2020-12-31")
))
```


```{r event_time_calculation, warning=FALSE}
# Calculate age at exit in years
vitamin_d_case_control_plus_abis$exit_age <- as.numeric(
  difftime(
    vitamin_d_case_control_plus_abis$exit_date,
    vitamin_d_case_control_plus_abis$child_dob,
    units = "days"
  )
) / 365.25

vitamin_d_case_control_plus_abis$event <- ifelse(
  !is.na(vitamin_d_case_control_plus_abis$INDATUMA) &
    vitamin_d_case_control_plus_abis$INDATUMA <= as.Date("2020-12-31"),
  1,
  0
)
```


```{r cox_data, warning=FALSE}
df_cox_full <- vitamin_d_case_control_plus_abis %>%
  dplyr::select(all_of(
    c(
      "ABISnr",
      "DBP",
      "conc_25OHD",
      "diagnosis",
      "child_dob",
      "First_degree_T1D_heredity_cat",
      "Preterm_group",
      "Preterm_dichotomous",
      "Delivery_mode",
      "Gestational_age",
      "Birth_weight",
      "Birth_weight_cat",
      "Birth_height",
      "Season",
      "Parity",
      "Smoking_during_pregnancy",
      "Vitamin_mineral_supplement_pregnancy",
      "Fish_lake_pregnancy",
      "Fish_Baltic_pregnancy",
      "Fish_other_pregnancy",
      "Fish_any_pregnancy",
      "Sex_cat",
      # "HDIA",
      "DIAGNOS",
      # "INDATUMA",
      # "UTDATUMA",
      # "SOURCE",
      "T1D_age_diagnosis",
      "onset_age_group",
      "Birth_weight_cat",
      "day_of_year",
      # "Ai",
      # "adjusted_25OHD",
      "start_age",
      "exit_date",
      "exit_age",
      "event"
    )
  ))
```

```{r}
df_cox_full <- df_cox_full %>%
  dplyr::mutate(
    Vitamin_mineral_supplement_pregnancy = ifelse(
      is.na(Vitamin_mineral_supplement_pregnancy) | Vitamin_mineral_supplement_pregnancy == "Do not know",
      NA_character_,
      Vitamin_mineral_supplement_pregnancy)
  )
```

0. Create free 25(OH)

assuming conc_25OHD is in nmol/L, DBP in µmol/L
convert DBP to nmol/L so units are consistent (1 µmol/L = 1000 nmol/L)
```{r}
df_cox_full$DBP_nmol <- df_cox_full$DBP * 1000
df_cox_full$free25OHD <- df_cox_full$conc_25OHD / df_cox_full$DBP_nmol
df_cox_full$free25OHD_scaled <- scale(df_cox_full$free25OHD)
```


```{r}
child_covariates <- c("Preterm_dichotomous","Birth_weight", "Season", "Sex_cat")

maternal_covariates <- c(
  "First_degree_T1D_heredity_cat",
  "Smoking_during_pregnancy",
  "Fish_any_pregnancy",
  "Parity",
  "Delivery_mode",
  "Vitamin_mineral_supplement_pregnancy"
)
```

1. Cox regression models

Model A: total 25(OH)D
```{r}
cox_total <- coxph(
  as.formula(
    paste0(
      "Surv(exit_age, event) ~ conc_25OHD + ",
      paste0(maternal_covariates, collapse = " + "),
      " + ",
      paste0(child_covariates, collapse = " + ")
  )
), data = df_cox_full)
```


Model B: DBP
```{r}
cox_dbp <- coxph(
  as.formula(
    paste0(
      "Surv(exit_age, event) ~ DBP_nmol + ",
      paste0(maternal_covariates, collapse = " + "),
      " + ",
      paste0(child_covariates, collapse = " + ")
    )
  ),
  data = df_cox_full
)
```


# Model C: both
```{r}
 cox_both <- coxph(
  as.formula(
    paste0(
      "Surv(exit_age, event) ~ conc_25OHD + DBP_nmol + ",
      paste0(maternal_covariates, collapse = " + "),
      " + ",
      paste0(child_covariates, collapse = " + ")
    )
  ),
  data = df_cox_full
)
```

# Model D: free 25(OH)D
```{r}
cox_free_scaled <- coxph(
  as.formula(
    paste0(
      "Surv(exit_age, event) ~ free25OHD_scaled + ",
      paste0(maternal_covariates, collapse = " + "),
      " + ",
      paste0(child_covariates, collapse = " + ")
    )
  ),
  data = df_cox_full
)
```

# Collect tidy results
```{r}
cox_results <- list(
  total = broom::tidy(cox_total, exponentiate=TRUE, conf.int=TRUE),
  dbp   = broom::tidy(cox_dbp, exponentiate=TRUE, conf.int=TRUE),
  both  = broom::tidy(cox_both, exponentiate=TRUE, conf.int=TRUE),
  free_scaled  = broom::tidy(cox_free_scaled, exponentiate=TRUE, conf.int=TRUE)
)
```

```{r}
cox_results
```
```{r}
#-----------------------------------------
# 2. Random Survival Forest (RSF)
#-----------------------------------------
rsf <- ranger(
  formula = Surv(exit_age, event) ~ conc_25OHD + DBP_nmol + free25OHD +
    Preterm_dichotomous + Birth_weight_cat + Season + Sex_cat +
    First_degree_T1D_heredity_cat + Smoking_during_pregnancy +
    Fish_any_pregnancy + Parity + Delivery_mode +
    Vitamin_mineral_supplement_pregnancy,
  data = df_cox_full %>%
    dplyr::select(
      exit_age,
      event,
      conc_25OHD,
      DBP_nmol,
      free25OHD,
      Preterm_dichotomous,
      Birth_weight_cat,
      Season,
      Sex_cat,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Vitamin_mineral_supplement_pregnancy,
    ) %>%
    tidyr::drop_na(),
  mtry = 3,
  importance = "permutation",
  splitrule = "extratrees",
  num.trees = 1000,
  seed = 123,
  respect.unordered.factors = "partition"
)

# Variable importance
print(rsf$variable.importance)
```


```{r}
library(dplyr)
library(tidyr)
library(iml)

#-----------------------------------------
# 3. SHAP analysis with iml
#-----------------------------------------
data <- df_cox_full %>%
  dplyr::select(
    exit_age,
    event,
    conc_25OHD,
    DBP_nmol,
    free25OHD,
    Preterm_dichotomous,
    Birth_weight_cat,
    Season,
    Sex_cat,
    First_degree_T1D_heredity_cat,
    Smoking_during_pregnancy,
    Fish_any_pregnancy,
    Parity,
    Delivery_mode,
    Vitamin_mineral_supplement_pregnancy
  ) %>%
  tidyr::drop_na()

#-----------------------------------------
# Function to convert survival curve to scalar
#-----------------------------------------
time_point <- 20  # e.g., 10-year risk

predict_survprob <- function(model, newdata) {
  surv <- predict(model, data = newdata)$survival
  times <- predict(model, data = newdata)$unique.death.times
  t_idx <- which.min(abs(times - time_point))
  event_prob <- 1 - surv[, t_idx]  # probability of event by 10 years
  return(event_prob)
}

#-----------------------------------------
# Prepare predictor for iml
#-----------------------------------------
predictor <- Predictor$new(
  model = rsf,
  data = data[, -c(which(names(data) %in% c("exit_age","event")))],
  y = data$event,
  predict.function = predict_survprob
)

#-----------------------------------------
# Global feature importance (based on prediction error)
#-----------------------------------------
feat_imp <- FeatureImp$new(predictor, loss = "ce")
plot(feat_imp)

#-----------------------------------------
# Partial dependence for vitamin D variables
#-----------------------------------------
pdp_total <- FeatureEffect$new(predictor, feature = "conc_25OHD", method = "pdp")
plot(pdp_total)

pdp_dbp <- FeatureEffect$new(predictor, feature = "DBP_nmol", method = "pdp")
plot(pdp_dbp)

pdp_free <- FeatureEffect$new(predictor, feature = "free25OHD", method = "pdp")
plot(pdp_free)

#-----------------------------------------
# Optional: SHAP for one subject
#-----------------------------------------
shap_one <- Shapley$new(
  predictor,
  x.interest = data[1, -c(which(names(data) %in% c("exit_age","event")))]
)
plot(shap_one)

```

```{r}
library(car)       # for VIF
library(corrplot)  # for correlation matrix

#-----------------------------------------
# 1. Correlation matrix
#-----------------------------------------
vitd_vars <- df_cox_full %>%
  dplyr::select(conc_25OHD, DBP_nmol, free25OHD)

cor_mat <- cor(vitd_vars, use = "pairwise.complete.obs")
print(cor_mat)

# visualize
corrplot(cor_mat, method = "ellipse", type = "upper",
         tl.col = "black", tl.srt = 45)

#-----------------------------------------
# 2. Variance Inflation Factor (VIF)
#-----------------------------------------
# Fit a linear model with all three predictors
lm_test <- lm(conc_25OHD ~ DBP_nmol + free25OHD, data = df_cox_full)
vif_all <- vif(lm_test)
print(vif_all)

# Alternatively, VIF from the Cox model itself:
vif_cox <- vif(cox_all)  # from your multivariable Cox
print(vif_cox)

```

