---
title: 'Cox Proportional Hazards model: quartiles of conc. 25(OH)D & DBP'
author: "Debojyoti Das"
date: "2025-08-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

set output directory
```{r output_dir, message=FALSE, warning=FALSE}
output_dir <- "/Users/debda22/OneDrive - Linköpings universitet/T1D_vitamin_D_results/cox_analysi_revision/quartiles"

if(!dir.exists(output_dir)){
  dir.create(output_dir, recursive = TRUE)
} else {
  cat(paste0(output_dir," is already present!!\n"))
}
```

load libraries
```{r load_libraries, message=FALSE, warning=FALSE}
pacman::p_load(
  haven,        
  zoo,
  sjlabelled,
  janitor,       
  data.table,   
  ggplot2,
  comprehenr,  
  stringr,     
  rstatix,
  dplyr,
  tidyr,
  forcats,
  officer,
  cowplot,
  ggExtra,
  ggpubr,
  GGally,
  janitor,
  ggrepel,
  ggtext,
  parameters,
  lme4,
  broom,
  pscl,
  tibble,
  knitr,
  ggeffects,
  cobalt,
  WeightIt,
  marginaleffects,
  survival,
  purrr,
  nnet
  )
```


```{r word_sect_properties}
sect_properties <- prop_section(
  page_size = page_size(
    orient = "landscape",
    height = 11.7,
    width = 8.3
  ),
  page_margins = page_mar(
    top = 0.5,
    bottom = 0.5,
    left = 0.5,
    right = 0.5,
    header = 0.3,
    footer = 0.3,
    gutter = 0
  ),
  type = "continuous"
)
```

define paths
```{r define_paths, eval=TRUE}
data_root <- "/Users/debda22/OneDrive - Linköpings universitet/"
data_root <- paste0(data_root,"Linkoping_projects/ABIS_RAW_DATA_SPSS/")
```

load parameters and helper functions
```{r}
source(paste0(data_root,"../../Rmd_and_R_files/T1D_projects/all_parameters_case_control_T1D_vitamin_D.R"))
source(paste0(data_root,"../../Rmd_and_R_files/T1D_projects/cox_T1D_helper.R"))
```


read/load vitamin_d_case_control_plus_abis for analysis
```{r}
vitamin_d_case_control_plus_abis <- read.table(paste0(data_root, "/vitamin_d/vitamin_d_case_control_plus_abis.csv"))
```


reorder factors
```{r}
vitamin_d_case_control_plus_abis$First_degree_T1D_heredity_cat <- factor(
  vitamin_d_case_control_plus_abis$First_degree_T1D_heredity_cat,
  levels = c("No", "Yes")
)

vitamin_d_case_control_plus_abis$Smoking_during_pregnancy <- factor(
  vitamin_d_case_control_plus_abis$Smoking_during_pregnancy,
  levels = c("No", "Yes")
)

vitamin_d_case_control_plus_abis$Parity <- factor(
  vitamin_d_case_control_plus_abis$Parity,
  levels = c("First parity", "Previous parity")
)

vitamin_d_case_control_plus_abis$Preterm_dichotomous <- factor(
  vitamin_d_case_control_plus_abis$Preterm_dichotomous,
  levels = c("Normal", "Preterm")
)

vitamin_d_case_control_plus_abis$Season <- factor(
  vitamin_d_case_control_plus_abis$Season,
  levels = c("Autumn", "Spring", "Summer", "Winter")
)

vitamin_d_case_control_plus_abis$Sex_cat <- factor(
  vitamin_d_case_control_plus_abis$Sex_cat,
  levels = c("Boy", "Girl")
)

vitamin_d_case_control_plus_abis$Fish_any_pregnancy <- factor(
  vitamin_d_case_control_plus_abis$Fish_any_pregnancy,
  levels = c("Seldom", "1–2 times week or more"))

vitamin_d_case_control_plus_abis$Delivery_mode <- factor(
  vitamin_d_case_control_plus_abis$Delivery_mode,
  levels = c("Normal","caesarean", "other problem"),
  labels = c("Normal","Caesarean", "Other problem"))
```


```{r}
qs_birth_weight <- quantile(df_cox_full$Birth_weight, probs = seq(0, 1, 0.25), na.rm = TRUE)

vitamin_d_case_control_plus_abis <- vitamin_d_case_control_plus_abis %>%
    dplyr::mutate(
        Birth_weight_cat = cut(
            Birth_weight,
            breaks = qs_birth_weight,
            include.lowest = TRUE,
            labels = c(
                paste0("≤", qs_birth_weight[2]),                                    
                paste0(qs_birth_weight[2], "–", qs_birth_weight[3]),                        
                paste0(qs_birth_weight[3], "–", qs_birth_weight[4]),                           
                paste0(">", qs_birth_weight[4])                                     
            )
        )
    )
```


Cox analysis

conducting a time-to-event Cox regression with:

 - Time scale: Age (from birth)

 - Exposure: Fixed (cord blood 25(OH)D)

 - Follow-up ends: December 31, 2020

Outcome: T1D diagnosis
```{r event_time, warning=FALSE}
vitamin_d_case_control_plus_abis$start_age <- 0  # since follow-up starts at birth

vitamin_d_case_control_plus_abis$INDATUMA <- as.Date(vitamin_d_case_control_plus_abis$INDATUMA, format = "%Y-%m-%d")
```


```{r event_time_formulation, warning=FALSE}
# Set exit date based on whether T1D occurred before 2020-12-31
vitamin_d_case_control_plus_abis$exit_date <- as.Date(ifelse(
  !is.na(vitamin_d_case_control_plus_abis$T1D_age_diagnosis) &
    vitamin_d_case_control_plus_abis$INDATUMA<= as.Date("2020-12-31"),
  vitamin_d_case_control_plus_abis$INDATUMA,
  as.Date("2020-12-31")
))
```


```{r event_time_calculation, warning=FALSE}
# Calculate age at exit in years
vitamin_d_case_control_plus_abis$exit_age <- as.numeric(
  difftime(
    vitamin_d_case_control_plus_abis$exit_date,
    vitamin_d_case_control_plus_abis$child_dob,
    units = "days"
  )
) / 365.25

vitamin_d_case_control_plus_abis$event <- ifelse(
  !is.na(vitamin_d_case_control_plus_abis$INDATUMA) &
    vitamin_d_case_control_plus_abis$INDATUMA <= as.Date("2020-12-31"),
  1,
  0
)
```


```{r cox_data, warning=FALSE}
df_cox_full <- vitamin_d_case_control_plus_abis %>%
  dplyr::select(all_of(
    c(
      "ABISnr",
      "DBP",
      "conc_25OHD",
      "diagnosis",
      "child_dob",
      "First_degree_T1D_heredity_cat",
      "Preterm_group",
      "Preterm_dichotomous",
      "Delivery_mode",
      "Gestational_age",
      "Birth_weight",
      "Birth_weight_cat",
      "Birth_height",
      "Season",
      "Parity",
      "Smoking_during_pregnancy",
      "Vitamin_mineral_supplement_pregnancy",
      "Fish_lake_pregnancy",
      "Fish_Baltic_pregnancy",
      "Fish_other_pregnancy",
      "Fish_any_pregnancy",
      "Sex_cat",
      # "HDIA",
      "DIAGNOS",
      # "INDATUMA",
      # "UTDATUMA",
      # "SOURCE",
      "T1D_age_diagnosis",
      "onset_age_group",
      "Birth_weight_cat",
      "day_of_year",
      # "Ai",
      # "adjusted_25OHD",
      "start_age",
      "exit_date",
      "exit_age",
      "event"
    )
  ))
```



```{r}
child_covariates <- c("Preterm_dichotomous","Birth_weight", "Season", "Sex_cat")

maternal_covariates <- c(
  "First_degree_T1D_heredity_cat",
  "Smoking_during_pregnancy",
  # "Vitamin_mineral_supplement_pregnancy",
  "Fish_any_pregnancy",
  "Parity",
  "Delivery_mode"
)
```

```{r remove_missing_rows}
df_cox_full <- df_cox_full %>%
  dplyr::select(all_of(
    c(
      "ABISnr",
      "diagnosis",
      "conc_25OHD",
      "DBP",
      "exit_age",
      "event",
      "Birth_weight_cat",
      child_covariates,
      maternal_covariates,
      "onset_age_group"
    )
  )) 
```

Use categorical (quartiles) for conc. 25(OH)D instead of continuous values


```{r}
qs_conc_25OHD <- quartile(df_cox_full$conc_25OHD, probs = seq(0, 1, 0.25), na.rm = TRUE)
qs_DBP <- quartile(df_cox_full$DBP, probs = seq(0, 1, 0.25), na.rm = TRUE)

df_cox_full <- df_cox_full %>%
  dplyr::mutate(
        quartile_25OHD = cut(
            conc_25OHD,
            breaks = qs_conc_25OHD,
            include.lowest = TRUE,
            labels = c(
                paste0("≤", qs_conc_25OHD[2]),                                    
                paste0(qs_conc_25OHD[2], "–", qs_conc_25OHD[3]),                             
                paste0(qs_conc_25OHD[3], "–", qs_conc_25OHD[4]),                             
                paste0(">", qs_conc_25OHD[4])                                     
            )
        ),
        quartile_DBP = cut(
            DBP,
            breaks = qs_DBP,
            include.lowest = TRUE,
            labels = c(
                paste0("≤", qs_DBP[2]),                                    
                paste0(qs_DBP[2], "–", qs_DBP[3]),                             
                paste0(qs_DBP[3], "–", qs_DBP[4]),                             
                paste0(">", qs_DBP[4])                                     
            )
        )        
    ) %>%
  dplyr::select(all_of(
    c(
      "ABISnr",
      "diagnosis",
      "quartile_25OHD",
      "quartile_DBP",
      "exit_age",
      "event",
      "Birth_weight_cat",
      child_covariates,
      maternal_covariates,
      "onset_age_group"
    )
  )) 
```

```{r cox_model_base}
cox_model_base_quartile_25OHD <- coxph(
  Surv(exit_age, event = event) ~ quartile_25OHD,
  data = df_cox_full %>% dplyr::select(-quartile_DBP)
)

cox_model_base_quartile_DBP <- coxph(
  Surv(exit_age, event = event) ~ quartile_DBP,
  data = df_cox_full %>% dplyr::select(-quartile_25OHD)
)
```


```{r cox_model_child}
cox_model_child_quartile_25OHD <- coxph(
  Surv(exit_age, event = event) ~ quartile_25OHD + Preterm_dichotomous + Season + 
    Sex_cat + Birth_weight_cat, data = df_cox_full %>% dplyr::select(-quartile_DBP)
)

cox_model_child_quartile_DBP <- coxph(
  Surv(exit_age, event = event) ~ quartile_DBP + Preterm_dichotomous + Season + 
    Sex_cat + Birth_weight_cat, data = df_cox_full %>% dplyr::select(-quartile_25OHD)
)
```



```{r cox_model_maternal}
cox_model_maternal_quartile_25OHD <- coxph(
  Surv(exit_age, event = event) ~ quartile_25OHD + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode,
  data = df_cox_full %>% dplyr::select(-quartile_DBP)
)

cox_model_maternal_quartile_DBP <- coxph(
  Surv(exit_age, event = event) ~ quartile_DBP + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode,
  data = df_cox_full %>% dplyr::select(-quartile_25OHD)
)
```



```{r cox_model_full}
cox_model_full_quartile_25OHD <- coxph(
  Surv(exit_age, event = event) ~ quartile_25OHD + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode +
    Preterm_dichotomous + Season + Sex_cat + Birth_weight_cat,
  data = df_cox_full %>% dplyr::select(-quartile_DBP)
)

cox_model_full_quartile_DBP <- coxph(
  Surv(exit_age, event = event) ~ quartile_DBP + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode +
    Preterm_dichotomous + Season + Sex_cat + Birth_weight_cat,
  data = df_cox_full %>% dplyr::select(-quartile_25OHD)
)
```




```{r}
child_levels <- paste0(paste0(
  unlist(lapply(c(child_covariates[-2], "Birth_weight_cat","quartile_25OHD","quartile_DBP"), function(x)
  levels(factor(df_cox_full[[x]]))))[-2], collapse = "|"), "|Preterm$")

child_pat_replace <- paste0(c(child_covariates[-2],"Birth_weight_cat"),collapse = "|")
```

```{r}
maternal_levels <- paste0(
  unique(
    unlist(
      lapply(
        c(maternal_covariates,"quartile_25OHD","quartile_DBP"), 
        function(x) levels(factor(df_cox_full[[x]]))
        )
      )
    ), 
  collapse = "|")

maternal_pat_replace <- paste0(maternal_covariates, collapse = "|")
```

```{r}
all_levels <- (paste0(paste0(
  unique(
    unlist(
      lapply(
        c(child_covariates[-2],"Birth_weight_cat","quartile_25OHD","quartile_DBP",maternal_covariates), 
        function(x) levels(factor(df_cox_full[[x]]))
        )
      )
    )[-2], 
  collapse = "|"), "|Preterm$"))

all_pat_replace <- paste0(
  c(child_covariates[-2],"Birth_weight_cat",maternal_covariates), collapse = "|")
```


```{r tidy_cox_model_base}
tidy_cox_model_base_quartile_25OHD <- tidy_base_model(cox_model_base_quartile_25OHD, "Base", "quartile_25OHD")
tidy_cox_model_base_quartile_DBP <- tidy_base_model(cox_model_base_quartile_DBP, "Base", "quartile_DBP")
```

```{r tidy_cox_model_child}
tidy_cox_model_child_quartile_25OHD <- tidy_model_specific_covariates(
  model = cox_model_child_quartile_25OHD,
  predictor = "quartile_25OHD",
  model_name = "Base + Child covariates", 
  covariates_as_pattern = child_pat_replace, 
  covariate_levels_as_pattern = child_levels
  )

tidy_cox_model_child_quartile_DBP <- tidy_model_specific_covariates(
  model = cox_model_child_quartile_DBP,
  predictor = "quartile_DBP",
  model_name = "Base + Child covariates", 
  covariates_as_pattern = child_pat_replace, 
  covariate_levels_as_pattern = child_levels
  )
```

```{r tidy_cox_model_maternal}
tidy_cox_model_maternal_quartile_25OHD <- tidy_model_specific_covariates(
  model = cox_model_maternal_quartile_25OHD,
  predictor = "quartile_25OHD",
  model_name = "Base + Maternal covariates", 
  covariates_as_pattern = maternal_pat_replace, 
  covariate_levels_as_pattern = maternal_levels
  )

tidy_cox_model_maternal_quartile_DBP <- tidy_model_specific_covariates(
  model = cox_model_maternal_quartile_DBP,
  predictor = "quartile_DBP",
  model_name = "Base + Maternal covariates", 
  covariates_as_pattern = maternal_pat_replace, 
  covariate_levels_as_pattern = maternal_levels
  )
```

```{r tidy_cox_model_full}
tidy_cox_model_full_quartile_25OHD <- tidy_model_specific_covariates(
  model = cox_model_full_quartile_25OHD,
  predictor = "quartile_25OHD",
  model_name = "Base + All covariates", 
  covariates_as_pattern = all_pat_replace, 
  covariate_levels_as_pattern = all_levels
  ) 

tidy_cox_model_full_quartile_DBP <- tidy_model_specific_covariates(
  model = cox_model_full_quartile_DBP,
  predictor = "quartile_DBP",
  model_name = "Base + All covariates", 
  covariates_as_pattern = all_pat_replace, 
  covariate_levels_as_pattern = all_levels
  ) 
```



```{r merged_all_model_quartile_results}
merged_all_model_results_quartile_25OHD <-  rbind(
  tidy_cox_model_base_quartile_25OHD, 
  tidy_cox_model_child_quartile_25OHD,
  tidy_cox_model_maternal_quartile_25OHD,
  tidy_cox_model_full_quartile_25OHD
  ) %>% 
  table2flex() %>%
  flextable::set_table_properties(width = 1, layout = "autofit") 

merged_all_model_results_quartile_DBP <-  rbind(
  tidy_cox_model_base_quartile_DBP, 
  tidy_cox_model_child_quartile_DBP,
  tidy_cox_model_maternal_quartile_DBP,
  tidy_cox_model_full_quartile_DBP
  ) %>% 
  table2flex() %>%
  flextable::set_table_properties(width = 1, layout = "autofit") 
```

```{r save_merged_all_model_quartile_results}
flextable::save_as_docx(
  `Cox models: quartile 25(OH)D ` = merged_all_model_results_quartile_25OHD,
  path = paste0(output_dir,"/flex_merged_all_model_results_quartile_25OHD.docx"),
  pr_section = sect_properties
)

flextable::save_as_docx(
  `Cox models: quartile DBP ` = merged_all_model_results_quartile_DBP,
  path = paste0(output_dir,"/flex_merged_all_model_results_quartile_DBP.docx"),
  pr_section = sect_properties
)
```

stratified cox models

```{r strata_wise_cox_base}
strata_wise_cox_base_quartile_25OHD <- run_stratified_cox_models(
  data = df_cox_full %>% dplyr::select(-quartile_DBP), 
  predictor = "quartile_25OHD",
  strata_vars = c(child_covariates[-2], "Birth_weight_cat", maternal_covariates),
  covariate_levels_as_pattern = paste0(levels(df_cox_full[["quartile_25OHD"]]), collapse = "|")
  )

strata_wise_cox_base_quartile_DBP <- run_stratified_cox_models(
  data = df_cox_full %>% dplyr::select(-quartile_25OHD), 
  predictor = "quartile_DBP",
  strata_vars = c(child_covariates[-2], "Birth_weight_cat", maternal_covariates),
  covariate_levels_as_pattern = paste0(levels(df_cox_full[["quartile_DBP"]]), collapse = "|")
  )
```

```{r strata_wise_cox_child_covariates}
strata_wise_cox_child_covariates_quartile_25OHD <- run_stratified_cox_models(
  data = df_cox_full %>% dplyr::select(-quartile_DBP), 
  predictor ="quartile_25OHD",
  strata_vars = maternal_covariates,
  covariate_list = c(child_covariates[-2], "Birth_weight_cat"),
  covariates_as_pattern = child_pat_replace,
  covariate_levels_as_pattern = child_levels
  )

strata_wise_cox_child_covariates_quartile_DBP <- run_stratified_cox_models(
  data = df_cox_full %>% dplyr::select(-quartile_25OHD), 
  predictor = "quartile_DBP",
  strata_vars = maternal_covariates,
  covariate_list = c(child_covariates[-2], "Birth_weight_cat"),
  covariates_as_pattern = child_pat_replace,
  covariate_levels_as_pattern = child_levels
  )
```

```{r strata_wise_cox_maternal_covariates}
strata_wise_cox_maternal_covariates_quartile_25OHD <- run_stratified_cox_models(
  data = df_cox_full %>% dplyr::select(-quartile_DBP), 
  predictor ="quartile_25OHD",
  strata_vars = c(child_covariates[-2], "Birth_weight_cat"),
  covariate_list =  maternal_covariates,
  covariates_as_pattern = maternal_pat_replace,
  covariate_levels_as_pattern = maternal_levels
  )

strata_wise_cox_maternal_covariates_quartile_DBP <- run_stratified_cox_models(
  data = df_cox_full %>% dplyr::select(-quartile_25OHD), 
  predictor = "quartile_DBP",
  strata_vars = c(child_covariates[-2], "Birth_weight_cat"),
  covariate_list =  maternal_covariates,
  covariates_as_pattern = maternal_pat_replace,
  covariate_levels_as_pattern = maternal_levels
  )
```



```{r strata_flex_tables}
flex_strata_wise_cox_base_quartile_25OHD <- table2flex(strata_wise_cox_base_quartile_25OHD)
flex_strata_wise_cox_base_quartile_DBP <- table2flex(strata_wise_cox_base_quartile_DBP)

flex_strata_wise_cox_child_covariates_quartile_25OHD <- table2flex(strata_wise_cox_child_covariates_quartile_25OHD)
flex_strata_wise_cox_child_covariates_quartile_DBP <- table2flex(strata_wise_cox_child_covariates_quartile_DBP)

flex_strata_wise_cox_maternal_covariates_quartile_25OHD <- table2flex(strata_wise_cox_maternal_covariates_quartile_25OHD)
flex_strata_wise_cox_maternal_covariates_quartile_DBP <- table2flex(strata_wise_cox_maternal_covariates_quartile_DBP)
```




```{r save_flex_strata_wise_cox_models}
flextable::save_as_docx(
  `Strata wise cox (base) model: 25(OH)D (quartile)` = flex_strata_wise_cox_base_quartile_25OHD,
  path = paste0(output_dir,"/flex_strata_wise_cox_base_quartile_25OHD.docx"),
  pr_section = sect_properties
)

flextable::save_as_docx(
  `Strata wise cox  (base) model: DBP (quartile)` = flex_strata_wise_cox_base_quartile_DBP,
  path = paste0(output_dir,"/flex_strata_wise_cox_base_quartile_DBP.docx"),
  pr_section = sect_properties
)

flextable::save_as_docx(
  `Strata wise cox model with child covariates: 25(OH)D (quartile)` = flex_strata_wise_cox_child_covariates_quartile_25OHD,
  path = paste0(output_dir,"/flex_strata_wise_cox_child_covariates_quartile_25OHD.docx"),
  pr_section = sect_properties
)

flextable::save_as_docx(
  `Strata wise cox model with child covariates: DBP (quartile)` = flex_strata_wise_cox_child_covariates_quartile_DBP,
  path = paste0(output_dir,"/flex_strata_wise_cox_child_covariates_quartile_DBP.docx"),
  pr_section = sect_properties
)

flextable::save_as_docx(
  `Strata wise cox model with maternal covariates: 25(OH)D (quartile)` = flex_strata_wise_cox_maternal_covariates_quartile_25OHD,
  path = paste0(output_dir,"/flex_strata_wise_cox_maternal_covariates_quartile_25OHD.docx"),
  pr_section = sect_properties
)

flextable::save_as_docx(
  `Strata wise cox model with maternal covariates: DBP (quartile)` = flex_strata_wise_cox_maternal_covariates_quartile_DBP,
  path = paste0(output_dir,"/flex_strata_wise_cox_maternal_covariates_quartile_DBP.docx"),
  pr_section = sect_properties
)
```



<!-- ```{r onset_group} -->
<!-- df_cox_full$onset_age_group <- factor(df_cox_full$onset_age_group) -->

<!-- multi_model <- multinom(onset_age_group ~ conc_25OHD + First_degree_T1D_heredity_cat + -->
<!--     Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode + -->
<!--     Preterm_dichotomous + Season + Sex_cat + Birth_weight_cat,  -->
<!--     data = df_cox_full %>% dplyr::filter(diagnosis==1)) -->
<!-- ``` -->


<!-- ```{r onset_group_tidy} -->
<!-- tidy_multi_model <- multi_model %>% -->
<!--   broom::tidy(exponentiate = TRUE, conf.int = TRUE)  %>% -->
<!--   dplyr::filter(term != "(Intercept)") %>% -->
<!--   dplyr::mutate( -->
<!--     Level = gsub(all_pat_replace, "",  -->
<!--                  gsub("conc_25OHD", "+1", term)),.before = 3) %>% -->
<!--   dplyr::mutate(term = gsub(all_levels, "", term)) %>% -->
<!--   dplyr::mutate( -->
<!--       y.level = ifelse( -->
<!--         row_number() > 1 & y.level == lag(y.level), -->
<!--         NA_character_,  -->
<!--         y.level) -->
<!--       ) %>% -->
<!--   dplyr::mutate( -->
<!--     p.value = ifelse( -->
<!--       p.value < 0.001, -->
<!--       formatC(p.value, format = "e", digits = 2), -->
<!--       formatC(p.value, format = "f", digits = 3) -->
<!--     )) %>% -->
<!--   dplyr::rename( -->
<!--         Term = term, -->
<!--         HR = estimate, -->
<!--         `P-value` = p.value, -->
<!--         `Onset Group` = y.level -->
<!--       ) %>% -->
<!--     dplyr::mutate( -->
<!--       Term = ifelse( -->
<!--         row_number() > 1 & Term == lag(Term), -->
<!--         NA_character_, -->
<!--         gsub("_", " ", gsub( -->
<!--           "_cat", "", gsub( -->
<!--             "_dichotomous", -->
<!--             "", -->
<!--             gsub("_during_pregnancy|_any_pregnancy", " (pregnancy)",  -->
<!--                             gsub("Preterm_dichotomous","",Term) -->
<!--                ) -->
<!--           ) -->
<!--       )) -->
<!--     )) %>% -->
<!--   dplyr::mutate( -->
<!--     `HR (95% CI)` = paste0(round(HR, 3)," (", round(conf.low, 3), "-", round(conf.high, 3), ")") -->
<!--     ) %>% -->
<!--   dplyr::select(`Onset Group`, Term, Level, `HR (95% CI)`, `P-value`) -->

<!-- flex_tidy_multi_model <- tidy_multi_model %>%  -->
<!--  flextable::flextable() %>%  -->
<!--  flextable::autofit() %>% -->
<!--     flextable::set_table_properties(width = 1, layout = "autofit") %>%  -->
<!--  flextable::color( -->
<!--      i = ~ as.numeric(`P-value`) < 0.05, -->
<!--      j = c("Term", "Level", "P-value"), -->
<!--      color = "red") -->
<!-- ``` -->


<!-- ```{r save_flex_cox_model_full} -->
<!-- flextable::save_as_docx( -->
<!--   `Multinomial model: Onset-age ~ conc. 25(OH)D` = flex_tidy_multi_model, -->
<!--   path = paste0(output_dir,"/flex_multinomial_model_onset_age_conc_25OHD.docx"), -->
<!--   pr_section = sect_properties -->
<!-- ) -->
<!-- ``` -->