---
title: 'Cox Proportional Hazards model: clinical categories of conc. 25(OH)D & DBP'
author: "Debojyoti Das"
date: "2025-08-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

set output directory
```{r output_dir, message=FALSE, warning=FALSE}
output_dir <- "/Users/debda22/OneDrive - Linköpings universitet/T1D_vitamin_D_results/cox_analysi_revision/clinical_categories"

if(!dir.exists(output_dir)){
  dir.create(output_dir, recursive = TRUE)
} else {
  cat(paste0(output_dir," is already present!!\n"))
}
```

load libraries
```{r load_libraries, message=FALSE, warning=FALSE}
pacman::p_load(
  haven,        
  zoo,
  sjlabelled,
  janitor,       
  data.table,   
  ggplot2,
  comprehenr,  
  stringr,     
  rstatix,
  dplyr,
  tidyr,
  forcats,
  officer,
  cowplot,
  ggExtra,
  ggpubr,
  GGally,
  janitor,
  ggrepel,
  ggtext,
  parameters,
  lme4,
  broom,
  pscl,
  tibble,
  knitr,
  ggeffects,
  cobalt,
  WeightIt,
  marginaleffects,
  survival,
  purrr,
  nnet,
  coxphf
  )
```


```{r word_sect_properties}
sect_properties <- prop_section(
  page_size = page_size(
    orient = "landscape",
    height = 11.7,
    width = 8.3
  ),
  page_margins = page_mar(
    top = 0.5,
    bottom = 0.5,
    left = 0.5,
    right = 0.5,
    header = 0.3,
    footer = 0.3,
    gutter = 0
  ),
  type = "continuous"
)
```

define paths
```{r define_paths, eval=TRUE}
data_root <- "/Users/debda22/OneDrive - Linköpings universitet/"
data_root <- paste0(data_root,"Linkoping_projects/ABIS_RAW_DATA_SPSS/")
```

load parameters and helper functions
```{r}
source(paste0(data_root,"../../Rmd_and_R_files/T1D_projects/all_parameters_case_control_T1D_vitamin_D.R"))
source(paste0(data_root,"../../Rmd_and_R_files/T1D_projects/cox_T1D_helper.R"))
```


read/load vitamin_d_case_control_plus_abis for analysis
```{r}
vitamin_d_case_control_plus_abis <- read.table(paste0(data_root, "/vitamin_d/vitamin_d_case_control_plus_abis.csv"))
```


reorder factors
```{r}
vitamin_d_case_control_plus_abis$First_degree_T1D_heredity_cat <- factor(
  vitamin_d_case_control_plus_abis$First_degree_T1D_heredity_cat,
  levels = c("No", "Yes")
)

vitamin_d_case_control_plus_abis$Smoking_during_pregnancy <- factor(
  vitamin_d_case_control_plus_abis$Smoking_during_pregnancy,
  levels = c("No", "Yes")
)

vitamin_d_case_control_plus_abis$Parity <- factor(
  vitamin_d_case_control_plus_abis$Parity,
  levels = c("First parity", "Previous parity")
)

vitamin_d_case_control_plus_abis$Preterm_dichotomous <- factor(
  vitamin_d_case_control_plus_abis$Preterm_dichotomous,
  levels = c("Normal", "Preterm")
)

vitamin_d_case_control_plus_abis$Season <- factor(
  vitamin_d_case_control_plus_abis$Season,
  levels = c("Autumn", "Spring", "Summer", "Winter")
)

vitamin_d_case_control_plus_abis$Sex_cat <- factor(
  vitamin_d_case_control_plus_abis$Sex_cat,
  levels = c("Boy", "Girl")
)

vitamin_d_case_control_plus_abis$Fish_any_pregnancy <- factor(
  vitamin_d_case_control_plus_abis$Fish_any_pregnancy,
  levels = c("Seldom", "1–2 times week or more"))

vitamin_d_case_control_plus_abis$Delivery_mode <- factor(
  vitamin_d_case_control_plus_abis$Delivery_mode,
  levels = c("Normal","caesarean", "other problem"),
  labels = c("Normal","Caesarean", "Other problem"))

vitamin_d_case_control_plus_abis$Birth_weight_cat <- factor(
  vitamin_d_case_control_plus_abis$Birth_weight_cat,
  levels = c("q1", "q2", "q3", "q4")
)
```


Cox analysis

conducting a time-to-event Cox regression with:

 - Time scale: Age (from birth)

 - Exposure: Fixed (cord blood 25(OH)D)

 - Follow-up ends: December 31, 2020

Outcome: T1D diagnosis
```{r event_time, warning=FALSE}
vitamin_d_case_control_plus_abis$start_age <- 0  # since follow-up starts at birth

vitamin_d_case_control_plus_abis$INDATUMA <- as.Date(vitamin_d_case_control_plus_abis$INDATUMA, format = "%Y-%m-%d")
```


```{r event_time_formulation, warning=FALSE}
# Set exit date based on whether T1D occurred before 2020-12-31
vitamin_d_case_control_plus_abis$exit_date <- as.Date(ifelse(
  !is.na(vitamin_d_case_control_plus_abis$T1D_age_diagnosis) &
    vitamin_d_case_control_plus_abis$INDATUMA<= as.Date("2020-12-31"),
  vitamin_d_case_control_plus_abis$INDATUMA,
  as.Date("2020-12-31")
))
```


```{r event_time_calculation, warning=FALSE}
# Calculate age at exit in years
vitamin_d_case_control_plus_abis$exit_age <- as.numeric(
  difftime(
    vitamin_d_case_control_plus_abis$exit_date,
    vitamin_d_case_control_plus_abis$child_dob,
    units = "days"
  )
) / 365.25

vitamin_d_case_control_plus_abis$event <- ifelse(
  !is.na(vitamin_d_case_control_plus_abis$INDATUMA) &
    vitamin_d_case_control_plus_abis$INDATUMA <= as.Date("2020-12-31"),
  1,
  0
)
```


```{r cox_data, warning=FALSE}
df_cox_full <- vitamin_d_case_control_plus_abis %>%
  dplyr::select(all_of(
    c(
      "ABISnr",
      "DBP",
      "conc_25OHD",
      "diagnosis",
      "child_dob",
      "First_degree_T1D_heredity_cat",
      "Preterm_group",
      "Preterm_dichotomous",
      "Delivery_mode",
      "Gestational_age",
      "Birth_weight",
      "Birth_weight_cat",
      "Birth_height",
      "Season",
      "Parity",
      "Smoking_during_pregnancy",
      "Vitamin_mineral_supplement_pregnancy",
      "Fish_lake_pregnancy",
      "Fish_Baltic_pregnancy",
      "Fish_other_pregnancy",
      "Fish_any_pregnancy",
      "Sex_cat",
      # "HDIA",
      "DIAGNOS",
      # "INDATUMA",
      # "UTDATUMA",
      # "SOURCE",
      "T1D_age_diagnosis",
      "onset_age_group",
      "Birth_weight_cat",
      "day_of_year",
      # "Ai",
      # "adjusted_25OHD",
      "start_age",
      "exit_date",
      "exit_age",
      "event"
    )
  ))
```



```{r}
child_covariates <- c("Preterm_dichotomous","Birth_weight", "Season", "Sex_cat")

maternal_covariates <- c(
  "First_degree_T1D_heredity_cat",
  "Smoking_during_pregnancy",
  # "Vitamin_mineral_supplement_pregnancy",
  "Fish_any_pregnancy",
  "Parity",
  "Delivery_mode"
)
```

```{r remove_missing_rows}
df_cox_full <- df_cox_full %>%
  dplyr::select(all_of(
    c(
      "ABISnr",
      "diagnosis",
      "conc_25OHD",
      "DBP",
      "exit_age",
      "event",
      "Birth_weight_cat",
      child_covariates,
      maternal_covariates,
      "onset_age_group"
    )
  )) 
```

check Proportinal Hazard Assumption
```{r}
fit_lin   <- coxph(
  Surv(exit_age, event) ~ conc_25OHD + Preterm_dichotomous + Birth_weight +
    Season + Sex_cat + First_degree_T1D_heredity_cat + Smoking_during_pregnancy +
    Fish_any_pregnancy + Parity + Delivery_mode,
  data = df_cox_full
)

fit_spline <- coxph(
  Surv(exit_age, event) ~ splines::ns(conc_25OHD, df = 3) +
    Preterm_dichotomous + Birth_weight + Season + Sex_cat +
    First_degree_T1D_heredity_cat + Smoking_during_pregnancy +
    Fish_any_pregnancy + Parity + Delivery_mode,
  data = df_cox_full
)

anova(fit_lin, fit_spline, test="LRT")
```


```{r}
zph_lin <- cox.zph(fit_lin)
zph_lin_table <- zph_lin$table %>% as.data.frame %>%
  tibble::rownames_to_column(var = "Term") %>%
  dplyr::mutate(Term = sapply(Term,clean_terms))

dimnames(zph_lin$y)[[2]] <- sapply(dimnames(zph_lin$y)[[2]],clean_terms)

print(zph_lin)           # identify offenders
          
jpeg(
  paste0(output_dir,"/zph_linear_beta_vs_time_PH_check.jpeg"),
  units = "in", 
  width = 11.67, 
  height = 8.67, 
  res = 300,
  type = "cairo"
  )
par(mfrow = c(4, 3))  # grid
print(plot(zph_lin, var = 1:10, plot = TRUE)) 
dev.off()
```

```{r}
zph_spline <- cox.zph(fit_spline)
zph_spline_table <- zph_spline$table %>% as.data.frame %>%
  tibble::rownames_to_column(var = "Term") %>%
  dplyr::mutate(Term = sapply(Term,clean_terms))

dimnames(zph_spline$y)[[2]] <- sapply(dimnames(zph_spline$y)[[2]],clean_terms)

print(zph_spline)           # identify offenders
          
jpeg(
  paste0(output_dir,"/zph_spline_beta_vs_time_PH_check.jpeg"),
  units = "in", 
  width = 11.67, 
  height = 8.67, 
  res = 300,
  type = "cairo"
  )

par(mfrow = c(4, 3))  # grid
plot(zph_spline, var = 1:10, plot = TRUE)
dev.off()
```

reset grid
```{r}
par(mfrow = c(1, 1))
```


Use categorical (quantiles) for conc. 25(OH)D instead of continuous values

A. 25-hydroxy Vitamin D 25(OH)D
Evaluation, Treatment, and Prevention of Vitamin D Deficiency: an Endocrine Society Clinical Practice Guideline
https://doi.org/10.1210/jc.2011-0385

"1.0 Diagnostic procedure
1.1 We recommend screening for vitamin D deficiency in individuals at risk for deficiency. We do not recommend population screening for vitamin D deficiency in individuals who are not at risk (1|⊕⊕⊕⊕).

1.2 We recommend using the serum circulating 25-hydroxy vitamin D [25(OH)D] level, measured by a reliable assay, to evaluate vitamin D status in patients who are at risk for vitamin D deficiency. Vitamin D deficiency is defined as a 25(OH)D below 20 ng/ml (50 nmol/liter), and vitamin D insufficiency as a 25(OH)D of 21–29 ng/ml (525–725 nmol/liter). We recommend against using the serum 1,25-dihydroxyvitamin D [1,25(OH)2D] assay for this purpose and are in favor of using it only in monitoring certain conditions, such as acquired and inherited disorders of vitamin D and phosphate metabolism."

“These thresholds are based on non-pregnant bone health outcomes; their applicability to gestational serum is uncertain given physiological changes during pregnancy.”

B. Vitamin D Binding Protein (DBP)
Significance of Vitamin D Binding Protein in Assessing Vitamin D Status Among Under-Five Children
https://doi.org/10.1007/s12291-020-00873-x

```{r}
df_cox_full$cat_25OHD <- cut(df_cox_full$conc_25OHD,
                 breaks = c(-Inf, 20, 40, Inf),
                 labels = c("<20","20–39","≥40"), right = FALSE)
df_cox_full$cat_25OHD <- relevel(df_cox_full$cat_25OHD, ref="≥40")
```


```{r}
fit_firth_base <- coxphf::coxphf(Surv(exit_age, event) ~ cat_25OHD, data=df_cox_full)

fit_firth_child_covariates <- coxphf::coxphf(Surv(exit_age, event) ~ cat_25OHD + Preterm_dichotomous + Birth_weight + Season + Sex_cat, data=df_cox_full)

fit_firth_child_covariates <- coxphf::coxphf(Surv(exit_age, event) ~ cat_25OHD + Preterm_dichotomous + Birth_weight + Season + Sex_cat, data=df_cox_full)


```
