---
title: "Cox Proportional Hazards model: quartlies of conc. 25(OH)D & DBP"
author: "Debojyoti Das"
date: "2025-07-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

set output directory
```{r output_dir, message=FALSE, warning=FALSE}
output_dir <- "/Users/debda22/OneDrive - Linköpings universitet/T1D_vitamin_D_results/cox_analysi_revision/quartiles"

if(!dir.exists(output_dir)){
  dir.create(output_dir, recursive = TRUE)
} else {
  cat(paste0(output_dir," is already present!!\n"))
}
```

load libraries
```{r load_libraries, message=FALSE, warning=FALSE}
pacman::p_load(
  haven,        
  zoo,
  sjlabelled,
  janitor,       
  data.table,   
  ggplot2,
  comprehenr,  
  stringr,     
  rstatix,
  dplyr,
  tidyr,
  forcats,
  officer,
  cowplot,
  ggExtra,
  ggpubr,
  GGally,
  janitor,
  ggrepel,
  ggtext,
  parameters,
  lme4,
  broom,
  pscl,
  tibble,
  knitr,
  ggeffects,
  cobalt,
  WeightIt,
  marginaleffects,
  splines,
  survival,
  survminer,
  purrr,
  nnet,
  magick
  )
```


```{r word_sect_properties}
sect_properties <- prop_section(
  page_size = page_size(
    orient = "landscape",
    height = 11.7,
    width = 8.3
  ),
  page_margins = page_mar(
    top = 0.5,
    bottom = 0.5,
    left = 0.5,
    right = 0.5,
    header = 0.3,
    footer = 0.3,
    gutter = 0
  ),
  type = "continuous"
)
```

define paths
```{r define_paths, eval=TRUE}
data_root <- "/Users/debda22/OneDrive - Linköpings universitet/"
data_root <- paste0(data_root,"Linkoping_projects/ABIS_RAW_DATA_SPSS/")
```

load parameters and helper functions
```{r}
source(paste0(data_root,"../../Rmd_and_R_files/T1D_projects/all_parameters_case_control_T1D_vitamin_D.R"))
source(paste0(data_root,"../../Rmd_and_R_files/T1D_projects/cox_T1D_helper.R"))
```


read/load vitamin_d_case_control_plus_abis for analysis
```{r}
vitamin_d_case_control_plus_abis <- read.table(paste0(data_root, "/vitamin_d/vitamin_d_case_control_plus_abis.csv"))
```

reorder factors
```{r}
vitamin_d_case_control_plus_abis$First_degree_T1D_heredity_cat <- factor(
  vitamin_d_case_control_plus_abis$First_degree_T1D_heredity_cat,
  levels = c("No", "Yes")
)

vitamin_d_case_control_plus_abis$Smoking_during_pregnancy <- factor(
  vitamin_d_case_control_plus_abis$Smoking_during_pregnancy,
  levels = c("No", "Yes")
)

vitamin_d_case_control_plus_abis$Parity <- factor(
  vitamin_d_case_control_plus_abis$Parity,
  levels = c("First parity", "Previous parity")
)

vitamin_d_case_control_plus_abis$Preterm_dichotomous <- factor(
  vitamin_d_case_control_plus_abis$Preterm_dichotomous,
  levels = c("Normal", "Preterm")
)

vitamin_d_case_control_plus_abis$Season <- factor(
  vitamin_d_case_control_plus_abis$Season,
  levels = c("Autumn", "Spring", "Summer", "Winter")
)

vitamin_d_case_control_plus_abis$Sex_cat <- factor(
  vitamin_d_case_control_plus_abis$Sex_cat,
  levels = c("Boy", "Girl")
)

vitamin_d_case_control_plus_abis$Fish_any_pregnancy <- factor(
  vitamin_d_case_control_plus_abis$Fish_any_pregnancy,
  levels = c("Seldom", "1–2 times week or more"))

vitamin_d_case_control_plus_abis$Delivery_mode <- factor(
  vitamin_d_case_control_plus_abis$Delivery_mode,
  levels = c("Normal","caesarean", "other problem"),
  labels = c("Normal","Caesarean", "Other problem"))
```


```{r}
qs_birth_weight <- round(quantile(vitamin_d_case_control_plus_abis$Birth_weight, probs = seq(0, 1, 0.25), na.rm = TRUE),2)

vitamin_d_case_control_plus_abis <- vitamin_d_case_control_plus_abis %>%
    dplyr::mutate(
        Birth_weight_cat = cut(
            Birth_weight,
            breaks = qs_birth_weight,
            include.lowest = TRUE,
            labels = c(
                paste0("≤", qs_birth_weight[2]),                                    
                paste0(qs_birth_weight[2], "–", qs_birth_weight[3]),                        
                paste0(qs_birth_weight[3], "–", qs_birth_weight[4]),                           
                paste0(">", qs_birth_weight[4])                                     
            )
        )
    )


```


Cox analysis

conducting a time-to-event Cox regression with:

 - Time scale: Age (from birth)

 - Exposure: Fixed (cord blood 25(OH)D)

 - Follow-up ends: December 31, 2020

Outcome: T1D diagnosis
```{r event_time, warning=FALSE}
vitamin_d_case_control_plus_abis$start_age <- 0  # since follow-up starts at birth

vitamin_d_case_control_plus_abis$INDATUMA <- as.Date(vitamin_d_case_control_plus_abis$INDATUMA, format = "%Y-%m-%d")
```


```{r event_time_formulation, warning=FALSE}
# Set exit date based on whether T1D occurred before 2020-12-31
vitamin_d_case_control_plus_abis$exit_date <- as.Date(ifelse(
  !is.na(vitamin_d_case_control_plus_abis$T1D_age_diagnosis) &
    vitamin_d_case_control_plus_abis$INDATUMA<= as.Date("2020-12-31"),
  vitamin_d_case_control_plus_abis$INDATUMA,
  as.Date("2020-12-31")
))
```


```{r event_time_calculation, warning=FALSE}
# Calculate age at exit in years
vitamin_d_case_control_plus_abis$exit_age <- as.numeric(
  difftime(
    vitamin_d_case_control_plus_abis$exit_date,
    vitamin_d_case_control_plus_abis$child_dob,
    units = "days"
  )
) / 365.25

vitamin_d_case_control_plus_abis$event <- ifelse(
  !is.na(vitamin_d_case_control_plus_abis$INDATUMA) &
    vitamin_d_case_control_plus_abis$INDATUMA <= as.Date("2020-12-31"),
  1,
  0
)
```


```{r cox_data, warning=FALSE}
df_cox_full <- vitamin_d_case_control_plus_abis %>%
  dplyr::select(all_of(
    c(
      "ABISnr",
      "DBP",
      "conc_25OHD",
      "diagnosis",
      "child_dob",
      "First_degree_T1D_heredity_cat",
      "Preterm_group",
      "Preterm_dichotomous",
      "Delivery_mode",
      "Gestational_age",
      "Birth_weight",
      "Birth_weight_cat",
      "Birth_height",
      "Season",
      "Parity",
      "Smoking_during_pregnancy",
      "Vitamin_mineral_supplement_pregnancy",
      "Fish_lake_pregnancy",
      "Fish_Baltic_pregnancy",
      "Fish_other_pregnancy",
      "Fish_any_pregnancy",
      "Sex_cat",
      # "HDIA",
      "DIAGNOS",
      # "INDATUMA",
      # "UTDATUMA",
      # "SOURCE",
      "T1D_age_diagnosis",
      "onset_age_group",
      "Birth_weight_cat",
      "day_of_year",
      # "Ai",
      # "adjusted_25OHD",
      "start_age",
      "exit_date",
      "exit_age",
      "event"
    )
  ))
```



```{r}
child_covariates <- c("Preterm_dichotomous","Birth_weight", "Season", "Sex_cat")

maternal_covariates <- c(
  "First_degree_T1D_heredity_cat",
  "Smoking_during_pregnancy",
  # "Vitamin_mineral_supplement_pregnancy",
  "Fish_any_pregnancy",
  "Parity",
  "Delivery_mode"
)
```

```{r remove_missing_rows}
df_cox_full <- df_cox_full %>%
  dplyr::select(all_of(
    c(
      "ABISnr",
      "diagnosis",
      "conc_25OHD",
      "DBP",
      "exit_age",
      "event",
      "Birth_weight_cat",
      child_covariates,
      maternal_covariates,
      "onset_age_group"
    )
  )) 
```

Use categorical (quartiles) for conc. 25(OH)D instead of continuous values


```{r}
qs_conc_25OHD <- quantile(df_cox_full$conc_25OHD, probs = seq(0, 1, 0.25), na.rm = TRUE)
qs_DBP <- quantile(df_cox_full$DBP, probs = seq(0, 1, 0.25), na.rm = TRUE)

df_cox_full <- df_cox_full %>%
  dplyr::mutate(
        quartile_25OHD = cut(
            conc_25OHD,
            breaks = qs_conc_25OHD,
            include.lowest = TRUE,
            labels = c(
                paste0("≤", qs_conc_25OHD[2]),                                    
                paste0(qs_conc_25OHD[2], "–", qs_conc_25OHD[3]),                             
                paste0(qs_conc_25OHD[3], "–", qs_conc_25OHD[4]),                             
                paste0(">", qs_conc_25OHD[4])                                     
            )
        ),
        quartile_DBP = cut(
            DBP,
            breaks = qs_DBP,
            include.lowest = TRUE,
            labels = c(
                paste0("≤", qs_DBP[2]),                                    
                paste0(qs_DBP[2], "–", qs_DBP[3]),                             
                paste0(qs_DBP[3], "–", qs_DBP[4]),                             
                paste0(">", qs_DBP[4])                                     
            )
        )        
    ) %>%
  dplyr::select(all_of(
    c(
      "ABISnr",
      "diagnosis",
      "quartile_25OHD",
      "quartile_DBP",
      "exit_age",
      "event",
      "Birth_weight_cat",
      child_covariates,
      maternal_covariates,
      "onset_age_group"
    )
  )) 
```

Check Proportinal Hazard Assumption
```{r}
fit_quartile_25OHD   <- coxph(
  Surv(exit_age, event) ~ quartile_25OHD + Preterm_dichotomous + Birth_weight_cat +
    Season + Sex_cat + First_degree_T1D_heredity_cat + Smoking_during_pregnancy +
    Fish_any_pregnancy + Parity + Delivery_mode,
  data = df_cox_full
)

fit_quartile_DBP <- coxph(
  Surv(exit_age, event) ~ quartile_DBP + Preterm_dichotomous + Birth_weight_cat +
    Season + Sex_cat + First_degree_T1D_heredity_cat + Smoking_during_pregnancy +
    Fish_any_pregnancy + Parity + Delivery_mode,
  data = df_cox_full
)
```


Schoenfeld residuals against the transformed time

```{r}
zph_quartile_25OHD <- cox.zph(fit_quartile_25OHD)
zph_quartile_DBP <- cox.zph(fit_quartile_DBP)

zph_quartile_25OHD_table <- zph_quartile_25OHD$table %>% 
  as.data.frame %>%
  tibble::rownames_to_column(var = "Term") %>%
  dplyr::mutate(Term = sapply(Term,clean_terms))

zph_quartile_DBP_table <- zph_quartile_DBP$table %>% 
  as.data.frame %>%
  tibble::rownames_to_column(var = "Term") %>%
  dplyr::mutate(Term = sapply(Term,clean_terms))

dimnames(zph_quartile_25OHD$y)[[2]] <- sapply(dimnames(zph_quartile_25OHD$y)[[2]], clean_terms)
dimnames(zph_quartile_DBP$y)[[2]] <- sapply(dimnames(zph_quartile_DBP$y)[[2]], clean_terms)

print(zph_quartile_25OHD)           
print(zph_quartile_DBP)
```


```{r}
pdf(
  paste0(output_dir,"/zph_quartile_25OHD_schoenfeld_residuals_vs_transformed_time_PH_check.pdf"),
  width = 11.67, 
  height = 8.67, 
  font = "Times"
  )

par(mfrow = c(2,2))
plot(zph_quartile_25OHD, var = 1:4, plot = TRUE)

par(mfrow = c(2,2))
plot(zph_quartile_25OHD, var = 5:8, plot = TRUE)

## page 3
par(mfrow = c(2,2))
plot(zph_quartile_25OHD, var = 9:10, plot = TRUE)

dev.off()

img <- image_read_pdf(paste0(output_dir, "/zph_quartile_25OHD_schoenfeld_residuals_vs_transformed_time_PH_check.pdf"), density = 300)
image_write(img, paste0(output_dir, "/zph_quartile_25OHD_schoenfeld_residuals_vs_transformed_time_PH_check_300dpi.pdf"), format = "pdf")
```


```{r}
pdf(
  paste0(output_dir,"/zph_quartile_DBP_schoenfeld_residuals_vs_transformed_time_PH_check.pdf"),
  width = 11.67, 
  height = 8.67, 
  font = "Times"
  )

par(mfrow = c(2,2))
plot(zph_quartile_DBP, var = 1:4, plot = TRUE)

par(mfrow = c(2,2))
plot(zph_quartile_DBP, var = 5:8, plot = TRUE)

## page 3
par(mfrow = c(2,2))
plot(zph_quartile_DBP, var = 9:10, plot = TRUE)

dev.off()

img <- image_read_pdf(paste0(output_dir, "/zph_quartile_DBP_schoenfeld_residuals_vs_transformed_time_PH_check.pdf"), density = 300)
image_write(img, paste0(output_dir, "/zph_quartile_DBP_schoenfeld_residuals_vs_transformed_time_PH_check_300dpi.pdf"), format = "pdf")
```

Cox regression	Hazard ratio: relative rate of developing T1D at any given age. 

```{r cox_model_base}
cox_model_base_quartile_25OHD <- survival::coxph(
  Surv(exit_age, event = event) ~ quartile_25OHD,
  data = df_cox_full %>% dplyr::select(exit_age, event,quartile_25OHD)
)

cox_model_base_quartile_DBP <- survival::coxph(
  Surv(exit_age, event = event) ~ quartile_DBP,
  data = df_cox_full %>% dplyr::select(exit_age, event, quartile_DBP)
)

firth_base_quartile_25OHD <- coxphf::coxphf(
  Surv(exit_age, event) ~ quartile_25OHD,
  data = df_cox_full %>% 
    dplyr::select(exit_age, event, quartile_25OHD) %>%
    tidyr::drop_na()
  )

firth_base_quartile_DBP <- coxphf::coxphf(
  Surv(exit_age, event) ~ quartile_DBP,
  data = df_cox_full %>% 
    dplyr::select(exit_age, event, quartile_DBP)  %>%
    tidyr::drop_na()
  )
```


```{r cox_model_child_confounders}
cox_model_child_confounders_quartile_25OHD <- survival::coxph(
  Surv(exit_age, event = event) ~ quartile_25OHD + Season + Sex_cat,
  data = df_cox_full %>%
    dplyr::select(
      exit_age,
      event,
      quartile_25OHD,
      Season,
      Sex_cat
    )
)

cox_model_child_confounders_quartile_DBP <- survival::coxph(
  Surv(exit_age, event = event) ~ quartile_DBP + Season + Sex_cat,
  data = df_cox_full %>%     dplyr::select(
      exit_age,
      event,
      quartile_DBP,
      Season,
      Sex_cat
    )
)
```


```{r firth_model_child_confounders}
firth_child_confounders_quartile_25OHD <- coxphf::coxphf(
  Surv(exit_age, event) ~ quartile_25OHD + Season + Sex_cat,
  data = df_cox_full %>%
    dplyr::select(
      exit_age,
      event,
      quartile_25OHD,
      Season,
      Sex_cat
    ) %>% tidyr::drop_na(),
  maxit = 1000,
  maxstep = 0.01
)

firth_child_confounders_quartile_DBP <- coxphf::coxphf(
  Surv(exit_age, event) ~ quartile_DBP + Season + Sex_cat,
  data = df_cox_full %>%
    dplyr::select(
      exit_age,
      event,
      quartile_DBP,
      Preterm_dichotomous,
      Season,
      Sex_cat,
      Birth_weight_cat
    ) %>% tidyr::drop_na(),
  maxit = 500,
  maxstep = 0.10
)
```



```{r cox_model_maternal_confounders}
cox_model_maternal_confounders_quartile_25OHD <- survival::coxph(
  Surv(exit_age, event = event) ~ quartile_25OHD + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode,
  data = df_cox_full %>%
    dplyr::select(
      exit_age,
      event,
      quartile_25OHD,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode
    )
)

cox_model_maternal_confounders_quartile_DBP <- survival::coxph(
  Surv(exit_age, event = event) ~ quartile_DBP + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode,
  data = df_cox_full %>% 
    dplyr::select(
      exit_age,
      event,
      quartile_DBP,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode
    )
)
```


```{r firth_model_maternal_confounders}
firth_maternal_confounders_quartile_25OHD <- coxphf::coxphf(
  Surv(exit_age, event = event) ~ quartile_25OHD + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode,
  data = df_cox_full %>% 
    dplyr::select(
      exit_age,
      event,
      quartile_25OHD,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode
    ) %>% 
    tidyr::drop_na()
)

firth_maternal_confounders_quartile_DBP <- coxphf::coxphf(
  Surv(exit_age, event = event) ~ quartile_DBP + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode,
  data = df_cox_full %>% 
    dplyr::select(
      exit_age,
      event,
      quartile_DBP,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode
    ) %>% 
    tidyr::drop_na()
)
```

```{r cox_model_maternal_and_child_confounders}
cox_model_maternal_and_child_confounders_quartile_25OHD <- survival::coxph(
  Surv(exit_age, event = event) ~ quartile_25OHD + First_degree_T1D_heredity_cat + 
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode + 
    Season + Sex_cat,
  data = df_cox_full %>%
    dplyr::select(
      exit_age,
      event,
      quartile_25OHD,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Season,
      Sex_cat
    )
)

cox_model_maternal_and_child_confounders_quartile_DBP <- survival::coxph(
  Surv(exit_age, event = event) ~ quartile_DBP + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode + 
    Season + Sex_cat,
  data = df_cox_full %>% 
    dplyr::select(
      exit_age,
      event,
      quartile_DBP,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Season,
      Sex_cat
    )
)
```

```{r firth_model_maternal_and_child_confounders}
firth_maternal_and_child_confounders_quartile_25OHD <- coxphf::coxphf(
  Surv(exit_age, event = event) ~ quartile_25OHD + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode + 
    Season + Sex_cat,
  data = df_cox_full %>% 
    dplyr::select(
      exit_age,
      event,
      quartile_25OHD,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Season,
      Sex_cat
    ) %>% 
    tidyr::drop_na()
)

firth_maternal_and_child_confounders_quartile_DBP <- coxphf::coxphf(
  Surv(exit_age, event = event) ~ quartile_DBP + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode + 
    Season + Sex_cat,
  data = df_cox_full %>% 
    dplyr::select(
      exit_age,
      event,
      quartile_DBP,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Season,
      Sex_cat
    ) %>% 
    tidyr::drop_na()
)
```


```{r cox_model_full}
cox_model_full_quartile_25OHD <- survival::coxph(
  Surv(exit_age, event = event) ~ quartile_25OHD + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode + 
    Season + Sex_cat + Preterm_dichotomous + Birth_weight_cat,
  data = df_cox_full %>%
    dplyr::select(
      exit_age,
      event,
      quartile_25OHD,
      Preterm_dichotomous,
      Season,
      Sex_cat,
      Birth_weight_cat,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode
    )
)

cox_model_full_quartile_DBP <- survival::coxph(
  Surv(exit_age, event = event) ~ quartile_DBP + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode +
    Season + Sex_cat + Preterm_dichotomous + Birth_weight_cat,
  data = df_cox_full %>%
    dplyr::select(
      exit_age,
      event,
      quartile_DBP,
      Preterm_dichotomous,
      Season,
      Sex_cat,
      Birth_weight_cat,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode
    )
)
```


```{r firth_model_full}
firth_full_quartile_25OHD <- coxphf::coxphf(
  Surv(exit_age, event = event) ~ quartile_25OHD + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode +
    Season + Sex_cat + Preterm_dichotomous + Birth_weight_cat,
  data = df_cox_full %>%
    dplyr::select(
      exit_age,
      event,
      quartile_25OHD,
      Preterm_dichotomous,
      Season,
      Sex_cat,
      Birth_weight_cat,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode
    ) %>%
    tidyr::drop_na(),
  maxit = 1000,
  maxstep = 0.01
)

firth_full_quartile_DBP <- coxphf::coxphf(
  Surv(exit_age, event = event) ~ quartile_DBP + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode +
    Season + Sex_cat + Preterm_dichotomous + Birth_weight_cat,
  data = df_cox_full %>%
    dplyr::select(
      exit_age,
      event,
      quartile_DBP,
      Preterm_dichotomous,
      Season,
      Sex_cat,
      Birth_weight_cat,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode
    ) %>%
    tidyr::drop_na(),
  maxit = 5000,
  maxstep = 0.01
)
```


create pattern strings to clean-tidy result tables 
```{r}
child_levels <- paste0(
  paste0(unlist(lapply(c(child_covariates[-2], "Birth_weight_cat", "quartile_25OHD", "quartile_DBP"), function(x)
  levels(factor(df_cox_full[[x]]))))[-2], collapse = "|"), "|^Preterm$")

child_pat_replace <- paste0(c(child_covariates[-2],"Birth_weight_cat"),collapse = "|")
```

```{r}
maternal_levels <- paste0(
  unique(
    unlist(
      lapply(
        maternal_covariates, 
        function(x) levels(factor(df_cox_full[[x]]))
        )
      )
    ), 
  collapse = "|")

maternal_pat_replace <- paste0(maternal_covariates, collapse = "|")
```

```{r}
all_levels <- (paste0(paste0(
  unique(
    unlist(
      lapply(
        c(child_covariates[-2],"Birth_weight_cat",maternal_covariates, "quartile_25OHD", "quartile_DBP"), 
        function(x) levels(factor(df_cox_full[[x]]))
        )
      )
    )[-2], 
  collapse = "|"), "|^Preterm$"))

all_pat_replace <- paste0(
  c(child_covariates[-2],"Birth_weight_cat",maternal_covariates), collapse = "|")
```

```{r tidy_model_base}
tidy_cox_model_base_quartile_25OHD <- tidy_base_model(cox_model_base_quartile_25OHD, "Base", "quartile_25OHD")
tidy_cox_model_base_quartile_DBP <- tidy_base_model(cox_model_base_quartile_DBP, "Base","quartile_DBP")

tidy_firth_base_quartile_25OHD <- tidy_base_model(firth_base_quartile_25OHD, "Base", "quartile_25OHD", firth_model = TRUE)
tidy_firth_base_quartile_DBP <- tidy_base_model(firth_base_quartile_DBP, "Base","quartile_DBP", firth_model = TRUE)
```


> tidy_firth_base_quartile_25OHD
# A tibble: 3 × 5
  Model Term           Level `HR (95% CI)`       `P-value`
  <chr> <chr>          <chr> <chr>               <chr>    
1 Base  quartile 25OHD 28–37 1.315 (0.708-2.444) 0.381    
2 NA    NA             37–52 1.1 (0.592-2.044)   0.760    
3 NA    NA             >52   1.332 (0.732-2.426) 0.342


```{r tidy_model_child_confounders}
tidy_cox_model_child_confounders_quartile_25OHD <- tidy_model_specific_covariates(
  model = cox_model_child_confounders_quartile_25OHD,
  predictor = "quartile_25OHD",
  model_name = "Base + Child confounders", 
  covariates_as_pattern = child_pat_replace, 
  covariate_levels_as_pattern = child_levels
  )

tidy_cox_model_child_confounders_quartile_DBP <- tidy_model_specific_covariates(
  model = cox_model_child_confounders_quartile_DBP,
  predictor = "quartile_DBP",
  model_name = "Base + Child confounders", 
  covariates_as_pattern = child_pat_replace, 
  covariate_levels_as_pattern = child_levels
  )

tidy_firth_child_confounders_quartile_25OHD <- tidy_model_specific_covariates(
  model = firth_child_confounders_quartile_25OHD,
  predictor = "quartile_25OHD",
  model_name = "Base + Child confounders", 
  covariates_as_pattern = child_pat_replace, 
  covariate_levels_as_pattern = child_levels, 
  firth_model = TRUE
  )

tidy_firth_child_confounders_quartile_DBP <- tidy_model_specific_covariates(
  model = firth_child_confounders_quartile_DBP,
  predictor = "quartile_DBP",
  model_name = "Base + Child confounders", 
  covariates_as_pattern = child_pat_replace, 
  covariate_levels_as_pattern = child_levels, 
  firth_model = TRUE
  )
```

```{r tidy_model_maternal_confounders}
tidy_cox_model_maternal_confounders_quartile_25OHD <- tidy_model_specific_covariates(
  model = cox_model_maternal_confounders_quartile_25OHD,
  predictor = "quartile_25OHD",
  model_name = "Base + Maternal confounders", 
  covariates_as_pattern = maternal_pat_replace, 
  covariate_levels_as_pattern = maternal_levels
  )

tidy_cox_model_maternal_confounders_quartile_DBP <- tidy_model_specific_covariates(
  model = cox_model_maternal_confounders_quartile_DBP,
  predictor = "quartile_DBP",
  model_name = "Base + Maternal confounders", 
  covariates_as_pattern = maternal_pat_replace, 
  covariate_levels_as_pattern = maternal_levels
  )

tidy_firth_maternal_confounders_quartile_25OHD <- tidy_model_specific_covariates(
  model = firth_maternal_confounders_quartile_25OHD,
  predictor = "quartile_25OHD",
  model_name = "Base + Maternal confounders", 
  covariates_as_pattern = maternal_pat_replace, 
  covariate_levels_as_pattern = maternal_levels, 
  firth_model = TRUE
  )

tidy_firth_maternal_confounders_quartile_DBP <- tidy_model_specific_covariates(
  model = firth_maternal_confounders_quartile_DBP,
  predictor = "quartile_DBP",
  model_name = "Base + Maternal confounders", 
  covariates_as_pattern = maternal_pat_replace, 
  covariate_levels_as_pattern = maternal_levels, 
  firth_model = TRUE
  )
```

```{r tidy_model_maternal_and_child_confounders}
tidy_cox_model_maternal_and_child_confounders_quartile_25OHD <- tidy_model_specific_covariates(
  model = cox_model_maternal_and_child_confounders_quartile_25OHD,
  predictor = "quartile_25OHD",
  model_name = "Base + Maternal\n + Child confounders", 
  covariates_as_pattern = all_pat_replace, 
  covariate_levels_as_pattern = all_levels
  )

tidy_cox_model_maternal_and_child_confounders_quartile_DBP <- tidy_model_specific_covariates(
  model = cox_model_maternal_and_child_confounders_quartile_DBP,
  predictor = "quartile_DBP",
  model_name = "Base + Maternal\n + Child confounders", 
  covariates_as_pattern = all_pat_replace, 
  covariate_levels_as_pattern = all_levels
  )

tidy_firth_maternal_and_child_confounders_quartile_25OHD <- tidy_model_specific_covariates(
  model = firth_maternal_and_child_confounders_quartile_25OHD,
  predictor = "quartile_25OHD",
  model_name = "Base + Maternal\n + Child confounders", 
  covariates_as_pattern = all_pat_replace, 
  covariate_levels_as_pattern = all_levels, 
  firth_model = TRUE
  )

tidy_firth_maternal_and_child_confounders_quartile_DBP <- tidy_model_specific_covariates(
  model = firth_maternal_and_child_confounders_quartile_DBP,
  predictor = "quartile_DBP",
  model_name = "Base + Maternal\n + Child confounders", 
  covariates_as_pattern = all_pat_replace, 
  covariate_levels_as_pattern = all_levels, 
  firth_model = TRUE
  )
```

```{r tidy_cox_model_full}
tidy_cox_model_full_quartile_25OHD <- tidy_model_specific_covariates(
  model = cox_model_full_quartile_25OHD,
  predictor = "quartile_25OHD",
  model_name = "Base + All confounders\n + Child mediators", 
  covariates_as_pattern = all_pat_replace, 
  covariate_levels_as_pattern = all_levels
  ) 

tidy_cox_model_full_quartile_DBP <- tidy_model_specific_covariates(
  model = cox_model_full_quartile_DBP,
  predictor = "quartile_DBP",
  model_name = "Base + All confounders\n + Child mediators", 
  covariates_as_pattern = all_pat_replace, 
  covariate_levels_as_pattern = all_levels
  ) 

tidy_firth_full_quartile_25OHD <- tidy_model_specific_covariates(
  model = firth_full_quartile_25OHD,
  predictor = "quartile_25OHD",
  model_name = "Base + All confounders\n + Child mediators", 
  covariates_as_pattern = all_pat_replace, 
  covariate_levels_as_pattern = all_levels, 
  firth_model = TRUE
  )

tidy_firth_full_quartile_DBP <- tidy_model_specific_covariates(
  model = firth_full_quartile_DBP,
  predictor = "quartile_DBP",
  model_name = "Base + All confounders\n + Child mediators", 
  covariates_as_pattern = all_pat_replace, 
  covariate_levels_as_pattern = all_levels, 
  firth_model = TRUE
  )
```




```{r merged_all_model_results_quartile_25OHD }
merged_all_cox_model_results_quartile_25OHD <-  rbind(
  tidy_cox_model_base_quartile_25OHD,
  tidy_cox_model_child_confounders_quartile_25OHD,
  tidy_cox_model_maternal_confounders_quartile_25OHD,
  tidy_cox_model_maternal_and_child_confounders_quartile_25OHD,
  tidy_cox_model_full_quartile_25OHD
) %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::set_table_properties(width = 1, layout = "autofit") %>%
  flextable::color(
    i = ~ as.numeric(`P-value`) < 0.05,
    j = c("Term", "Level", "P-value"),
    color = "red"
  )

merged_all_firth_model_results_quartile_25OHD <-  rbind(
  tidy_firth_base_quartile_25OHD,
  tidy_firth_child_confounders_quartile_25OHD,
  tidy_firth_maternal_confounders_quartile_25OHD,
  tidy_firth_maternal_and_child_confounders_quartile_25OHD,
  tidy_firth_full_quartile_25OHD
) %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::set_table_properties(width = 1, layout = "autofit") %>%
  flextable::color(
    i = ~ as.numeric(`P-value`) < 0.05,
    j = c("Term", "Level", "P-value"),
    color = "red"
  )
```

```{r merged_all_model_results_quartile_DBP}
merged_all_cox_model_results_quartile_DBP <-  rbind(
  tidy_cox_model_base_quartile_DBP,
  tidy_cox_model_child_confounders_quartile_DBP,
  tidy_cox_model_maternal_confounders_quartile_DBP, 
  tidy_cox_model_maternal_and_child_confounders_quartile_DBP,
  tidy_cox_model_full_quartile_DBP
) %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::set_table_properties(width = 1, layout = "autofit") %>%
  flextable::color(
    i = ~ as.numeric(`P-value`) < 0.05,
    j = c("Term", "Level", "P-value"),
    color = "red"
  )

merged_all_firth_model_results_quartile_DBP <-  rbind(
  tidy_firth_base_quartile_DBP,
  tidy_firth_child_confounders_quartile_DBP,
  tidy_firth_maternal_confounders_quartile_DBP,
  tidy_firth_maternal_and_child_confounders_quartile_DBP,
  tidy_firth_full_quartile_DBP
) %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::set_table_properties(width = 1, layout = "autofit") %>%
  flextable::color(
    i = ~ as.numeric(`P-value`) < 0.05,
    j = c("Term", "Level", "P-value"),
    color = "red"
  )
```


```{r save_merged_all_model_results}
flextable::save_as_docx(
  `Cox proportional hazards models: conc. 25(OH)D` = merged_all_cox_model_results_quartile_25OHD,
  path = paste0(output_dir,"/flex_merged_all_cox_model_results_quartile_25OHD.docx"),
  pr_section = sect_properties
)

flextable::save_as_docx(
  `Cox proportional hazards models: quartile_DBP` = merged_all_cox_model_results_quartile_DBP,
  path = paste0(output_dir,"/flex_merged_all_cox_model_results_quartile_DBP.docx"),
  pr_section = sect_properties
)

flextable::save_as_docx(
  `Cox proportional hazards models with Firth's Penalized Likelihood: conc. 25(OH)D` = merged_all_firth_model_results_quartile_25OHD,
  path = paste0(output_dir,"/flex_merged_all_firth_model_results_quartile_25OHD.docx"),
  pr_section = sect_properties
)

flextable::save_as_docx(
  `Cox proportional hazards models with Firth's Penalized Likelihood: quartile_DBP` = merged_all_firth_model_results_quartile_DBP,
  path = paste0(output_dir,"/flex_merged_all_firth_model_results_quartile_DBP.docx"),
  pr_section = sect_properties
)
```

Cox models to test effect modification by heredity

1. Interaction model (primary test)

```{r}
cox_model_quartile_25OHD_interaction_heredity <- coxph(Surv(exit_age, event = event) ~ quartile_25OHD * First_degree_T1D_heredity_cat + Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode + Season + Sex_cat,
  data = df_cox_full %>% 
    dplyr::select(
      exit_age,
      event,
      quartile_25OHD,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Season,
      Sex_cat
    ) %>% 
    tidyr::drop_na()
)

cox_model_quartile_25OHD_interaction_season <- coxph(Surv(exit_age, event = event) ~ quartile_25OHD * Season + Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode + First_degree_T1D_heredity_cat + Sex_cat,
  data = df_cox_full %>% 
    dplyr::select(
      exit_age,
      event,
      quartile_25OHD,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Season,
      Sex_cat
    ) %>% 
    tidyr::drop_na()
)
```

```{r}
cox_model_quartile_DBP_interaction_heredity <- coxph(Surv(exit_age, event = event) ~ quartile_DBP * First_degree_T1D_heredity_cat + Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode + Season + Sex_cat,
  data = df_cox_full %>% 
    dplyr::select(
      exit_age,
      event,
      quartile_DBP,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Season,
      Sex_cat
    ) %>% 
    tidyr::drop_na()
)

cox_model_quartile_DBP_interaction_season <- coxph(Surv(exit_age, event = event) ~ quartile_DBP * Season + Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode + First_degree_T1D_heredity_cat + Sex_cat,
  data = df_cox_full %>% 
    dplyr::select(
      exit_age,
      event,
      quartile_DBP,
      First_degree_T1D_heredity_cat,
      Smoking_during_pregnancy,
      Fish_any_pregnancy,
      Parity,
      Delivery_mode,
      Season,
      Sex_cat
    ) %>% 
    tidyr::drop_na()
)
```


```{r}
tab_quartile_25OHD_hered <- make_comp_table(
  cox_model_maternal_and_child_confounders_quartile_25OHD,
  cox_model_quartile_25OHD_interaction_heredity,
  "quartile 25(OH)D × First-degree T1D Heredity"
)

tab_quartile_25OHD_season <- make_comp_table(
  cox_model_maternal_and_child_confounders_quartile_25OHD,
  cox_model_quartile_25OHD_interaction_season,
  "quartile 25(OH)D × Season"
)

tab_quartile_DBP_hered <- make_comp_table(
  cox_model_maternal_and_child_confounders_quartile_DBP,
  cox_model_quartile_DBP_interaction_heredity,
  "quartile DBP × First-degree T1D Heredity"
)

tab_quartile_DBP_season <- make_comp_table(
  cox_model_maternal_and_child_confounders_quartile_DBP,
  cox_model_quartile_DBP_interaction_season,
  "quartile DBP × Season"
)

anova_interaction_quartile <- rbind(
  tab_quartile_25OHD_hered,
  tab_quartile_25OHD_season,
  tab_quartile_DBP_hered,
  tab_quartile_DBP_season
  ) %>%
  flextable::flextable() %>%
  flextable::autofit()

flextable::save_as_docx(
  `Analysis of Deviance Table: Cox model: response is  Surv(time, event)` = anova_interaction_quartile,
  path = paste0(output_dir,"/anova_interaction_quartile.docx"),
  pr_section = sect_properties
)
```


survival curves

```{r}
plot_base_quartile_25OHD <- survminer::ggadjustedcurves(
  cox_model_base_quartile_25OHD, 
  data = df_cox_full %>% dplyr::select(exit_age, event, quartile_25OHD), 
  variable = "quartile_25OHD",
  size = 1,                 # change line size
  palette = c("#E60000", "#E69F00", "#56B4E9", "#009E73"),
  conf.int = TRUE,          # Add confidence interval
  pval = TRUE,              # Add p-value
  risk.table = TRUE,        # Add risk table
  risk.table.col = "strata",# Risk table color by groups
  # legend.labs = 
  #   c("Male", "Female"),    # Change legend labels
  risk.table.height = 0.25, # Useful to change when you have multiple groups
  ggtheme = theme_survminer()      # Change ggplot2 theme
)
```

