---
title: "Perinatal 25-OH-vitamin D as a risk factor for developing Type 1 Diabetes: ABIS cohort"
author: "Debojyoti Das"
date: "2024-01-31"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
pacman::p_load(
  haven,        # File import
  # here,         # File locator
  # sjlabelled,
  janitor,      # 
  data.table,   #
  dplyr,
  tidyverse,
  janitor,
  zoo,            # for date format
  comprehenr      # List comprehension in R
  )
```

directory paths
```{r}
project_root <- "/Users/debda22/OneDrive - Linköpings universitet/"
output_dir <- paste0(project_root,"cox_results_vitamin_d")

if(!dir.exists(output_dir)){
  dir.create(output_dir)  
}
```

load helping functions

```{r, include=FALSE}
# helper functions for imputation
src_path <- paste0(project_root,"Rmd_and_R_files/T1D_projects/")
# source(paste0(src_path, "\\Functions_rmph.R"))
# source(paste0(src_path, "\\serious_life_events_helper_functions.R"))
# source(paste0(src_path, "\\censor_data.R"))
# source(paste0(src_path, "\\all_parameters_serious_life_events_ABIS.R"))
```

Read cohort data from SAV format file

```{r}
data_path <- paste0(project_root,"Linkoping_projects/ABIS_RAW_DATA_SPSS")
abis_table <- read_sav(paste0(data_path, "/ABIS_data_base.sav"))
```

!! We use the entire data for now. We will subset it later in the analysis pipeline.

save ABISnr who participated in first year
```{r}
year_1_participants_ABISnr <- (abis_table %>% dplyr::filter(par_1==1))$ABISnr
year_3_participants_ABISnr <- (abis_table %>% dplyr::filter(par_2==1))$ABISnr
year_5_participants_ABISnr <- (abis_table %>% dplyr::filter(par_5==1))$ABISnr
year_8_participants_ABISnr <- (abis_table %>% dplyr::filter(par_8==1))$ABISnr
year_10_12_participants_ABISnr <- (abis_table %>% dplyr::filter(ParP_11==1))$ABISnr
```


```{r}
abis_table <- abis_table %>% 
  mutate(
    child_dob = as.Date(as.yearmon(as.character(barn_år_mån), format = "%Y%m"))
    )

i=0
Name <- c()
Label <- c()
for (col in names(abis_table)){
  
  i=i+1
  Name <- c(Name,col)
  
  curr_col_attribute <- attr(abis_table[[col]], "label")
  
  if (is.null(curr_col_attribute)||(length(curr_col_attribute)>1)){
    Label <- c(Label, col)
    } else {
    Label <- c(Label, curr_col_attribute)
    }
} 

abis_table_labels <- 
  data.frame("Name" = Name, "Label" = Label)
```


```{r}
psoriasis_diagonistic_table.labels <- 
  to_vec(for (col in names(psoriasis_diagonistic_table)) 
    if(is.null(attr(psoriasis_diagonistic_table[[col]], "label"))) col 
    else attr(psoriasis_diagonistic_table[[col]], "label"))




psoriasis_diagonistic_labels <- 
  data.frame("Name" = names(psoriasis_diagonistic_table), 
             "Label" = psoriasis_diagonistic_table.labels)

```

```{r}
date_10_12_qa_child = c()
for (v in seq_along(abis_table$gb_1)) {
  
  yr <- abis_table$gb_1[[v]]
  mn <- abis_table$gb_2[[v]]
  
  if(!is.na(mn)&(mn>12)){
    while(mn>12){
      mn <- floor(mn/10) 
    }
  }
  
  if(is.na(yr)|is.na(mn)){
    date_10_12_qa_child <- c(date_10_12_qa_child, NA_character_)
  } else {
    year_value <- sjlabelled::get_labels(
      sjlabelled::as_character(abis_table$gb_1[[v]])
      )
    if(is.na(year_value)){
      date_10_12_qa_child <- c(date_10_12_qa_child, NA_character_)
    } else {
          date_10_12_qa_child <- c(
          date_10_12_qa_child,
          paste0(year_value,"-",
            ifelse(
              nDigits2(mn)==1,
              paste0("0",as.character(mn)),
              as.character(mn)
              )
            )
          )
    }
  }
}
```

follow-up date for 10-13 year questionnaires
```{r}
abis_table["date_10_12_qa_child"] <- date_10_12_qa_child 

format_date <- function(my_string){
  return(paste0(substr(my_string, 1, 4), "-", substr(my_string, 5, 6), "-", substr(my_string, 7, nchar(my_string))))
}

first_digit <- function(my_number){
  return(
    as.character(
      str_split(as.character(as.integer(my_number)), pattern = "")[[1]][1]
      )
    )
}
  
abis_table <- abis_table %>%
        dplyr::mutate(
            gv_1 = ifelse(is.na(gv_1), NA_character_, gsub("000","0",gsub("0000","00",as.character(gv_1)))),
            date_10_12_qa_parent = case_when(
                is.na(gv_1) ~ NA_character_, 
                (nchar(gv_1) == 6 & grepl("^99|^98|^91|^81|^41", gv_1)) ~ format_date(paste0(gsub("-","",date_10_12_qa_child),"01")),
                (nchar(gv_1) == 6 & grepl("^00|00$",substr(gv_1,3,6))) ~ format_date(paste0(gsub("-","",date_10_12_qa_child),"01")),   
                nchar(gv_1) == 6 ~ format_date(paste0("20",as.character(gv_1))),
                (nchar(gv_1) == 5 & substr(gv_1,1,3) == "100") ~ format_date(paste0("20", gv_1, "0")),   
                (nchar(gv_1) == 5 & substr(gv_1,1,2) %in% c("10", "11")) ~ format_date(paste0("201", gv_1)),   
                (nchar(gv_1) == 5 & substr(gv_1,1,1) %in% c("4","5")) ~ format_date(paste0(gsub("-","",date_10_12_qa_child),"01")),   
                (nchar(gv_1) == 5 & substr(gv_1,1,1) == "2") ~ format_date(paste0("201", gv_1)),   
                (nchar(gv_1) == 5 & substr(gv_1,1,1) == "9") ~ format_date(paste0("200", gv_1)),   
                (nchar(gv_1) == 4 & as.integer(substr(gv_1,3,4)) > 12) ~ format_date(paste0(gsub("-","",date_10_12_qa_child),"01")),   
                nchar(gv_1) == 4 ~ format_date(paste0("20", gv_1,"01")),   
                nchar(gv_1) == 3 ~ NA_character_,
                TRUE ~ gv_1)
        )
```


```{r}
abis_table <- abis_table %>%  
  dplyr::mutate( 
    date_10_12_qa_parent = lubridate::as_date(date_10_12_qa_parent, format = "%Y-%m-%d"),
    ageyC_10_12 = round(lubridate::time_length(difftime(date_10_12_qa_parent,child_dob), "years"), digits = 2)
)
```
setting up SLE Child for year-1 data
```{r}
year_1_sle_dictionary <- year_1_sle_codes$b_44txt
names(year_1_sle_dictionary) <- year_1_sle_codes$category
```


setting up SLE Child for year-3 data
```{r}
year_3_sle_dictionary <- year_3_sle_codes$c_99txt
names(year_3_sle_dictionary) <- year_3_sle_codes$category
```


Merge the data to include psoriasis diagonistics

```{r}
psoriasis_data_full <- left_join(abis_table, 
                             psoriasis_diagonistic_table,
                             by = "ABISnr") %>% 
  mutate(Diagnos = replace(Diagnos, is.na(Diagnos), "")) %>%
  # reformat the date (of detection)
  mutate(
    first_detection = as.Date(as.character(övFörsta), format = "%Y%m%d"),
    second_detection = as.Date(as.character(övSenaste), format = "%Y%m%d")
    ) %>%
  mutate(
    age_first_detection = lubridate::time_length(
      difftime(first_detection,child_dob), "years"),
    age_second_detection = lubridate::time_length(
      difftime(second_detection,child_dob), "years")
  ) %>% mutate_if(is.numeric, round,2)

```

Remove attributes

```{r}

for (col in names(psoriasis_data_full)) {
  if(!(is.null(attr(psoriasis_data_full[[col]], "label")))){
    attr(psoriasis_data_full[[col]], "label") <- NULL
  }
  if(!(is.null(attr(psoriasis_data_full[[col]], "labels")))){
    attr(psoriasis_data_full[[col]], "labels") <- NULL
  }
  if(!(is.null(attr(psoriasis_data_full[[col]], "format.spss")))){
    attr(psoriasis_data_full[[col]], "format.spss") <- NULL
  }  
  if(!(is.null(attr(psoriasis_data_full[[col]], "names")))){
    attr(psoriasis_data_full[[col]], "names") <- NULL
  } 
  if(!(is.null(attr(psoriasis_data_full[[col]], "display_width")))){
    attr(psoriasis_data_full[[col]], "display_width") <- NULL
  }   
}

attr(psoriasis_data_full, "notes") <- NULL


for (col in names(psoriasis_data_full)) {
  if(is.labelled(psoriasis_data_full[[col]])){
    
    psoriasis_data_full[col] <- unlabel(psoriasis_data_full[[col]])
    
  }
  
}


psoriasis_data_full <- as.data.frame(psoriasis_data_full)

```

```{r}
psoriasis_data_full <- psoriasis_data_full %>%
  mutate(
    composite_psoriasis_heredity = pmin(
      e_34p_1,
      e_34p_2,
      e_34p_3,
      f_28o_1,
      f_28o_2,
      f_28o_3,
      gv_21o1,
      gv_21o2,
      gv_21o3, 
      na.rm = TRUE)
    ) %>% 
  mutate(
    first_degree_psoriasis_heredity = ifelse(
      is.na(composite_psoriasis_heredity),0,1)
  ) %>% 
  dplyr::select(-c(composite_psoriasis_heredity,
            barn_år_mån,
            e_34p_1,e_34p_2,e_34p_3,
            f_28o_1,f_28o_2,f_28o_3,
            gv_21o1,gv_21o2,gv_21o3,
            övAntal,övFörsta,övSenaste,
            first_detection,second_detection
            )
         )
```


```{r}
psoriasis_data_full <- psoriasis_data_full %>%
  dplyr::mutate(
   eduM_10_12_3kat = case_when(
      ParP_11 == 1 & gv_52 == 1 ~ 0,
      ParP_11 == 1 & gv_52 %in% c(2,3) ~ 1,
      ParP_11 == 1 & gv_52 %in% c(4,5,6) ~ 2,      
      TRUE ~ NA
    )
 )
```


convert dichotomous variables to 0/1

```{r}
psoriasis_data_full <- psoriasis_data_full %>%  
  mutate(across(                                      
    .cols = all_of(c("Diagnos")),  ## for each column listed and "outcome"
    .fns = ~case_when(                              
      . %in% c("L40")   ~ 1,  ## recode L40 (psoriasis diagnosis) to 1
      . %in% c("") ~ 0,       ## not present  to 0
      TRUE ~ NA_real_)        ## otherwise set to missing
    )
  ) 
  
```



```{r}
assign_slekind <- function(par_col, sle_txt_column, sle_dictionary, category2check){

  slekind <- c()
  
  # double check that a list is being read using [[]]
  for (i in seq_along(psoriasis_data_full[[sle_txt_column]])){
    
    sle <- psoriasis_data_full[[sle_txt_column]][[i]]
    par <- psoriasis_data_full[[par_col]][[i]]
    
    if((!is.na(par))&(par == 1)){
        if(sle %in% sle_dictionary){
              category <- names(sle_dictionary[sle_dictionary == sle])
              if(category == category2check){
                slekind <- c(slekind,1)
              } else{
                slekind <- c(slekind,0)
              }
        } else {
          slekind <- c(slekind,0)
        }
    }
    else{
      slekind <- c(slekind,NA)
    }
  }
  
  return(slekind)
}
```


```{r}
psoriasis_data_full["slekindC1_1"] <- assign_slekind(par_col = "par_1", sle_txt_column = "b_44txt", sle_dictionary = year_1_sle_dictionary, category2check = "SLE KIND Child death 1 years")
psoriasis_data_full["slekindC2_1"] <- assign_slekind(par_col = "par_1", sle_txt_column = "b_44txt", sle_dictionary = year_1_sle_dictionary, category2check = "SLE KIND Child death of relative 1 years")
psoriasis_data_full["slekindC3_1"] <- assign_slekind(par_col = "par_1", sle_txt_column = "b_44txt", sle_dictionary = year_1_sle_dictionary, category2check = "SLE KIND Child illness 1 years")
psoriasis_data_full["slekindC4_1"] <- assign_slekind(par_col = "par_1", sle_txt_column = "b_44txt", sle_dictionary = year_1_sle_dictionary, category2check = "SLE KIND Child relative illness 1 years")
psoriasis_data_full["slekindC5_1"] <- assign_slekind(par_col = "par_1", sle_txt_column = "b_44txt", sle_dictionary = year_1_sle_dictionary, category2check = "SLE KIND Child accident 1 years")
################################################################################
psoriasis_data_full["slekindC6_1"] <- assign_slekind(par_col = "par_1", sle_txt_column = "b_44txt", sle_dictionary = year_1_sle_dictionary, category2check = "SLE KIND Child divorce 1 years")
psoriasis_data_full["slekindC7_1"] <- assign_slekind(par_col = "par_1", sle_txt_column = "b_44txt", sle_dictionary = year_1_sle_dictionary, category2check = "SLE KIND Child exposed to violence 1 years")
################################################################################
psoriasis_data_full["slekindC8_1"] <- assign_slekind(par_col = "par_1", sle_txt_column = "b_44txt", sle_dictionary = year_1_sle_dictionary, category2check = "SLE KIND Child unemployed 1 years")
################################################################################
psoriasis_data_full["slekindC9_1"] <- assign_slekind(par_col = "par_1", sle_txt_column = "b_44txt", sle_dictionary = year_1_sle_dictionary, category2check = "SLE KIND Child New family structure 1 years")
################################################################################
psoriasis_data_full["slekindC10_1"] <- assign_slekind(par_col = "par_1", sle_txt_column = "b_44txt", sle_dictionary = year_1_sle_dictionary, category2check = "SLE KIND Child other 1 years")
psoriasis_data_full["unknown_sle_1"] <- assign_slekind(par_col = "par_1", sle_txt_column = "b_44txt", sle_dictionary = year_1_sle_dictionary, category2check = "Unknown")
psoriasis_data_full["no_sle_1"] <- assign_slekind(par_col = "par_1", sle_txt_column = "b_44txt", sle_dictionary = year_1_sle_dictionary, category2check = "No SLE")
```

```{r}
psoriasis_data_full["subjective_slekindC1_3"] <- assign_slekind(par_col = "par_2", sle_txt_column = "c_99txt", sle_dictionary = year_3_sle_dictionary, category2check = "SLE KIND Child death 3 years")
psoriasis_data_full["subjective_slekindC2_3"] <- assign_slekind(par_col = "par_2", sle_txt_column = "c_99txt", sle_dictionary = year_3_sle_dictionary, category2check = "SLE KIND Child death of close person 3 years")
################################################################################
psoriasis_data_full["subjective_slekindC3_3"] <- assign_slekind(par_col = "par_2", sle_txt_column = "c_99txt", sle_dictionary = year_3_sle_dictionary, category2check = "SLE KIND Child illness 3 years")
psoriasis_data_full["subjective_slekindC4_3"] <- assign_slekind(par_col = "par_2", sle_txt_column = "c_99txt", sle_dictionary = year_3_sle_dictionary, category2check = "SLE KIND Child illness special group 3 years")
################################################################################
psoriasis_data_full["subjective_slekindC5_3"] <- assign_slekind(par_col = "par_2", sle_txt_column = "c_99txt", sle_dictionary = year_3_sle_dictionary, category2check = "SLE KIND Child sorrow in the family 3 years")
################################################################################
psoriasis_data_full["subjective_slekindC6_3"] <- assign_slekind(par_col = "par_2", sle_txt_column = "c_99txt", sle_dictionary = year_3_sle_dictionary, category2check = "SLE KIND Child divorce 3 years")
psoriasis_data_full["subjective_slekindC7_3"] <- assign_slekind(par_col = "par_2", sle_txt_column = "c_99txt", sle_dictionary = year_3_sle_dictionary, category2check = "SLE KIND Child exposed to violence 3 years")
################################################################################
psoriasis_data_full["subjective_slekindC8_3"] <- assign_slekind(par_col = "par_2", sle_txt_column = "c_99txt", sle_dictionary = year_3_sle_dictionary, category2check = "SLE KIND Child family problems 3 years")
################################################################################
psoriasis_data_full["subjective_nosle_3"] <- assign_slekind(par_col = "par_2", sle_txt_column = "c_99txt", sle_dictionary = year_3_sle_dictionary, category2check = "No SLE")
```


```{r}
recode_slekind_yes_year_3_5_8 <- function(dataset, par, slekind){
  dataset <- dataset %>%
  dplyr::mutate(
    !!sym(slekind):= case_when(
      !!sym(par) == 1 & !!sym(slekind) == 1 ~ 1,
      !!sym(par) == 1 & is.na(!!sym(slekind)) ~ 0,
      TRUE ~ NA
    )
  )
}

recode_slekind_binary_year_3_5_8 <- function(dataset, par, slekind){
  dataset <- dataset %>%
  dplyr::mutate(
    !!sym(slekind):= case_when(
      !!sym(par) == 1 & !!sym(slekind) == 1 ~ 1,
      !!sym(par) == 1 & !!sym(slekind) == 0 ~ 0,
      !!sym(par) == 1 & is.na(!!sym(slekind)) ~ 0,
      TRUE ~ NA
    )
  )
}

recode_slekind_year_10_12 <- function(dataset, par, slekind_new, slekind){
  dataset <- dataset %>%
  dplyr::mutate(
    !!sym(slekind_new):= case_when(
      !!sym(par) == 1 & !!sym(slekind) == 2 ~ 1,
      !!sym(par) == 1 & is.na(!!sym(slekind)) ~ 0,
      TRUE ~ NA
    )
  )
}
```


```{r}
SLEs <- names(psoriasis_data_full %>%  dplyr::select(starts_with("slekind")&ends_with("_2")))

for (sle in SLEs){
  if(grepl("slekindC", sle, ignore.case = TRUE)){
    psoriasis_data_full <- recode_slekind_binary_year_3_5_8(dataset = psoriasis_data_full, par = "par_2", slekind = sle)
  } else {
    psoriasis_data_full <- recode_slekind_yes_year_3_5_8(dataset = psoriasis_data_full, par = "par_2", slekind = sle)
  }
}
```


child SLE
```{r}
psoriasis_data_full <- psoriasis_data_full %>% 
  dplyr::mutate(
    # SLE KIND Child death
    slekindC1_3 = case_when(
      par_2 == 1 & slekind1_2 == 1 & subjective_nosle_3 == 0 ~ 1,
      par_2 == 1 & slekind1_2 == 1 & subjective_nosle_3 == 1 ~ 0,
      par_2 == 1 & slekind1_2 == 0 ~ 0,
      TRUE ~ NA),
    # SLE KIND Child illness
    slekindC2_3 = case_when(
      par_2 == 1 & slekind2_2 == 1 & subjective_nosle_3 == 0 ~ 1,
      par_2 == 1 & slekind2_2 == 1 & subjective_nosle_3 == 1 ~ 0,
      par_2 == 1 & slekind2_2 == 0 ~ 0,
      TRUE ~ NA), 
    # SLE KIND Child accident
    slekindC3_3 = case_when(
      par_2 == 1 & slekind3_2 == 1 & subjective_nosle_3 == 0 ~ 1,
      par_2 == 1 & slekind3_2 == 1 & subjective_nosle_3 == 1 ~ 0,
      par_2 == 1 & slekind3_2 == 0 ~ 0,
      TRUE ~ NA),     
    # SLE KIND Child divorce
    slekindC4_3 = case_when(
      par_2 == 1 & slekind4_2 == 1 & subjective_nosle_3 == 0 ~ 1,
      par_2 == 1 & slekind4_2 == 1 & subjective_nosle_3 == 1 ~ 0,
      par_2 == 1 & slekind4_2 == 0 ~ 0,
      TRUE ~ NA), 
    # SLE KIND Child violence
    slekindC5_3 = case_when(
      par_2 == 1 & slekind5_2 == 1 & subjective_nosle_3 == 0 ~ 1,
      par_2 == 1 & slekind5_2 == 1 & subjective_nosle_3 == 1 ~ 0,
      par_2 == 1 & slekind5_2 == 0 ~ 0,
      TRUE ~ NA),     
    # SLE KIND Child unemployment
    slekindC6_3 = case_when(
      par_2 == 1 & (slekind6_2 == 1|(slekind7_2 == 1)) & subjective_nosle_3 == 0 ~ 1,
      par_2 == 1 & (slekind6_2 == 1|(slekind7_2 == 1)) & subjective_nosle_3 == 1 ~ 0,
      par_2 == 1 & slekind6_2 == 0 & slekind7_2 == 0 ~ 0,
      TRUE ~ NA)    
  ) 
```


```{r}
SLEs <- names(psoriasis_data_full %>%  dplyr::select(starts_with("slekind")&ends_with("_5")))

for (sle in SLEs){
  if(grepl("slekindC", sle, ignore.case = TRUE)){
    psoriasis_data_full <- recode_slekind_binary_year_3_5_8(dataset = psoriasis_data_full, par = "par_5", slekind = sle)
  } else {
    psoriasis_data_full <- recode_slekind_yes_year_3_5_8(dataset = psoriasis_data_full, par = "par_5", slekind = sle)
  }
}
```


```{r}
psoriasis_data_full["slekindC11_8"] <- psoriasis_data_full$SlekindC11_8
psoriasis_data_full$SlekindC11_8 <- NULL

SLEs <- names(psoriasis_data_full %>% dplyr::select(starts_with("slekind")&ends_with("_8")))

for (sle in SLEs){
  if(grepl("slekindC", sle, ignore.case = TRUE)){
    psoriasis_data_full <- recode_slekind_binary_year_3_5_8(dataset = psoriasis_data_full, par = "par_8", slekind = sle)
  } else {
    psoriasis_data_full <- recode_slekind_yes_year_3_5_8(dataset = psoriasis_data_full, par = "par_8", slekind = sle)
  }
}
```


```{r}
SLEs <- names(psoriasis_data_full %>% dplyr::select(starts_with("gv_64")&ends_with("1")))

for (isle in seq_along(SLEs)){
  
  sle <- SLEs[isle]
  slekind_new <- paste0("slekindC",isle,"_10_12")
  
  psoriasis_data_full <- recode_slekind_year_10_12(
    dataset = psoriasis_data_full, par = "ParP_11", 
    slekind_new = slekind_new, slekind = sle)
}

SLEs <- names(psoriasis_data_full %>% dplyr::select(starts_with("gv_65")&ends_with("1")))

for (isle in seq_along(SLEs)){
  
  sle <- SLEs[isle]
  slekind_new <- paste0("slekind",isle,"_10_12")
  
  psoriasis_data_full <- recode_slekind_year_10_12(
    dataset = psoriasis_data_full, par = "ParP_11", 
    slekind_new = slekind_new, slekind = sle)
}
```

remove cases with other auto immune diseases other than psoriasis

```{r}
print("number of participants with other autoimmune diseases")
print(
  nrow(
    psoriasis_data_full[
      which((
        (psoriasis_data_full$AutoimmDisease == 1)&
          (psoriasis_data_full$Diagnos != 1))),]))

print("number of participants before filtering")
print(nrow(psoriasis_data_full))

print("number of first order heredity psoriasis cases before filtering")
print(psoriasis_data_full %>% 
        filter(first_degree_psoriasis_heredity == 1) %>%
        nrow()
      )

psoriasis_data_full <- psoriasis_data_full %>% 
  filter(!((AutoimmDisease == 1)&(Diagnos != 1)))

# remove rows with missing sex information
psoriasis_data_full <- psoriasis_data_full %>% filter(!is.na(Gender))

print("number of participants after filtering")
print(nrow(psoriasis_data_full))

print("number of first order heredity psoriasis cases after filtering")
print(psoriasis_data_full %>% 
        filter(first_degree_psoriasis_heredity == 1) %>%
        nrow()
      )
```
keep only ABISnr that participated

```{r}
print("number of participants")
print(nrow(psoriasis_data_full))

print("number of first order heredity psoriasis cases")
print(psoriasis_data_full %>% 
        filter(first_degree_psoriasis_heredity == 1) %>%
        nrow()
      )


print("number of psoriasis cases")
print(psoriasis_data_full %>% 
        filter(Diagnos == 1) %>%
        nrow()
      )
```



```{r}
psoriasis_data_cox <- psoriasis_data_full %>%
  dplyr::select(ABISnr, child_dob,  
    # year-1 variables
    ageyC_1, par_1,
    slekindC1_1, slekindC2_1, slekindC3_1, slekindC4_1,  # Death illness
    slekindC1_1, slekindC2_1, slekindC3_1, slekindC4_1,  # accidents
    slekindC6_1, slekindC7_1,                            # Divorce_and_violence
    slekindC8_1,                                         # Unemployment
    slekindC9_1,                                         # New family structure
    # year-3 variables
    ageyC_2, par_2,
    slekind1_2,                                          # SLE KIND Parent death 2 years
    slekind2_2,                                          # SLE KIND Parent illness 2 years
    slekind3_2,                                          # SLE KIND Parent accident 2 years 
    slekind4_2,                                          # SLE KIND Parent divorce 2 years
    slekind5_2,                                          # SLE KIND Parent exposed to violence 2 years
    slekind6_2,                                          # SLE KIND Parent unemployed 2 years
    slekind7_2,                                          # SLE KIND Parent unemployed partner 2 years
    slekind8_2,                                          # SLE KIND Parent other 2 years
    ############################################################################
    slekindC1_3,                                         # SLE KIND Child death
    slekindC2_3,                                         # SLE KIND Child illness
    slekindC3_3,                                         # SLE KIND Child accident
    slekindC4_3,                                         # SLE KIND Child divorce
    slekindC5_3,                                         # SLE KIND Child violence
    slekindC6_3,                                         # SLE KIND Child unemployment
    # year-5 variables
    ageyC_5, par_5,
    slekindC1_5,                                         # Death and illness
    slekindC2_5, slekindC3_5, slekindC4_5,               # New family structure
    slekindC5_5,                                         # conflict at home
    slekindC6_5, slekindC7_5, slekindC8_5,               # Intervention by social services
    slekind1_5, slekind2_5, slekind3_5,                  # Death illness accidents
    slekind4_5, slekind5_5,                              # conflict
    slekind6_5, slekind7_5,                              # unemployment
    # year-8 variables
    par_8, ageyC_8, 
    slekindC1_8, slekindC2_8, slekindC3_8, slekindC4_8,                                                     # Death and illness
    slekindC6_8, slekindC10_8, slekindC11_8, slekindC7_8, slekindC8_8, slekindC9_8,                         # New family structure
    slekindC5_8,                                                                                            # conflict at home
    slekindC12_8, slekindC13_8, slekindC14_8,                                                               # Intervention by social services
    slekind1_8, slekind2_8, slekind3_8, slekind4_8, slekind5_8,                                             # Death illness accidents
    slekind6_8, slekind7_8, slekind8_8,                                                                     # conflict
    slekind9_8, slekind10_8,                                                                                # unemployment
    # year-10-12 variables
    ParP_11, ageyC_10_12, date_10_12_qa_parent,      
    slekindC1_10_12,slekindC2_10_12,slekindC3_10_12,slekindC4_10_12,                                        # Death and illness
    slekindC6_10_12,slekindC10_10_12,slekindC11_10_12, slekindC7_10_12, slekindC8_10_12, slekindC9_10_12,   # New family structure
    slekindC5_10_12,                                                                                        # conflict at home
    slekindC12_10_12,slekindC13_10_12,slekindC14_10_12,                                                     # Intervention by social services
    slekind1_10_12,slekind2_10_12,slekind3_10_12,slekind4_10_12,slekind5_10_12,                             # Death illness accidents
    slekind6_10_12,slekind7_10_12,slekind8_10_12,                                                           # conflict
    slekind9_10_12,slekind10_10_12,                                                                         # unemployment            
    Diagnos, age_first_detection, 
    Gender, first_degree_psoriasis_heredity, originM,
    eduM_b_3kat, eduM_2_3kat, eduM_5_3kat, eduM_8_3kat, eduM_10_12_3kat
    )
```


```{r}
child_psoriasis_data_cox <- psoriasis_data_cox %>%
  dplyr::mutate(
    ############################################################################
    edu_mother_1 =  eduM_b_3kat,
    SLE_KIND_Child_death_illness_1 = pmax(slekindC1_1, slekindC2_1, slekindC3_1, slekindC4_1, na.rm = TRUE), # Death and illness
    SLE_KIND_Child_conflict_at_home_1 = pmax(slekindC6_1, slekindC7_1, na.rm = TRUE),                        # Conflict at home: divorce and/or violence      
    SLE_KIND_Child_new_family_structure_1 = pmax(slekindC6_1, slekindC9_1, na.rm = TRUE),                    # New family structure
    ############################################################################
    edu_mother_3 = pmax(eduM_b_3kat, eduM_2_3kat, na.rm = TRUE),
    SLE_KIND_Child_death_illness_3 = pmax(slekindC1_3, slekindC2_3, na.rm = TRUE),
    SLE_KIND_Child_new_family_structure_3 = slekindC4_3, 
    SLE_KIND_Child_conflict_at_home_3 = pmax(slekindC4_3, slekindC5_3, na.rm = TRUE),
    ############################################################################
    edu_mother_5 = pmax(eduM_b_3kat, eduM_2_3kat, eduM_5_3kat, na.rm = TRUE),
    SLE_KIND_Child_death_illness_5 = slekindC1_5,           
    SLE_KIND_Child_new_family_structure_5 = pmax(slekindC2_5, slekindC3_5, slekindC4_5, na.rm = TRUE), 
    SLE_KIND_Child_conflict_at_home_5 = slekindC5_5,            
    SLE_KIND_Child_social_service_intervention_5 = pmax(slekindC6_5, slekindC7_5, slekindC8_5, na.rm = TRUE),
    ############################################################################
    edu_mother_8 =  pmax(eduM_b_3kat, eduM_2_3kat, eduM_5_3kat, eduM_8_3kat, na.rm = TRUE),
    SLE_KIND_Child_death_illness_8 = pmax(slekindC1_8, slekindC2_8, slekindC3_8, slekindC4_8, na.rm = TRUE),           
    SLE_KIND_Child_new_family_structure_8 = pmax(slekindC6_8, slekindC10_8, slekindC11_8, slekindC7_8, slekindC8_8, slekindC9_8, na.rm = TRUE), 
    SLE_KIND_Child_conflict_at_home_8 = slekindC5_8,            
    SLE_KIND_Child_social_service_intervention_8 = pmax(slekindC12_8, slekindC13_8, slekindC14_8, na.rm = TRUE),
    ############################################################################
    edu_mother_10_12 =  pmax(eduM_b_3kat, eduM_2_3kat, eduM_5_3kat, eduM_8_3kat, eduM_10_12_3kat, na.rm = TRUE),
    SLE_KIND_Child_death_illness_10_12 = pmax(slekindC1_10_12, slekindC2_10_12, slekindC3_10_12, slekindC4_10_12, na.rm = TRUE),           
    SLE_KIND_Child_new_family_structure_10_12 = pmax(slekindC6_10_12, slekindC7_10_12, slekindC8_10_12, slekindC9_10_12, slekindC10_10_12, slekindC11_10_12, na.rm = TRUE), 
    SLE_KIND_Child_conflict_at_home_10_12 = slekindC5_10_12,            
    SLE_KIND_Child_social_service_intervention_10_12 = pmax(slekindC12_10_12, slekindC13_10_12, slekindC14_10_12, na.rm = TRUE)
    )
```

```{r}
parent_psoriasis_data_cox <- psoriasis_data_cox %>%
  dplyr::mutate(
    ############################################################################
    edu_mother_3 = pmax(eduM_b_3kat, eduM_2_3kat, na.rm = TRUE),
    SLE_KIND_Parent_death_illness_accident_3 = pmax(slekind1_2, slekind2_2, slekind3_2, na.rm = TRUE),
    SLE_KIND_Parent_conflict_3 = pmax(slekind4_2, slekind5_2, na.rm = TRUE),
    SLE_KIND_Parent_unemployment_3 = pmax(slekind6_2, slekind7_2, na.rm = TRUE),
    ############################################################################
    edu_mother_5 = pmax(eduM_b_3kat, eduM_2_3kat, eduM_5_3kat, na.rm = TRUE),
    SLE_KIND_Parent_death_illness_accident_5 = pmax(slekind1_5, slekind2_5, slekind3_5, na.rm =TRUE),
    SLE_KIND_Parent_conflict_5 = pmax(slekind4_5, slekind5_5, na.rm =TRUE),               
    SLE_KIND_Parent_unemployment_5 = pmax(slekind6_5, slekind7_5, na.rm = TRUE),               
    ######################################
    edu_mother_8 =  pmax(eduM_b_3kat, eduM_2_3kat, eduM_5_3kat, eduM_8_3kat, na.rm = TRUE),
    SLE_KIND_Parent_death_illness_accident_8 = pmax(slekind1_8, slekind2_8, slekind3_8, slekind4_8, slekind5_8, na.rm =TRUE),
    SLE_KIND_Parent_conflict_8 = pmax(slekind6_8, slekind7_8, slekind8_8, na.rm =TRUE),               
    SLE_KIND_Parent_unemployment_8 = pmax(slekind9_8, slekind10_8, na.rm = TRUE),   
    ############################################################################
    edu_mother_10_12 =  pmax(eduM_b_3kat, eduM_2_3kat, eduM_5_3kat, eduM_8_3kat, eduM_10_12_3kat, na.rm = TRUE),
    SLE_KIND_Parent_death_illness_accident_10_12 = pmax(slekind1_10_12, slekind2_10_12, slekind3_10_12, slekind4_10_12, slekind5_10_12, na.rm =TRUE),
    SLE_KIND_Parent_conflict_10_12 = pmax(slekind6_10_12, slekind7_10_12, slekind8_10_12, na.rm =TRUE),               
    SLE_KIND_Parent_unemployment_10_12 = pmax(slekind9_10_12, slekind10_10_12, na.rm = TRUE),    
    )
```

filter out cases before 1-year follow-up
```{r}
child_psoriasis_data_cox <- child_psoriasis_data_cox %>% 
  dplyr::filter((par_1 == 1)&(is.na(date_10_12_qa_parent)|date_10_12_qa_parent < as.Date("2012-12-31"))) %>%
  dplyr::filter(!(Diagnos == 1 & ageyC_1 > age_first_detection)) 

print("Number of cases in participants post 1-year follow-up")
print(nrow(child_psoriasis_data_cox %>% dplyr::filter(Diagnos == 1)))
```
filter out cases before 3-year follow-up
```{r}
parent_psoriasis_data_cox <- parent_psoriasis_data_cox %>% 
  dplyr::filter((par_2 == 1)&(is.na(date_10_12_qa_parent)|date_10_12_qa_parent < as.Date("2012-12-31"))) %>%
  dplyr::filter(!(Diagnos == 1 & ageyC_2 > age_first_detection)) 

print("Number of cases in participants post 3-year follow-up")
print(nrow(parent_psoriasis_data_cox %>% dplyr::filter(Diagnos == 1)))
```

```{r}
child_year_1_data <- child_psoriasis_data_cox %>%
  dplyr::filter(par_1 == 1) %>%
  dplyr::mutate(
    follow_up_time = ageyC_1,
    event = ifelse(Diagnos == 0| is.na(age_first_detection)|age_first_detection > follow_up_time, 0,1),
    edu_mother =  edu_mother_1,
    Death_and_illness = SLE_KIND_Child_death_illness_1,                         # Death and illness
    New_family_structure = SLE_KIND_Child_new_family_structure_1,               # New family structure
    Conflict_at_home = SLE_KIND_Child_conflict_at_home_1,                       # conflict at home
    Intervention_by_social_services = NA,                                       # Intervention by social services
  ) %>%
  dplyr::select(ABISnr, 
                par_1, par_2, par_5, par_8, ParP_11,
                follow_up_time,
                event,
                ######################################
                Death_and_illness,   
                New_family_structure,
                Conflict_at_home,     
                Intervention_by_social_services, 
                ######################################
                Gender, 
                originM,
                edu_mother,
                first_degree_psoriasis_heredity,
                age_first_detection
                ) %>%
  dplyr::filter(!is.na(follow_up_time))
```

```{r}
child_year_3_data <- child_psoriasis_data_cox %>%
  dplyr::filter(par_2 == 1) %>%
  dplyr::mutate(
    follow_up_time = ageyC_2,
    event = ifelse(Diagnos == 0| is.na(age_first_detection)|age_first_detection > follow_up_time, 0,1),
    edu_mother =  edu_mother_3,
    Death_and_illness = SLE_KIND_Child_death_illness_3,                         # Death and illness
    New_family_structure = SLE_KIND_Child_new_family_structure_3,               # New family structure
    Conflict_at_home = SLE_KIND_Child_conflict_at_home_3,                       # conflict at home
    Intervention_by_social_services = NA,                                       # Intervention by social services
  ) %>%
  dplyr::select(ABISnr, 
                par_1, par_2, par_5, par_8, ParP_11,
                follow_up_time,
                event,
                ######################################
                Death_and_illness,   
                New_family_structure,
                Conflict_at_home,     
                Intervention_by_social_services, 
                ######################################
                Gender, 
                originM,
                edu_mother,
                first_degree_psoriasis_heredity,
                age_first_detection
                ) %>%
  dplyr::filter(!is.na(follow_up_time))
```

```{r}
child_year_5_data <- child_psoriasis_data_cox %>%
  dplyr::filter(par_5 == 1) %>%
  dplyr::mutate(
    follow_up_time = ageyC_5,
    event = ifelse(Diagnos == 0| is.na(age_first_detection)|age_first_detection > follow_up_time, 0,1),
    edu_mother =  edu_mother_5,
    Death_and_illness = SLE_KIND_Child_death_illness_5,                              # Death and illness
    New_family_structure = SLE_KIND_Child_new_family_structure_5,                    # New family structure
    Conflict_at_home = SLE_KIND_Child_conflict_at_home_5,                            # conflict at home
    Intervention_by_social_services = SLE_KIND_Child_social_service_intervention_5,  # Intervention by social services
  ) %>%
  dplyr::select(ABISnr, 
                par_1, par_2, par_5, par_8, ParP_11,
                follow_up_time,
                event,
                ######################################
                Death_and_illness,   
                New_family_structure,
                Conflict_at_home,     
                Intervention_by_social_services, 
                ######################################
                Gender, 
                originM,
                edu_mother,
                first_degree_psoriasis_heredity,
                age_first_detection
                ) %>%
  dplyr::filter(!is.na(follow_up_time))
```

```{r}
child_year_5_data <- child_psoriasis_data_cox %>%
  dplyr::filter(par_5 == 1) %>%
  dplyr::mutate(
    follow_up_time = ageyC_5,
    event = ifelse(Diagnos == 0| is.na(age_first_detection)|age_first_detection > follow_up_time, 0,1),
    edu_mother =  edu_mother_5,
    Death_and_illness = SLE_KIND_Child_death_illness_5,                              # Death and illness
    New_family_structure = SLE_KIND_Child_new_family_structure_5,                    # New family structure
    Conflict_at_home = SLE_KIND_Child_conflict_at_home_5,                            # conflict at home
    Intervention_by_social_services = SLE_KIND_Child_social_service_intervention_5,  # Intervention by social services
  ) %>%
  dplyr::select(ABISnr, 
                par_1, par_2, par_5, par_8, ParP_11,
                follow_up_time,
                event,
                ######################################
                Death_and_illness,   
                New_family_structure,
                Conflict_at_home,     
                Intervention_by_social_services, 
                ######################################
                Gender, 
                originM,
                edu_mother,
                first_degree_psoriasis_heredity,
                age_first_detection
                ) %>%
  dplyr::filter(!is.na(follow_up_time))
```

```{r}
child_year_8_data <- child_psoriasis_data_cox %>%
  dplyr::filter(par_8 == 1) %>%
  dplyr::mutate(
    follow_up_time = ageyC_8,
    event = ifelse(Diagnos == 0| is.na(age_first_detection)|age_first_detection > follow_up_time, 0,1),
    edu_mother =  edu_mother_8,
    Death_and_illness = SLE_KIND_Child_death_illness_8,                              # Death and illness
    New_family_structure = SLE_KIND_Child_new_family_structure_8,                    # New family structure
    Conflict_at_home = SLE_KIND_Child_conflict_at_home_8,                            # conflict at home
    Intervention_by_social_services = SLE_KIND_Child_social_service_intervention_8,  # Intervention by social services
  ) %>%
  dplyr::select(ABISnr, 
                par_1, par_2, par_5, par_8, ParP_11,
                follow_up_time,
                event,
                ######################################
                Death_and_illness,   
                New_family_structure,
                Conflict_at_home,     
                Intervention_by_social_services, 
                ######################################
                Gender, 
                originM,
                edu_mother,
                first_degree_psoriasis_heredity,
                age_first_detection
                ) %>%
  dplyr::filter(!is.na(follow_up_time))
```


```{r}
child_year_10_12_data <- child_psoriasis_data_cox %>%
  dplyr::filter(ParP_11 == 1) %>%
  dplyr::mutate(
    follow_up_time = ageyC_10_12,
    event = ifelse(Diagnos == 0| is.na(age_first_detection)|age_first_detection > follow_up_time, 0,1),
    edu_mother =  edu_mother_10_12,
    Death_and_illness = SLE_KIND_Child_death_illness_10_12,                              # Death and illness
    New_family_structure = SLE_KIND_Child_new_family_structure_10_12,                    # New family structure
    Conflict_at_home = SLE_KIND_Child_conflict_at_home_10_12,                            # conflict at home
    Intervention_by_social_services = SLE_KIND_Child_social_service_intervention_10_12,  # Intervention by social services
  ) %>%
  dplyr::select(ABISnr, 
                par_1, par_2, par_5, par_8, ParP_11,
                follow_up_time,
                event,
                ######################################
                Death_and_illness,   
                New_family_structure,
                Conflict_at_home,     
                Intervention_by_social_services, 
                ######################################
                Gender, 
                originM,
                edu_mother,
                first_degree_psoriasis_heredity,
                age_first_detection
                ) %>%
  dplyr::filter(!is.na(follow_up_time))
```


```{r}
child_cox_data <- rbind(
  child_year_1_data,
  child_year_3_data,
  child_year_5_data,
  child_year_8_data,
  child_year_10_12_data
)

names(child_cox_data) <- c("ABISnr", "par_1", "par_2", "par_5", "par_8", "ParP_11", 
                           "follow_up_time", "event", 
                           "Death_and_illness", "New_family_structure", 
                           "Conflict_at_home", "Intervention_by_social_services", 
                           "Sex", "Origin_mother", "Education_mother", 
                           "First_degree_psoriasis_heredity", "age_first_detection")
```


clean cox data
```{r}
child_cox_data_censored <- delete_data_after_censor_date(child_cox_data)
```

remove variables that will not be used
```{r}
child_cox_data_censored_test <- child_cox_data_censored %>%
  dplyr::select(-c(par_1,par_2,par_5,par_8,ParP_11, age_first_detection))
```

```{r}
# factorize categorical variables
for (cv in names(child_cox_data_censored_test)[-c(2:3)]){
  print(cv)
  if(cv == "ABISnr"){
    next
  }
  if (cv == "Education_mother"){
    child_cox_data_censored_test[[cv]] <- factor(child_cox_data_censored_test[[cv]], levels = c(1,2,0))
  } else{
    child_cox_data_censored_test[[cv]] <- factor(child_cox_data_censored_test[[cv]])    
  }
  
  if(cv == "Education_mother"){
    curr_levels <- c("Gymnasium or folk high school", "College at least 1 year", "9 year primary school")
  } else if(cv == "Origin_Mother"){
    curr_levels <- c("Born in Sweden", "Born outside Sweden")
  } else if(cv == "Sex"){
    curr_levels <- c("Boy", "Girl")
  } else{
    curr_levels <- c("No", "Yes")
  }
  
  levels(child_cox_data_censored_test[[cv]]) <- curr_levels
  child_cox_data_censored_test[[cv]] <- droplevels(child_cox_data_censored_test[[cv]])
}

#  child_cox_data_censored <- child_cox_data_censored[complete.cases(child_cox_data_censored),]
```

```{r}
cox_model_1 <- function(cox_data, predictor_var){
  
  model <- survival::coxph(
    as.formula(paste0("survival::Surv(follow_up_time, event) ~ ",predictor_var)), 
    data = cox_data)
  return(model)
}

cox_model_2 <- function(cox_data, predictor_var){
  
  model <- survival::coxph(
    as.formula(paste0("survival::Surv(follow_up_time, event) ~ ",predictor_var, " + Sex + Origin_mother + Education_mother + First_degree_psoriasis_heredity")), 
    data = cox_data)
  return(model)
  }
```



```{r}
i = 1

child_unadjusted_models <- list()
child_unadjusted_survival_curves <- list()

factors_to_test <- c("Death_and_illness", 
                     "New_family_structure",
                     "Conflict_at_home", 
                     "Intervention_by_social_services")

for (predictor_var in names(child_cox_data_censored_test)[-c(1:3)]){
  
  child_model_1 <- cox_model_1(child_cox_data_censored_test, predictor_var)
  
  if(predictor_var %in% factors_to_test){
    
      child_unadjusted_models[[predictor_var]] <- child_model_1
      
      survival_object <- survival::survfit(
        child_model_1,
        data = child_cox_data_censored_test)
      
      child_unadjusted_survival_curves[[predictor_var]] <- survminer::ggsurvplot(
        survival_object,  
        title = paste0("Unadjusted Survival Curve for ", gsub("_", " ",predictor_var)), 
        xlab = "Time", ylab = "Survival Probability")
  }
  
  curr_summary <- broom::tidy(child_model_1, exponentiate = TRUE, conf.int = TRUE) %>%
    dplyr::mutate_if(is.numeric, round, 2)
  
  if(i==1){
    child_model_1_summary <- curr_summary
  } else {
    child_model_1_summary <- rbind(child_model_1_summary, curr_summary)
  }
  i = i + 1
}

child_model_1_summary <- child_model_1_summary %>%
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>%
  dplyr::mutate(
    term = gsub("_", " ",
        gsub("9 year primary school", " (9 year primary school)", 
        gsub("Gymnasium or folk high school", " (Gymnasium or folk high school)",
        gsub("Born outside Sweden", " (Born outside Sweden)",
        gsub("Girl", " (Girl)",gsub("Yes", " (Yes)",term))))))
    )
```
 

```{r}
i = 1

child_adjusted_models <- list()
child_adjusted_survival_curves <- list()

factors_to_test <- c("Death_and_illness", 
                     "New_family_structure",
                     "Conflict_at_home", 
                     "Intervention_by_social_services")

for (predictor_var in names(child_cox_data_censored_test)[-c(1:3,8:11)]){
  
  child_model_2 <- cox_model_2(child_cox_data_censored_test, predictor_var)
  
    if(predictor_var %in% factors_to_test){
      child_adjusted_models[[predictor_var]] <- child_model_2
      
      survival_object <- survival::survfit(
      child_model_2,
      data = child_cox_data_censored_test)
      
      child_adjusted_survival_curves[[predictor_var]] <- survminer::ggsurvplot(
      survival_object,  
      title = paste0("Adjusted Survival Curve for ", gsub("_", " ",predictor_var)), 
      xlab = "Time", ylab = "Survival Probability")
  }
  
  curr_summary <- broom::tidy(child_model_2, exponentiate = TRUE, conf.int = TRUE) %>%
    dplyr::mutate_if(is.numeric, round, 2)
  
  if(i==1){
    child_model_2_summary <- curr_summary[1,]
  } else {
    child_model_2_summary <- rbind(child_model_2_summary, curr_summary[1,])
  }
  i = i + 1
}

child_model_2_summary <- child_model_2_summary %>%
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>%
  dplyr::mutate(
    term = gsub("_", " ",
        gsub("9 year primary school", " (9 year primary school)", 
        gsub("Gymnasium or folk high school", " (Gymnasium or folk high school)",
        gsub("Born outside Sweden", " (Born outside Sweden)",
        gsub("Girl", " (Girl)",gsub("Yes", " (Yes)",term))))))
    )

# names(child_model_2_summary) <- c("Feature", "aHR", "Conf.low", "Conf.high", "P-value")
```
forest plot
```{r}
child_model_2_plot <- child_model_2_summary %>% dplyr::mutate(conf.high = ifelse(is.infinite(conf.high)|conf.high>10.0,10,conf.high))

child_plot <- ggplot(child_model_2_plot[complete.cases(child_model_2_plot),], 
            aes(x = term, y = estimate,
                ymin = conf.low, ymax = conf.high, fill = "black", color = term)) +
  #specify position here
  geom_linerange(linewidth = 5,position=position_dodge(width = 0.25)) +
  geom_hline(yintercept=1, lty=2) +
  #specify position here too
  geom_point(size=3, shape=21, colour="black", stroke = 0.5, fill = "black",
             position=position_dodge(width = 0.25), ) +
  guides(fill = FALSE, color = FALSE) +
  scale_x_discrete(name="Serious Life Event (Child)") +
  scale_y_continuous(name="Hazard Ratio") +#, limits = c(0.0, 5.6)) +
  coord_flip() +
  theme_minimal()

child_plot <- child_plot + theme(legend.position=c(0.75,0.65)) +
  annotate("text", x = 0.5, y = 0.25, label = "  Protective   ") +
  annotate("text", x = 0.5, y = 1.75, label = "  Harmful")

filename <- paste0(output_dir,"\\child_SLE_adjusted_cox_hazard_model.jpeg")
ggsave(filename,child_plot)
```


```{r}
parent_year_3_data <- parent_psoriasis_data_cox %>%
  dplyr::filter(par_2 == 1) %>%
  dplyr::mutate(
    follow_up_time = ageyC_2,
    event = ifelse(Diagnos == 0| is.na(age_first_detection)|age_first_detection > follow_up_time, 0,1),
    edu_mother =  edu_mother_3,
    Death_illness_accidents = SLE_KIND_Parent_death_illness_accident_3,         # Death illness accidents
    Conflict = SLE_KIND_Parent_conflict_3,                                      # conflict
    Unemployment = SLE_KIND_Parent_unemployment_3                               # unemployment
  ) %>%
  dplyr::select(ABISnr, 
                par_1, par_2, par_5, par_8, ParP_11,
                follow_up_time,
                event,
                ######################################
                Death_illness_accidents,
                Conflict,
                Unemployment,
                ######################################
                Gender, 
                originM,
                edu_mother,
                first_degree_psoriasis_heredity,
                age_first_detection
                ) %>%
  dplyr::filter(!is.na(follow_up_time))
```


```{r}
parent_year_5_data <- parent_psoriasis_data_cox %>%
  dplyr::filter(par_5 == 1) %>%
  dplyr::mutate(
    follow_up_time = ageyC_5,
    event = ifelse(Diagnos == 0| is.na(age_first_detection)|age_first_detection > follow_up_time, 0,1),
    edu_mother =  edu_mother_5,
    Death_illness_accidents = SLE_KIND_Parent_death_illness_accident_5,         # Death illness accidents
    Conflict = SLE_KIND_Parent_conflict_5,                                      # conflict
    Unemployment = SLE_KIND_Parent_unemployment_5                               # unemployment
  ) %>%
  dplyr::select(ABISnr, 
                par_1, par_2, par_5, par_8, ParP_11,
                follow_up_time,
                event,
                ######################################
                Death_illness_accidents,
                Conflict,
                Unemployment,
                ######################################
                Gender, 
                originM,
                edu_mother,
                first_degree_psoriasis_heredity,
                age_first_detection
                ) %>%
  dplyr::filter(!is.na(follow_up_time))
```


```{r}
parent_year_8_data <- parent_psoriasis_data_cox %>%
  dplyr::filter(par_8 == 1) %>%
  dplyr::mutate(
    follow_up_time = ageyC_8,
    event = ifelse(Diagnos == 0| is.na(age_first_detection)|age_first_detection > follow_up_time, 0,1),
    edu_mother =  edu_mother_8,
    Death_illness_accidents = SLE_KIND_Parent_death_illness_accident_8,         # Death illness accidents
    Conflict = SLE_KIND_Parent_conflict_8,                                      # conflict
    Unemployment = SLE_KIND_Parent_unemployment_8                               # unemployment
  ) %>%
  dplyr::select(ABISnr, 
                par_1, par_2, par_5, par_8, ParP_11,
                follow_up_time,
                event,
                ######################################
                Death_illness_accidents,
                Conflict,
                Unemployment,
                ######################################
                Gender, 
                originM,
                edu_mother,
                first_degree_psoriasis_heredity,
                age_first_detection
                ) %>%
  dplyr::filter(!is.na(follow_up_time))
```


```{r}
parent_year_10_12_data <- parent_psoriasis_data_cox %>%
  dplyr::filter(ParP_11 == 1) %>%
  dplyr::mutate(
    follow_up_time = ageyC_10_12,
    event = ifelse(Diagnos == 0| is.na(age_first_detection)|age_first_detection > follow_up_time, 0,1),
    edu_mother =  edu_mother_10_12,
    Death_illness_accidents = SLE_KIND_Parent_death_illness_accident_10_12,     # Death illness accidents
    Conflict = SLE_KIND_Parent_conflict_10_12,                                  # conflict
    Unemployment = SLE_KIND_Parent_unemployment_10_12                           # unemployment
  ) %>%
  dplyr::select(ABISnr, 
                par_1, par_2, par_5, par_8, ParP_11,
                follow_up_time,
                event,
                ######################################
                Death_illness_accidents,
                Conflict,
                Unemployment,
                ######################################
                Gender, 
                originM,
                edu_mother,
                first_degree_psoriasis_heredity,
                age_first_detection
                ) %>%
  dplyr::filter(!is.na(follow_up_time))
```


```{r}
parent_cox_data <- rbind(
  parent_year_3_data,
  parent_year_5_data,
  parent_year_8_data,
  parent_year_10_12_data
)
names(parent_cox_data) <- c("ABISnr", "par_1", "par_2", "par_5", "par_8", "ParP_11", 
                           "follow_up_time", "event", 
                           "Death_illness_accidents", "Conflict","Unemployment", 
                           "Sex", "Origin_mother", "Education_mother", 
                           "First_degree_psoriasis_heredity", "age_first_detection")
```


clean cox data
```{r}
parent_cox_data_censored <- delete_data_after_censor_date(parent_cox_data)
```

remove variables that will not be used
```{r}
parent_cox_data_censored_test <- parent_cox_data_censored %>%
  dplyr::select(-c(par_1,par_2,par_5,par_8,ParP_11, age_first_detection))
```

```{r}
# factorize categorical variables
for (cv in names(parent_cox_data_censored_test)[-c(2:3)]){
  print(cv)
  if(cv == "ABISnr"){
    next
  }
  if (cv == "Education_mother"){
    parent_cox_data_censored_test[[cv]] <- factor(parent_cox_data_censored_test[[cv]], levels = c(1,2,0))
  } else{
    parent_cox_data_censored_test[[cv]] <- factor(parent_cox_data_censored_test[[cv]])    
  }
  
  if(cv == "Education_mother"){
    curr_levels <- c("Gymnasium or folk high school", "College at least 1 year", "9 year primary school")
  } else if(cv == "Origin_Mother"){
    curr_levels <- c("Born in Sweden", "Born outside Sweden")
  } else if(cv == "Sex"){
    curr_levels <- c("Boy", "Girl")
  } else{
    curr_levels <- c("No", "Yes")
  }
  
  levels(parent_cox_data_censored_test[[cv]]) <- curr_levels
  parent_cox_data_censored_test[[cv]] <- droplevels(parent_cox_data_censored_test[[cv]])
}

#  parent_cox_data_censored <- parent_cox_data_censored[complete.cases(parent_cox_data_censored),]
```

```{r}
cox_model_1 <- function(cox_data, predictor_var){
  
  model <- survival::coxph(
    as.formula(paste0("survival::Surv(follow_up_time, event) ~ ",predictor_var)), 
    data = cox_data)
  return(model)
}

cox_model_2 <- function(cox_data, predictor_var){
  
  model <- survival::coxph(
    as.formula(paste0("survival::Surv(follow_up_time, event) ~ ",predictor_var, " + Sex + Origin_mother + Education_mother + First_degree_psoriasis_heredity")), 
    data = cox_data)
  return(model)
  }
```

```{r}
i = 1
for (predictor_var in names(parent_cox_data_censored_test)[-c(1:3)]){
  
  parent_model_1 <- cox_model_1(parent_cox_data_censored_test, predictor_var)
  curr_summary <- broom::tidy(parent_model_1, exponentiate = TRUE, conf.int = TRUE) %>%
    dplyr::mutate_if(is.numeric, round, 2)
  
  if(i==1){
    parent_model_1_summary <- curr_summary
  } else {
    parent_model_1_summary <- rbind(parent_model_1_summary, curr_summary)
  }
  i = i + 1
}

parent_model_1_summary <- parent_model_1_summary %>%
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>%
  dplyr::mutate(
    term = gsub("_", " ",
        gsub("9 year primary school", " (9 year primary school)", 
        gsub("Gymnasium or folk high school", " (Gymnasium or folk high school)",
        gsub("Born outside Sweden", " (Born outside Sweden)",
        gsub("Girl", " (Girl)",gsub("Yes", " (Yes)",term))))))
    )

```

```{r}
i = 1
for (predictor_var in names(parent_cox_data_censored_test)[-c(1:3,7:11)]){
  
  parent_model_2 <- cox_model_2(parent_cox_data_censored_test, predictor_var)
  curr_summary <- broom::tidy(parent_model_2, exponentiate = TRUE, conf.int = TRUE) %>%
    dplyr::mutate_if(is.numeric, round, 2)
  
  if(i==1){
    parent_model_2_summary <- curr_summary[1,]
  } else {
    parent_model_2_summary <- rbind(parent_model_2_summary, curr_summary[1,])
  }
  i = i + 1
}

parent_model_2_summary <- parent_model_2_summary %>%
  dplyr::select(term, estimate, conf.low, conf.high, p.value) %>%
  dplyr::mutate(
    term = gsub("_", " ",
        gsub("9 year primary school", " (9 year primary school)", 
        gsub("Gymnasium or folk high school", " (Gymnasium or folk high school)",
        gsub("Born outside Sweden", " (Born outside Sweden)",
        gsub("Girl", " (Girl)",gsub("Yes", " (Yes)",term))))))
    )

# names(parent_model_2_summary) <- c("Feature", "aHR", "Conf.low", "Conf.high", "P-value")
```

forest plot
```{r}
parent_model_2_plot <- parent_model_2_summary %>% dplyr::mutate(conf.high = ifelse(is.infinite(conf.high)|conf.high>10.0,10,conf.high))

parent_plot <- ggplot(parent_model_2_plot[complete.cases(parent_model_2_plot),], 
            aes(x = term, y = estimate,
                ymin = conf.low, ymax = conf.high, fill = "black", color = term)) +
  #specify position here
  geom_linerange(linewidth = 5,position=position_dodge(width = 0.25)) +
  geom_hline(yintercept=1, lty=2) +
  #specify position here too
  geom_point(size=3, shape=21, colour="black", stroke = 0.5, fill = "black",
             position=position_dodge(width = 0.25), ) +
  guides(fill = FALSE, color = FALSE) +
  scale_x_discrete(name="Serious Life Event (Parent)") +
  scale_y_continuous(name="Hazard Ratio") +#, limits = c(0.0, 5.6)) +
  coord_flip() +
  theme_minimal()

parent_plot <- parent_plot + theme(legend.position=c(0.75,0.65)) +
  annotate("text", x = 0.5, y = 0.25, label = "  Protective   ") +
  annotate("text", x = 0.5, y = 1.75, label = "  Harmful")


filename <- paste0(output_dir,"\\parent_SLE_adjusted_cox_hazard_model.jpeg")
ggsave(filename,parent_plot)
```


