---
title: "Cox Proportional Hazards model"
author: "Debojyoti Das"
date: "2025-07-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

set output directory
```{r output_dir, message=FALSE, warning=FALSE}
output_dir <- "/Users/debda22/OneDrive - Linköpings universitet/T1D_vitamin_D_results/cox_analysi_revision"

if(!dir.exists(output_dir)){
  dir.create(output_dir, recursive = TRUE)
} else {
  cat(paste0(output_dir," is already present!!\n"))
}
```

load libraries
```{r load_libraries, message=FALSE, warning=FALSE}
pacman::p_load(
  haven,        
  zoo,
  sjlabelled,
  janitor,       
  data.table,   
  ggplot2,
  comprehenr,  
  stringr,     
  rstatix,
  dplyr,
  tidyr,
  forcats,
  officer,
  cowplot,
  ggExtra,
  ggpubr,
  GGally,
  janitor,
  ggrepel,
  ggtext,
  parameters,
  lme4,
  broom,
  pscl,
  tibble,
  knitr,
  ggeffects,
  cobalt,
  WeightIt,
  marginaleffects,
  survival,
  purrr,
  nnet
  )
```


```{r word_sect_properties}
sect_properties <- prop_section(
  page_size = page_size(
    orient = "landscape",
    height = 11.7,
    width = 8.3
  ),
  page_margins = page_mar(
    top = 0.5,
    bottom = 0.5,
    left = 0.5,
    right = 0.5,
    header = 0.3,
    footer = 0.3,
    gutter = 0
  ),
  type = "continuous"
)
```

define paths
```{r define_paths, eval=TRUE}
data_root <- "/Users/debda22/OneDrive - Linköpings universitet/"
data_root <- paste0(data_root,"Linkoping_projects/ABIS_RAW_DATA_SPSS/")
```

load parameters and helper functions
```{r}
source(paste0(data_root,"../../Rmd_and_R_files/T1D_projects/all_parameters_case_control_T1D_vitamin_D.R"))
source(paste0(data_root,"../../Rmd_and_R_files/T1D_projects/cox_T1D_helper.R"))
```


read/load vitamin_d_case_control_plus_abis for analysis
```{r}
vitamin_d_case_control_plus_abis <- read.table(paste0(data_root, "/vitamin_d/vitamin_d_case_control_plus_abis.csv"))
```

reorder factors
```{r}
vitamin_d_case_control_plus_abis$First_degree_T1D_heredity_cat <- factor(
  vitamin_d_case_control_plus_abis$First_degree_T1D_heredity_cat,
  levels = c("No", "Yes")
)

vitamin_d_case_control_plus_abis$Smoking_during_pregnancy <- factor(
  vitamin_d_case_control_plus_abis$Smoking_during_pregnancy,
  levels = c("No", "Yes")
)

vitamin_d_case_control_plus_abis$Parity <- factor(
  vitamin_d_case_control_plus_abis$Parity,
  levels = c("First parity", "Previous parity")
)

vitamin_d_case_control_plus_abis$Preterm_dichotomous <- factor(
  vitamin_d_case_control_plus_abis$Preterm_dichotomous,
  levels = c("Normal", "Preterm")
)

vitamin_d_case_control_plus_abis$Season <- factor(
  vitamin_d_case_control_plus_abis$Season,
  levels = c("Autumn", "Spring", "Summer", "Winter")
)

vitamin_d_case_control_plus_abis$Sex_cat <- factor(
  vitamin_d_case_control_plus_abis$Sex_cat,
  levels = c("Boy", "Girl")
)

vitamin_d_case_control_plus_abis$Fish_any_pregnancy <- factor(
  vitamin_d_case_control_plus_abis$Fish_any_pregnancy,
  levels = c("Seldom", "1–2 times week or more"))

vitamin_d_case_control_plus_abis$Delivery_mode <- factor(
  vitamin_d_case_control_plus_abis$Delivery_mode,
  levels = c("Normal","caesarean", "other problem"),
  labels = c("Normal","Caesarean", "Other problem"))

vitamin_d_case_control_plus_abis$Birth_weight_cat <- factor(
  vitamin_d_case_control_plus_abis$Birth_weight_cat,
  levels = c("q1", "q2", "q3", "q4")
)
```


Cox analysis

conducting a time-to-event Cox regression with:

 - Time scale: Age (from birth)

 - Exposure: Fixed (cord blood 25(OH)D)

 - Follow-up ends: December 31, 2020

Outcome: T1D diagnosis
```{r event_time, warning=FALSE}
vitamin_d_case_control_plus_abis$start_age <- 0  # since follow-up starts at birth

vitamin_d_case_control_plus_abis$INDATUMA <- as.Date(vitamin_d_case_control_plus_abis$INDATUMA, format = "%Y-%m-%d")
```


```{r event_time_formulation, warning=FALSE}
# Set exit date based on whether T1D occurred before 2020-12-31
vitamin_d_case_control_plus_abis$exit_date <- as.Date(ifelse(
  !is.na(vitamin_d_case_control_plus_abis$T1D_age_diagnosis) &
    vitamin_d_case_control_plus_abis$INDATUMA<= as.Date("2020-12-31"),
  vitamin_d_case_control_plus_abis$INDATUMA,
  as.Date("2020-12-31")
))
```


```{r event_time_calculation, warning=FALSE}
# Calculate age at exit in years
vitamin_d_case_control_plus_abis$exit_age <- as.numeric(
  difftime(
    vitamin_d_case_control_plus_abis$exit_date,
    vitamin_d_case_control_plus_abis$child_dob,
    units = "days"
  )
) / 365.25

vitamin_d_case_control_plus_abis$event <- ifelse(
  !is.na(vitamin_d_case_control_plus_abis$INDATUMA) &
    vitamin_d_case_control_plus_abis$INDATUMA <= as.Date("2020-12-31"),
  1,
  0
)
```


```{r cox_data, warning=FALSE}
df_cox_full <- vitamin_d_case_control_plus_abis %>%
  dplyr::select(all_of(
    c(
      "ABISnr",
      "DBP",
      "conc_25OHD",
      "diagnosis",
      "child_dob",
      "First_degree_T1D_heredity_cat",
      "Preterm_group",
      "Preterm_dichotomous",
      "Delivery_mode",
      "Gestational_age",
      "Birth_weight",
      "Birth_weight_cat",
      "Birth_height",
      "Season",
      "Parity",
      "Smoking_during_pregnancy",
      "Vitamin_mineral_supplement_pregnancy",
      "Fish_lake_pregnancy",
      "Fish_Baltic_pregnancy",
      "Fish_other_pregnancy",
      "Fish_any_pregnancy",
      "Sex_cat",
      # "HDIA",
      "DIAGNOS",
      # "INDATUMA",
      # "UTDATUMA",
      # "SOURCE",
      "T1D_age_diagnosis",
      "onset_age_group",
      "Birth_weight_cat",
      "day_of_year",
      # "Ai",
      # "adjusted_25OHD",
      "start_age",
      "exit_date",
      "exit_age",
      "event"
    )
  ))
```



```{r}
child_covariates <- c("Preterm_dichotomous","Birth_weight", "Season", "Sex_cat")

maternal_covariates <- c(
  "First_degree_T1D_heredity_cat",
  "Smoking_during_pregnancy",
  # "Vitamin_mineral_supplement_pregnancy",
  "Fish_any_pregnancy",
  "Parity",
  "Delivery_mode"
)
```

```{r remove_missing_rows}
df_cox_full <- df_cox_full %>%
  dplyr::select(all_of(
    c(
      "ABISnr",
      "diagnosis",
      "conc_25OHD",
      "DBP",
      "exit_age",
      "event",
      "Birth_weight_cat",
      child_covariates,
      maternal_covariates,
      "onset_age_group"
    )
  )) 
```

Cox regression	Hazard ratio: relative rate of developing T1D at any given age. 

```{r cox_model_base}
cox_model_base_conc_25OHD <- coxph(
  Surv(exit_age, event = event) ~ conc_25OHD,
  data = df_cox_full %>% dplyr::select(-DBP)
)

cox_model_base_DBP <- coxph(
  Surv(exit_age, event = event) ~ DBP,
  data = df_cox_full %>% dplyr::select(-conc_25OHD)
)
```


```{r cox_model_child}
cox_model_child_conc_25OHD <- coxph(
  Surv(exit_age, event = event) ~ conc_25OHD + Preterm_dichotomous + Season + 
    Sex_cat + Birth_weight_cat, data = df_cox_full %>% dplyr::select(-DBP)
)

cox_model_child_DBP <- coxph(
  Surv(exit_age, event = event) ~ DBP + Preterm_dichotomous + Season + 
    Sex_cat + Birth_weight_cat, data = df_cox_full %>% dplyr::select(-conc_25OHD)
)
```


```{r cox_model_child_interaction}
cox_model_interaction_child_conc_25OHD <- coxph(
  Surv(exit_age, event = event) ~ conc_25OHD * (Preterm_dichotomous + Season + 
    Sex_cat + Birth_weight_cat), data = df_cox_full %>% dplyr::select(-DBP)
)

cox_model_interaction_child_DBP <- coxph(
  Surv(exit_age, event = event) ~ DBP * (Preterm_dichotomous + Season + 
    Sex_cat + Birth_weight_cat), data = df_cox_full %>% dplyr::select(-conc_25OHD)
)
```



```{r cox_model_maternal}
cox_model_maternal_conc_25OHD <- coxph(
  Surv(exit_age, event = event) ~ conc_25OHD + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode,
  data = df_cox_full %>% dplyr::select(-DBP)
)

cox_model_maternal_DBP <- coxph(
  Surv(exit_age, event = event) ~ DBP + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode,
  data = df_cox_full %>% dplyr::select(-conc_25OHD)
)
```

```{r cox_model_maternal_interaction}
cox_model_interaction_maternal_conc_25OHD <- coxph(
  Surv(exit_age, event = event) ~ conc_25OHD *(First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode),
  data = df_cox_full %>% dplyr::select(-DBP)
)

cox_model_interaction_maternal_DBP <- coxph(
  Surv(exit_age, event = event) ~ DBP *(First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode),
  data = df_cox_full %>% dplyr::select(-conc_25OHD)
)
```

```{r cox_model_full}
cox_model_full_conc_25OHD <- coxph(
  Surv(exit_age, event = event) ~ conc_25OHD + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode +
    Preterm_dichotomous + Season + Sex_cat + Birth_weight_cat,
  data = df_cox_full %>% dplyr::select(-DBP)
)

cox_model_full_DBP <- coxph(
  Surv(exit_age, event = event) ~ DBP + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode +
    Preterm_dichotomous + Season + Sex_cat + Birth_weight_cat,
  data = df_cox_full %>% dplyr::select(-conc_25OHD)
)
```

```{r cox_model_full_interaction}
cox_model_interaction_full_conc_25OHD <- coxph(
  Surv(exit_age, event = event) ~ conc_25OHD *(First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode +
    Preterm_dichotomous + Season + Sex_cat + Birth_weight_cat),
  data = df_cox_full %>% dplyr::select(-DBP)
)

cox_model_interaction_full_DBP <- coxph(
  Surv(exit_age, event = event) ~ DBP *(First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode +
    Preterm_dichotomous + Season + Sex_cat + Birth_weight_cat),
  data = df_cox_full %>% dplyr::select(-conc_25OHD)
)
```

create pattern strings to clean-tidy result tables 
```{r}
child_levels <- paste0(
  unlist(lapply(c(child_covariates[-2], "Birth_weight_cat"), function(x)
  levels(factor(df_cox_full[[x]]))))[-2], collapse = "|")

child_pat_replace <- paste0(c(child_covariates[-2],"Birth_weight_cat"),collapse = "|")
```

```{r}
maternal_levels <- paste0(
  unique(
    unlist(
      lapply(
        maternal_covariates, 
        function(x) levels(factor(df_cox_full[[x]]))
        )
      )
    ), 
  collapse = "|")

maternal_pat_replace <- paste0(maternal_covariates, collapse = "|")
```

```{r}
all_levels <- (paste0(
  unique(
    unlist(
      lapply(
        c(child_covariates[-2],"Birth_weight_cat",maternal_covariates), 
        function(x) levels(factor(df_cox_full[[x]]))
        )
      )
    )[-2], 
  collapse = "|"))

all_pat_replace <- paste0(
  c(child_covariates[-2],"Birth_weight_cat",maternal_covariates), collapse = "|")
```

```{r tidy_cox_model_base}
tidy_cox_model_base_conc_25OHD <- tidy_base_model(cox_model_base_conc_25OHD, "Base")
tidy_cox_model_base_DBP <- tidy_base_model(cox_model_base_DBP, "Base")
```

```{r tidy_cox_model_child}
tidy_cox_model_child_conc_25OHD <- tidy_model_specific_covariates(
  model = cox_model_child_conc_25OHD,
  predictor = "conc_25OHD",
  model_name = "Base + Child covariates", 
  covariates_as_pattern = child_pat_replace, 
  covariate_levels_as_pattern = child_levels
  )

tidy_cox_model_child_DBP <- tidy_model_specific_covariates(
  model = cox_model_child_DBP,
  predictor = "DBP",
  model_name = "Base + Child covariates", 
  covariates_as_pattern = child_pat_replace, 
  covariate_levels_as_pattern = child_levels
  )
```

```{r tidy_cox_model_maternal}
tidy_cox_model_maternal_conc_25OHD <- tidy_model_specific_covariates(
  model = cox_model_maternal_conc_25OHD,
  predictor = "conc_25OHD",
  model_name = "Base + Maternal covariates", 
  covariates_as_pattern = maternal_pat_replace, 
  covariate_levels_as_pattern = maternal_levels
  )

tidy_cox_model_maternal_DBP <- tidy_model_specific_covariates(
  model = cox_model_maternal_DBP,
  predictor = "DBP",
  model_name = "Base + Maternal covariates", 
  covariates_as_pattern = maternal_pat_replace, 
  covariate_levels_as_pattern = maternal_levels
  )
```

```{r tidy_cox_model_full}
tidy_cox_model_full_conc_25OHD <- tidy_model_specific_covariates(
  model = cox_model_full_conc_25OHD,
  predictor = "conc_25OHD",
  model_name = "Base + All covariates", 
  covariates_as_pattern = all_pat_replace, 
  covariate_levels_as_pattern = all_levels
  ) 

tidy_cox_model_full_DBP <- tidy_model_specific_covariates(
  model = cox_model_full_DBP,
  predictor = "DBP",
  model_name = "Base + All covariates", 
  covariates_as_pattern = all_pat_replace, 
  covariate_levels_as_pattern = all_levels
  ) 
```

```{r tidy_cox_model_base_conc_25OHD}
# tidy_cox_model_base_conc_25OHD <- cox_model_base_conc_25OHD %>%
#   broom::tidy(exponentiate = TRUE, conf.int = TRUE)  %>%
#   dplyr::mutate(p.value = ifelse(
#     p.value < 0.001,
#     formatC(p.value, format = "e", digits = 2),
#     formatC(p.value, format = "f", digits = 3)
#   )) %>%
#   dplyr::mutate(Model = "Base", .before = 1) %>%
#   dplyr::mutate(`HR (95% CI)` = paste0(
#     round(estimate, 3),
#     " (",
#     round(conf.low, 3),
#     "-",
#     round(conf.high, 3),
#     ")"
#   )) %>%
#   dplyr::rename(Term = term, `P-value` = p.value) %>%
#   dplyr::mutate(Level = "+1") %>%
#   dplyr::select(Model, Term, Level, `HR (95% CI)`, `P-value`)
```


```{r tidy_cox_model_base_DBP}
# tidy_cox_model_base_DBP <- cox_model_base_DBP %>%
#   broom::tidy(exponentiate = TRUE, conf.int = TRUE)  %>%
#   dplyr::mutate(p.value = ifelse(
#     p.value < 0.001,
#     formatC(p.value, format = "e", digits = 2),
#     formatC(p.value, format = "f", digits = 3)
#   )) %>%
#   dplyr::mutate(Model = "Base", .before = 1) %>%
#   dplyr::mutate(`HR (95% CI)` = paste0(
#     round(estimate, 3),
#     " (",
#     round(conf.low, 3),
#     "-",
#     round(conf.high, 3),
#     ")"
#   )) %>%
#   dplyr::rename(Term = term, `P-value` = p.value) %>%
#   dplyr::mutate(Level = "+1") %>%
#   dplyr::select(Model, Term, Level, `HR (95% CI)`, `P-value`)
```




```{r tidy_cox_model_child_conc_25OHD}
# Base + child covariates
# tidy_cox_model_child_conc_25OHD <- cox_model_child_conc_25OHD %>%
#   broom::tidy(exponentiate = TRUE, conf.int = TRUE)  %>%
#   dplyr::mutate(p.value = ifelse(
#     p.value < 0.001,
#     formatC(p.value, format = "e", digits = 2),
#     formatC(p.value, format = "f", digits = 3)
#   )) %>%
#   dplyr::mutate(Model = "Base + Child covariates", .before = 1) %>%
#   dplyr::mutate(`HR (95% CI)` = paste0(
#     round(estimate, 3),
#     " (",
#     round(conf.low, 3),
#     "-",
#     round(conf.high, 3),
#     ")"
#   )) %>%
#   dplyr::rename(Term = term, `P-value` = p.value) %>%
#   dplyr::select(Model, Term, `HR (95% CI)`, `P-value`) %>%
#   dplyr::mutate(Level = gsub(child_pat_replace, "", gsub("conc_25OHD", "+1", Term)),
#                 .before = 3) %>%
#   dplyr::mutate(Term = gsub(child_levels, "", Term)) %>%
#   dplyr::mutate(
#     Term = ifelse(
#       row_number() > 1 & Term == lag(Term),
#       NA_character_,
#       gsub("_", " ", gsub(
#         "_cat", "", gsub(
#           "_dichotomous",
#           "",
#           gsub(
#             "Birth_weight_cat",
#             "Birth weight category",
#             gsub("Preterm_dichotomous", "", Term)
#           )
#         )
#       ))
#     ),
#     Model = ifelse(row_number() > 1 &
#                      Model == lag(Model), NA_character_, Model)
#   )
```


```{r tidy_cox_model_child_DBP}
# tidy_cox_model_child_DBP <- cox_model_child_DBP %>%
#   broom::tidy(exponentiate = TRUE, conf.int = TRUE)  %>%
#   dplyr::mutate(p.value = ifelse(
#     p.value < 0.001,
#     formatC(p.value, format = "e", digits = 2),
#     formatC(p.value, format = "f", digits = 3)
#   )) %>%
#   dplyr::mutate(Model = "Base + Child covariates", .before = 1) %>%
#   dplyr::mutate(`HR (95% CI)` = paste0(
#     round(estimate, 3),
#     " (",
#     round(conf.low, 3),
#     "-",
#     round(conf.high, 3),
#     ")"
#   )) %>%
#   dplyr::rename(Term = term, `P-value` = p.value) %>%
#   dplyr::select(Model, Term, `HR (95% CI)`, `P-value`) %>%
#   dplyr::mutate(Level = gsub(child_pat_replace, "", gsub("DBP", "+1", Term)),
#                 .before = 3) %>%
#   dplyr::mutate(Term = gsub(child_levels, "", Term)) %>%
#   dplyr::mutate(
#     Term = ifelse(
#       row_number() > 1 & Term == lag(Term),
#       NA_character_,
#       gsub("_", " ", gsub(
#         "_cat", "", gsub(
#           "_dichotomous",
#           "",
#           gsub(
#             "Birth_weight_cat",
#             "Birth weight category",
#             gsub("Preterm_dichotomous", "", Term)
#           )
#         )
#       ))
#     ),
#     Model = ifelse(row_number() > 1 &
#                      Model == lag(Model), NA_character_, Model)
#   )
```


```{r tidy_cox_model_maternal_conc_25OHD }
# Base + Maternal covariates
# tidy_cox_model_maternal_conc_25OHD <- cox_model_maternal_conc_25OHD %>%
#   broom::tidy(exponentiate = TRUE, conf.int = TRUE)  %>%
#   dplyr::mutate(p.value = ifelse(
#     p.value < 0.001,
#     formatC(p.value, format = "e", digits = 2),
#     formatC(p.value, format = "f", digits = 3)
#   )) %>%
#   dplyr::mutate(Model = "Base + Maternal covariates", .before = 1) %>%
#   dplyr::mutate(`HR (95% CI)` = paste0(
#     round(estimate, 3),
#     " (",
#     round(conf.low, 3),
#     "-",
#     round(conf.high, 3),
#     ")"
#   )) %>%
#   dplyr::rename(Term = term, `P-value` = p.value) %>%
#   dplyr::select(Model, Term, `HR (95% CI)`, `P-value`) %>%
#   dplyr::mutate(Level = gsub(maternal_pat_replace, "", gsub("conc_25OHD", "+1", Term)),
#                 .before = 3) %>%
#   dplyr::mutate(Term = gsub(maternal_levels, "", Term)) %>%
#   dplyr::mutate(
#     Term = ifelse(
#       row_number() > 1 & Term == lag(Term),
#       NA_character_,
#       gsub("_", " ", gsub(
#         "_cat", "", gsub(
#           "_dichotomous",
#           "",
#           gsub(
#             "Birth_weight_cat",
#             "Birth weight category",
#             gsub("Preterm_dichotomous", "", Term)
#           )
#         )
#       ))
#     ),
#     Model = ifelse(row_number() > 1 &
#                      Model == lag(Model), NA_character_, Model)
#   )
```


```{r tidy_cox_model_maternal_DBP}
# tidy_cox_model_maternal_DBP <- cox_model_maternal_DBP %>%
#   broom::tidy(exponentiate = TRUE, conf.int = TRUE)  %>%
#   dplyr::mutate(p.value = ifelse(
#     p.value < 0.001,
#     formatC(p.value, format = "e", digits = 2),
#     formatC(p.value, format = "f", digits = 3)
#   )) %>%
#   dplyr::mutate(Model = "Base + Maternal covariates", .before = 1) %>%
#   dplyr::mutate(`HR (95% CI)` = paste0(
#     round(estimate, 3),
#     " (",
#     round(conf.low, 3),
#     "-",
#     round(conf.high, 3),
#     ")"
#   )) %>%
#   dplyr::rename(Term = term, `P-value` = p.value) %>%
#   dplyr::select(Model, Term, `HR (95% CI)`, `P-value`) %>%
#   dplyr::mutate(Level = gsub(maternal_pat_replace, "", gsub("DBP", "+1", Term)),
#                 .before = 3) %>%
#   dplyr::mutate(Term = gsub(maternal_levels, "", Term)) %>%
#   dplyr::mutate(
#     Term = ifelse(
#       row_number() > 1 & Term == lag(Term),
#       NA_character_,
#       gsub("_", " ", gsub(
#         "_cat", "", gsub(
#           "_dichotomous",
#           "",
#           gsub(
#             "Birth_weight_cat",
#             "Birth weight category",
#             gsub("Preterm_dichotomous", "", Term)
#           )
#         )
#       ))
#     ),
#     Model = ifelse(row_number() > 1 &
#                      Model == lag(Model), NA_character_, Model)
#   )
```


```{r tidy_cox_model_full_conc_25OHD}
# tidy_cox_model_full_conc_25OHD <- cox_model_full_conc_25OHD %>%
#   broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%
#   as.data.frame() %>%
#   dplyr::mutate(across(where(is.numeric) &
#                          !matches("p.value"), ~ round(.x, 3))) %>%
#   dplyr::mutate(p.value = ifelse(
#     p.value < 0.001,
#     formatC(p.value, format = "e", digits = 2),
#     formatC(p.value, format = "f", digits = 3)
#   )) %>%
#   dplyr::rename(Term = term,
#                 HR = estimate,
#                 `P-value` = p.value) %>%
#   dplyr::mutate(`HR (95% CI)` = paste0(
#     round(HR, 3),
#     " (",
#     round(conf.low, 3),
#     "-",
#     round(conf.high, 3),
#     ")"
#   )) %>%
#   dplyr::select(Term, `HR (95% CI)`, `P-value`) %>%
#   dplyr::mutate(Level = gsub(all_pat_replace, "", gsub("conc_25OHD", "+1", Term)),
#                 .before = 3) %>%
#   dplyr::mutate(Term = gsub(all_levels, "", Term)) %>%
#   dplyr::mutate(Term = ifelse(
#     row_number() > 1 & Term == lag(Term),
#     NA_character_,
#     gsub("_", " ", gsub(
#       "_cat", "", gsub(
#         "_dichotomous",
#         "",
#         gsub(
#           "_during_pregnancy|_any_pregnancy",
#           " (pregnancy)",
#           gsub("Preterm_dichotomous", "", Term)
#         )
#       )
#     ))
#   )) %>%
#   dplyr::mutate(Model = "Base + All covariates", .before = 1) %>%
#   dplyr::mutate(Model = ifelse(row_number() > 1 &
#                                  Model == lag(Model), NA_character_, Model))
```


```{r tidy_cox_model_full_DBP}
# tidy_cox_model_full_DBP <- cox_model_full_DBP %>%
#   broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%
#   as.data.frame() %>%
#   dplyr::mutate(across(where(is.numeric) &
#                          !matches("p.value"), ~ round(.x, 3))) %>%
#   dplyr::mutate(p.value = ifelse(
#     p.value < 0.001,
#     formatC(p.value, format = "e", digits = 2),
#     formatC(p.value, format = "f", digits = 3)
#   )) %>%
#   dplyr::rename(Term = term,
#                 HR = estimate,
#                 `P-value` = p.value) %>%
#   dplyr::mutate(`HR (95% CI)` = paste0(
#     round(HR, 3),
#     " (",
#     round(conf.low, 3),
#     "-",
#     round(conf.high, 3),
#     ")"
#   )) %>%
#   dplyr::select(Term, `HR (95% CI)`, `P-value`) %>%
#   dplyr::mutate(Level = gsub(all_pat_replace, "", gsub("DBP", "+1", Term)),
#                 .before = 3) %>%
#   dplyr::mutate(Term = gsub(all_levels, "", Term)) %>%
#   dplyr::mutate(Term = ifelse(
#     row_number() > 1 & Term == lag(Term),
#     NA_character_,
#     gsub("_", " ", gsub(
#       "_cat", "", gsub(
#         "_dichotomous",
#         "",
#         gsub(
#           "_during_pregnancy|_any_pregnancy",
#           " (pregnancy)",
#           gsub("Preterm_dichotomous", "", Term)
#         )
#       )
#     ))
#   )) %>%
#   dplyr::mutate(Model = "Base + All covariates", .before = 1) %>%
#   dplyr::mutate(Model = ifelse(row_number() > 1 &
#                                  Model == lag(Model), NA_character_, Model))
```


```{r merged_all_model_results_conc_25OHD }
merged_all_model_results_conc_25OHD <-  rbind(
  tidy_cox_model_base_conc_25OHD,
  tidy_cox_model_child_conc_25OHD,
  tidy_cox_model_maternal_conc_25OHD,
  tidy_cox_model_full_conc_25OHD
) %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::set_table_properties(width = 1, layout = "autofit") %>%
  flextable::color(
    i = ~ as.numeric(`P-value`) < 0.05,
    j = c("Term", "Level", "P-value"),
    color = "red"
  ) 
```

```{r merged_all_model_results_DBP}
merged_all_model_results_DBP <-  rbind(
  tidy_cox_model_base_DBP,
  tidy_cox_model_child_DBP,
  tidy_cox_model_maternal_DBP,
  tidy_cox_model_full_DBP
) %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::set_table_properties(width = 1, layout = "autofit") %>%
  flextable::color(
    i = ~ as.numeric(`P-value`) < 0.05,
    j = c("Term", "Level", "P-value"),
    color = "red"
  ) 
```

```{r save_merged_all_model_results}
flextable::save_as_docx(
  `Cox proportional hazards models: DBP` = merged_all_model_results_DBP,
  path = paste0(output_dir,"/flex_merged_all_model_results_DBP.docx"),
  pr_section = sect_properties
)
```


```{r strata_wise_base}
# run_stratified_cox_base_models <- function(data, predictor, strata_vars) {
#   results <- list()
#   
#   for (strata_var in strata_vars) {
#     # Construct formula
#     fmla <- as.formula(paste0("Surv(exit_age, event = event) ~ ",predictor," + strata(", strata_var, ")"))
#     
#     # Fit Cox model
#     cox_model <- coxph(fmla, data = data)
#     
#     # Tidy and format
#     tidy_result <- tidy(cox_model, exponentiate = TRUE, conf.int = TRUE) %>%
#       dplyr::rename(
#         Term = term,
#         OR = estimate,
#         `P-value` = p.value
#       ) %>%
#       dplyr::mutate(`HR (95% CI)` = paste0(
#         round(OR, 3),
#         " (", round(conf.low, 3), "-", round(conf.high, 3), ")"
#       )) %>%
#       dplyr::mutate(Strata = strata_var) %>%
#       dplyr::select(Term, Strata, `HR (95% CI)`, `P-value`) 
#     
#     # Store in list
#     results[[strata_var]] <- tidy_result
#   }
#   
#   # Combine all results
#   final_result <- bind_rows(results)
#   return(final_result)
# }
```

```{r strata_wise_cox_base}
strata_wise_cox_base_conc_25OHD <- run_stratified_cox_models(
  data = df_cox_full %>% dplyr::select(-DBP), 
  predictor = "conc_25OHD",
  strata_vars = c(child_covariates[-2], "Birth_weight_cat", maternal_covariates)
  )

strata_wise_cox_base_DBP <- run_stratified_cox_models(
  data = df_cox_full %>% dplyr::select(-conc_25OHD), 
  predictor = "DBP",
  strata_vars = c(child_covariates[-2], "Birth_weight_cat", maternal_covariates)
  )
```

```{r strata_wise_cox_child_covariates}
strata_wise_cox_child_covariates_conc_25OHD <- run_stratified_cox_models(
  data = df_cox_full %>% dplyr::select(-DBP), 
  predictor ="conc_25OHD",
  strata_vars = maternal_covariates,
  covariate_list = c(child_covariates[-2], "Birth_weight_cat"),
  covariates_as_pattern = child_pat_replace,
  covariate_levels_as_pattern = child_levels
  )

strata_wise_cox_child_covariates_DBP <- run_stratified_cox_models(
  data = df_cox_full %>% dplyr::select(-conc_25OHD), 
  predictor = "DBP",
  strata_vars = maternal_covariates,
  covariate_list = c(child_covariates[-2], "Birth_weight_cat"),
  covariates_as_pattern = child_pat_replace,
  covariate_levels_as_pattern = child_levels
  )
```

```{r strata_wise_cox_maternal_covariates}
strata_wise_cox_maternal_covariates_conc_25OHD <- run_stratified_cox_models(
  data = df_cox_full %>% dplyr::select(-DBP), 
  predictor ="conc_25OHD",
  strata_vars = c(child_covariates[-2], "Birth_weight_cat"),
  covariate_list =  maternal_covariates,
  covariates_as_pattern = maternal_pat_replace,
  covariate_levels_as_pattern = maternal_levels
  )

strata_wise_cox_maternal_covariates_DBP <- run_stratified_cox_models(
  data = df_cox_full %>% dplyr::select(-conc_25OHD), 
  predictor = "DBP",
  strata_vars = c(child_covariates[-2], "Birth_weight_cat"),
  covariate_list =  maternal_covariates,
  covariates_as_pattern = maternal_pat_replace,
  covariate_levels_as_pattern = maternal_levels
  )
```


```{r strata_flex_tables}
flex_strata_wise_cox_base_conc_25OHD <- table2flex(strata_wise_cox_base_conc_25OHD)
flex_strata_wise_cox_base_DBP <- table2flex(strata_wise_cox_base_DBP)

flex_strata_wise_cox_child_covariates_conc_25OHD <- table2flex(strata_wise_cox_child_covariates_conc_25OHD)
flex_strata_wise_cox_child_covariates_DBP <- table2flex(strata_wise_cox_child_covariates_DBP)

flex_strata_wise_cox_maternal_covariates_conc_25OHD <- table2flex(strata_wise_cox_maternal_covariates_conc_25OHD)
flex_strata_wise_cox_maternal_covariates_DBP <- table2flex(strata_wise_cox_maternal_covariates_DBP)
```


```{r strata_wise_child}
# run_stratified_cox_child_models <- function(data, predictor, strata_vars) {
#   results <- list()
#   
#   for (strata_var in strata_vars) {
#     # Construct formula
#     fmla <- as.formula(paste0("Surv(exit_age, event = event) ~ ",predictor," + Preterm_dichotomous + Season + 
#     Sex_cat + Birth_weight_cat + strata(", strata_var, ")"))
#     
#     # Fit Cox model
#     cox_model <- coxph(fmla, data = data)
#     
#     # Tidy and format
#     tidy_result <- tidy(cox_model, exponentiate = TRUE, conf.int = TRUE) %>%
#       dplyr::mutate(
#         p.value = ifelse(
#           p.value < 0.001,
#           formatC(p.value, format = "e", digits = 2),
#           formatC(p.value, format = "f", digits = 3)
#           )) %>%
#       dplyr::rename(
#         Term = term,
#         HR = estimate,
#         `P-value` = p.value
#       ) %>%
#       dplyr::mutate(
#         `HR (95% CI)` = paste0(
#         round(HR, 3),
#         " (", round(conf.low, 3), "-", round(conf.high, 3), ")"
#       )) %>%
#       dplyr::mutate(Strata = strata_var) %>%
#       dplyr::select(Strata, Term, `HR (95% CI)`, `P-value`) 
#     
#     # Store in list
#     results[[strata_var]] <- tidy_result
#   }
#   
#   # Combine all results
#   final_result <- bind_rows(results)
#   return(final_result)
# }
```

```{r strata_wise_cox_child}
# strata_wise_cox_child_conc_25OHD <- run_stratified_cox_child_models(
#   df_cox_full %>% dplyr::select(-DBP), 
#   "conc_25OHD", 
#   maternal_covariates
#   )
# 
# strata_wise_cox_child_DBP <- run_stratified_cox_child_models(
#   df_cox_full %>% dplyr::select(-conc_25OHD), 
#   "DBP", 
#   maternal_covariates
#   )
```



```{r flex_strata_wise_cox_child_conc_25OHD}
# strata_wise_cox_child_conc_25OHD <- strata_wise_cox_child_conc_25OHD %>%
#   dplyr::mutate(Level = gsub(child_pat_replace, "", gsub("conc_25OHD", "+1", Term)),
#                 .before = 3) %>%
#   dplyr::mutate(Term = gsub(child_levels, "", Term)) %>%
#   dplyr::mutate(Term = ifelse(
#     row_number() > 1 & Term == lag(Term),
#     NA_character_,
#     gsub("_", " ", gsub(
#       "_cat", "", gsub(
#         "_dichotomous",
#         "",
#         gsub(
#           "Birth_weight_cat",
#           "Birth weight category",
#           gsub("Preterm_dichotomous", "", Term)
#         )
#       )
#     ))
#   )) %>%
#   dplyr::mutate(Strata = ifelse(
#     row_number() > 1 & Strata == lag(Strata),
#     NA_character_,
#     gsub("_", " ", gsub(
#       "_cat", "", gsub(
#         "_dichotomous",
#         "",
#         gsub("_during_pregnancy|_any_pregnancy", " (pregnancy)", Strata)
#       )
#     ))
#   ))


# flex_strata_wise_cox_child_conc_25OHD <- strata_wise_cox_child_conc_25OHD %>% 
#  flextable::flextable() %>% 
#  flextable::autofit() %>% 
#   flextable::set_table_properties(width = 1, layout = "autofit") %>% 
#  flextable::color(
#      i = ~ `P-value` < 0.05,
#      j = c("Term", "Level", "P-value"),
#      color = "red") 
```


```{r flex_strata_wise_cox_child_DBP}
# strata_wise_cox_child_DBP <- strata_wise_cox_child_DBP %>%
#   dplyr::mutate(Level = gsub(child_pat_replace, "", gsub("DBP", "+1", Term)),
#                 .before = 3) %>%
#   dplyr::mutate(Term = gsub(child_levels, "", Term)) %>%
#   dplyr::mutate(Term = ifelse(
#     row_number() > 1 & Term == lag(Term),
#     NA_character_,
#     gsub("_", " ", gsub(
#       "_cat", "", gsub(
#         "_dichotomous",
#         "",
#         gsub(
#           "Birth_weight_cat",
#           "Birth weight category",
#           gsub("Preterm_dichotomous", "", Term)
#         )
#       )
#     ))
#   )) %>%
#   dplyr::mutate(Strata = ifelse(
#     row_number() > 1 & Strata == lag(Strata),
#     NA_character_,
#     gsub("_", " ", gsub(
#       "_cat", "", gsub(
#         "_dichotomous",
#         "",
#         gsub("_during_pregnancy|_any_pregnancy", " (pregnancy)", Strata)
#       )
#     ))
#   ))
# 
# flex_strata_wise_cox_child_DBP <- strata_wise_cox_child_DBP %>% 
#  flextable::flextable() %>% 
#  flextable::autofit() %>% 
#   flextable::set_table_properties(width = 1, layout = "autofit") %>% 
#  flextable::color(
#      i = ~ `P-value` < 0.05,
#      j = c("Term", "Level", "P-value"),
#      color = "red") 
```


```{r save_flex_strata_wise_cox_models}
flextable::save_as_docx(
  `Strata wise cox (base) model: conc. 25(OH)D` = flex_strata_wise_cox_base_conc_25OHD,
  path = paste0(output_dir,"/flex_strata_wise_cox_base_conc_25OHD.docx"),
  pr_section = sect_properties
)

flextable::save_as_docx(
  `Strata wise cox  (base) model: DBP` = flex_strata_wise_cox_base_DBP,
  path = paste0(output_dir,"/flex_strata_wise_cox_base_DBP.docx"),
  pr_section = sect_properties
)

flextable::save_as_docx(
  `Strata wise cox model with child covariates: conc. 25(OH)D` = flex_strata_wise_cox_child_covariates_conc_25OHD,
  path = paste0(output_dir,"/flex_strata_wise_cox_child_covariates_conc_25OHD.docx"),
  pr_section = sect_properties
)

flextable::save_as_docx(
  `Strata wise cox model with child covariates: DBP` = flex_strata_wise_cox_child_covariates_DBP,
  path = paste0(output_dir,"/flex_strata_wise_cox_child_covariates_DBP.docx"),
  pr_section = sect_properties
)

flextable::save_as_docx(
  `Strata wise cox model with maternal covariates: conc. 25(OH)D` = flex_strata_wise_cox_maternal_covariates_conc_25OHD,
  path = paste0(output_dir,"/flex_strata_wise_cox_maternal_covariates_conc_25OHD.docx"),
  pr_section = sect_properties
)

flextable::save_as_docx(
  `Strata wise cox model with maternal covariates: DBP` = flex_strata_wise_cox_maternal_covariates_DBP,
  path = paste0(output_dir,"/flex_strata_wise_cox_maternal_covariates_DBP.docx"),
  pr_section = sect_properties
)
```


```{r strata_wise_maternal}
# run_stratified_cox_maternal_models <- function(data, predictor, strata_vars) {
#   results <- list()
#   
#   for (strata_var in strata_vars) {
#     # Construct formula
#     fmla <- as.formula(paste0("Surv(exit_age, event = event) ~ ", predictor," + First_degree_T1D_heredity_cat +
#     Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode + strata(", strata_var, ")"))
#     
#     # Fit Cox model
#     cox_model <- coxph(fmla, data = data)
#     
#     # Tidy and format
#     tidy_result <- tidy(cox_model, exponentiate = TRUE, conf.int = TRUE)  %>%
#       dplyr::mutate(
#         p.value = ifelse(
#           p.value < 0.001,
#           formatC(p.value, format = "e", digits = 2),
#           formatC(p.value, format = "f", digits = 3)
#           )) %>%
#       dplyr::rename(
#         Term = term,
#         HR = estimate,
#         `P-value` = p.value
#       ) %>%
#       dplyr::mutate(`HR (95% CI)` = paste0(
#         round(HR, 3),
#         " (", round(conf.low, 3), "-", round(conf.high, 3), ")"
#       )) %>%
#       dplyr::mutate(Strata = strata_var, .before = 1) %>%
#       dplyr::select(Strata, Term, `HR (95% CI)`, `P-value`) 
#     
#     # Store in list
#     results[[strata_var]] <- tidy_result
#   }
#   
#   # Combine all results
#   final_result <- bind_rows(results)
#   return(final_result)
# }
```

```{r strata_wise_cox_maternal}
# strata_wise_cox_maternal_conc_25OHD <- run_stratified_cox_maternal_models(
#   df_cox_full %>% dplyr::select(-DBP),
#   "conc_25OHD", 
#   c(child_covariates[-2], "Birth_weight_cat")
#   )
# 
# strata_wise_cox_maternal_DBP <- run_stratified_cox_maternal_models(
#   df_cox_full %>% dplyr::select(-conc_25OHD),
#   "DBP", 
#   c(child_covariates[-2], "Birth_weight_cat")
#   )
```




```{r}
# strata_wise_cox_maternal <- strata_wise_cox_maternal %>%
#     dplyr::mutate(
#     Level = gsub(maternal_pat_replace, "", 
#                  gsub("conc_25OHD", "+1", Term)),.before = 3) %>%
#   dplyr::mutate(Term = gsub(maternal_levels, "", Term)) %>%
#   dplyr::mutate(Term = ifelse(
#     row_number() > 1 & Term == lag(Term),
#     NA_character_,
#     gsub("_", " ", gsub(
#       "_cat", "", gsub(
#         "_dichotomous",
#         "",
#         gsub("_during_pregnancy|_any_pregnancy", " (pregnancy)", Term)
#       )
#     ))
#   )) %>%
#   dplyr::mutate(Strata = ifelse(
#     row_number() > 1 & Strata == lag(Strata),
#     NA_character_,
#     gsub("_", " ", gsub(
#       "_cat", "", gsub(
#         "_dichotomous",
#         "",
#         gsub("_during_pregnancy|_any_pregnancy", " (pregnancy)", Strata)
#       )
#     ))
#   ))
# 
# flex_strata_wise_cox_maternal <- strata_wise_cox_maternal %>% 
#  flextable::flextable() %>% 
#  flextable::autofit() %>% 
#   flextable::set_table_properties(width = 1, layout = "autofit") %>% 
#  flextable::color(
#      i = ~ as.numeric(`P-value`) < 0.05,
#      j = c("Term", "Level", "P-value"),
#      color = "red")
```

```{r save_flex_strata_wise_cox_maternal}
# flextable::save_as_docx(
#   `Strata wise cox model with maternal covariates` = flex_strata_wise_cox_maternal,
#   path = paste0(output_dir,"/flex_strata_wise_cox_maternal.docx"),
#   pr_section = sect_properties
# )
```


```{r}
df_cox_full <- df_cox_full %>%
  dplyr::mutate(
    quantile_25OHD = factor(ntile(conc_25OHD, 4), labels = c("Q1", "Q2", "Q3", "Q4"))
  )
```

```{r cox_model_quantile_base}
cox_model_quantile_base <- coxph(
  Surv(exit_age, event = event) ~ quantile_25OHD,
  data = df_cox_full
)
```


```{r cox_model_quantile_child}
cox_model_quantile_child <- coxph(
  Surv(exit_age, event = event) ~ quantile_25OHD + Preterm_dichotomous + Season + 
    Sex_cat + Birth_weight_cat, data = df_cox_full
)
```



```{r cox_model_quantile_maternal}
cox_model_quantile_maternal <- coxph(
  Surv(exit_age, event = event) ~ quantile_25OHD + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode,
  data = df_cox_full
)
```

```{r cox_model_quantile_full}
cox_model_quantile_full <- coxph(
  Surv(exit_age, event = event) ~ quantile_25OHD + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode +
    Preterm_dichotomous + Season + Sex_cat + Birth_weight_cat,
  data = df_cox_full
)
```


```{r}
child_quantile_levels <- paste0(
  unlist(lapply(c(child_covariates[-2], "Birth_weight_cat","quantile_25OHD"), function(x)
  levels(factor(df_cox_full[[x]]))))[-2], collapse = "|")

child_quantile_pat_replace <- paste0(c(child_covariates[-2],"Birth_weight_cat"),collapse = "|")
```

```{r}
maternal_quantile_levels <- paste0(
  unique(
    unlist(
      lapply(
        c(maternal_covariates,"quantile_25OHD"), 
        function(x) levels(factor(df_cox_full[[x]]))
        )
      )
    ), 
  collapse = "|")

maternal_quantile_pat_replace <- paste0(maternal_covariates, collapse = "|")
```

```{r}
all_quantile_levels <- (paste0(
  unique(
    unlist(
      lapply(
        c(child_covariates[-2],"Birth_weight_cat","quantile_25OHD",maternal_covariates), 
        function(x) levels(factor(df_cox_full[[x]]))
        )
      )
    )[-2], 
  collapse = "|"))

all_quantile_pat_replace <- paste0(
  c(child_covariates[-2],"Birth_weight_cat","quantile_25OHD",maternal_covariates), collapse = "|")
```


```{r tidy_cox_model_quantile_base}
tidy_cox_model_quantile_base <- cox_model_quantile_base %>% 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)  %>%
      dplyr::mutate(
        p.value = ifelse(
          p.value < 0.001,
          formatC(p.value, format = "e", digits = 2),
          formatC(p.value, format = "f", digits = 3)
          )) %>%
  dplyr::mutate(Model = "Base", .before=1) %>%
  dplyr::mutate(
    `HR (95% CI)` = paste0(
        round(estimate, 3),
        " (", round(conf.low, 3), "-", round(conf.high, 3), ")"
      )
  ) %>%
  dplyr::rename(
    Term = term, `P-value` = p.value
  ) %>%
  dplyr::mutate(Level= gsub("quantile_25OHD","",Term)) %>%
  dplyr::select(Model, Term, Level,`HR (95% CI)`, `P-value`) %>%
  dplyr::mutate(Term = gsub(child_quantile_levels, "", Term)) %>%
  dplyr::mutate(
    Model = ifelse(
      row_number() > 1 & Model == lag(Model),
      NA_character_,
      Model),
    Term = ifelse(
    row_number() > 1 & Term == lag(Term),
    NA_character_,
    Term)
  )
```


```{r tidy_cox_model_quantile_child}
# Base + child covariates
tidy_cox_model_quantile_child <- cox_model_quantile_child %>% 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)  %>%
      dplyr::mutate(
        p.value = ifelse(
          p.value < 0.001,
          formatC(p.value, format = "e", digits = 2),
          formatC(p.value, format = "f", digits = 3)
          )) %>%
  dplyr::mutate(Model = "Base + Child covariates", .before=1) %>%
  dplyr::mutate(
    `HR (95% CI)` = paste0(
        round(estimate, 3),
        " (", round(conf.low, 3), "-", round(conf.high, 3), ")"
      )
  ) %>%
  dplyr::rename(
    Term = term, `P-value` = p.value
  ) %>%
  dplyr::select(Model, Term,`HR (95% CI)`, `P-value`) %>%
    dplyr::mutate(
        Level = gsub(child_pat_replace, "", 
                     gsub("quantile_25OHD", "", Term)),.before = 3) %>%
    dplyr::mutate(Term = gsub(child_quantile_levels, "", Term)) %>%
    dplyr::mutate(
      Term = ifelse(
        row_number() > 1 & Term == lag(Term),
        NA_character_,
        gsub("_", " ", gsub(
            "_cat", "", gsub(
                "_dichotomous",
                "",
                gsub("Birth_weight_cat", "Birth weight category",
                     gsub("Preterm_dichotomous","",Term)
                )
            )
        ))
    ),
    Model = ifelse(
      row_number() > 1 & Model == lag(Model), 
      NA_character_, 
      Model)
    )
```


```{r tidy_cox_model_quantile_maternal}
# Base + Maternal covariates
tidy_cox_model_quantile_maternal <- cox_model_quantile_maternal %>% 
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)  %>%
      dplyr::mutate(
        p.value = ifelse(
          p.value < 0.001,
          formatC(p.value, format = "e", digits = 2),
          formatC(p.value, format = "f", digits = 3)
          )) %>%
  dplyr::mutate(Model = "Base + Maternal covariates", .before=1) %>%
  dplyr::mutate(
    `HR (95% CI)` = paste0(
        round(estimate, 3),
        " (", round(conf.low, 3), "-", round(conf.high, 3), ")"
      )
  ) %>%
  dplyr::rename(
    Term = term, `P-value` = p.value
  ) %>%
  dplyr::select(Model, Term,`HR (95% CI)`, `P-value`) %>%
    dplyr::mutate(
        Level = gsub(maternal_pat_replace, "", 
                     gsub("quantile_25OHD", "", Term)),.before = 3) %>%
    dplyr::mutate(Term = gsub(maternal_quantile_levels, "", Term)) %>%
    dplyr::mutate(
      Term = ifelse(
        row_number() > 1 & Term == lag(Term),
        NA_character_,
        gsub("_", " ", gsub(
            "_cat", "", gsub(
                "_dichotomous",
                "",
                gsub("Birth_weight_cat", "Birth weight category",
                     gsub("Preterm_dichotomous","",Term)
                )
            )
        ))
    ),
    Model = ifelse(row_number() > 1 & Model == lag(Model), NA_character_, Model)
    )
```


```{r tidy_cox_model_quantile_full}
tidy_cox_model_quantile_full <- cox_model_quantile_full %>% 
    broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>% 
    as.data.frame() %>%   
    dplyr::mutate(
      across(where(is.numeric) & !matches("p.value"), ~ round(.x, 3))) %>%
      dplyr::mutate(
        p.value = ifelse(
          p.value < 0.001,
          formatC(p.value, format = "e", digits = 2),
          formatC(p.value, format = "f", digits = 3)
          )) %>%
      dplyr::rename(
        Term = term,
        HR = estimate,
        `P-value` = p.value
      ) %>%
      dplyr::mutate(`HR (95% CI)` = paste0(
        round(HR, 3),
        " (", round(conf.low, 3), "-", round(conf.high, 3), ")"
      )) %>%
      dplyr::select(Term, `HR (95% CI)`, `P-value`) %>%
      dplyr::mutate(
      Level = gsub(all_quantile_pat_replace, "", 
                   gsub("quantile_25OHD", "", Term)),.before = 2) %>%
    dplyr::mutate(Term = gsub(all_quantile_levels, "", Term)) %>%
    dplyr::mutate(
      Term = ifelse(
      row_number() > 1 & Term == lag(Term),
      NA_character_,
      gsub("_", " ", gsub(
        "_cat", "", gsub(
          "_dichotomous",
          "",
          gsub("_during_pregnancy|_any_pregnancy", " (pregnancy)", 
                            gsub("Preterm_dichotomous","",Term)))))
      )
    ) %>% 
  dplyr::mutate(
    Model = "Base + All covariates", .before =1) %>%
  dplyr::mutate(
    Model = ifelse(
      row_number() > 1 & Model == lag(Model),
      NA_character_,
      Model)
  )
```


```{r merged_all_model_quantile_results}
merged_all_model_quantile_results <-  rbind(
  tidy_cox_model_quantile_base, 
  tidy_cox_model_quantile_child,
  tidy_cox_model_quantile_maternal,
  tidy_cox_model_quantile_full
  ) %>% 
  flextable::flextable() %>% 
  flextable::autofit() %>%
    flextable::set_table_properties(width = 1, layout = "autofit") %>% 
 flextable::color(
     i = ~ as.numeric(`P-value`) < 0.05,
     j = c("Term", "Level", "P-value"),
     color = "red") 
```

```{r save_merged_all_model_quantile_results}
flextable::save_as_docx(
  `Cox models: quantile_25(OH)D ` = merged_all_model_quantile_results,
  path = paste0(output_dir,"/flex_merged_all_model_quantile_results.docx"),
  pr_section = sect_properties
)
```

```{r onset_group}
df_cox_full$onset_age_group <- factor(df_cox_full$onset_age_group)

multi_model <- multinom(onset_age_group ~ conc_25OHD + First_degree_T1D_heredity_cat +
    Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode +
    Preterm_dichotomous + Season + Sex_cat + Birth_weight_cat, 
    data = df_cox_full %>% dplyr::filter(diagnosis==1))
```


```{r onset_group_tidy}
tidy_multi_model <- multi_model %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE)  %>%
  dplyr::filter(term != "(Intercept)") %>%
  dplyr::mutate(
    Level = gsub(all_pat_replace, "", 
                 gsub("conc_25OHD", "+1", term)),.before = 3) %>%
  dplyr::mutate(term = gsub(all_levels, "", term)) %>%
  dplyr::mutate(
      y.level = ifelse(
        row_number() > 1 & y.level == lag(y.level),
        NA_character_, 
        y.level)
      ) %>%
  dplyr::mutate(
    p.value = ifelse(
      p.value < 0.001,
      formatC(p.value, format = "e", digits = 2),
      formatC(p.value, format = "f", digits = 3)
    )) %>%
  dplyr::rename(
        Term = term,
        HR = estimate,
        `P-value` = p.value,
        `Onset Group` = y.level
      ) %>%
    dplyr::mutate(
      Term = ifelse(
        row_number() > 1 & Term == lag(Term),
        NA_character_,
        gsub("_", " ", gsub(
          "_cat", "", gsub(
            "_dichotomous",
            "",
            gsub("_during_pregnancy|_any_pregnancy", " (pregnancy)", 
                            gsub("Preterm_dichotomous","",Term)
               )
          )
      ))
    )) %>%
  dplyr::mutate(
    `HR (95% CI)` = paste0(round(HR, 3)," (", round(conf.low, 3), "-", round(conf.high, 3), ")")
    ) %>%
  dplyr::select(`Onset Group`, Term, Level, `HR (95% CI)`, `P-value`)

flex_tidy_multi_model <- tidy_multi_model %>% 
 flextable::flextable() %>% 
 flextable::autofit() %>%
    flextable::set_table_properties(width = 1, layout = "autofit") %>% 
 flextable::color(
     i = ~ as.numeric(`P-value`) < 0.05,
     j = c("Term", "Level", "P-value"),
     color = "red")
```


```{r save_flex_cox_model_full}
flextable::save_as_docx(
  `Multinomial model: Onset-age ~ conc. 25(OH)D` = flex_tidy_multi_model,
  path = paste0(output_dir,"/flex_multinomial_model_onset_age_conc_25OHD.docx"),
  pr_section = sect_properties
)
```





```{r}
# Prepare the data
df_cox_case <- df_cox_full %>%
  dplyr::filter(event == 1) %>%
  dplyr::mutate(
    quantile_25OHD = factor(ntile(conc_25OHD, 4), labels = c("Q1", "Q2", "Q3", "Q4"))
    #,
    # onset_age_group = factor(onset_age_group, levels = c("Early onset", "School-age", "Puberty-onset", "Young-adult"))
  )

# Summarise the counts
plot_data <- df_cox_case %>%
  dplyr::count(quantile_25OHD, onset_age_group) %>%
  dplyr::group_by(quantile_25OHD) %>%
  dplyr::mutate(prop = n / sum(n))

# Plot
plot_vitamin_d_quartile_vs_onset_age <-
  ggplot(plot_data, aes(x = onset_age_group, y = prop, fill = quantile_25OHD)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.8)) +
  labs(
    x = "Onset Age Group",
    y = "Proportion within Quartile",
    fill = "25OHD Quartile",
    title = "Distribution of Onset Age Groups by Vitamin D Quartile"
  ) +
  scale_y_continuous(labels = scales::percent_format()) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 30, hjust = 1))
```

```{r}
suppressWarnings(ggplot2::ggsave(
  filename = paste0(output_dir, "/vitamin_d_quantile_vs_onset_age.jpeg"),
  plot = plot_vitamin_d_quartile_vs_onset_age,
  dpi = 300
))
```


```{r}
fit_cause_specific_cox <- function(data, onset_var = "onset_age_group",   event_var = "event", exposure = "conc_25OHD", covariates = NULL) {
  onset_levels <- unique(na.omit(data[[onset_var]]))
  
  # Optional: Include covariates
  covariate_terms <- if (!is.null(covariates)) paste(covariates, collapse = " + ") else ""
  
  # Run model for each onset group
  results <- map_dfr(onset_levels, function(onset_type) {
    # Create binary event variable for this onset group
    data_mod <- data %>%
      mutate(event_temp = case_when(
        !!sym(onset_var) == onset_type ~ 1,
        !!sym(event_var) == 1 ~ 0,
        TRUE ~ 0
      ))
    
    # Create formula
    formula_str <- paste0("Surv(exit_age, event_temp) ~ ", exposure,
                          if (covariate_terms != "") paste0(" + ", covariate_terms) else "")
    cox_formula <- as.formula(formula_str)
    
    print(cox_formula)
    
    # Fit Cox model
    model <- coxph(cox_formula, data = data_mod)
    
    # Extract tidy results
    tidy(model, exponentiate = TRUE, conf.int = TRUE) %>%
      filter(term == exposure) %>%
      mutate(
        Onset_Group = onset_type,
        `HR (95% CI)` = paste0(round(estimate, 3), " (", round(conf.low, 3), "-", round(conf.high, 3), ")"),
        `P-value` = ifelse(
          p.value < 0.001, 
          formatC(p.value, format = "e", digits = 2), 
          formatC(p.value, format = "f", digits = 3)
          )) %>%
      select(Onset_Group, `HR (95% CI)`, `P-value`)
  })
  
  return(results)
}
```

```{r}
cox_results_onset_base_conc_25OHD <- fit_cause_specific_cox(
  data = df_cox_full,
  onset_var = "onset_age_group",
  event_var = "event",
  exposure = "conc_25OHD",
  covariates = NULL #c("Sex_cat", "Season")  # optional
)

cox_results_onset_base_conc_25OHD
```

```{r}
cox_results_onset_base_conc_25OHD_child_covariates <- fit_cause_specific_cox(
  data = df_cox_full,
  onset_var = "onset_age_group",
  event_var = "event",
  exposure = "conc_25OHD",
  covariates = c(child_covariates[-2],"Birth_weight_cat")
)

cox_results_onset_base_conc_25OHD_child_covariates
```

```{r}
cox_results_onset_base_conc_25OHD_maternal_covariates <- fit_cause_specific_cox(
  data = df_cox_full,
  onset_var = "onset_age_group",
  event_var = "event",
  exposure = "conc_25OHD",
  covariates = maternal_covariates
)

cox_results_onset_base_conc_25OHD_maternal_covariates
```

```{r}
cox_results_onset_base_conc_25OHD_all_covariates <- fit_cause_specific_cox(
  data = df_cox_full,
  onset_var = "onset_age_group",
  event_var = "event",
  exposure = "conc_25OHD",
  covariates = c(child_covariates[-2],"Birth_weight_cat", maternal_covariates)
)

cox_results_onset_base_conc_25OHD_all_covariates
```


```{r}
# Function to fit Cox model in each onset group
fit_cox_by_onset_group <- function(group) {
  data_group <- df_cox_case %>% filter(onset_age_group == group)

  if (n_distinct(data_group$quantile_25OHD) < 2) return(NULL)

  model <- coxph(Surv(exit_age, event) ~ quantile_25OHD, data = data_group)
  broom::tidy(model, exponentiate = TRUE, conf.int = TRUE) %>%
  dplyr::mutate(Onset_Group = group)
}

# Apply to each onset group
stratified_models <- levels(df_cox_case$onset_age_group)

cox_results <- map_dfr(stratified_models, fit_cox_by_onset_group)

# Clean and format
formatted_cox <- cox_results %>%
  dplyr::filter(term != "quantile_25OHDQ1") %>%  # Q1 is reference
  dplyr::mutate(
    `HR (95% CI)` = paste0(round(estimate, 2), " (", round(conf.low, 2), "–", round(conf.high, 2), ")"),
    `P-value` = ifelse(
      p.value < 0.001,
      formatC(p.value, format = "e", digits = 2),
      formatC(p.value, format = "f", digits = 3)
      )
  ) %>%
  dplyr::select(Onset_Group, Term = term, `HR (95% CI)`, `P-value`)  %>%
  dplyr::mutate(
    Term = gsub("quantile_25OHD","",Term),
    Onset_Group = ifelse(
      row_number() >1 & Onset_Group== lag(Onset_Group), 
      NA_character_, 
      Onset_Group)
    ) 

flex_cox_onset_age <- formatted_cox %>%
  dplyr::rename(`Onset Group` = Onset_Group) %>%
  flextable::flextable() %>% 
  flextable::autofit() %>%  
  flextable::set_table_properties(width = 1, layout = "autofit") %>% 
  flextable::color(
     i = ~ as.numeric(`P-value`) < 0.05,
     j = c("Term", "Onset Group", "P-value"),
     color = "red")
```


```{r save_flex_cox_onset_age}
flextable::save_as_docx(
  `Cox model: quantile 25(OH)D (Onset age group-wise)` = flex_cox_onset_age,
  path = paste0(output_dir,"/flex_cox_quantile_25ohd_onset_age_group_wise.docx"),
  pr_section = sect_properties
)
```


<!-- ```{r cox_results} -->
<!-- tidy_cox_base <- tidy(cox_model_base, exponentiate = TRUE, conf.int = TRUE)  %>% -->
<!--   dplyr::rename( -->
<!--     Term = term, -->
<!--     HR = estimate, -->
<!--     `P-value` = p.value -->
<!--   ) %>% -->
<!--   dplyr::mutate(`HR (95% CI)` = paste0( -->
<!--     as.character(round(HR,3)), -->
<!--     " (", -->
<!--     as.character(round(conf.low,3)), -->
<!--     "-", -->
<!--     as.character(round(conf.high,3)), -->
<!--     ")" -->
<!--   ))  %>% -->
<!--   dplyr::select(all_of(c("Term", "HR (95% CI)", "P-value"))) -->
<!-- ``` -->


<!-- ```{r cox_results} -->
<!-- tidy_cox_base  %>% -->
<!--   as.data.frame() %>% -->
<!--   dplyr::mutate_if(is.numeric, round, 3) %>% -->
<!--   flextable::flextable() %>% -->
<!--   flextable::autofit() %>% -->
<!--   flextable::color( -->
<!--     i = ~ as.numeric(`P-value`) < 0.05, -->
<!--     j = c("Term", "P-value"), -->
<!--     color = "red") -->
<!-- ``` -->








<!-- weightIt to balance out covariates across conc_25OHD exposure/treatment  -->
<!-- ```{r} -->
<!-- w_25OHD_cox <- weightit( -->
<!--   conc_25OHD ~  Preterm_dichotomous + Season + First_degree_T1D_heredity_cat + -->
<!--     Sex_cat + Smoking_during_pregnancy + Fish_any_pregnancy + Delivery_mode + -->
<!--     Parity + Birth_weight_cat, -->
<!--   data = df_cox_full, -->
<!--   method = "ebal", -->
<!--   maxit = 50000 -->
<!-- ) -->
<!-- ``` -->

<!-- Cox regression (weighted)	Hazard ratio: relative rate of developing T1D at any given age.  -->
<!-- ```{r cox_model_weighted} -->
<!-- cox_model_weighted <- coxph( -->
<!--   Surv(exit_age, event = event) ~ conc_25OHD + First_degree_T1D_heredity_cat + -->
<!--     Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode + -->
<!--     Preterm_dichotomous + Season + Sex_cat + Birth_weight_cat, -->
<!--   data = df_cox_full, weights = w_25OHD_cox$weights -->
<!-- ) -->
<!-- ``` -->

<!-- ```{r cox_weighted_results} -->
<!-- tidy_cox_weighted <- tidy(cox_model, exponentiate = TRUE, conf.int = TRUE) -->
<!-- tidy_cox_weighted  %>% -->
<!--   as.data.frame() %>% -->
<!--   dplyr::mutate_if(is.numeric, round, 3) %>% -->
<!--   flextable::flextable() %>% -->
<!--   flextable::autofit() %>% -->
<!--   flextable::color( -->
<!--     i = ~ p.value < 0.05, -->
<!--     j = c("term", "p.value"), -->
<!--     color = "red") -->
<!-- ``` -->