---
title: "logistic Regression"
author: "Debojyoti Das"
date: "2025-07-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

set output directory
```{r output_dir, message=FALSE, warning=FALSE}
output_dir <- "/Users/debda22/OneDrive - Linköpings universitet/T1D_vitamin_D_result/logistic_regression"

if(!dir.exists(output_dir)){
  dir.create(output_dir, recursive = TRUE)
} else {
  cat(paste0(output_dir," is already present!!\n"))
}
```

load libraries
```{r load_libraries, message=FALSE, warning=FALSE}
pacman::p_load(
  haven,        
  zoo,
  sjlabelled,
  janitor,       
  data.table,   
  ggplot2,
  comprehenr,  
  stringr,     
  rstatix,
  dplyr,
  tidyr,
  forcats,
  officer,
  cowplot,
  ggExtra,
  ggpubr,
  GGally,
  janitor,
  ggrepel,
  ggtext,
  parameters,
  lme4,
  broom,
  pscl,
  tibble,
  knitr,
  ggeffects,
  cobalt,
  WeightIt,
  marginaleffects,
  survival
  )
```

define paths
```{r define_paths, eval=TRUE}
data_root <- "/Users/debda22/OneDrive - Linköpings universitet/"
data_root <- paste0(data_root,"Linkoping_projects/ABIS_RAW_DATA_SPSS/")
```

access helper functions
```{r load_helper_function}
source(paste0(data_root,"../../Rmd_and_R_files/T1D_projects/case_control_T1D_vitamin_D_helper.R"))
```

load factor labels
```{r load_factor_labels}
source(paste0(data_root,"../../Rmd_and_R_files/T1D_projects/all_parameters_case_control_T1D_vitamin_D.R"))
```

read/load vitamin_d_case_control_plus_abis for analysis
```{r}
vitamin_d_case_control_plus_abis <- read.table(
  paste0(
    data_root,
    "/vitamin_d/vitamin_d_case_control_plus_abis.csv"
    )
)
```

```{r features2check}
# categorical features
features2check <- names(vitamin_d_case_control_plus_abis)[c(6,8:9,13:21)]

# continuous features
features2check_numeric <- names(vitamin_d_case_control_plus_abis)[c(10:12)]
```

model baseline factors ~ diagnosis
```{r}
univariate_models_25OHD <- complete_case_univariate_logistic_regression(
  c(features2check, features2check_numeric),
  vitamin_d_case_control_plus_abis)
```
```{r}
univariate_models_25OHD_flex <- univariate_models_25OHD %>%
  as.data.frame() %>% dplyr::mutate(term = ifelse(
    term %in% c("Season", unlist(all_parameters$category_variable), features2check_numeric),
    gsub("_", " ", gsub(
      "_pregnancy", " (pregnancy)", gsub("_cat", "", gsub(
        "_dichotomous",
        "",
        gsub(
          "Gestational_age",
          "Gestational age (weeks)",
          gsub(
            "Birth_height",
            "Birth height (cm)",
            gsub("Birth_weight", "Birth weight (kg)", term)
          )
        )
      ))
    )),
    paste0("\t", term)
  )) %>%
  dplyr::rename(
    Term = term,
    No = no,
    Yes = yes,
    OR = estimate,
    `P-value` = p.value,
    Conf.low = conf.low,
    Conf.high = conf.high
  ) %>%
  dplyr::mutate( 
    Term = if_else(Term == lag(Term) & row_number() > 1, "", Term)
  ) %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::color(i = ~ !grepl("\t", Term), ~ Term, color = "blue")  %>%
  flextable::color(i = ~ `P-value` < 0.05,
                   j = c(1, 4, 5),
                   color = "red")

univariate_models_25OHD_flex
```

Fit models: Diagnosis  vitamin D

```{r vitamin_d_models_summary, message=FALSE, warning=FALSE}
model_base <- glm(diagnosis ~ conc_25OHD, family = "binomial", data = vitamin_d_case_control_plus_abis)

model_maternal <- glm(diagnosis ~ conc_25OHD +
                        First_degree_T1D_heredity_cat +
                        Smoking_during_pregnancy +
                        Vitamin_mineral_supplement_pregnancy +
                        Fish_any_pregnancy +
                        Parity +
                        Delivery_mode,
                      family = "binomial", data = vitamin_d_case_control_plus_abis)

model_child <- glm(diagnosis ~ conc_25OHD +
                     Preterm_dichotomous +
                     Season +
                     Sex_cat,
                   family = "binomial", data = vitamin_d_case_control_plus_abis)

model_full <- glm(diagnosis ~ conc_25OHD +
                    First_degree_T1D_heredity_cat +
                    Smoking_during_pregnancy +
                    Vitamin_mineral_supplement_pregnancy +
                    Fish_any_pregnancy +
                    Parity +
                    Delivery_mode +
                    Preterm_dichotomous +
                    Season +
                    Sex_cat,
                  family = "binomial", data = vitamin_d_case_control_plus_abis)
```

Fit interaction models
```{r vitamin_d_interaction_models_summary, message=FALSE, warning=FALSE}
model_maternal_interaction <- glm(diagnosis ~ conc_25OHD *(
                        First_degree_T1D_heredity_cat +
                        Smoking_during_pregnancy +
                        Vitamin_mineral_supplement_pregnancy +
                        Fish_any_pregnancy +
                        Parity +
                        Delivery_mode),
                      family = "binomial", data = vitamin_d_case_control_plus_abis)

model_child_interaction <- glm(diagnosis ~ conc_25OHD *(
                     Preterm_dichotomous +
                     Season +
                     Sex_cat),
                   family = "binomial", data = vitamin_d_case_control_plus_abis)

model_full_interaction <- glm(diagnosis ~ conc_25OHD *(
                    First_degree_T1D_heredity_cat +
                    Smoking_during_pregnancy +
                    Vitamin_mineral_supplement_pregnancy +
                    Fish_any_pregnancy +
                    Parity +
                    Delivery_mode +
                    Preterm_dichotomous +
                    Season +
                    Sex_cat),
                  family = "binomial", data = vitamin_d_case_control_plus_abis)
```

```{r vitamin_d_models_summary_merge, message=FALSE, warning=FALSE}
# Create model list
model_list <- list(model_base, model_maternal, model_maternal_interaction, model_child, model_child_interaction, model_full, model_full_interaction)
model_names <- c("Base (25OHD only)",
                 "Base + Maternal covariates",
                 "Base * Maternal covariates (interaction model)",
                 "Base + Child/Perinatal covariates",
                 "Base * Child/Perinatal covariates (interaction model)",
                 "Full model",
                 "Full model (interaction model)")

# Extract OR, CI, R², and AIC
suppressMessages(summary_table <- lapply(seq_along(model_list), function(i) {
  model <- model_list[[i]]
  or_info <- tidy(model, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(term == "conc_25OHD") %>%
    dplyr::mutate(OR = estimate, CI = paste0("(", round(conf.low, 2), "–", round(conf.high, 2), ")")) %>%
    dplyr::select(OR, CI)
  
  r2_val <- pR2(model)["McFadden"]
  aic_val <- AIC(model)
  
  tibble(
    Model = model_names[i],
    `Odds Ratio (25OHD)` = round(or_info$OR, 2),
    `95% CI (25OHD)` = or_info$CI,
    `Pseudo-R² (McFadden)` = round(r2_val, 2),
    AIC = round(aic_val, 1)
  )
}) %>%
  bind_rows())
```

Summary of Logistic Regression Models for 25(OH)D

```{r vitamin_d_models_summary_display, message=FALSE, warning=FALSE}
# kable(summary_table, caption = "Summary of Logistic Regression Models for 25OHD")
summary_table %>%
  as.data.frame() %>%
  flextable::flextable() %>%
  flextable::autofit() 
```

Prepare data for plotting
```{r}
plot_data <- summary_table %>%
  dplyr::mutate(
    Model = factor(Model, levels = Model), # preserve order
    OR = `Odds Ratio (25OHD)`,
    R2 = `Pseudo-R² (McFadden)`,
    # Extract lower and upper bounds from CI string "(x–y)"
    CI_lower = as.numeric(sub(".*\\((.*)–.*", "\\1", `95% CI (25OHD)`)),
    CI_upper = as.numeric(sub(".*–(.*)\\)", "\\1", `95% CI (25OHD)`))
  )

plot_data %>%
  as.data.frame() %>%
  flextable::flextable() %>%
  flextable::autofit() 
```


```{r}
model_plot <- plot_data %>%
  ggplot(aes(y = Model, x = OR, size = R2, label = Model)) +
  geom_point(aes(color = Model, size = R2), alpha = 0.7) +
  geom_pointrange(aes(xmin = CI_lower, xmax = CI_upper, color = Model),alpha = 0.7) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "red") +
  labs(
    x = "Odds Ratio (25OHD)",
    y = "",
    size = "R²"
  )  +
  scale_size_continuous(
    breaks = c(0.02, 0.04, 0.06,0.08, 0.10, 0.140),
    range = c(1, 2.2),
    guide = guide_axis(check.overlap = TRUE)
  ) +
  xlim(c(0.70,1.10)) +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),   
    axis.text.y = element_text(hjust = 0),
    legend.position = "none"
)

model_plot 
```

save plots
```{r warning=FALSE}
suppressWarnings(
  ggsave(
  filename = paste0(output_dir,"/model_OR_25OHD.jpeg"),
  plot = model_plot,
  dpi = 300,
  width = 8.3,
  height = 11.7,
  units = "in"
  )
)
```

group prediction probabilities
```{r}
groups <- names(model_full_interaction$model)[-c(1, 2, 11)]

plot_group_probs <- vector("list", length = length(groups))

for (iv in seq_along(groups)) {
  v <- groups[[iv]]
  
  pred <- ggpredict(model_full_interaction,
                    terms = c("conc_25OHD [all]", v, "Sex_cat"))
  color_text <- gsub("_", " ", gsub("_cat", "", gsub(
    "_dichotomous",
    "",
    gsub(
      "_during_pregnancy|_any_pregnancy", " (pregnancy)",
           gsub("Vitamin_mineral_supplement_pregnancy", " Vitamin/mineral\n(pregnancy)",
                 gsub("First_degree_T1D_heredity_cat", "T1D heredity",v)
    ))
  )))
  title_text <- paste0("Interaction: 25(OH)D × ", color_text, " × Sex")
  # Plot manually
  plot_group_probs[[iv]] <- ggplot(pred, aes(x = x, y = predicted, color = group)) +
    geom_line(size = 1) +
    facet_wrap(~ facet) +  # Will use v here
    labs(
      # x = "Centered 25OHD (nmol/L from mean)",
      x = "25(OH)D (nmol/L)",
      y = "Predicted probability of T1D",
      color = color_text
      # title = title_text
    ) +
    xlim(c(0, 100)) +
    theme_minimal()
}
```

```{r}
for(iv in seq_along(plot_group_probs)){
  print(plot_group_probs[[iv]])
}
```

save plots
```{r warning=FALSE}
cairo_pdf(paste0(output_dir,"/predicted_prob_full_interaction_model.pdf"))
for(iv in seq_along(plot_group_probs)){
  print(plot_group_probs[[iv]])
}
dev.off()
```

DBP: Fit models
```{r DBP_models, message=FALSE, warning=FALSE}
model_dbp_base <- glm(diagnosis ~ DBP, family = "binomial", data = vitamin_d_case_control_plus_abis)

model_dbp_maternal <- glm(diagnosis ~ DBP +
                        First_degree_T1D_heredity_cat +
                        Smoking_during_pregnancy +
                        Vitamin_mineral_supplement_pregnancy +
                        Fish_any_pregnancy +
                        Parity +
                        Delivery_mode,
                      family = "binomial", data = vitamin_d_case_control_plus_abis)

model_dbp_child <- glm(diagnosis ~ DBP +
                     Preterm_dichotomous +
                     Season +
                     Sex_cat,
                   family = "binomial", data = vitamin_d_case_control_plus_abis)

model_dbp_full <- glm(diagnosis ~ DBP +
                    First_degree_T1D_heredity_cat +
                    Smoking_during_pregnancy +
                    Vitamin_mineral_supplement_pregnancy +
                    Fish_any_pregnancy +
                    Parity +
                    Delivery_mode +
                    Preterm_dichotomous +
                    Season +
                    Sex_cat,
                  family = "binomial", data = vitamin_d_case_control_plus_abis)
```

DBP : Interaction models
```{r DBP_interaction_models_summary, message=FALSE, warning=FALSE}
model_dbp_maternal_interaction <- glm(diagnosis ~ DBP *(
                        First_degree_T1D_heredity_cat +
                        Smoking_during_pregnancy +
                        Vitamin_mineral_supplement_pregnancy +
                        Fish_any_pregnancy +
                        Parity +
                        Delivery_mode),
                      family = "binomial", data = vitamin_d_case_control_plus_abis)

model_dbp_child_interaction <- glm(diagnosis ~ DBP *(
                     Preterm_dichotomous +
                     Season +
                     Sex_cat),
                   family = "binomial", data = vitamin_d_case_control_plus_abis)

model_dbp_full_interaction <- glm(diagnosis ~ DBP *(
                    First_degree_T1D_heredity_cat +
                    Smoking_during_pregnancy +
                    Vitamin_mineral_supplement_pregnancy +
                    Fish_any_pregnancy +
                    Parity +
                    Delivery_mode +
                    Preterm_dichotomous +
                    Season +
                    Sex_cat),
                  family = "binomial", data = vitamin_d_case_control_plus_abis)
```

DBP: Create model list
```{r DBP_models_summary_merge, message=FALSE, warning=FALSE}
model_dbp_list <- list(model_dbp_base, model_dbp_maternal, model_dbp_maternal_interaction, model_dbp_child, model_dbp_child_interaction, model_dbp_full, model_dbp_full_interaction)
model_dbp_names <- c("Base (DBP only)",
                 "Base + Maternal covariates",
                 "Base * Maternal covariates (interaction model)",
                 "Base + Child/Perinatal covariates",
                 "Base * Child/Perinatal covariates (interaction model)",
                 "Full model",
                 "Full model (interaction model)")

# Extract OR, CI, R², and AIC
suppressWarnings(
  summary_dbp_table <- lapply(seq_along(model_dbp_list), function(i) {
  model <- model_dbp_list[[i]]
  or_info <- tidy(model, conf.int = TRUE, exponentiate = TRUE) %>%
    dplyr::filter(term == "DBP") %>%
    dplyr::mutate(
      OR = estimate,
      CI = paste0("(", round(conf.low, 2), "–", round(conf.high, 2), ")")
    ) %>%
    dplyr::select(OR, CI)
  
  r2_val <- pR2(model)["McFadden"]
  aic_val <- AIC(model)
  
  tibble(
    Model = model_dbp_names[i],
    `Odds Ratio (DBP)` = round(or_info$OR, 2),
    `95% CI (DBP)` = or_info$CI,
    `Pseudo-R² (McFadden)` = round(r2_val, 2),
    AIC = round(aic_val, 1)
  )
}) %>%
  bind_rows()
)
```
Summary of Logistic Regression Models for DBP
```{r dbp_models_summary_display, message=FALSE, warning=FALSE}
# kable(summary_dbp_table, caption = "Summary of Logistic Regression Models for DBP")
summary_dbp_table %>%
  as.data.frame() %>%
  flextable::flextable() %>%
  flextable::autofit() 
```


Prepare data for plotting
```{r}
plot_data_dbp <- summary_dbp_table %>%
  dplyr::mutate(
    Model = factor(Model, levels = Model), # preserve order
    OR = `Odds Ratio (DBP)`,
    R2 = `Pseudo-R² (McFadden)`,
    # Extract lower and upper bounds from CI string "(x–y)"
    CI_lower = as.numeric(sub(".*\\((.*)–.*", "\\1", `95% CI (DBP)`)),
    CI_upper = as.numeric(sub(".*–(.*)\\)", "\\1", `95% CI (DBP)`))
  )

plot_data_dbp  %>%
  as.data.frame() %>%
  flextable::flextable() %>%
  flextable::autofit() 
```


```{r}
model_dbp_plot <- plot_data_dbp %>%
  ggplot(aes(y = Model, x = OR, size = R2, label = Model)) +
  geom_point(aes(color = Model, size = R2), alpha = 0.7) +
  geom_pointrange(
    aes(xmin = CI_lower, xmax =  ifelse(CI_upper>3,OR+0.15,CI_upper), 
        color = Model),alpha = 0.7) +
  geom_vline(xintercept = 1, linetype = "dashed", color = "red") +
  labs(
    x = "Odds Ratio (DBP)",
    y = "",
    size = "R²"
  )  +
  scale_size_continuous(
    breaks = c(0.02, 0.04, 0.06,0.08, 0.10, 0.120),
    range = c(1, 2.2) #,
    # guide = guide_axis(check.overlap = TRUE)
  ) +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),   
    axis.text.y = element_text(hjust = 0),
    legend.position = "none"
)

model_dbp_plot 
```


save plots
```{r}
ggsave(
  filename = paste0(output_dir,"/model_OR_DBP.jpeg"),
  plot = model_dbp_plot,
  dpi = 300,
  width = 8.3,
  height = 11.7,
  units = "in"
  )
```


group prediction probabilities DBP
```{r}
groups <- names(model_dbp_full_interaction$model)[-c(1, 2, 11)]
print(groups)
```


```{r}
plot_group_probs_dbp <- vector("list", length = length(groups))

for (iv in seq_along(groups)) {
  v <- groups[[iv]]
  
  pred_dbp <- ggpredict(model_dbp_full_interaction,
                    terms = c("DBP [all]", v, "Sex_cat"))
  color_text <- gsub("_", " ", gsub("_cat", "", gsub(
    "_dichotomous",
    "",
    gsub(
      "_during_pregnancy|_any_pregnancy", " (pregnancy)",
           gsub("Vitamin_mineral_supplement_pregnancy", " Vitamin/mineral\n(pregnancy)",
                 gsub("First_degree_T1D_heredity_cat", "T1D heredity",v)
    ))
  )))
  title_text <- paste0("Interaction: DBP × ", color_text, " × Sex")
  # Plot manually
  plot_group_probs_dbp[[iv]] <- ggplot(pred_dbp, aes(x = x, y = predicted, color = group)) +
    geom_line(size = 1) +
    facet_wrap(~ facet) +  # Will use v here
    labs(
      x = "DBP (\u03bcmol/L)",
      y = "Predicted probability of T1D",
      color = color_text
      # title = title_text
    ) +
    xlim(c(0, 10)) +
    theme_minimal()
}
```

```{r}
for(iv in seq_along(plot_group_probs_dbp)){
  print(plot_group_probs_dbp[[iv]])
}
```

save plots
```{r}
cairo_pdf(paste0(output_dir,"/predicted_prob_full_interaction_model_DBP.pdf"))
for(iv in seq_along(plot_group_probs_dbp)){
  print(plot_group_probs_dbp[[iv]])
}
dev.off()
```



glm diagnosis ~ conc_25OHD
```{r}
fit_25OHD_no_covariate <- glm(
    formula = diagnosis ~ conc_25OHD,
    data = vitamin_d_case_control_plus_abis,
    family = binomial
)

results_25OHD_diagnosis_no_covariate <- fit_25OHD_no_covariate %>%
  broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>%
  dplyr::mutate_if(is.numeric, function(x)
    ifelse(x < 0.001, formatC(x, format = "e", 3), round(x, 3))) %>%
  dplyr::filter(term != "(Intercept)")  %>%
  dplyr::select(-c("std.error", "statistic")) %>%
  dplyr::rename(
    Term = term,
    OR = estimate,
    `P-value` = p.value,
    Conf.low = conf.low,
    Conf.high =conf.high
    )


results_25OHD_diagnosis_no_covariate  %>%
  as.data.frame() %>%
  flextable::flextable() %>%
  flextable::autofit() 
```
