---
title: "Add diagnosis (T1D)  from medical records and ABIS info to vitamin D data"
author: "Debojyoti Das"
date: "2025-05-26"
output: html_document
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

set output directory
```{r output_dir, message=FALSE, warning=FALSE}
output_dir <- "/Users/debda22/OneDrive - Linköpings universitet/T1D_vitamin_D_results/baseline_characteristics"

if(!dir.exists(output_dir)){
  dir.create(output_dir, recursive = TRUE)
} else {
  cat(paste0(output_dir," is already present!!\n"))
}
```

load libraries
```{r load_libraries, message=FALSE, warning=FALSE}
pacman::p_load(
  haven,        
  zoo,
  sjlabelled,
  janitor,       
  data.table,   
  ggplot2,
  comprehenr,  
  stringr,     
  rstatix,
  dplyr,
  tidyr,
  forcats,
  officer,
  cowplot,
  ggExtra,
  ggpubr,
  GGally,
  janitor,
  ggrepel,
  ggtext,
  parameters,
  lme4,
  broom,
  pscl,
  tibble,
  knitr,
  ggeffects,
  cobalt,
  WeightIt,
  marginaleffects,
  survival
  )
```

define paths
```{r define_paths, eval=TRUE}
data_root <- "/Users/debda22/OneDrive - Linköpings universitet/"
data_root <- paste0(data_root,"Linkoping_projects/ABIS_RAW_DATA_SPSS/")
```

load parameters
```{r}
source(paste0(data_root,"../../Rmd_and_R_files/T1D_projects/all_parameters_case_control_T1D_vitamin_D.R"))
```


```{r read_vitamin_D_data, message=FALSE, warning=FALSE }
suppressMessages(
  vitamin_d <- readxl::read_xlsx(
  paste0(data_root,"vitamin_d/Vitamin_D_16_2023_ABIS.xlsx"))
)
vitamin_d <- vitamin_d %>% as.data.frame()
```


```{r vitamin_D_data_rename_columns}
vitamin_d <- vitamin_d[-c(1,2),-c(1,3,6,7)]

column_names <- c("ABISnr", "DBP", "conc_25OHD")
names(vitamin_d) <-column_names

head(vitamin_d,4)
```


Read cohort data from SAV format file
```{r read_abis_data}
abis <- read_sav(paste0(data_root, "Updated registries/ABIS_Database_Updated_240124.sav"))
```


Variable names and labels
```{r extract_abis_colname_description}
column_names <- names(abis)
label <- to_vec(
  for (i in seq_along(column_names))
    if (
      (!is.null(
        attr(abis[[column_names[i]]], "label")
        )
       )&(
         length(attr(abis[[column_names[i]]], "label")) == 1)
      ) attr(abis[[column_names[i]]], "label")
  else column_names[i])

abis_info <- data.frame(
  name = column_names,
  label = label
)
```


```{r extract_T1D_question_columns}
T1D_first_order_heridity <- abis_info %>%
  dplyr::mutate(
    T1D_info = ifelse(
      grepl("type 1 Diabetes",label,ignore.case = TRUE),label,NA)) %>%
  dplyr::filter(
    (!is.na(T1D_info))&(!grepl("grand", T1D_info, ignore.case=TRUE))
    ) #%>%
  # dplyr::filter(
  #   (!is.na(T1D_info)) & grepl("suffer", T1D_info) &
  #     (!grepl("screen|test", T1D_info, ignore.case = TRUE) &
  #        (!grepl("does not know", T1D_info, ignore.case = TRUE))
  #      )
  #   )

str(T1D_first_order_heridity)
```
season of birth
```{r add_season_of_birth}
abis <- abis %>%
  dplyr::mutate(
    child_dob = as.Date(
      as.yearmon(as.character(barn_år_mån), format = "%Y%m")
      )
    ) %>%
  dplyr::mutate(
    month_num = lubridate::month(child_dob),
    Season = dplyr::case_when(
      month_num %in% c(12, 1, 2)  ~ "Winter",
      month_num %in% c(3, 4, 5)   ~ "Spring",
      month_num %in% c(6, 7, 8)   ~ "Summer",
      month_num %in% c(9, 10, 11) ~ "Autumn"
   )
  ) %>%
  dplyr::select(-all_of(c("barn_år_mån", "month_num")))
```

<!-- composite_T1D_heredity = fr_15f_1 mother with T1D -->
<!-- composite_T1D_heredity = pmax(fr_15f_1, fr_15f_2, fr_15f_3, na.rm = TRUE)) mother/ father/ sibling -->

```{r add_T1D_heredity}
abis <- abis %>% 
  dplyr::mutate(across(c(fr_15f_1, fr_15f_2, fr_15f_3), ~ as.numeric(.))) %>%
  dplyr::mutate(
    composite_T1D_heredity = pmax(fr_15f_1, fr_15f_2, fr_15f_3, na.rm = TRUE)
    ) %>%
    # composite_T1D_heredity = fr_15f_1) %>% 
  dplyr::mutate(
    first_degree_T1D_heredity = ifelse(!is.na(composite_T1D_heredity),composite_T1D_heredity,NA_integer_)
  ) 
```

```{r remove_attributes_to_preserve_data_frame_class, echo=FALSE}
for(col in names(abis)){
  for(atr in c("label","format.spss","display_width","labels")){
    attr(abis[[col]], atr) <- NULL
  }
}
```

```{r select_columns_of_interest}
abis <- abis %>% dplyr::select(
  ABISnr,
  child_dob,
  first_degree_T1D_heredity,
  # fr_15f_1, # Inheritance, T1D mom
  # fr_15f_2, # Inheritance, T1D dad
  # fr_15f_3, # Inheritance, T1D siblings
  fr_18,    # Did you smoke during your pregnancy
  # fr_18a,  # If yes, how many cigarettes per day?
  parity,  # 0 = "first parity", 1 = "previous parity"
  fr_81,   # Did you use any vitamins or mineral supplement during your pregnancy?
  fr_91,   # How often did you eat lake fish?
  fr_92,   # How often did you eat Baltic fish ?
  fr_93,   # How often did you eat other fish?
  # fr_92_93_sum_recoded,  # mother during pregnancy: index of long chain omega3 intake based on fish intake
  fr_12,   # Week of delivery
  fr_13,   # The way of delivery
  fr_14,   # birth weight
  fr_14a,  # birth length
  Season,
  Gender
) %>%
  dplyr::mutate(across(c(fr_91, fr_92, fr_93), ~ as.numeric(.))) %>%
  dplyr::mutate(
    fish_intake_frequency = pmin(
      fr_91, 
      fr_92, 
      fr_93, 
      na.rm = TRUE)
  )
```


access helper functions
```{r load_helper_function}
source("case_control_T1D_vitamin_D_helper.R")
```

load factor labels
```{r load_factor_labels}
source("all_parameters_case_control_T1D_vitamin_D.R")
```

remove vitamin supplement "Do not know"
```{r}
abis <- abis %>%
  dplyr::mutate(
    fr_81 = ifelse(
      is.na(fr_81) | fr_81 == 3,
      NA,
      fr_81)
  )
```


```{r categorise_variables_with_proper_labels}
for (cv in unlist(all_parameters$category_variable)){
  
  curr_parameters <- all_parameters[all_parameters$category_variable == cv,]
  abis <- categorise_variable(abis,
                          variable2categorise = curr_parameters$variable2categorise[[1]],
                          category_variable = curr_parameters$category_variable[[1]],
                          breaks = curr_parameters$breaks[[1]],
                          labels = curr_parameters$labels[[1]]
)
  abis[[cv]] <- droplevels(abis[[cv]])
}
```

reorder some factors
```{r reorder_some_factors1}
abis$Fish_lake_pregnancy <- factor(
  abis$Fish_lake_pregnancy,
  levels = rev(c("Every day", "3\u20135 times/week", "1\u20132 times week", "Seldom")))

abis$Fish_Baltic_pregnancy <- factor(
  abis$Fish_Baltic_pregnancy,
  levels = rev(c("Every day", "3\u20135 times/week", "1\u20132 times week", "Seldom")))

abis$Fish_other_pregnancy <- factor(
  abis$Fish_other_pregnancy,
  levels = rev(c("Every day", "3\u20135 times/week", "1\u20132 times week", "Seldom")))

abis$Fish_any_pregnancy <- factor(
  abis$Fish_any_pregnancy,
  levels = rev(c("Every day", "3\u20135 times/week", "1\u20132 times week", "Seldom")))
```

```{r reorder_some_factors2}
abis$Fish_lake_pregnancy <- forcats::fct_collapse(
  abis$Fish_lake_pregnancy,
  "1–2 times week or more" = c("Every day", "3–5 times/week", "1–2 times week")
)

abis$Fish_Baltic_pregnancy <- forcats::fct_collapse(
  abis$Fish_Baltic_pregnancy,
  "1–2 times week or more" = c("Every day", "3–5 times/week", "1–2 times week")
)

abis$Fish_other_pregnancy <- forcats::fct_collapse(
  abis$Fish_other_pregnancy,
  "1–2 times week or more" = c("Every day", "3–5 times/week", "1–2 times week")
)

abis$Fish_any_pregnancy <- forcats::fct_collapse(
  abis$Fish_any_pregnancy,
  "1–2 times week or more" = c("Every day", "3–5 times/week", "1–2 times week")
)
```


```{r rename_columns}
abis["Gestational_age"] <- abis$fr_12 
abis["Birth_weight"] <- abis$fr_14 
abis["Birth_length"] <- abis$fr_14a 
```

```{r select_renamed_columns}
abis <- abis %>% dplyr::select(all_of(
  c(
    "ABISnr",
    "child_dob",
    "First_degree_T1D_heredity_cat",
    "Preterm_group",
    "Preterm_dichotomous",
    "Delivery_mode",
    "Gestational_age",
    "Birth_weight",
    "Birth_length",
    "Season",
    "Parity",
    "Smoking_during_pregnancy",
    "Vitamin_mineral_supplement_pregnancy",
    "Fish_lake_pregnancy",
    "Fish_Baltic_pregnancy",
    "Fish_other_pregnancy",
    "Fish_any_pregnancy",
    "Sex_cat"
  )
))

abis <- as.data.frame(abis)
```


```{r set_season_reference_level}
abis$Season <- factor(
  abis$Season, levels = c("Autumn", "Spring", "Summer", "Winter")
  )
```

Diagnosis from NPR
```{r read_diagnosis_NPR}
OV_NPR_ORIGINAL <- read_sav(paste0(
  data_root,
  "Updated registries/OV_ABISnr_updated_20240301.sav"
))
SV_NPR_ORIGINAL <- read_sav(paste0(
  data_root,
  "Updated registries/SV_ABISnr_updated _20240301.sav"
))
```


```{r reformat_diagnosis_data_NPR}
SV_NPR <- SV_NPR_ORIGINAL %>%  
  dplyr::mutate(daywise_all_diagnosis = paste(HDIA,DIAGNOS, sep = " ")) %>%
  dplyr::mutate(daywise_all_diagnosis = gsub(replacement = ",",
                                      pattern = " ",
                                      x = daywise_all_diagnosis)) %>% 
  dplyr::group_by(ABISnr) %>% 
  dplyr::summarise(ABISnr_all_diagnosis = paste0(daywise_all_diagnosis, 
                                         collapse = ",")) %>%
  dplyr::select(ABISnr,ABISnr_all_diagnosis)

OV_NPR <- OV_NPR_ORIGINAL %>%  
  dplyr::mutate(daywise_all_diagnosis = paste(HDIA,DIAGNOS, sep = " ")) %>%
  dplyr::mutate(daywise_all_diagnosis = gsub(replacement = ",",
                                      pattern = " ",
                                      x = daywise_all_diagnosis)) %>% 
  dplyr::group_by(ABISnr) %>% 
  dplyr::summarise(ABISnr_all_diagnosis = paste0(daywise_all_diagnosis, 
                                         collapse = ","))  %>%
  dplyr::select(ABISnr,ABISnr_all_diagnosis)

diagnosis_SV_OV <- dplyr::full_join(SV_NPR, 
                             OV_NPR, 
                             by = "ABISnr",
                             suffix = c("_SV","_OV")
                             ) %>%
  dplyr::mutate(all_diagnosis = 
           paste(ABISnr_all_diagnosis_SV, 
                 ABISnr_all_diagnosis_OV, ",")) %>%
  dplyr::mutate(all_diagnosis = gsub(replacement = "",
                                      pattern = "NA ",
                                      x = all_diagnosis))
```



```{r get_diagnosis}
complete_diagnosis_SV_OV_psoriasis_or_ibd <- diagnosis_SV_OV %>%
dplyr::filter(grepl("L40|K50|K51|K523", all_diagnosis, ignore.case = TRUE)) %>%
dplyr::select(ABISnr, all_diagnosis) %>%
dplyr::mutate(Diagnos = sapply(all_diagnosis, function(x) {
    y <- strsplit(gsub(" ",",",x), ",")[[1]]
    match <- grep("L40|K50|K51|K523", y, ignore.case = TRUE)
if (length(match) > 0) return(y[match]) else return(NA)
}),
count =sapply(Diagnos, function(x) length(x)),
Autoimmune_diagnoses = sapply(Diagnos, function(x) paste0(x, collapse = "|")),
Diagnos = ifelse(count>0,"psoriasis_or_ibd",NA_character_),
all_diagnosis = sapply(all_diagnosis, function(x) paste0(base::strsplit(x,",")[[1]], collapse = "|"))
)
```

Type 1 Diabetes

```{r get_T1D_diagnosis}
complete_diagnosis_SV_OV_T1D <- diagnosis_SV_OV %>%
dplyr::filter(grepl("E10", all_diagnosis, ignore.case = TRUE)) %>%
dplyr::select(ABISnr, all_diagnosis) %>%
dplyr::mutate(Diagnos = sapply(all_diagnosis, function(x) {
    y <- strsplit(gsub(" ",",",x), ",")[[1]]
    match <- grep("E10", y, ignore.case = TRUE)
if (length(match) > 0) return(y[match]) else return(NA)
}),
count =sapply(Diagnos, function(x) length(x)),
Diabetes_diagnoses = sapply(Diagnos, function(x) paste0(x, collapse = "|")),
Diagnos = ifelse(count>0,"E10",NA_character_),
all_diagnosis = sapply(all_diagnosis, function(x) paste0(base::strsplit(x,",")[[1]], collapse = "|"))
)
```

<!-- save: do not delete -->
```{r save_diagnosis, include=FALSE}
# writexl::write_xlsx(
#   complete_abis_diagnosis_T1D,
#   paste0(
#     data_root,
#     "vitamin_d/t1d_diagnosis_from_sv_and_ov_npr.xlsx"
#   )
# )
```

Type 2 Diabetes

```{r get_T2D_diagnosis}
complete_diagnosis_SV_OV_T2D <- diagnosis_SV_OV %>%
  dplyr::filter(grepl("E11", all_diagnosis, ignore.case = TRUE)) %>%
  dplyr::select(ABISnr, all_diagnosis) %>%
  dplyr::mutate(
    Diagnos = sapply(all_diagnosis, function(x) {
      y <- strsplit(gsub(" ", ",", x), ",")[[1]]
      match <- grep("E11", y, ignore.case = TRUE)
      if (length(match) > 0)
        return(y[match])
      else
        return(NA)
    }),
    count = sapply(Diagnos, function(x)
      length(x)),
    Diabetes_diagnoses = sapply(Diagnos, function(x)
      paste0(x, collapse = "|")),
    Diagnos = ifelse(count > 0, "E11", NA_character_),
    all_diagnosis = sapply(all_diagnosis, function(x)
      paste0(base::strsplit(x, ",")[[1]], collapse = "|"))
  )
```

T1D cases from sv/OV NPR no overlapping T2D/ psoriais/ IBD
```{r get_individual_diagnosis_list}
T1D_abis_list <- complete_diagnosis_SV_OV_T1D$ABISnr
T2D_abis_list <- complete_diagnosis_SV_OV_T2D$ABISnr
psoriais_or_ibd_abis_list <- complete_diagnosis_SV_OV_psoriasis_or_ibd$ABISnr
```


```{r diagnosis_list_overlap_check}
T1D_abis_list_clean <- T1D_abis_list[!(T1D_abis_list %in% c(T2D_abis_list,psoriais_or_ibd_abis_list))]
T2D_abis_list_clean <- T2D_abis_list[!(T2D_abis_list %in% c(T1D_abis_list,psoriais_or_ibd_abis_list))]
psoriais_or_ibd_abis_list_clean <- psoriais_or_ibd_abis_list[!(psoriais_or_ibd_abis_list %in% c(T1D_abis_list,T2D_abis_list))]
```


```{r extract_T1D_first_diagnosis}
T1D_first_diagnosis <- rbind(
  OV_NPR_ORIGINAL %>%
    dplyr::mutate(UTDATUMA = NA_character_, SOURCE = "OV_20240301") %>%
    dplyr::select(all_of(
      c("ABISnr", "HDIA", "DIAGNOS", "INDATUMA", "UTDATUMA", "SOURCE")
    ))
  ,
  SV_NPR_ORIGINAL %>%
    dplyr::mutate(SOURCE = "SV_20240301") %>%
    dplyr::select(all_of(
      c("ABISnr", "HDIA", "DIAGNOS", "INDATUMA", "UTDATUMA", "SOURCE")
    ))
) %>%
  dplyr::filter(grepl("E10", HDIA) | grepl("E10", DIAGNOS)) %>%
  dplyr::mutate(INDATUMA = as.Date(INDATUMA, format = "%Y%m%d")) %>%
  dplyr::arrange(ABISnr, INDATUMA) %>%
  dplyr::group_by(ABISnr) %>%
  dplyr::slice_head(n = 1) %>%
  dplyr::ungroup()
```


```{r remove_attributes_from_T1D_first_diagnosis}
for(col in names(T1D_first_diagnosis)){
  for(atr in c("label","format.spss","display_width")){
    attr(T1D_first_diagnosis[[col]], atr) <- NULL
  }
}
T1D_first_diagnosis <- as.data.frame(T1D_first_diagnosis)
```


```{r check_str_of_T1D_first_diagnosis}
str(T1D_first_diagnosis)
```

```{r include=FALSE}
# E10_and_sv_ov_npr <- E10 %>% 
#   dplyr::mutate(
#     t1d_same_diagnosis_in_ov_and_sv_npr = ifelse(
#       `ABISnr T1D(Cord Blood)` %in% complete_abis_diagnosis$ABISnr, TRUE, FALSE)
#     )
```

<!-- do not delete -->
```{r include=FALSE}
# writexl::write_xlsx(
#   E10_and_sv_ov_npr,
#   paste0(data_root,"vitamin_d/comparison_sv_ov_npr_vs_t1d_abis_indusmita.xlsx")
# )
```

```{r classify_samples_in_vitamin_D_data_by_diagnosis}
vitamin_d_case_control <- vitamin_d %>%
  dplyr::mutate(diagnosis = ifelse(
    !(
      ABISnr %in% c(T1D_abis_list, T2D_abis_list, psoriais_or_ibd_abis_list)
    ),
    "control",
    ifelse(
      ABISnr %in% T1D_abis_list &
        ABISnr %in% T2D_abis_list,
      "T1D and T2D",
      ifelse(
        ABISnr %in% T1D_abis_list &
          ABISnr %in% psoriais_or_ibd_abis_list,
        "IBD/PSORIASIS and T1D",
        ifelse(
          ABISnr %in% T2D_abis_list &
            ABISnr %in% psoriais_or_ibd_abis_list,
          "IBD/PSORIASIS and T2D",
          ifelse(
            ABISnr %in% T1D_abis_list_clean,
            "T1D",
            ifelse(
              ABISnr %in% T2D_abis_list_clean,
              "T2D",
              ifelse(
                ABISnr %in% psoriais_or_ibd_abis_list_clean,
                "IBD/PSORIASIS",
                NA_character_
              )
            )
          )
        )
      )
    )
  ))
```

convert2numeric
```{r function_to_convert_entries_with_symbols_to_numbers}
convert2numeric <- function(v) {
  if (grepl("<", v)) {
    v <- gsub("<", "", v)
    return(as.numeric(v) / sqrt(2))
  } else {
    return(as.numeric(v))
  }
}
```

```{r convert_entries_with_symbols_to_numbers}
vitamin_d_case_control <- vitamin_d_case_control %>% 
  dplyr::mutate(
    conc_25OHD = sapply(conc_25OHD, convert2numeric), 
    DBP = sapply(DBP, convert2numeric)
    )
```

add abis data to vitamin d data
```{r join_abis_data_to_vitamin_D_data}
vitamin_d_case_control_plus_abis <- dplyr::left_join(
  vitamin_d_case_control,
  abis %>% dplyr::mutate(ABISnr=as.character(ABISnr)),
  by = "ABISnr"
)
```


```{r extract_only_T1D_and_controls}
vitamin_d_case_control_plus_abis <- vitamin_d_case_control_plus_abis %>% 
    dplyr::filter(diagnosis %in% c("T1D","control"))
```


```{r remove_unnecessary_attributes}
for(col in c("DBP", "25OHD")){
  for(atr in c("names")){
    attr(vitamin_d_case_control_plus_abis[[col]], atr) <- NULL
  }
}
```

add first diagnosis dates
```{r add_first_diagnosis_dates}
vitamin_d_case_control_plus_abis <- dplyr::left_join(
  vitamin_d_case_control_plus_abis,
  T1D_first_diagnosis %>% 
    dplyr::mutate(ABISnr=as.character(ABISnr)),
  by = "ABISnr"
)
```

clear space
```{r clean_space}
objs <- ls(envir = .GlobalEnv)
obj_sizes <- sapply(objs, function(x) object.size(get(x, envir = .GlobalEnv)))
obj_sizes_sorted <- sort(obj_sizes, decreasing = TRUE)

# Print nicely
print(data.frame(
  Object = names(obj_sizes_sorted),
  Size_MB = round(obj_sizes_sorted / 1024^2, 2)
), row.names = FALSE)
```

save some space
```{r save_some_space}
rm(
  list = c(
    "T1D_first_diagnosis",
    "abis",
    "OV_NPR_ORIGINAL",
    "SV_NPR_ORIGINAL",
    "SV_NPR",
    "OV_NPR",
    "abis_info"
  ),
  envir = .GlobalEnv
)
```


```{r onset_age_group}
vitamin_d_case_control_plus_abis <- vitamin_d_case_control_plus_abis %>%
  dplyr::mutate(
    T1D_age_diagnosis = round(
      lubridate::time_length(difftime(INDATUMA,child_dob), "years"),
      2),
    onset_age_group = dplyr::case_when(
      T1D_age_diagnosis < 6 ~ "Early onset",
      T1D_age_diagnosis >= 6 &  T1D_age_diagnosis <= 12 ~  "School-age",
      T1D_age_diagnosis > 12 &  T1D_age_diagnosis < 16 ~  "Puberty-onset",
      T1D_age_diagnosis > 16 ~ "Young-adult",
      TRUE ~ "Healthy"
    )
  ) %>%
  dplyr::mutate(
    Birth_weight = Birth_weight/1000
  ) %>%
  dplyr::rename(
    Birth_height = Birth_length
  )
```

```{r}
birth_weight_quantiles <- quantile(vitamin_d_case_control_plus_abis$Birth_weight, na.rm=TRUE)

vitamin_d_case_control_plus_abis$Birth_weight_cat <- cut(
  vitamin_d_case_control_plus_abis$Birth_weight,
  breaks = c(
    birth_weight_quantiles[[1]],
    birth_weight_quantiles[[2]],
    birth_weight_quantiles[[3]],
    birth_weight_quantiles[[4]],
    Inf
  ),
  labels = c("q1", "q2", "q3", "q4"),
  include.lowest = TRUE
)
```

remove attributes
```{r remove_attributes_from_data}
for(col in c("conc_25OHD", "adjusted_25OHD")){
  attr(vitamin_d_case_control_plus_abis[[col]], "names") <- NULL  
}

vitamin_d_case_control_plus_abis <- as.data.frame(vitamin_d_case_control_plus_abis)
```

```{r set_reference_level_onset_age_group}
vitamin_d_case_control_plus_abis$onset_age_group <- factor(
  vitamin_d_case_control_plus_abis$onset_age_group, 
  levels = c("Early onset", "School-age", "Puberty-onset", "Young-adult")) 
```

```{r features2check}
features2check <- names(vitamin_d_case_control_plus_abis)[c(6,8:9,13:21)]
```

categorical features
```{r, categorical_features, message=FALSE, warning=FALSE}
for(m in seq_along(features2check)){
  
    v <- features2check[[m]]
    
    curr_count_raw <- vitamin_d_case_control_plus_abis %>%
      janitor::tabyl(!!sym(v), diagnosis) 
    
    # Transpose to make rows = group, columns = levels
    contingency <- as.table(as.matrix(curr_count_raw[, c("control", "T1D")]))
    rownames(contingency) <- curr_count_raw[[v]]
    contingency <- t(contingency)
    # chi-squared test
    result <- chisq.test(contingency)
    curr_count_raw["P.value"] <- NA_real_
    if(is.na(result$p.value)){
      curr_count_raw[["P.value"]][1] <- NA_real_
    } else {
          curr_count_raw[["P.value"]][1] <- if (result$p.value < 0.001) {
            formatC(result$p.value, format = "e", digits = 2)  
          } else {
              round(result$p.value,3)
            }
    }

    curr_count <- curr_count_raw %>% 
      dplyr::mutate(
        control_perc = round(control/sum(control)*100,2), 
        T1D_perc = round(T1D/sum(T1D)*100,2)) %>% 
      dplyr::mutate(
        control=paste0(control," (",control_perc,")"), 
        T1D = paste0(T1D," (",T1D_perc,")")
        ) %>%
      dplyr::select(-all_of(c("control_perc","T1D_perc")))
    
    names(curr_count) <- c("Feature", "control", "T1D","P.value")
    
    curr_table <- data.frame(
      Feature = character(), 
      control = character(), 
      T1D = character(), 
      P.value = double()
      )
    
    curr_table[1,] <- c(v,NA_character_,NA_character_,NA_real_)
    table2add <-rbind(curr_table, curr_count)
    if(m==1){
      final_table_cat <- table2add
    } else {
      final_table_cat <- rbind(final_table_cat,table2add)
    }
}
```

continuous features
```{r continuous_features}
features2check_numeric <- names(vitamin_d_case_control_plus_abis)[c(10:12)]
```

```{r, final_table_numeric, message=FALSE, warning=FALSE}
for(m in seq_along(features2check_numeric)){
  
    v <- features2check_numeric[[m]]
    
    curr_summary_raw <- vitamin_d_case_control_plus_abis %>% 
      dplyr::group_by(diagnosis) %>% 
      dplyr::summarise(
        mean = mean(!!sym(v), na.rm = TRUE), 
        sd = sd(!!sym(v), na.rm = TRUE)
        ) %>% 
      as.data.frame() %>%
      dplyr::mutate(
        stats = paste0(as.character(round(mean,2))," (",as.character(round(sd,2)),")")
      ) %>%
      dplyr::select(-all_of(c("mean","sd")))
    
    curr_summary <- data.frame(
      Feature=v,
      control=curr_summary_raw$stats[2],
      T1D=curr_summary_raw$stats[1]
      )
    
    # wilcox test
    formula2use = paste0(v," ~ diagnosis")
    df <- vitamin_d_case_control_plus_abis %>%
      dplyr::select(diagnosis, !!sym(v)) 
                    
    wilcox_result <- rstatix::wilcox_test(
      data = df,
      formula = as.formula(formula2use),
      p.adjust.method = "BH"
      )
    
    
    curr_summary["P.value"] <- NA_real_
    if(is.na(wilcox_result$p)){
      curr_summary[["P.value"]][1] <- NA_real_
    } else {
          curr_summary[["P.value"]][1] <- if (wilcox_result$p < 0.001) {
            formatC(wilcox_result$p, format = "e", digits = 2)  
          } else {
              round(wilcox_result$p,3)
            }
    }

    if(m==1){
      final_table_numeric <- curr_summary
    } else {
      final_table_numeric <- rbind(final_table_numeric,curr_summary)
    }
}
```

```{r final_table_cat}
final_table_cat <- final_table_cat %>%
  dplyr::mutate(Feature = ifelse(
    Feature %in% features2check,
    Feature,
    ifelse(is.na(Feature), "\tMissing", paste0("\t", Feature))
  )) %>%
  dplyr::mutate(Feature = gsub("_", " ", gsub(
    "_pregnancy", " (pregnancy)", gsub("_cat", "", 
                                        gsub("_dichotomous", "", Feature)
              )
          )
      )
  )
```


concatenate tables for both set of features
```{r final_table_cat_numeric}
final_table_cat_numeric <- rbind(
  data.frame(
    Feature = "Categorical variable count (%)",
    control = NA_character_,
    T1D = NA_character_,
    P.value = NA_real_
  ),
  final_table_cat,
  data.frame(
    Feature = "Continuous variable mean (sd)",
    control = NA_character_,
    T1D = NA_character_,
    P.value = NA_real_
  ),
  final_table_numeric
)  %>%
  dplyr::mutate(Feature = gsub(
    "Gestational_age",
    "Gestational age (weeks)",
    gsub(
      "Birth_height",
      "Birth height (cm)",
      gsub("Birth_weight", "Birth weight (kg)", Feature)
    )
  )) %>%
  flextable::flextable() %>%
  flextable::color(i = ~ Feature == "\tMissing",
                   j = "Feature",
                   color = "red") %>%
  flextable::color(
    i = ~ grepl("Continuous variable|Categorical variable", Feature),
    j = "Feature",
    color = "blue"
  ) %>%
  flextable::autofit()
```

```{r word_sect_properties}
# sect_properties <- prop_section(
#   page_size = page_size(
#     orient = "landscape",
#     width = 8.3,
#     height = 11.7
#   ),
#   type = "continuous",
#   page_margins = page_mar()
# )

sect_properties <- prop_section(
  page_size = page_size(
    orient = "landscape",
    width = 11.7,
    height = 8.3
  ),
  page_margins = page_mar(
    top = 0.5,
    bottom = 0.5,
    left = 0.5,
    right = 0.5,
    header = 0.3,
    footer = 0.3,
    gutter = 0
  ),
  type = "continuous"
)
```


```{r save_baseline_characteristics_table}
flextable::save_as_docx(
  `baseline characteristics table` = final_table_cat_numeric,
  path = paste0(output_dir,"/baseline_characteristics_table.docx"),
  pr_section = sect_properties
)
```

plot
```{r df_height_weight}
df_height_weight <- vitamin_d_case_control_plus_abis %>% 
  dplyr::select(diagnosis, Birth_weight, Birth_height) %>%
  dplyr::rename(
    `Height (cm) at Birth` = Birth_height,
    `Weight (Kg) at Birth` = Birth_weight
    )
```

```{r plot_df_height_weight}
suppressWarnings(plot_height_weight <- df_height_weight %>% 
  ggplot2::ggplot(
    aes(`Height (cm) at Birth`, 
        `Weight (Kg) at Birth`, 
        fill = diagnosis, 
        color = diagnosis
        )
    ) + 
  geom_point(aes(fill = diagnosis, color = diagnosis)) +
  theme(
    legend.title = element_blank(),
    legend.position = "bottom")
  )
suppressWarnings(
  legend <- ggpubr::get_legend(plot_height_weight)
)
```

```{r marginal_plot_height_weight}
suppressWarnings(
  marginal_plot_height_weight <- ggExtra::ggMarginal(
    plot_height_weight + theme(legend.position = "none"),
    groupColour =  TRUE,
    groupFill = TRUE,
    type = "density"
  )
)
```

```{r display_marginal_plot_height_weight}
marginal_plot_height_weight_final <- cowplot::plot_grid(
marginal_plot_height_weight,
legend,
ncol = 1,
rel_heights = c(1, 0.05)  
)

marginal_plot_height_weight_final
```

save box-plot
```{r save_marginal_plot_height_weight}
suppressMessages(
  ggplot2::ggsave(
    filename = paste0(output_dir, "/birth_height_weight_plot.jpeg"),
    plot = marginal_plot_height_weight_final,
    dpi = 300
  )
)
```

pair-wise plots
```{r pairwise_plot_numric}
pairwise_plot_numric <- GGally::ggpairs(
  vitamin_d_case_control_plus_abis, 
  columns = c("Gestational_age", "Birth_weight", "Birth_height"), 
  mapping = aes(color = diagnosis), 
  columnLabels = c("Gestational age (weeks)", 
                   "Birth weight(Kg)", 
                   "Birth height(cm)")
  )

pairwise_plot_numric
```

save pairwise-plot
```{r save_pairwise_plot_numric, warning=FALSE}
suppressWarnings(ggplot2::ggsave(
  filename = paste0(output_dir, "/numeric_features_pairwiseplot.jpeg"),
  plot = pairwise_plot_numric,
  dpi = 300
))
```

Box-violin plot
```{r features2check_numeric_box_plot}
features2check_numeric_box_plot <- vitamin_d_case_control_plus_abis %>%
    tidyr::pivot_longer(cols = c(Gestational_age, Birth_weight, Birth_height), 
                 names_to = "Variable", values_to = "Value") %>%
    dplyr::mutate(Variable = stringr::str_replace_all(Variable, "_", " ")) %>%
    ggplot(aes(x = diagnosis, y = Value, fill = diagnosis)) +
    geom_violin(trim = FALSE, alpha = 0.5) +
    geom_boxplot(width = 0.1, outlier.shape = NA) +
    xlab("Group") + 
    facet_wrap(~ Variable, scales = "free_y") +
    theme_minimal() +
    theme(legend.position = "none")

features2check_numeric_box_plot
```

save box-plot
```{r save_features2check_numeric_box_plot, warning=FALSE}
suppressWarnings(ggplot2::ggsave(
  filename = paste0(output_dir, "/numeric_features_boxplot.jpeg"),
  plot = features2check_numeric_box_plot,
  dpi = 300
))
```

day of year
```{r day_of_year}
vitamin_d_case_control_plus_abis$day_of_year <- lubridate::yday(
  vitamin_d_case_control_plus_abis$child_dob
  )
```


Estimate seasonal component to subtract
```{r estimate_seasonal_component_to_subtract}
df <- vitamin_d_case_control_plus_abis %>% 
  dplyr::select(day_of_year, conc_25OHD) %>% 
  dplyr::arrange(day_of_year) %>% 
  dplyr::group_by(day_of_year) %>% 
  dplyr::summarise(
    max_val=max(conc_25OHD, na.rm = TRUE),
    min_val=min(conc_25OHD, na.rm = TRUE),
    mean_val=mean(conc_25OHD, na.rm = TRUE),
    median_val=median(conc_25OHD, na.rm = TRUE)
    )
```


```{r seasonal_component_fit_max}
fit_max <- nls(
  max_val ~ mu + A * sin(2 * pi * day_of_year / 365.0),
  data = df,
  start = list(
    mu = mean(df$max_val, na.rm = TRUE),
    A = (max(df$max_val, na.rm = TRUE) - min(df$max_val, na.rm = TRUE)) / 2
  )
)

coef(fit_max)
```

```{r seasonal_component_fit_median}
fit_median <- nls(
  median_val ~ mu + A * sin(2 * pi * day_of_year / 365.0),
  data = df,
  start = list(
    mu = median(df$median_val, na.rm = TRUE),
    A = (max(df$median_val, na.rm = TRUE) - min(df$median_val, na.rm = TRUE)) / 2
  )
)

coef(fit_median)
```

```{r seasonal_component_fit_min}
fit_min <- nls(
  min_val ~ mu + A * sin(2 * pi * day_of_year / 365.0),
  data = df,
  start = list(
    mu = mean(df$min_val, na.rm = TRUE),
    A = (max(df$min_val, na.rm = TRUE) - min(df$min_val, na.rm = TRUE)) / 2
  )
)

coef(fit_min)
```

```{r seasonal_component_fit_mean}
fit_mean <- nls(
  mean_val ~ mu + A * sin(2 * pi * day_of_year / 365.0),
  data = df,
  start = list(
    mu = mean(df$mean_val, na.rm = TRUE),
    A = (max(df$mean_val, na.rm = TRUE) - min(df$mean_val, na.rm = TRUE)) / 2
  )
)

coef(fit_mean)
```

```{r seasonal_component_fit_data}
df$fit_mean <- predict(fit_mean)
df$fit_max <- predict(fit_max)
df$fit_min <- predict(fit_min)
df$fit_median <- predict(fit_median)

df <- df %>% tidyr::pivot_longer(
  c("fit_mean", "fit_median", "fit_max", "fit_min"),
  names_to = "fit_type", values_to = "fit_val"
)
```


```{r seasonal_component_fit_plot}
ggplot(df, aes(x = day_of_year)) +
  geom_point(aes(day_of_year, fit_val, color = fit_type)) +
  geom_line(aes(day_of_year, fit_val, color = fit_type)) +
  facet_wrap(~ fit_type, ncol = 1)
  # geom_point(aes(y = mean_val), color ="blue", shape = 0) +
  # geom_line(aes(y = fit_mean), color = "blue") +
  #   geom_point(aes(y = max_val), color ="red", shape = 6) +
  # geom_line(aes(y = fit_max), color = "red") +
  #   geom_point(aes(y = min_val), color ="black", shape = 12) +
  # geom_line(aes(y = fit_min), color = "black") +
  #   geom_point(aes(y = median_val), color ="green", shape = 18) +
  # geom_line(aes(y = fit_median), color = "green") +
  # labs(title = "Sine Fit to mean/ max/ min /median values by Day of Year")
```

```{r sampling_amplitude_of_seasonal_variation}
set.seed(123)  # Set the random seed to make results reproducible

A_i <- rnorm(n = nrow(vitamin_d_case_control_plus_abis), mean = 1.34, sd = 1)
```

<!-- mui = true 25(OH)D of subject i -->
<!-- mui_ti = mesured 25(OH)D of subject i on calendar day t_i -->
<!-- Ai = subject-specific amplitude  -->

```{r function_to_get_season_adjusted_25OHD}
season_adjusted_25OHD <- function(mui_ti,Ai,t_i){
  return(mui_ti - Ai*sin(2*pi*t_i/365))
}
```

```{r add_amplitude_of_variation_to_data_frame}
vitamin_d_case_control_plus_abis["Ai"] <- A_i
```

```{r season_adjusted_25OHD}
vitamin_d_case_control_plus_abis <- vitamin_d_case_control_plus_abis %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    adjusted_25OHD = season_adjusted_25OHD(conc_25OHD,Ai,day_of_year)
  ) %>%
  dplyr::ungroup()
```

```{r transformed_season_adjusted_25OHD}
x_25OHD <- vitamin_d_case_control_plus_abis$conc_25OHD
x_25OHD_log <- log(x_25OHD[x_25OHD>0])
normal_p_25OHD <- shapiro.test(x_25OHD)$p.value
lognormal_p_25OHD <- shapiro.test(x_25OHD_log)$p.value
```

```{r plot_season_adjusted_25OHD}
suppressWarnings(
  hist_25OHD <- ggplot(vitamin_d_case_control_plus_abis, aes(x = conc_25OHD)) +
  geom_histogram(
    aes(y = ..density..),
    bins = 100,
    fill = "lightblue",
    color = "black",
    alpha = 0.6
  ) +
  # Normal distribution envelope
  stat_function(
    fun = dnorm,
    args = list(
      mean = mean(vitamin_d_case_control_plus_abis$conc_25OHD, na.rm = TRUE),
      sd = sd(vitamin_d_case_control_plus_abis$conc_25OHD, na.rm = TRUE)
    ),
    color = "blue",
    size = 1.2,
    linetype = "dashed",
    show.legend = TRUE
  ) +
  
  # Log-normal distribution envelope
  stat_function(
    fun = dlnorm,
    args = list(
      meanlog = mean(log(
        vitamin_d_case_control_plus_abis$conc_25OHD[vitamin_d_case_control_plus_abis$conc_25OHD > 0]
      ), na.rm = TRUE),
      sdlog = sd(log(
        vitamin_d_case_control_plus_abis$conc_25OHD[vitamin_d_case_control_plus_abis$conc_25OHD > 0]
      ), na.rm = TRUE)
    ),
    color = "red",
    size = 1.2,
    linetype = "solid",
    show.legend = TRUE
  ) +
  labs(x = "25(OH)D", y = "Density") +  # No title
  theme_minimal()
)
```



```{r define_plot_variables}
x_pos <- Inf
y_pos <- Inf
hjust_val <- 1.05
vjust_val <- 1.5
```

```{r plot_transformed_data}
hist_25OHD <- hist_25OHD +
    annotate(
        "text",
        x = x_pos,
        y = y_pos,
        hjust = hjust_val,
        vjust = vjust_val,
        label = paste0(
            "Normal P-value = ",
            formatC(normal_p_25OHD, format = "e", digits = 2)
        ),
        color = "blue",
        size = 4
    ) +
    annotate(
        "text",
        x = x_pos,
        y = y_pos,
        hjust = hjust_val,
        vjust = vjust_val + 2.50,
        label = paste0(
           "Log-normal P-value = ",
            formatC(lognormal_p_25OHD, format = "e", digits = 2)
        ),
        color = "red",
        size = 4
    )  

hist_25OHD
```

```{r save_plot_transformed_data, warning=FALSE}
suppressWarnings(
  ggplot2::ggsave(
    filename = paste0(output_dir,"/distribution_of_25OHD.jpeg"),
    plot = hist_25OHD,
    dpi = 300
  )
  )
```


<!-- > summary(vitamin_d_case_control_plus_abis$adjusted_25OHD) -->
<!--    Min. 1st Qu.  Median    Mean 3rd Qu.    Max.  -->
<!--   4.364  26.143  36.981  40.605  52.009 116.037  -->
  
```{r transform_adjusted_data}
x_adj_25OHD <- vitamin_d_case_control_plus_abis$adjusted_25OHD
x_adj_25OHD_log <- log(x_adj_25OHD[x_adj_25OHD>0])
normal_p_adj_25OHD <- shapiro.test(x_adj_25OHD)$p.value
lognormal_p_adj_25OHD <- shapiro.test(x_adj_25OHD_log)$p.value
```

```{r plot_transformed_adjusted_data}
hist_adjusted_25OHD <- ggplot(vitamin_d_case_control_plus_abis, aes(x = adjusted_25OHD)) +
  geom_histogram(
    aes(y = ..density..),
    bins = 100,
    fill = "lightblue",
    color = "black",
    alpha = 0.6
  ) +
  # Normal distribution envelope
  stat_function(
    fun = dnorm,
    args = list(
      mean = mean(vitamin_d_case_control_plus_abis$adjusted_25OHD, na.rm = TRUE),
      sd = sd(vitamin_d_case_control_plus_abis$adjusted_25OHD, na.rm = TRUE)
    ),
    color = "blue",
    size = 1.2,
    linetype = "dashed"
  ) +
  
  # Log-normal distribution envelope
  stat_function(
    fun = dlnorm,
    args = list(
      meanlog = mean(log(
        vitamin_d_case_control_plus_abis$adjusted_25OHD[vitamin_d_case_control_plus_abis$adjusted_25OHD > 0]
      ), na.rm = TRUE),
      sdlog = sd(log(
        vitamin_d_case_control_plus_abis$adjusted_25OHD[vitamin_d_case_control_plus_abis$adjusted_25OHD > 0]
      ), na.rm = TRUE)
    ),
    color = "red",
    size = 1.2,
    linetype = "solid"
  ) +
  # Set color labels manually
  scale_color_manual(
    name = "Distribution",
    values = c("Normal" = "blue", "Log-normal" = "red")
  ) +

  labs(x = "25(OH)D", y = "Density") +  # No title
  theme_minimal() +
  theme(
    legend.position = "right"
  )
```


```{r plot_histogram_transformed_adjusted_data}
hist_adjusted_25OHD <- hist_adjusted_25OHD +
    annotate(
        "text",
        x = x_pos,
        y = y_pos,
        hjust = hjust_val,
        vjust = vjust_val,
        label = paste0(
            "Normal P-value = ",
            formatC(normal_p_adj_25OHD, format = "e", digits = 2)
        ),
        color = "blue",
        size = 4
    ) +
    annotate(
        "text",
        x = x_pos,
        y = y_pos,
        hjust = hjust_val,
        vjust = vjust_val + 2.50,
        label = paste0(
           "Log-normal P-value = ",
            formatC(lognormal_p_adj_25OHD, format = "e", digits = 2)
        ),
        color = "red",
        size = 4
    )  

hist_adjusted_25OHD
```


```{r save_plot_transformed_adjusted_data, warning=FALSE}
suppressWarnings(
  ggplot2::ggsave(
    filename = paste0(output_dir,"/distribution_of_adjusted_25OHD.jpeg"),
    plot = hist_adjusted_25OHD,
    dpi = 300
  )
)
```


convert dichotomous variables to 0/1
```{r cnvert_to_binary}
vitamin_d_case_control_plus_abis <- vitamin_d_case_control_plus_abis %>%  
  mutate(across(                                      
    .cols = all_of(c("diagnosis")),  ## for each column listed and "outcome"
    .fns = ~case_when(                              
      . %in% c("T1D")   ~ 1,  
      . %in% c("control") ~ 0,  
      TRUE ~ NA_real_)        
    )
  ) 
```

save vitamin_d_case_control_plus_abis for analysis
```{r}
write.table(
  as.data.frame(vitamin_d_case_control_plus_abis),
  file = paste0(
    data_root,
    "/vitamin_d/vitamin_d_case_control_plus_abis.csv"
  )
)
```


<!-- model baseline factors ~ diagnosis -->
<!-- ```{r} -->
<!-- univariate_models_25OHD <- complete_case_univariate_logistic_regression( -->
<!--   c(features2check, features2check_numeric), -->
<!--   vitamin_d_case_control_plus_abis) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- univariate_models_25OHD_flex <- univariate_models_25OHD %>% -->
<!--   as.data.frame() %>% dplyr::mutate(term = ifelse( -->
<!--     term %in% c("Season", unlist(all_parameters$category_variable), features2check_numeric), -->
<!--     gsub("_", " ", gsub( -->
<!--       "_pregnancy", " (pregnancy)", gsub("_cat", "", gsub( -->
<!--         "_dichotomous", -->
<!--         "", -->
<!--         gsub( -->
<!--           "Gestational_age", -->
<!--           "Gestational age (weeks)", -->
<!--           gsub( -->
<!--             "Birth_height", -->
<!--             "Birth height (cm)", -->
<!--             gsub("Birth_weight", "Birth weight (kg)", term) -->
<!--           ) -->
<!--         ) -->
<!--       )) -->
<!--     )), -->
<!--     paste0("\t", term) -->
<!--   )) %>% -->
<!--   dplyr::rename( -->
<!--     Term = term, -->
<!--     No = no, -->
<!--     Yes = yes, -->
<!--     OR = estimate, -->
<!--     `P-value` = p.value, -->
<!--     Conf.low = conf.low, -->
<!--     Conf.high = conf.high -->
<!--   ) %>% -->
<!--   dplyr::mutate(  -->
<!--     Term = if_else(Term == lag(Term) & row_number() > 1, "", Term) -->
<!--   ) %>% -->
<!--   flextable::flextable() %>% -->
<!--   flextable::autofit() %>% -->
<!--   flextable::color(i = ~ !grepl("\t", Term), ~ Term, color = "blue")  %>% -->
<!--   flextable::color(i = ~ `P-value` < 0.05, -->
<!--                    j = c(1, 4, 5), -->
<!--                    color = "red") -->

<!-- univariate_models_25OHD_flex -->
<!-- ``` -->

<!-- Fit models: Diagnosis  vitamin D -->

<!-- ```{r vitamin_d_models_summary, message=FALSE, warning=FALSE} -->
<!-- model_base <- glm(diagnosis ~ conc_25OHD, family = "binomial", data = vitamin_d_case_control_plus_abis) -->

<!-- model_maternal <- glm(diagnosis ~ conc_25OHD + -->
<!--                         First_degree_T1D_heredity_cat + -->
<!--                         Smoking_during_pregnancy + -->
<!--                         Vitamin_mineral_supplement_pregnancy + -->
<!--                         Fish_any_pregnancy + -->
<!--                         Parity + -->
<!--                         Delivery_mode, -->
<!--                       family = "binomial", data = vitamin_d_case_control_plus_abis) -->

<!-- model_child <- glm(diagnosis ~ conc_25OHD + -->
<!--                      Preterm_dichotomous + -->
<!--                      Season + -->
<!--                      Sex_cat, -->
<!--                    family = "binomial", data = vitamin_d_case_control_plus_abis) -->

<!-- model_full <- glm(diagnosis ~ conc_25OHD + -->
<!--                     First_degree_T1D_heredity_cat + -->
<!--                     Smoking_during_pregnancy + -->
<!--                     Vitamin_mineral_supplement_pregnancy + -->
<!--                     Fish_any_pregnancy + -->
<!--                     Parity + -->
<!--                     Delivery_mode + -->
<!--                     Preterm_dichotomous + -->
<!--                     Season + -->
<!--                     Sex_cat, -->
<!--                   family = "binomial", data = vitamin_d_case_control_plus_abis) -->
<!-- ``` -->

<!-- Fit interaction models -->
<!-- ```{r vitamin_d_interaction_models_summary, message=FALSE, warning=FALSE} -->
<!-- model_maternal_interaction <- glm(diagnosis ~ conc_25OHD *( -->
<!--                         First_degree_T1D_heredity_cat + -->
<!--                         Smoking_during_pregnancy + -->
<!--                         Vitamin_mineral_supplement_pregnancy + -->
<!--                         Fish_any_pregnancy + -->
<!--                         Parity + -->
<!--                         Delivery_mode), -->
<!--                       family = "binomial", data = vitamin_d_case_control_plus_abis) -->

<!-- model_child_interaction <- glm(diagnosis ~ conc_25OHD *( -->
<!--                      Preterm_dichotomous + -->
<!--                      Season + -->
<!--                      Sex_cat), -->
<!--                    family = "binomial", data = vitamin_d_case_control_plus_abis) -->

<!-- model_full_interaction <- glm(diagnosis ~ conc_25OHD *( -->
<!--                     First_degree_T1D_heredity_cat + -->
<!--                     Smoking_during_pregnancy + -->
<!--                     Vitamin_mineral_supplement_pregnancy + -->
<!--                     Fish_any_pregnancy + -->
<!--                     Parity + -->
<!--                     Delivery_mode + -->
<!--                     Preterm_dichotomous + -->
<!--                     Season + -->
<!--                     Sex_cat), -->
<!--                   family = "binomial", data = vitamin_d_case_control_plus_abis) -->
<!-- ``` -->

<!-- ```{r vitamin_d_models_summary_merge, message=FALSE, warning=FALSE} -->
<!-- # Create model list -->
<!-- model_list <- list(model_base, model_maternal, model_maternal_interaction, model_child, model_child_interaction, model_full, model_full_interaction) -->
<!-- model_names <- c("Base (25OHD only)", -->
<!--                  "Base + Maternal covariates", -->
<!--                  "Base * Maternal covariates (interaction model)", -->
<!--                  "Base + Child/Perinatal covariates", -->
<!--                  "Base * Child/Perinatal covariates (interaction model)", -->
<!--                  "Full model", -->
<!--                  "Full model (interaction model)") -->

<!-- # Extract OR, CI, R², and AIC -->
<!-- suppressMessages(summary_table <- lapply(seq_along(model_list), function(i) { -->
<!--   model <- model_list[[i]] -->
<!--   or_info <- tidy(model, conf.int = TRUE, exponentiate = TRUE) %>% -->
<!--     dplyr::filter(term == "conc_25OHD") %>% -->
<!--     dplyr::mutate(OR = estimate, CI = paste0("(", round(conf.low, 2), "–", round(conf.high, 2), ")")) %>% -->
<!--     dplyr::select(OR, CI) -->

<!--   r2_val <- pR2(model)["McFadden"] -->
<!--   aic_val <- AIC(model) -->

<!--   tibble( -->
<!--     Model = model_names[i], -->
<!--     `Odds Ratio (25OHD)` = round(or_info$OR, 2), -->
<!--     `95% CI (25OHD)` = or_info$CI, -->
<!--     `Pseudo-R² (McFadden)` = round(r2_val, 2), -->
<!--     AIC = round(aic_val, 1) -->
<!--   ) -->
<!-- }) %>% -->
<!--   bind_rows()) -->
<!-- ``` -->

<!-- Summary of Logistic Regression Models for 25(OH)D -->

<!-- ```{r vitamin_d_models_summary_display, message=FALSE, warning=FALSE} -->
<!-- # kable(summary_table, caption = "Summary of Logistic Regression Models for 25OHD") -->
<!-- summary_table %>% -->
<!--   as.data.frame() %>% -->
<!--   flextable::flextable() %>% -->
<!--   flextable::autofit()  -->
<!-- ``` -->

<!-- Prepare data for plotting -->
<!-- ```{r} -->
<!-- plot_data <- summary_table %>% -->
<!--   dplyr::mutate( -->
<!--     Model = factor(Model, levels = Model), # preserve order -->
<!--     OR = `Odds Ratio (25OHD)`, -->
<!--     R2 = `Pseudo-R² (McFadden)`, -->
<!--     # Extract lower and upper bounds from CI string "(x–y)" -->
<!--     CI_lower = as.numeric(sub(".*\\((.*)–.*", "\\1", `95% CI (25OHD)`)), -->
<!--     CI_upper = as.numeric(sub(".*–(.*)\\)", "\\1", `95% CI (25OHD)`)) -->
<!--   ) -->

<!-- plot_data %>% -->
<!--   as.data.frame() %>% -->
<!--   flextable::flextable() %>% -->
<!--   flextable::autofit()  -->
<!-- ``` -->


<!-- ```{r} -->
<!-- model_plot <- plot_data %>% -->
<!--   ggplot(aes(y = Model, x = OR, size = R2, label = Model)) + -->
<!--   geom_point(aes(color = Model, size = R2), alpha = 0.7) + -->
<!--   geom_pointrange(aes(xmin = CI_lower, xmax = CI_upper, color = Model),alpha = 0.7) + -->
<!--   geom_vline(xintercept = 1, linetype = "dashed", color = "red") + -->
<!--   labs( -->
<!--     x = "Odds Ratio (25OHD)", -->
<!--     y = "", -->
<!--     size = "R²" -->
<!--   )  + -->
<!--   scale_size_continuous( -->
<!--     breaks = c(0.02, 0.04, 0.06,0.08, 0.10, 0.140), -->
<!--     range = c(1, 2.2), -->
<!--     guide = guide_axis(check.overlap = TRUE) -->
<!--   ) + -->
<!--   xlim(c(0.70,1.10)) + -->
<!--   theme_minimal() + -->
<!--   theme( -->
<!--     axis.text.x = element_blank(),    -->
<!--     axis.text.y = element_text(hjust = 0), -->
<!--     legend.position = "none" -->
<!-- ) -->

<!-- model_plot  -->
<!-- ``` -->

<!-- save plots -->
<!-- ```{r warning=FALSE} -->
<!-- suppressWarnings( -->
<!--   ggsave( -->
<!--   filename = paste0(output_dir,"/model_OR_25OHD.jpeg"), -->
<!--   plot = model_plot, -->
<!--   dpi = 300, -->
<!--   width = 8.3, -->
<!--   height = 11.7, -->
<!--   units = "in" -->
<!--   ) -->
<!-- ) -->
<!-- ``` -->

<!-- group prediction probabilities -->
<!-- ```{r} -->
<!-- groups <- names(model_full_interaction$model)[-c(1, 2, 11)] -->

<!-- plot_group_probs <- vector("list", length = length(groups)) -->

<!-- for (iv in seq_along(groups)) { -->
<!--   v <- groups[[iv]] -->

<!--   pred <- ggpredict(model_full_interaction, -->
<!--                     terms = c("conc_25OHD [all]", v, "Sex_cat")) -->
<!--   color_text <- gsub("_", " ", gsub("_cat", "", gsub( -->
<!--     "_dichotomous", -->
<!--     "", -->
<!--     gsub( -->
<!--       "_during_pregnancy|_any_pregnancy", " (pregnancy)", -->
<!--            gsub("Vitamin_mineral_supplement_pregnancy", " Vitamin/mineral\n(pregnancy)", -->
<!--                  gsub("First_degree_T1D_heredity_cat", "T1D heredity",v) -->
<!--     )) -->
<!--   ))) -->
<!--   title_text <- paste0("Interaction: 25(OH)D × ", color_text, " × Sex") -->
<!--   # Plot manually -->
<!--   plot_group_probs[[iv]] <- ggplot(pred, aes(x = x, y = predicted, color = group)) + -->
<!--     geom_line(size = 1) + -->
<!--     facet_wrap(~ facet) +  # Will use v here -->
<!--     labs( -->
<!--       # x = "Centered 25OHD (nmol/L from mean)", -->
<!--       x = "25(OH)D (nmol/L)", -->
<!--       y = "Predicted probability of T1D", -->
<!--       color = color_text -->
<!--       # title = title_text -->
<!--     ) + -->
<!--     xlim(c(0, 100)) + -->
<!--     theme_minimal() -->
<!-- } -->
<!-- ``` -->

<!-- ```{r} -->
<!-- for(iv in seq_along(plot_group_probs)){ -->
<!--   print(plot_group_probs[[iv]]) -->
<!-- } -->
<!-- ``` -->

<!-- save plots -->
<!-- ```{r warning=FALSE} -->
<!-- cairo_pdf(paste0(output_dir,"/predicted_prob_full_interaction_model.pdf")) -->
<!-- for(iv in seq_along(plot_group_probs)){ -->
<!--   print(plot_group_probs[[iv]]) -->
<!-- } -->
<!-- dev.off() -->
<!-- ``` -->

<!-- DBP: Fit models -->
<!-- ```{r DBP_models, message=FALSE, warning=FALSE} -->
<!-- model_dbp_base <- glm(diagnosis ~ DBP, family = "binomial", data = vitamin_d_case_control_plus_abis) -->

<!-- model_dbp_maternal <- glm(diagnosis ~ DBP + -->
<!--                         First_degree_T1D_heredity_cat + -->
<!--                         Smoking_during_pregnancy + -->
<!--                         Vitamin_mineral_supplement_pregnancy + -->
<!--                         Fish_any_pregnancy + -->
<!--                         Parity + -->
<!--                         Delivery_mode, -->
<!--                       family = "binomial", data = vitamin_d_case_control_plus_abis) -->

<!-- model_dbp_child <- glm(diagnosis ~ DBP + -->
<!--                      Preterm_dichotomous + -->
<!--                      Season + -->
<!--                      Sex_cat, -->
<!--                    family = "binomial", data = vitamin_d_case_control_plus_abis) -->

<!-- model_dbp_full <- glm(diagnosis ~ DBP + -->
<!--                     First_degree_T1D_heredity_cat + -->
<!--                     Smoking_during_pregnancy + -->
<!--                     Vitamin_mineral_supplement_pregnancy + -->
<!--                     Fish_any_pregnancy + -->
<!--                     Parity + -->
<!--                     Delivery_mode + -->
<!--                     Preterm_dichotomous + -->
<!--                     Season + -->
<!--                     Sex_cat, -->
<!--                   family = "binomial", data = vitamin_d_case_control_plus_abis) -->
<!-- ``` -->

<!-- DBP : Interaction models -->
<!-- ```{r DBP_interaction_models_summary, message=FALSE, warning=FALSE} -->
<!-- model_dbp_maternal_interaction <- glm(diagnosis ~ DBP *( -->
<!--                         First_degree_T1D_heredity_cat + -->
<!--                         Smoking_during_pregnancy + -->
<!--                         Vitamin_mineral_supplement_pregnancy + -->
<!--                         Fish_any_pregnancy + -->
<!--                         Parity + -->
<!--                         Delivery_mode), -->
<!--                       family = "binomial", data = vitamin_d_case_control_plus_abis) -->

<!-- model_dbp_child_interaction <- glm(diagnosis ~ DBP *( -->
<!--                      Preterm_dichotomous + -->
<!--                      Season + -->
<!--                      Sex_cat), -->
<!--                    family = "binomial", data = vitamin_d_case_control_plus_abis) -->

<!-- model_dbp_full_interaction <- glm(diagnosis ~ DBP *( -->
<!--                     First_degree_T1D_heredity_cat + -->
<!--                     Smoking_during_pregnancy + -->
<!--                     Vitamin_mineral_supplement_pregnancy + -->
<!--                     Fish_any_pregnancy + -->
<!--                     Parity + -->
<!--                     Delivery_mode + -->
<!--                     Preterm_dichotomous + -->
<!--                     Season + -->
<!--                     Sex_cat), -->
<!--                   family = "binomial", data = vitamin_d_case_control_plus_abis) -->
<!-- ``` -->

<!-- DBP: Create model list -->
<!-- ```{r DBP_models_summary_merge, message=FALSE, warning=FALSE} -->
<!-- model_dbp_list <- list(model_dbp_base, model_dbp_maternal, model_dbp_maternal_interaction, model_dbp_child, model_dbp_child_interaction, model_dbp_full, model_dbp_full_interaction) -->
<!-- model_dbp_names <- c("Base (DBP only)", -->
<!--                  "Base + Maternal covariates", -->
<!--                  "Base * Maternal covariates (interaction model)", -->
<!--                  "Base + Child/Perinatal covariates", -->
<!--                  "Base * Child/Perinatal covariates (interaction model)", -->
<!--                  "Full model", -->
<!--                  "Full model (interaction model)") -->

<!-- # Extract OR, CI, R², and AIC -->
<!-- suppressWarnings( -->
<!--   summary_dbp_table <- lapply(seq_along(model_dbp_list), function(i) { -->
<!--   model <- model_dbp_list[[i]] -->
<!--   or_info <- tidy(model, conf.int = TRUE, exponentiate = TRUE) %>% -->
<!--     dplyr::filter(term == "DBP") %>% -->
<!--     dplyr::mutate( -->
<!--       OR = estimate, -->
<!--       CI = paste0("(", round(conf.low, 2), "–", round(conf.high, 2), ")") -->
<!--     ) %>% -->
<!--     dplyr::select(OR, CI) -->

<!--   r2_val <- pR2(model)["McFadden"] -->
<!--   aic_val <- AIC(model) -->

<!--   tibble( -->
<!--     Model = model_dbp_names[i], -->
<!--     `Odds Ratio (DBP)` = round(or_info$OR, 2), -->
<!--     `95% CI (DBP)` = or_info$CI, -->
<!--     `Pseudo-R² (McFadden)` = round(r2_val, 2), -->
<!--     AIC = round(aic_val, 1) -->
<!--   ) -->
<!-- }) %>% -->
<!--   bind_rows() -->
<!-- ) -->
<!-- ``` -->
<!-- Summary of Logistic Regression Models for DBP -->
<!-- ```{r dbp_models_summary_display, message=FALSE, warning=FALSE} -->
<!-- # kable(summary_dbp_table, caption = "Summary of Logistic Regression Models for DBP") -->
<!-- summary_dbp_table %>% -->
<!--   as.data.frame() %>% -->
<!--   flextable::flextable() %>% -->
<!--   flextable::autofit()  -->
<!-- ``` -->


<!-- Prepare data for plotting -->
<!-- ```{r} -->
<!-- plot_data_dbp <- summary_dbp_table %>% -->
<!--   dplyr::mutate( -->
<!--     Model = factor(Model, levels = Model), # preserve order -->
<!--     OR = `Odds Ratio (DBP)`, -->
<!--     R2 = `Pseudo-R² (McFadden)`, -->
<!--     # Extract lower and upper bounds from CI string "(x–y)" -->
<!--     CI_lower = as.numeric(sub(".*\\((.*)–.*", "\\1", `95% CI (DBP)`)), -->
<!--     CI_upper = as.numeric(sub(".*–(.*)\\)", "\\1", `95% CI (DBP)`)) -->
<!--   ) -->

<!-- plot_data_dbp  %>% -->
<!--   as.data.frame() %>% -->
<!--   flextable::flextable() %>% -->
<!--   flextable::autofit()  -->
<!-- ``` -->


<!-- ```{r} -->
<!-- model_dbp_plot <- plot_data_dbp %>% -->
<!--   ggplot(aes(y = Model, x = OR, size = R2, label = Model)) + -->
<!--   geom_point(aes(color = Model, size = R2), alpha = 0.7) + -->
<!--   geom_pointrange( -->
<!--     aes(xmin = CI_lower, xmax =  ifelse(CI_upper>3,OR+0.15,CI_upper),  -->
<!--         color = Model),alpha = 0.7) + -->
<!--   geom_vline(xintercept = 1, linetype = "dashed", color = "red") + -->
<!--   labs( -->
<!--     x = "Odds Ratio (DBP)", -->
<!--     y = "", -->
<!--     size = "R²" -->
<!--   )  + -->
<!--   scale_size_continuous( -->
<!--     breaks = c(0.02, 0.04, 0.06,0.08, 0.10, 0.120), -->
<!--     range = c(1, 2.2) #, -->
<!--     # guide = guide_axis(check.overlap = TRUE) -->
<!--   ) + -->
<!--   theme_minimal() + -->
<!--   theme( -->
<!--     axis.text.x = element_blank(),    -->
<!--     axis.text.y = element_text(hjust = 0), -->
<!--     legend.position = "none" -->
<!-- ) -->

<!-- model_dbp_plot  -->
<!-- ``` -->


<!-- save plots -->
<!-- ```{r} -->
<!-- ggsave( -->
<!--   filename = paste0(output_dir,"/model_OR_DBP.jpeg"), -->
<!--   plot = model_dbp_plot, -->
<!--   dpi = 300, -->
<!--   width = 8.3, -->
<!--   height = 11.7, -->
<!--   units = "in" -->
<!--   ) -->
<!-- ``` -->


<!-- group prediction probabilities DBP -->
<!-- ```{r} -->
<!-- groups <- names(model_dbp_full_interaction$model)[-c(1, 2, 11)] -->
<!-- print(groups) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- plot_group_probs_dbp <- vector("list", length = length(groups)) -->

<!-- for (iv in seq_along(groups)) { -->
<!--   v <- groups[[iv]] -->

<!--   pred_dbp <- ggpredict(model_dbp_full_interaction, -->
<!--                     terms = c("DBP [all]", v, "Sex_cat")) -->
<!--   color_text <- gsub("_", " ", gsub("_cat", "", gsub( -->
<!--     "_dichotomous", -->
<!--     "", -->
<!--     gsub( -->
<!--       "_during_pregnancy|_any_pregnancy", " (pregnancy)", -->
<!--            gsub("Vitamin_mineral_supplement_pregnancy", " Vitamin/mineral\n(pregnancy)", -->
<!--                  gsub("First_degree_T1D_heredity_cat", "T1D heredity",v) -->
<!--     )) -->
<!--   ))) -->
<!--   title_text <- paste0("Interaction: DBP × ", color_text, " × Sex") -->
<!--   # Plot manually -->
<!--   plot_group_probs_dbp[[iv]] <- ggplot(pred_dbp, aes(x = x, y = predicted, color = group)) + -->
<!--     geom_line(size = 1) + -->
<!--     facet_wrap(~ facet) +  # Will use v here -->
<!--     labs( -->
<!--       x = "DBP (\u03bcmol/L)", -->
<!--       y = "Predicted probability of T1D", -->
<!--       color = color_text -->
<!--       # title = title_text -->
<!--     ) + -->
<!--     xlim(c(0, 10)) + -->
<!--     theme_minimal() -->
<!-- } -->
<!-- ``` -->

<!-- ```{r} -->
<!-- for(iv in seq_along(plot_group_probs_dbp)){ -->
<!--   print(plot_group_probs_dbp[[iv]]) -->
<!-- } -->
<!-- ``` -->

<!-- save plots -->
<!-- ```{r} -->
<!-- cairo_pdf(paste0(output_dir,"/predicted_prob_full_interaction_model_DBP.pdf")) -->
<!-- for(iv in seq_along(plot_group_probs_dbp)){ -->
<!--   print(plot_group_probs_dbp[[iv]]) -->
<!-- } -->
<!-- dev.off() -->
<!-- ``` -->



<!-- glm diagnosis ~ conc_25OHD -->
<!-- ```{r} -->
<!-- fit_25OHD_no_covariate <- glm( -->
<!--     formula = diagnosis ~ conc_25OHD, -->
<!--     data = vitamin_d_case_control_plus_abis, -->
<!--     family = binomial -->
<!-- ) -->

<!-- results_25OHD_diagnosis_no_covariate <- fit_25OHD_no_covariate %>% -->
<!--   broom::tidy(exponentiate = TRUE, conf.int = TRUE) %>% -->
<!--   dplyr::mutate_if(is.numeric, function(x) -->
<!--     ifelse(x < 0.001, formatC(x, format = "e", 3), round(x, 3))) %>% -->
<!--   dplyr::filter(term != "(Intercept)")  %>% -->
<!--   dplyr::select(-c("std.error", "statistic")) %>% -->
<!--   dplyr::rename( -->
<!--     Term = term, -->
<!--     OR = estimate, -->
<!--     `P-value` = p.value, -->
<!--     Conf.low = conf.low, -->
<!--     Conf.high =conf.high -->
<!--     ) -->


<!-- results_25OHD_diagnosis_no_covariate  %>% -->
<!--   as.data.frame() %>% -->
<!--   flextable::flextable() %>% -->
<!--   flextable::autofit()  -->
<!-- ``` -->

<!-- conditional regression - case-control matching   -->


<!-- Assessing Balance: A. Check Initial Imbalance -->

<!-- ```{r} -->
<!-- child_covariates <- c("Preterm_dichotomous","Birth_weight", "Season", "Sex_cat") -->

<!-- maternal_covariates <- c( -->
<!--   "First_degree_T1D_heredity_cat", -->
<!--   "Smoking_during_pregnancy", -->
<!--   # "Vitamin_mineral_supplement_pregnancy", -->
<!--   "Fish_any_pregnancy", -->
<!--   "Parity", -->
<!--   "Delivery_mode" -->
<!-- ) -->
<!-- ``` -->


<!-- all covariates -->
<!-- ```{r} -->
<!-- df_all <- vitamin_d_case_control_plus_abis %>% -->
<!--   dplyr::select( -->
<!--     all_of(c("ABISnr", "diagnosis", "conc_25OHD", child_covariates, maternal_covariates))) %>% -->
<!--   tidyr::drop_na() -->
<!-- ``` -->

<!-- ```{r} -->
<!-- all_covariates_formula <- paste0(c(child_covariates, maternal_covariates),collapse = " + ") -->
<!-- all_covariates_formula  -->
<!-- ``` -->


<!-- ```{r echo=FALSE} -->
<!-- w_25OHD_all <- weightit( -->
<!--   as.formula(paste0("conc_25OHD ~  ", all_covariates_formula)), -->
<!--   data = df_all, -->
<!--   method = "ebal" -->
<!-- ) -->
<!-- ``` -->


<!-- ```{r echo=FALSE} -->
<!-- suppressMessages( -->
<!--   fit_25OHD_all <- glm_weightit( -->
<!--   as.formula(paste0("diagnosis ~ conc_25OHD * (",all_covariates_formula,")")), -->
<!--   data = df_all, -->
<!--   weightit = w_25OHD_all, -->
<!--   family = binomial -->
<!-- ) -->
<!-- ) -->
<!-- ``` -->


<!-- ```{r echo=FALSE} -->
<!-- avg_25OHD_all <- avg_comparisons( -->
<!--   fit_25OHD_all, -->
<!--   type = "link", -->
<!--   comparison = "lnor", -->
<!--   transform = "exp" -->
<!-- ) %>% -->
<!--   broom::tidy() %>% -->
<!--   as.data.frame() %>% -->
<!--   dplyr::mutate( -->
<!--     term = ifelse(row_number()> 1 & term == lag(term), -->
<!--                     NA_character_, -->
<!--                     gsub("_"," ",gsub("_cat","",gsub("_dichotomous","",gsub("Birth_weight_cat","Birth weight category",term)))) -->
<!--                     ) -->
<!--     ) %>% -->
<!--   dplyr::mutate_if(is.numeric, function(x) ifelse(x<0.001, formatC(x, format = "e", 3), round(x,3))) %>% -->
<!--   dplyr::rename( -->
<!--     Term = term, -->
<!--     Contrast = contrast, -->
<!--     OR = estimate, -->
<!--     `P-value` = p.value -->
<!--     ) %>% -->
<!--   dplyr::mutate( -->
<!--         `OR (95% CI)` = paste0(as.character(OR)," (",as.character(conf.low),"-",as.character(conf.high),")") -->
<!--   ) %>% -->
<!--   dplyr::select(all_of(c("Term", "Contrast","OR (95% CI)", "P-value")))  -->


<!-- first_row <- avg_25OHD_all %>% dplyr::filter(Term == "conc 25OHD") -->

<!-- other_rows <- avg_25OHD_all %>% dplyr::filter(Term != "conc 25OHD" | is.na(Term)) -->

<!-- avg_25OHD_all <- rbind(first_row, other_rows) -->
<!-- ``` -->

<!-- Diagnosis ~ conc. 25(OH)D * (all covariates) : weighted data -->
<!-- ```{r echo=FALSE} -->
<!-- # kable(avg_25OHD_all, caption = "Diagnosis ~ conc. 25(OH)D * (all covariates) : weighted data") -->
<!-- avg_25OHD_all %>% -->
<!--   as.data.frame() %>% -->
<!--   flextable::flextable() %>% -->
<!--   flextable::autofit()  -->
<!-- ``` -->

<!-- ```{r} -->
<!-- birth_weight_quantiles_all <- quantile(df_all$Birth_weight) -->

<!-- df_all$Birth_weight_cat <- cut( -->
<!--   df_all$Birth_weight, -->
<!--   breaks = c( -->
<!--     birth_weight_quantiles_all[[1]], -->
<!--     birth_weight_quantiles_all[[2]], -->
<!--     birth_weight_quantiles_all[[3]], -->
<!--     birth_weight_quantiles_all[[4]], -->
<!--     Inf -->
<!--   ), -->
<!--   labels = c("q1", "q2", "q3", "q4"), -->
<!--   include.lowest = TRUE -->
<!-- ) -->

<!-- df_all_all_cat <- df_all %>% -->
<!--   dplyr::select(-Birth_weight) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- all_covariates_cat_formula <- paste0(names(df_all_all_cat)[-c(1:3)],collapse = " + ") -->
<!-- all_covariates_cat_formula -->
<!-- ``` -->


<!-- ```{r echo=FALSE} -->
<!-- w_25OHD_all_cat <- weightit( -->
<!--   as.formula(paste0("conc_25OHD ~  ", all_covariates_cat_formula)), -->
<!--   data = df_all_all_cat, -->
<!--   method = "ebal" -->
<!-- ) -->

<!-- fit_25OHD_all_cat <- glm_weightit( -->
<!--   as.formula(paste0("diagnosis ~ conc_25OHD * (",all_covariates_cat_formula,")")), -->
<!--   data = df_all_all_cat, -->
<!--   weightit = w_25OHD_all_cat, -->
<!--   family = binomial -->
<!-- ) -->

<!-- avg_25OHD_all_cat <- avg_comparisons( -->
<!--   fit_25OHD_all_cat, -->
<!--   type = "link", -->
<!--   comparison = "lnor", -->
<!--   transform = "exp" -->
<!-- ) %>% -->
<!--   broom::tidy() %>% -->
<!--   as.data.frame() %>% -->
<!--   dplyr::mutate( -->
<!--     term = ifelse(row_number()> 1 & term == lag(term), -->
<!--                     NA_character_, -->
<!--                     gsub("_"," ",gsub("_cat","",gsub("_dichotomous","",gsub("Birth_weight_cat","Birth weight category",term)))) -->
<!--                     ) -->
<!--     ) %>% -->
<!--   dplyr::mutate_if(is.numeric, function(x) ifelse(x<0.001, formatC(x, format = "e", 3), round(x,3))) %>% -->
<!--   dplyr::rename( -->
<!--     Term = term, -->
<!--     Contrast = contrast, -->
<!--     OR = estimate, -->
<!--     `P-value` = p.value -->
<!--     ) %>% -->
<!--   dplyr::mutate( -->
<!--         `OR (95% CI)` = paste0(as.character(OR)," (",as.character(conf.low),"-",as.character(conf.high),")") -->
<!--   ) %>% -->
<!--   dplyr::select(all_of(c("Term", "Contrast","OR (95% CI)", "P-value")))  -->

<!-- first_row <- avg_25OHD_all_cat %>% dplyr::filter(Term == "conc 25OHD") -->
<!-- other_rows <- avg_25OHD_all_cat %>% dplyr::filter(Term != "conc 25OHD" | is.na(Term)) -->
<!-- avg_25OHD_all_cat <- rbind(first_row, other_rows) -->
<!-- ``` -->

<!-- Diagnosis ~ conc. 25(OH)D * (all categorical covariates) : weighted data -->
<!-- ```{r echo=FALSE} -->
<!-- # kable(avg_25OHD_all_cat, caption = "Diagnosis ~ conc. 25(OH)D * (all categorical covariates) : weighted data") -->
<!-- avg_25OHD_all_cat %>% -->
<!--   as.data.frame() %>% -->
<!--   flextable::flextable() %>% -->
<!--   flextable::autofit() %>% -->
<!--   flextable::color( -->
<!--     i = ~ `P-value` < 0.05, -->
<!--     j = c("Contrast", "P-value"), -->
<!--     color = "red") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_child <- vitamin_d_case_control_plus_abis %>% -->
<!--   dplyr::select(ABISnr, diagnosis,conc_25OHD,Preterm_dichotomous,Birth_weight,Season,Sex_cat) %>% -->
<!--   tidyr::drop_na() -->
<!-- ``` -->

<!-- ```{r} -->
<!-- fit_25OHD_child_no_match_no_interaction <- glm_weightit( -->
<!--     diagnosis ~ conc_25OHD,  -->
<!--     data = df_child, -->
<!--     family = binomial -->
<!--   ) -->

<!-- avg_25OHD_child_no_match_no_interaction <- avg_comparisons( -->
<!--   fit_25OHD_child_no_match_no_interaction, -->
<!--   type = "link", -->
<!--   comparison = "lnor", -->
<!--   transform = "exp" -->
<!-- ) %>% -->
<!--   broom::tidy() %>% -->
<!--   dplyr::select(term, estimate, p.value, conf.low, conf.high) %>% -->
<!--   dplyr::mutate(term = ifelse( -->
<!--     row_number() > 1 & term == lag(term), -->
<!--     NA_character_, -->
<!--     gsub("_", " ", gsub( -->
<!--       "_cat", "", gsub( -->
<!--         "_dichotomous", -->
<!--         "", gsub("Birth_weight_cat", "Birth weight category", term))))) -->
<!--     ) %>% -->
<!--   dplyr::mutate_if(is.numeric, function(x) -->
<!--     ifelse(x < 0.001, formatC(x, format = "e", 3), round(x, 3))) %>% -->
<!--   dplyr::rename( -->
<!--     Term = term, -->
<!--     OR = estimate, -->
<!--     `P-value` = p.value -->
<!--   ) %>% -->
<!--   dplyr::mutate( -->
<!--         `OR (95% CI)` = paste0(as.character(OR)," (",as.character(conf.low),"-",as.character(conf.high),")") -->
<!--   ) %>% -->
<!--   dplyr::select(-all_of(c("OR","conf.low", "conf.high"))) %>% -->
<!--   dplyr::select(all_of(c("Term","OR (95% CI)", "P-value"))) -->
<!-- ``` -->

<!-- Diagnosis ~ conc. 25(OH)D * (child covariates) : not weighted nointeraction -->
<!-- ```{r} -->
<!-- # kable( -->
<!-- #   avg_25OHD_child_no_match_no_interaction,  -->
<!-- #   caption = "Diagnosis ~ conc. 25(OH)D * (child covariates) : not weighted nointeraction") -->

<!-- avg_25OHD_child_no_match_no_interaction %>% -->
<!--   as.data.frame() %>% -->
<!--   flextable::flextable() %>% -->
<!--   flextable::autofit() -->
<!-- ``` -->

<!-- glm diagnosis ~ conc_25OHD; case-control matched by child (birth) covariates -->
<!-- ```{r} -->
<!-- w_25OHD_child <- weightit( -->
<!--     conc_25OHD ~ Preterm_dichotomous + Season + Sex_cat + Birth_weight, -->
<!--     data = df_child, -->
<!--     method = "ebal" -->
<!--   ) -->

<!-- fit_25OHD_child_cov_matched_no_interaction <- glm_weightit( -->
<!--     diagnosis ~ conc_25OHD,  -->
<!--     data = df_child, -->
<!--     weightit = w_25OHD_child, -->
<!--     family = binomial -->
<!--   ) -->

<!-- avg_25OHD_child_cov_matched_no_interaction <- avg_comparisons( -->
<!--   fit_25OHD_child_cov_matched_no_interaction, -->
<!--   type = "link", -->
<!--   comparison = "lnor", -->
<!--   transform = "exp" -->
<!-- ) %>% -->
<!--   broom::tidy() %>% -->
<!--   dplyr::select(term, estimate, p.value, conf.low, conf.high) %>% -->
<!--   dplyr::mutate(term = ifelse( -->
<!--     row_number() > 1 & term == lag(term), -->
<!--     NA_character_, -->
<!--     gsub("_", " ", gsub( -->
<!--       "_cat", "", gsub( -->
<!--         "_dichotomous", -->
<!--         "", gsub("Birth_weight_cat", "Birth weight category", term))))) -->
<!--     ) %>%   -->
<!--   dplyr::mutate_if(is.numeric, function(x) -->
<!--     ifelse(x < 0.001, formatC(x, format = "e", 3), round(x, 3))) %>% -->
<!--   dplyr::rename( -->
<!--     Term = term, -->
<!--     OR = estimate, -->
<!--     `P-value` = p.value -->
<!--   ) %>% -->
<!--   dplyr::mutate( -->
<!--         `OR (95% CI)` = paste0(as.character(OR)," (",as.character(conf.low),"-",as.character(conf.high),")") -->
<!--   ) %>% -->
<!--   dplyr::select(-all_of(c("OR","conf.low", "conf.high"))) %>% -->
<!--   dplyr::select(all_of(c("Term","OR (95% CI)", "P-value"))) -->
<!-- ``` -->

<!-- Diagnosis ~ conc. 25(OH)D * (child covariates) : weighted but no interaction -->
<!-- ```{r} -->
<!-- # kable( -->
<!-- #   avg_25OHD_child_cov_matched_no_interaction,  -->
<!-- #   caption = "Diagnosis ~ conc. 25(OH)D * (child covariates) : weighted but no interaction") -->

<!-- avg_25OHD_child_cov_matched_no_interaction %>% -->
<!--   as.data.frame() %>% -->
<!--   flextable::flextable() %>% -->
<!--   flextable::autofit() -->
<!-- ``` -->


<!-- ```{r} -->
<!-- fit_25OHD_child_cov_matched_and_with_interaction <- glm_weightit( -->
<!--     diagnosis ~ conc_25OHD*(Preterm_dichotomous + Season + Sex_cat + Birth_weight), -->
<!--     data = df_child, -->
<!--     weightit = w_25OHD_child, -->
<!--     family = binomial -->
<!--   ) -->

<!-- avg_25OHD_child_cov_matched_and_with_interaction <- avg_comparisons( -->
<!--   fit_25OHD_child_cov_matched_and_with_interaction, -->
<!--   type = "link", -->
<!--   comparison = "lnor", -->
<!--   transform = "exp" -->
<!-- ) %>% -->
<!--   broom::tidy() %>% -->
<!--   dplyr::select(term, contrast, estimate, p.value, conf.low, conf.high) %>% -->
<!--   dplyr::mutate(term = ifelse( -->
<!--     row_number() > 1 & term == lag(term), -->
<!--     NA_character_, -->
<!--     gsub("_", " ", gsub( -->
<!--       "_cat", "", gsub( -->
<!--         "_dichotomous", -->
<!--         "", gsub("Birth_weight_cat", "Birth weight category", term))))) -->
<!--     ) %>%   -->
<!--   dplyr::mutate_if(is.numeric, function(x) -->
<!--     ifelse(x < 0.001, formatC(x, format = "e", 3), round(x, 3))) %>% -->
<!--   dplyr::rename( -->
<!--     Term = term, -->
<!--     Contrast = contrast, -->
<!--     OR = estimate, -->
<!--     `P-value` = p.value -->
<!--   ) %>% -->
<!--   dplyr::mutate( -->
<!--         `OR (95% CI)` = paste0(as.character(OR)," (",as.character(conf.low),"-",as.character(conf.high),")") -->
<!--   ) %>% -->
<!--   dplyr::select(-all_of(c("OR","conf.low", "conf.high"))) %>% -->
<!--   dplyr::select(all_of(c("Term", "Contrast","OR (95% CI)", "P-value"))) %>% -->
<!--   dplyr::mutate( -->
<!--     Term = ifelse( -->
<!--       row_number() > 1 & lag(Term) == Term, -->
<!--       NA_character_, -->
<!--       gsub("_"," ",gsub("_cat","",gsub("_dichotomous","",Term)))) -->
<!--     ) -->

<!-- first_row <- avg_25OHD_child_cov_matched_and_with_interaction %>%  -->
<!--   dplyr::filter(Term == "conc 25OHD") -->
<!-- other_rows <- avg_25OHD_child_cov_matched_and_with_interaction %>%  -->
<!--   dplyr::filter(Term != "conc 25OHD" | is.na(Term)) -->
<!-- avg_25OHD_child_cov_matched_and_with_interaction <- rbind(first_row, other_rows) -->
<!-- ``` -->

<!-- Diagnosis ~ conc. 25(OH)D * (child covariates) : weighted and with interaction -->
<!-- ```{r} -->
<!-- # kable( -->
<!-- #   avg_25OHD_child_cov_matched_and_with_interaction,  -->
<!-- #   caption = "Diagnosis ~ conc. 25(OH)D * (child covariates) : weighted and with interaction") -->

<!-- avg_25OHD_child_cov_matched_and_with_interaction %>% -->
<!--   as.data.frame() %>% -->
<!--   flextable::flextable() %>% -->
<!--   flextable::autofit() %>% -->
<!--   flextable::color( -->
<!--     i = ~ `P-value` < 0.05, -->
<!--     j = c("Contrast", "P-value"), -->
<!--     color = "red") -->
<!-- ``` -->



<!-- ```{r} -->
<!-- birth_weight_quantiles <- quantile(df_child$Birth_weight) -->

<!-- df_child$Birth_weight_cat <- cut( -->
<!--   df_child$Birth_weight, -->
<!--   breaks = c(birth_weight_quantiles[[1]],birth_weight_quantiles[[2]],birth_weight_quantiles[[3]],birth_weight_quantiles[[4]], Inf), -->
<!--   labels = c("q1", "q2", "q3", "q4"), -->
<!--   include.lowest = TRUE -->
<!-- ) -->

<!-- df_child_all_cat <- df_child %>% -->
<!--   dplyr::select(-Birth_weight) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- results_list_child <- lapply(c("Season", "Sex_cat","Birth_weight_cat", "Preterm_dichotomous"), function(stratum) { -->
<!--   w <- weightit( -->
<!--     conc_25OHD ~ Preterm_dichotomous + Season + Sex_cat + Birth_weight_cat, -->
<!--     data = df_child, -->
<!--     by = as.formula(paste0("~", stratum)), -->
<!--     method = "ebal" -->
<!--   ) -->
<!--   fit <- glm_weightit( -->
<!--     diagnosis ~ conc_25OHD*(Preterm_dichotomous + Season + Sex_cat + Birth_weight_cat), -->
<!--     data = df_child, -->
<!--     weightit = w, -->
<!--     family = binomial -->
<!--   ) -->
<!--   avg <- avg_comparisons( -->
<!--     fit, -->
<!--     variables = "conc_25OHD", -->
<!--     by = stratum, -->
<!--     type = "link", -->
<!--     comparison = "lnor", -->
<!--     transform = "exp" -->
<!--   ) %>% -->
<!--     broom::tidy() %>% -->
<!--     dplyr::rename(level = !!stratum) %>%          -->
<!--     dplyr::mutate(strata = stratum) %>%           -->
<!--     dplyr::select(strata, level, estimate, p.value, conf.low, conf.high)     -->
<!--   return(list(avg,w)) -->
<!-- }) -->
<!-- names(results_list_child) <- c("Season", "Sex_cat", "Birth_weight_cat", "Preterm_dichotomous") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- cowplot::plot_grid( -->
<!--     love.plot(results_list_child[[1]][[2]], title = ""), -->
<!--     love.plot(results_list_child[[2]][[2]], title = ""), -->
<!--     love.plot(results_list_child[[3]][[2]], title = ""), -->
<!--     love.plot(results_list_child[[4]][[2]], title = ""), -->
<!--     labels = c("A. Season",  -->
<!--                "B. Sex",  -->
<!--                "C. Birth weight category", -->
<!--                "D. Preterm-birth"), -->
<!--     nrow = 2, -->
<!--     ncol = 2 -->
<!--     ) -->
<!-- ``` -->

<!-- save stratum stratified weight matched  -->
<!-- ```{r} -->
<!-- ggsave( -->
<!--   filename = paste0(output_dir, "/conc_25OHD_child_covariate_correlations_entropy_balanced.jpeg"), -->
<!--   plot = cowplot::plot_grid( -->
<!--     love.plot(results_list_child[[1]][[2]], title = ""), -->
<!--     love.plot(results_list_child[[2]][[2]], title = ""), -->
<!--     love.plot(results_list_child[[3]][[2]], title = ""), -->
<!--     love.plot(results_list_child[[4]][[2]], title = ""), -->
<!--     labels = c("A. Season",  -->
<!--                "B. Sex",  -->
<!--                "C. Birth weight category", -->
<!--                "D. Preterm-birth") -->
<!--     ), -->
<!--   dpi = 300, -->
<!--   units = "in", -->
<!--   width = 10.3, -->
<!--   height = 14.7 -->
<!-- ) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- for(ii in seq_along(names(results_list_child))) { -->
<!--   if (ii == 1) { -->
<!--     child_25OHD_diagnosis_results <- results_list_child[[ii]][[1]] -->
<!--   } else { -->
<!--     child_25OHD_diagnosis_results <- rbind(child_25OHD_diagnosis_results, results_list_child[[ii]][[1]]) -->
<!--   } -->
<!-- } -->
<!-- ``` -->


<!-- ```{r} -->
<!-- child_25OHD_diagnosis_results <- child_25OHD_diagnosis_results %>% -->
<!--   dplyr::mutate( -->
<!--     strata = ifelse(row_number()> 1 & strata == lag(strata), -->
<!--                     NA_character_, -->
<!--                     gsub("_"," ",gsub("_cat","",gsub("_dichotomous","",gsub("Birth_weight_cat","Birth weight category",strata)))) -->
<!--                     ) -->
<!--     ) %>% -->
<!--   dplyr::mutate_if(is.numeric, function(x) ifelse(x<0.001, formatC(x, format = "e", 3), round(x,3))) %>% -->
<!--   dplyr::rename( -->
<!--     Strata = strata, -->
<!--     Level = level, -->
<!--     OR = estimate, -->
<!--     `P-value` = p.value -->
<!--     ) %>% -->
<!--   dplyr::mutate( -->
<!--         `OR (95% CI)` = paste0(as.character(OR)," (",as.character(conf.low),"-",as.character(conf.high),")") -->
<!--   ) %>% -->
<!--   dplyr::select(-all_of(c("OR","conf.low", "conf.high"))) %>% -->
<!--   dplyr::select(all_of(c("Strata", "Level","OR (95% CI)", "P-value")))  -->

<!-- child_25OHD_diagnosis_results %>% -->
<!--   as.data.frame() %>% -->
<!--   flextable::flextable() %>% -->
<!--   flextable::autofit() %>% -->
<!--   flextable::color( -->
<!--     i = ~ `P-value` < 0.05, -->
<!--     j = c("Level", "P-value"), -->
<!--     color = "red") -->
<!-- ``` -->

<!-- ```{r} -->
<!-- maternal_covariates <- c( -->
<!--   "First_degree_T1D_heredity_cat", -->
<!--   "Smoking_during_pregnancy", -->
<!--   # "Vitamin_mineral_supplement_pregnancy", -->
<!--   "Fish_any_pregnancy", -->
<!--   "Parity", -->
<!--   "Delivery_mode" -->
<!-- ) -->
<!-- ``` -->

<!-- ```{r} -->
<!-- df_maternal <- vitamin_d_case_control_plus_abis %>% -->
<!--   dplyr::select( -->
<!--     all_of(c("ABISnr", "diagnosis", "conc_25OHD", maternal_covariates))) %>% -->
<!--   tidyr::drop_na() -->
<!-- ``` -->

<!-- no stratification -->
<!-- ```{r echo=FALSE} -->
<!-- w_25OHD_maternal <- weightit( -->
<!--   conc_25OHD ~  First_degree_T1D_heredity_cat + Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode, -->
<!--   data = df_maternal, -->
<!--   method = "ebal" -->
<!-- ) -->


<!-- fit_25OHD_maternal <- glm_weightit( -->
<!--   diagnosis ~ conc_25OHD * ( -->
<!--     First_degree_T1D_heredity_cat + Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode -->
<!--   ), -->
<!--   data = df_maternal, -->
<!--   weightit = w_25OHD_maternal, -->
<!--   family = binomial -->
<!-- )   -->

<!-- avg_25OHD_maternal <- avg_comparisons( -->
<!--   fit_25OHD_maternal, -->
<!--   type = "link", -->
<!--   comparison = "lnor", -->
<!--   transform = "exp" -->
<!-- ) %>% -->
<!--   broom::tidy() %>% -->
<!--   as.data.frame() %>% -->
<!--   dplyr::mutate(term = ifelse( -->
<!--     row_number() > 1 & term == lag(term), -->
<!--     NA_character_, -->
<!--     gsub("_", " ", gsub( -->
<!--       "_cat", "", gsub( -->
<!--         "_dichotomous", -->
<!--         "", -->
<!--         gsub("Birth_weight_cat", "Birth weight category", term) -->
<!--       ) -->
<!--     )) -->
<!--   )) %>%  -->
<!--   dplyr::mutate_if(is.numeric, function(x) ifelse(x<0.001, formatC(x, format = "e", 3), round(x,3))) %>% -->
<!--   dplyr::rename( -->
<!--     Term = term, -->
<!--     Contrast = contrast, -->
<!--     OR = estimate, -->
<!--     `P-value` = p.value -->
<!--     ) %>% -->
<!--   dplyr::mutate( -->
<!--         `OR (95% CI)` = paste0(as.character(OR)," (",as.character(conf.low),"-",as.character(conf.high),")") -->
<!--   ) %>% -->
<!--   dplyr::select(all_of(c("Term", "Contrast","OR (95% CI)", "P-value")))  -->

<!-- avg_25OHD_maternal %>% -->
<!--   as.data.frame() %>% -->
<!--   flextable::flextable() %>% -->
<!--   flextable::autofit() %>% -->
<!--   flextable::color( -->
<!--     i = ~ `P-value` < 0.05, -->
<!--     j = c("Contrast", "P-value"), -->
<!--     color = "red") -->
<!-- ``` -->

<!-- diagnosis ~ conc_25OHD with stratified maternal covariates -->
<!-- ```{r warning=FALSE} -->
<!-- results_list_maternal <- lapply(maternal_covariates[c(2:3)], function(stratum) { -->

<!--     curr_strata <- gsub("_", " ", gsub( -->
<!--     "_during_pregnancy", -->
<!--     " (pregnancy) ", -->
<!--     gsub("_any_pregnancy", " any (pregnancy) ", stratum) -->
<!--     )) -->

<!--     w <- weightit( -->
<!--     conc_25OHD ~ First_degree_T1D_heredity_cat + Smoking_during_pregnancy +  -->
<!--     Fish_any_pregnancy + Parity + Delivery_mode, -->
<!--     data = df_maternal, -->
<!--     by = as.formula(paste0("~", stratum)), -->
<!--     method = "ebal" -->
<!--     ) -->

<!--     fit <- glm_weightit( -->
<!--     diagnosis ~ conc_25OHD*(First_degree_T1D_heredity_cat + Smoking_during_pregnancy +  -->
<!--     Fish_any_pregnancy + Parity + Delivery_mode), -->
<!--     data = df_maternal, -->
<!--     weightit = w, -->
<!--     family = binomial -->
<!--     ) -->

<!--     avg <- avg_comparisons( -->
<!--     fit, -->
<!--     variables = "conc_25OHD", -->
<!--     by = stratum, -->
<!--     type = "link", -->
<!--     comparison = "lnor", -->
<!--     transform = "exp" -->
<!--     ) %>% -->
<!--     broom::tidy() %>% -->
<!--     dplyr::mutate_if(is.numeric, function(x) ifelse(x<0.001, formatC(x, format = "e", 3), round(x,3))) %>% -->
<!--     dplyr::mutate( -->
<!--       Term = ifelse(row_number()>1 & term == lag(term), NA_character_, gsub("conc_25OHD","conc 25(OH)D",term)), -->
<!--        `OR (95% CI)` = paste0(as.character(estimate)," (",as.character(conf.low),"-",as.character(conf.high),")"), -->
<!--        `P-value` = p.value, -->
<!--       Stratum = curr_strata, -->
<!--       Stratum = ifelse(row_number()==1, Stratum,NA_character_), -->
<!--     ) %>% -->
<!--     dplyr::rename( -->
<!--       Contrast = contrast, -->
<!--       Level = sym(stratum) -->
<!--      ) %>% -->
<!--     dplyr::select(Term, Contrast, Stratum, Level,`OR (95% CI)`,  `P-value`) -->

<!--   return(list(avg,w)) -->
<!-- }) -->
<!-- names(results_list_maternal) <- maternal_covariates[c(2:3)] -->
<!-- ``` -->


<!-- ```{r} -->
<!-- for(ii in seq_along(names(results_list_maternal))) { -->
<!--   if (ii == 1) { -->
<!--     maternal_25OHD_diagnosis_results <- results_list_maternal[[ii]][[1]] -->
<!--   } else { -->
<!--     maternal_25OHD_diagnosis_results <- rbind(maternal_25OHD_diagnosis_results, results_list_maternal[[ii]][[1]]) -->
<!--   } -->
<!-- } -->

<!-- maternal_25OHD_diagnosis_results <- as.data.frame(maternal_25OHD_diagnosis_results) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- maternal_25OHD_diagnosis_results %>% -->
<!--   as.data.frame() %>% -->
<!--   flextable::flextable() %>% -->
<!--   flextable::autofit() %>% -->
<!--   flextable::color( -->
<!--     i = ~ `P-value` < 0.05, -->
<!--     j = c("Contrast", "P-value"), -->
<!--     color = "red") -->
<!-- ``` -->

<!-- Cox analysis -->

<!-- conducting a time-to-event Cox regression with: -->

<!--  - Time scale: Age (from birth) -->

<!--  - Exposure: Fixed (cord blood 25(OH)D) -->

<!--  - Follow-up ends: December 31, 2020 -->

<!-- Outcome: T1D diagnosis -->
<!-- ```{r event_time_formulation, warning=FALSE} -->
<!-- vitamin_d_case_control_plus_abis$start_age <- 0  # since follow-up starts at birth -->

<!-- # Set exit date based on whether T1D occurred before 2020-12-31 -->
<!-- vitamin_d_case_control_plus_abis$exit_date <- ifelse( -->
<!--   !is.na(vitamin_d_case_control_plus_abis$T1D_age_diagnosis) & -->
<!--     vitamin_d_case_control_plus_abis$INDATUMA <= as.Date("2020-12-31"), -->
<!--   vitamin_d_case_control_plus_abis$INDATUMA, -->
<!--   as.Date("2020-12-31") -->
<!-- ) -->

<!-- vitamin_d_case_control_plus_abis$exit_date <- as.Date(vitamin_d_case_control_plus_abis$exit_date) -->

<!-- # Calculate age at exit in years -->
<!-- vitamin_d_case_control_plus_abis$exit_age <- as.numeric( -->
<!--   difftime( -->
<!--     vitamin_d_case_control_plus_abis$exit_date, -->
<!--     vitamin_d_case_control_plus_abis$child_dob, -->
<!--     units = "days" -->
<!--   ) -->
<!-- ) / 365.25 -->

<!-- vitamin_d_case_control_plus_abis$event <- ifelse( -->
<!--   !is.na(vitamin_d_case_control_plus_abis$INDATUMA) & -->
<!--     vitamin_d_case_control_plus_abis$INDATUMA <= as.Date("2020-12-31"), -->
<!--   1, -->
<!--   0 -->
<!-- ) -->
<!-- ``` -->


<!-- ```{r cox_data, warning=FALSE} -->
<!-- df_cox_full <- vitamin_d_case_control_plus_abis %>% -->
<!--   dplyr::select(all_of( -->
<!--     c( -->
<!--       "ABISnr", -->
<!--       "DBP", -->
<!--       "conc_25OHD", -->
<!--       "diagnosis", -->
<!--       "child_dob", -->
<!--       "First_degree_T1D_heredity_cat", -->
<!--       "Preterm_group", -->
<!--       "Preterm_dichotomous", -->
<!--       "Delivery_mode", -->
<!--       "Gestational_age", -->
<!--       "Birth_weight", -->
<!--       "Birth_weight_cat", -->
<!--       "Birth_height", -->
<!--       "Season", -->
<!--       "Parity", -->
<!--       "Smoking_during_pregnancy", -->
<!--       "Vitamin_mineral_supplement_pregnancy", -->
<!--       "Fish_lake_pregnancy", -->
<!--       "Fish_Baltic_pregnancy", -->
<!--       "Fish_other_pregnancy", -->
<!--       "Fish_any_pregnancy", -->
<!--       "Sex_cat", -->
<!--       # "HDIA", -->
<!--       "DIAGNOS", -->
<!--       # "INDATUMA", -->
<!--       # "UTDATUMA", -->
<!--       # "SOURCE", -->
<!--       "T1D_age_diagnosis", -->
<!--       "onset_age_group", -->
<!--       "Birth_weight_cat", -->
<!--       "day_of_year", -->
<!--       # "Ai", -->
<!--       # "adjusted_25OHD", -->
<!--       "start_age", -->
<!--       "exit_date", -->
<!--       "exit_age", -->
<!--       "event" -->
<!--     ) -->
<!--   )) -->
<!-- ``` -->

<!-- ```{r remove_missing_rows} -->
<!-- df_cox_full <- df_cox_full %>% -->
<!--   dplyr::select(all_of( -->
<!--     c( -->
<!--       "ABISnr", -->
<!--       "diagnosis", -->
<!--       "conc_25OHD", -->
<!--       "exit_age", -->
<!--       "event", -->
<!--       "Birth_weight_cat", -->
<!--       child_covariates, -->
<!--       maternal_covariates -->
<!--     ) -->
<!--   )) %>% -->
<!--   tidyr::drop_na() -->
<!-- ``` -->

<!-- Cox regression	Hazard ratio: relative rate of developing T1D at any given age.  -->

<!-- ```{r cox_model} -->
<!-- cox_model <- coxph( -->
<!--   Surv(exit_age, event = event) ~ conc_25OHD + First_degree_T1D_heredity_cat + -->
<!--     Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode + -->
<!--     Preterm_dichotomous + Season + Sex_cat + Birth_weight_cat, -->
<!--   data = df_cox_full -->
<!-- ) -->
<!-- ``` -->

<!-- ```{r cox_results} -->
<!-- tidy_cox <- tidy(cox_model, exponentiate = TRUE, conf.int = TRUE) -->
<!-- tidy_cox  %>% -->
<!--   as.data.frame() %>% -->
<!--   dplyr::mutate_if(is.numeric, round, 3) %>% -->
<!--   flextable::flextable() %>% -->
<!--   flextable::autofit() %>% -->
<!--   flextable::color( -->
<!--     i = ~ p.value < 0.05, -->
<!--     j = c("term", "p.value"), -->
<!--     color = "red") -->
<!-- ``` -->

<!-- weightIt to balance out covariates across conc_25OHD exposure/treatment -->
<!-- ```{r} -->
<!-- w_25OHD_cox <- weightit( -->
<!--   conc_25OHD ~  Preterm_dichotomous + Season + First_degree_T1D_heredity_cat + -->
<!--     Sex_cat + Smoking_during_pregnancy + Fish_any_pregnancy + Delivery_mode + -->
<!--     Parity + Birth_weight_cat, -->
<!--   data = df_cox_full, -->
<!--   method = "ebal", -->
<!--   maxit = 50000 -->
<!-- ) -->
<!-- ``` -->

<!-- Cox regression (weighted)	Hazard ratio: relative rate of developing T1D at any given age. -->
<!-- ```{r cox_model_weighted} -->
<!-- cox_model_weighted <- coxph( -->
<!--   Surv(exit_age, event = event) ~ conc_25OHD + First_degree_T1D_heredity_cat + -->
<!--     Smoking_during_pregnancy + Fish_any_pregnancy + Parity + Delivery_mode + -->
<!--     Preterm_dichotomous + Season + Sex_cat + Birth_weight_cat, -->
<!--   data = df_cox_full, weights = w_25OHD_cox$weights -->
<!-- ) -->
<!-- ``` -->

<!-- ```{r cox_weighted_results} -->
<!-- tidy_cox_weighted <- tidy(cox_model, exponentiate = TRUE, conf.int = TRUE) -->
<!-- tidy_cox_weighted  %>% -->
<!--   as.data.frame() %>% -->
<!--   dplyr::mutate_if(is.numeric, round, 3) %>% -->
<!--   flextable::flextable() %>% -->
<!--   flextable::autofit() %>% -->
<!--   flextable::color( -->
<!--     i = ~ p.value < 0.05, -->
<!--     j = c("term", "p.value"), -->
<!--     color = "red") -->
<!-- ``` -->
