---
title: "Conditional regression with balancing"
author: "Debojyoti Das"
date: "2025-07-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

set output directory
```{r output_dir, message=FALSE, warning=FALSE}
output_dir <- "/Users/debda22/Projects/T1D_vitamin_D_results/cox_analysis/conditional_regression/new"

if(!dir.exists(output_dir)){
  dir.create(output_dir, recursive = TRUE)
} else {
  cat(paste0(output_dir," is already present!!\n"))
}
```

load libraries
```{r load_libraries, message=FALSE, warning=FALSE}
pacman::p_load(
  haven,        
  zoo,
  sjlabelled,
  janitor,       
  data.table,   
  ggplot2,
  comprehenr,  
  stringr,     
  rstatix,
  dplyr,
  tidyr,
  forcats,
  officer,
  cowplot,
  ggExtra,
  ggpubr,
  GGally,
  janitor,
  ggrepel,
  ggtext,
  parameters,
  lme4,
  broom,
  pscl,
  tibble,
  knitr,
  ggeffects,
  cobalt,
  WeightIt,
  marginaleffects,
  survival,
  purrr
  )
```

define paths
```{r define_paths, eval=TRUE}
data_root <- "/Users/debda22/Projects/"
data_root <- paste0(data_root,"Linkoping_projects/ABIS_RAW_DATA_SPSS/")
```

load parameters
```{r}
source(paste0(data_root,"../../Rmd_and_R_files/T1D_projects/all_parameters_case_control_T1D_vitamin_D.R"))
```


read/load vitamin_d_case_control_plus_abis for analysis
```{r}
vitamin_d_case_control_plus_abis <- read.table(paste0(
  data_root,
  "/vitamin_d/vitamin_d_case_control_plus_abis.csv"
  )
)
```

reorder factors
```{r}
vitamin_d_case_control_plus_abis$First_degree_T1D_heredity_cat <- factor(
  vitamin_d_case_control_plus_abis$First_degree_T1D_heredity_cat,
  levels = c("No", "Yes")
)

vitamin_d_case_control_plus_abis$Smoking_during_pregnancy <- factor(
  vitamin_d_case_control_plus_abis$Smoking_during_pregnancy,
  levels = c("No", "Yes")
)

vitamin_d_case_control_plus_abis$Parity <- factor(
  vitamin_d_case_control_plus_abis$Parity,
  levels = c("First parity", "Previous parity")
)

vitamin_d_case_control_plus_abis$Preterm_dichotomous <- factor(
  vitamin_d_case_control_plus_abis$Preterm_dichotomous,
  levels = c("Normal", "Preterm")
)

vitamin_d_case_control_plus_abis$Season <- factor(
  vitamin_d_case_control_plus_abis$Season,
  levels = c("Autumn", "Spring", "Summer", "Winter")
)

vitamin_d_case_control_plus_abis$Sex_cat <- factor(
  vitamin_d_case_control_plus_abis$Sex_cat,
  levels = c("Boy", "Girl")
)

vitamin_d_case_control_plus_abis$Fish_any_pregnancy <- factor(
  vitamin_d_case_control_plus_abis$Fish_any_pregnancy,
  levels = c("Seldom", "1–2 times week or more"))

vitamin_d_case_control_plus_abis$Delivery_mode <- factor(
  vitamin_d_case_control_plus_abis$Delivery_mode,
  levels = c("Normal","caesarean", "other problem"),
  labels = c("Normal","Caesarean", "Vaginal delivery with problem"))
```

```{r}
vitamin_d_case_control_plus_abis <- vitamin_d_case_control_plus_abis %>%
  dplyr::mutate(
    Vitamin_mineral_supplement_pregnancy = ifelse(
      is.na(Vitamin_mineral_supplement_pregnancy) | Vitamin_mineral_supplement_pregnancy == "Do not know",
      NA_character_,
      Vitamin_mineral_supplement_pregnancy)
  )
```

```{r}
vitamin_d_case_control_plus_abis$Vitamin_mineral_supplement_pregnancy <- factor(vitamin_d_case_control_plus_abis$Vitamin_mineral_supplement_pregnancy)
```


```{r}
qs_birth_weight <- round(quantile(vitamin_d_case_control_plus_abis$Birth_weight, probs = seq(0, 1, 0.25), na.rm = TRUE),2)

vitamin_d_case_control_plus_abis <- vitamin_d_case_control_plus_abis %>%
    dplyr::mutate(
        Birth_weight_cat = cut(
            Birth_weight,
            breaks = qs_birth_weight,
            include.lowest = TRUE,
            labels = c(
                paste0("≤", qs_birth_weight[2]),                                    
                paste0(qs_birth_weight[2], "–", qs_birth_weight[3]),                        
                paste0(qs_birth_weight[3], "–", qs_birth_weight[4]),                           
                paste0(">", qs_birth_weight[4])                                     
            )
        )
    )


```


```{r word_sect_properties}
sect_properties <- prop_section(
  page_size = page_size(
    orient = "landscape",
    width = 8.3,
    height = 11.7
  ),
  type = "continuous",
  page_margins = page_mar()
)

```


conditional regression - case-control matching


Assessing Balance: A. Check Initial Imbalance

```{r}
child_covariates <- c("Preterm_dichotomous",
                      "Birth_weight",
                      "Birth_weight_cat",
                      "Season", 
                      "Sex_cat")

maternal_covariates <- c(
  "First_degree_T1D_heredity_cat",
  "Smoking_during_pregnancy",
  "Fish_any_pregnancy",
  "Parity",
  "Delivery_mode",
  "Vitamin_mineral_supplement_pregnancy"
)
```


all covariates
```{r}
df_all <- vitamin_d_case_control_plus_abis %>%
  dplyr::select(
    all_of(c("diagnosis", "conc_25OHD", child_covariates, maternal_covariates))) %>%
  tidyr::drop_na() 
```

```{r}
all_covariates_formula <- paste0(c(child_covariates[-3], maternal_covariates),collapse = " + ")
all_covariates_formula
```


```{r echo=FALSE}
w_25OHD_all <- weightit(
  as.formula(paste0("conc_25OHD ~  ", all_covariates_formula)),
  data = df_all,
  method = "ebal"
)
```


```{r echo=FALSE}
suppressMessages(
  fit_25OHD_all <- glm_weightit(
  as.formula(paste0("diagnosis ~ conc_25OHD * (",all_covariates_formula,")")),
  data = df_all,
  weightit = w_25OHD_all,
  family = binomial
)
)
```


```{r echo=FALSE}
avg_25OHD_all <- avg_comparisons(
  fit_25OHD_all,
  type = "link",
  comparison = "lnor",
  transform = "exp"
) %>%
  broom::tidy() %>%
  as.data.frame() %>%
  dplyr::mutate(
    term = ifelse(row_number()> 1 & term == lag(term),
                    NA_character_,
                    gsub("_"," ",gsub("_cat","",gsub("_dichotomous","",gsub("Birth_weight_cat","Birth weight category",term))))
                    )
    ) %>%
  dplyr::mutate_if(is.numeric, function(x) ifelse(x<0.001, formatC(x, format = "e", 3), round(x,3))) %>%
  dplyr::rename(
    Term = term,
    Contrast = contrast,
    OR = estimate,
    `P-value` = p.value
    ) %>%
  dplyr::mutate(
        `OR (95% CI)` = paste0(as.character(OR)," (",as.character(conf.low),"-",as.character(conf.high),")")
  ) %>%
  dplyr::select(all_of(c("Term", "Contrast","OR (95% CI)", "P-value")))


first_row <- avg_25OHD_all %>% dplyr::filter(Term == "conc 25OHD")

other_rows <- avg_25OHD_all %>% dplyr::filter(Term != "conc 25OHD" | is.na(Term))

avg_25OHD_all <- rbind(first_row, other_rows)
```

Diagnosis ~ conc. 25(OH)D * (all covariates) : weighted data
```{r echo=FALSE}
# kable(avg_25OHD_all, caption = "Diagnosis ~ conc. 25(OH)D * (all covariates) : weighted data")
flex_all_cov_interaction_model <- avg_25OHD_all %>%
  as.data.frame() %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::color(
    i = ~ `P-value` < 0.05,
    j = c("Contrast", "P-value"),
    color = "red")
```


```{r save_}
flextable::save_as_docx(
  `Conditional regression interaction model with covariate balancing` = flex_all_cov_interaction_model,
  path = paste0(output_dir,"/flex_conditional_regression_all_cov_interaction_model.docx"),
  pr_section = sect_properties
)
```
<!-- ```{r} -->
<!-- birth_weight_quantiles_all <- quantile(df_all$Birth_weight) -->

<!-- df_all$Birth_weight_cat <- cut( -->
<!--   df_all$Birth_weight, -->
<!--   breaks = c( -->
<!--     birth_weight_quantiles_all[[1]], -->
<!--     birth_weight_quantiles_all[[2]], -->
<!--     birth_weight_quantiles_all[[3]], -->
<!--     birth_weight_quantiles_all[[4]], -->
<!--     Inf -->
<!--   ), -->
<!--   labels = c("q1", "q2", "q3", "q4"), -->
<!--   include.lowest = TRUE -->
<!-- ) -->
<!-- ``` -->


```{r}
df_all_all_cat <- df_all %>% 
  dplyr::select(-Birth_weight) 
```


```{r}
all_covariates_cat_formula <- paste0(names(df_all_all_cat)[-c(1:3)],collapse = " + ")
all_covariates_cat_formula
```



```{r echo=FALSE}
w_25OHD_all_cat <- weightit(
  as.formula(paste0("conc_25OHD ~  ", all_covariates_cat_formula)),
  data = df_all_all_cat
  #,
  # method = "ebal"
)
```


```{r echo=FALSE}
fit_25OHD_all_cat <- glm_weightit(
  as.formula(paste0("diagnosis ~ conc_25OHD * (",all_covariates_cat_formula,")")),
  data = df_all_all_cat,
  weightit = w_25OHD_all_cat,
  family = binomial
)

avg_25OHD_all_cat <- avg_comparisons(
  fit_25OHD_all_cat,
  type = "link",
  comparison = "lnor",
  transform = "exp"
) %>%
  broom::tidy() %>%
  as.data.frame() %>%
  dplyr::mutate(
    term = ifelse(row_number()> 1 & term == lag(term),
                    NA_character_,
                    gsub("_"," ",gsub("_cat","",gsub("_dichotomous","",gsub("Birth_weight_cat","Birth weight category",term))))
                    )
    ) %>%
  dplyr::mutate_if(is.numeric, function(x) ifelse(x<0.001, formatC(x, format = "e", 3), round(x,3))) %>%
  dplyr::rename(
    Term = term,
    Contrast = contrast,
    OR = estimate,
    `P-value` = p.value
    ) %>%
  dplyr::mutate(
        `OR (95% CI)` = paste0(as.character(OR)," (",as.character(conf.low),"-",as.character(conf.high),")")
  ) %>%
  dplyr::select(all_of(c("Term", "Contrast","OR (95% CI)", "P-value")))

first_row <- avg_25OHD_all_cat %>% dplyr::filter(Term == "conc 25OHD")
other_rows <- avg_25OHD_all_cat %>% dplyr::filter(Term != "conc 25OHD" | is.na(Term))
avg_25OHD_all_cat <- rbind(first_row, other_rows)
```

Diagnosis ~ conc. 25(OH)D * (all categorical covariates) : weighted data
```{r echo=FALSE}
# kable(avg_25OHD_all_cat, caption = "Diagnosis ~ conc. 25(OH)D * (all categorical covariates) : weighted data")
flex_avg_25OHD_all_cat_interaction_model <- avg_25OHD_all_cat %>%
  as.data.frame() %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::color(
    i = ~ `P-value` < 0.05,
    j = c("Contrast", "P-value"),
    color = "red")
```



```{r save_flex_avg_25OHD_all_cat_interaction_model}
flextable::save_as_docx(
  `Conditional regression interaction model with covariate (all categorical) balancing` = flex_avg_25OHD_all_cat_interaction_model,
  path = paste0(output_dir,"/flex_conditional_regression_all_categorical_cov_interaction_model.docx"),
  pr_section = sect_properties
)
```

```{r}
df_child <- vitamin_d_case_control_plus_abis %>%
  dplyr::select(ABISnr, diagnosis,conc_25OHD,Preterm_dichotomous,Birth_weight_cat,Season,Sex_cat) %>%
  tidyr::drop_na()
```

```{r}
# fit_25OHD_child_no_match_no_interaction <- glm_weightit(
#     diagnosis ~ conc_25OHD,
#     data = df_child,
#     family = binomial
#   )
# 
# avg_25OHD_child_no_match_no_interaction <- avg_comparisons(
#   fit_25OHD_child_no_match_no_interaction,
#   type = "link",
#   comparison = "lnor",
#   transform = "exp"
# ) %>%
#   broom::tidy() %>%
#   dplyr::select(term, estimate, p.value, conf.low, conf.high) %>%
#   dplyr::mutate(term = ifelse(
#     row_number() > 1 & term == lag(term),
#     NA_character_,
#     gsub("_", " ", gsub(
#       "_cat", "", gsub(
#         "_dichotomous",
#         "", gsub("Birth_weight_cat", "Birth weight category", term)))))
#     ) %>%
#   dplyr::mutate_if(is.numeric, function(x)
#     ifelse(x < 0.001, formatC(x, format = "e", 3), round(x, 3))) %>%
#   dplyr::rename(
#     Term = term,
#     OR = estimate,
#     `P-value` = p.value
#   ) %>%
#   dplyr::mutate(
#         `OR (95% CI)` = paste0(as.character(OR)," (",as.character(conf.low),"-",as.character(conf.high),")")
#   ) %>%
#   dplyr::select(-all_of(c("OR","conf.low", "conf.high"))) %>%
#   dplyr::select(all_of(c("Term","OR (95% CI)", "P-value")))
```

<!-- Diagnosis ~ conc. 25(OH)D * (child covariates) : not weighted nointeraction -->
```{r}
# kable(
#   avg_25OHD_child_no_match_no_interaction,
#   caption = "Diagnosis ~ conc. 25(OH)D * (child covariates) : not weighted nointeraction")

# avg_25OHD_child_no_match_no_interaction %>%
#   as.data.frame() %>%
#   flextable::flextable() %>%
#   flextable::autofit()
```

glm diagnosis ~ conc_25OHD; case-control matched by child (birth) covariates
```{r}
w_25OHD_child <- weightit(
    conc_25OHD ~ Preterm_dichotomous + Season + Sex_cat + Birth_weight_cat,
    data = df_child
    #,
    # method = "ebal"
  )

# fit_25OHD_child_cov_matched_no_interaction <- glm_weightit(
#     diagnosis ~ conc_25OHD,
#     data = df_child,
#     weightit = w_25OHD_child,
#     family = binomial
#   )
# 
# avg_25OHD_child_cov_matched_no_interaction <- avg_comparisons(
#   fit_25OHD_child_cov_matched_no_interaction,
#   type = "link",
#   comparison = "lnor",
#   transform = "exp"
# ) %>%
#   broom::tidy() %>%
#   dplyr::select(term, estimate, p.value, conf.low, conf.high) %>%
#   dplyr::mutate(term = ifelse(
#     row_number() > 1 & term == lag(term),
#     NA_character_,
#     gsub("_", " ", gsub(
#       "_cat", "", gsub(
#         "_dichotomous",
#         "", gsub("Birth_weight_cat", "Birth weight category", term)))))
#     ) %>%
#   dplyr::mutate_if(is.numeric, function(x)
#     ifelse(x < 0.001, formatC(x, format = "e", 3), round(x, 3))) %>%
#   dplyr::rename(
#     Term = term,
#     OR = estimate,
#     `P-value` = p.value
#   ) %>%
#   dplyr::mutate(
#         `OR (95% CI)` = paste0(as.character(OR)," (",as.character(conf.low),"-",as.character(conf.high),")")
#   ) %>%
#   dplyr::select(-all_of(c("OR","conf.low", "conf.high"))) %>%
#   dplyr::select(all_of(c("Term","OR (95% CI)", "P-value")))
```

<!-- Diagnosis ~ conc. 25(OH)D * (child covariates) : weighted but no interaction -->
<!-- ```{r} -->
<!-- # kable( -->
<!-- #   avg_25OHD_child_cov_matched_no_interaction, -->
<!-- #   caption = "Diagnosis ~ conc. 25(OH)D * (child covariates) : weighted but no interaction") -->

<!-- avg_25OHD_child_cov_matched_no_interaction %>% -->
<!--   as.data.frame() %>% -->
<!--   flextable::flextable() %>% -->
<!--   flextable::autofit() -->
<!-- ``` -->


```{r}
fit_25OHD_child_cov_matched_and_with_interaction <- glm_weightit(
    diagnosis ~ conc_25OHD*(Preterm_dichotomous + Season + Sex_cat + Birth_weight_cat),
    data = df_child,
    weightit = w_25OHD_child,
    family = binomial
  )

avg_25OHD_child_cov_matched_and_with_interaction <- avg_comparisons(
  fit_25OHD_child_cov_matched_and_with_interaction,
  type = "link",
  comparison = "lnor",
  transform = "exp"
) %>%
  broom::tidy() %>%
  dplyr::select(term, contrast, estimate, p.value, conf.low, conf.high) %>%
  dplyr::mutate(term = ifelse(
    row_number() > 1 & term == lag(term),
    NA_character_,
    gsub("_", " ", gsub(
      "_cat", "", gsub(
        "_dichotomous",
        "", gsub("Birth_weight_cat", "Birth weight category", term)))))
    ) %>%
  dplyr::mutate_if(
    is.numeric, function(x)
    ifelse(x < 0.001, 
           formatC(x, format = "e", 2), 
           formatC(x, format = "f", 3))
    ) %>%
  dplyr::rename(
    Term = term,
    Contrast = contrast,
    OR = estimate,
    `P-value` = p.value
  ) %>%
  dplyr::mutate(
        `OR (95% CI)` = paste0(as.character(OR)," (",as.character(conf.low),"-",as.character(conf.high),")")
  ) %>%
  dplyr::select(all_of(c("Term", "Contrast","OR (95% CI)", "P-value"))) 

first_row <- avg_25OHD_child_cov_matched_and_with_interaction %>%
  dplyr::filter(Term == "conc 25OHD")
other_rows <- avg_25OHD_child_cov_matched_and_with_interaction %>%
  dplyr::filter(Term != "conc 25OHD" | is.na(Term))

avg_25OHD_child_cov_matched_and_with_interaction <- rbind(first_row, other_rows)
```

Diagnosis ~ conc. 25(OH)D * (child covariates) : weighted and with interaction
```{r}
flex_avg_25OHD_child_cov_matched_and_with_interaction <- avg_25OHD_child_cov_matched_and_with_interaction %>%
  as.data.frame() %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::color(
    i = ~ as.numeric(`P-value`) < 0.05,
    j = c("Contrast", "P-value"),
    color = "red")
```

```{r save_flex_avg_25OHD_child_cov_matched_and_with_interaction}
flextable::save_as_docx(
  `Conditional regression interaction model with child covariate balancing` = flex_avg_25OHD_child_cov_matched_and_with_interaction,
  path = paste0(output_dir,"/flex_conditional_regression_child_cov_interaction_model.docx"),
  pr_section = sect_properties
)
```

```{r}
results_list_child <- lapply(c("Season", "Sex_cat", "Birth_weight_cat", "Preterm_dichotomous"), 
  function(stratum) {
    covariates2match <- paste0(setdiff(c("Season", "Sex_cat", "Birth_weight_cat", "Preterm_dichotomous"),stratum), collapse = " + ")
    formula2use <- paste0("conc_25OHD ~ ", covariates2match)
    
    print(covariates2match)
    print(formula2use)
    
    w <- weightit(
    as.formula(formula2use),
    data = df_child,
    by = as.formula(paste0("~", stratum))
    #,
    # method = "ebal"
  )
    formula2fit <- paste0("diagnosis ~ conc_25OHD * (",covariates2match,")")
    print(formula2fit)
    print("###")
    
  fit <- glm_weightit(
    as.formula(formula2fit),
    data = df_child,
    weightit = w,
    family = binomial
  )
  avg <- avg_comparisons(
    fit,
    variables = "conc_25OHD",
    by = stratum,
    type = "link",
    comparison = "lnor",
    transform = "exp"
  ) %>%
    broom::tidy() %>%
    dplyr::mutate_if(
      is.numeric, function(x) ifelse(x<0.001, 
                                     formatC(x, format = "e", 2),  
                                     formatC(x, format = "f", 3))) %>%
    dplyr::mutate(
      Term = ifelse(row_number()>1 & term == lag(term), NA_character_, gsub("conc_25OHD","conc 25(OH)D",term)),
       `OR (95% CI)` = paste0(as.character(estimate)," (",as.character(conf.low),"-",as.character(conf.high),")"),
       `P-value` = p.value,
      Stratum = gsub("_", " ", 
                     gsub("_cat", "", 
                          gsub("_dichotomous","", 
                               gsub("Birth_weight_cat", 
                                    "Birth weight category", stratum)
                               )
                          )
                     ),
      Stratum = ifelse(row_number()==1, Stratum,NA_character_),
    ) %>%
    dplyr::rename(
      Contrast = contrast,
      Level = sym(stratum)
     ) %>%
    dplyr::select(Term, Contrast, Stratum, Level,`OR (95% CI)`,  `P-value`)
    # dplyr::rename(level = !!stratum) %>%
    # dplyr::mutate(strata = stratum) %>%
    # dplyr::select(Term, Contrast, strata, level, estimate, p.value, conf.low, conf.high)
  return(list(avg, w, fit))
})
names(results_list_child) <- c("Season", "Sex_cat", "Birth_weight_cat", "Preterm_dichotomous")
```

```{r}
cowplot::plot_grid(
    love.plot(results_list_child[[1]][[2]], title = ""),
    love.plot(results_list_child[[2]][[2]], title = ""),
    love.plot(results_list_child[[3]][[2]], title = ""),
    love.plot(results_list_child[[4]][[2]], title = ""),
    labels = c("A. Season",
               "B. Sex",
               "C. Birth weight category",
               "D. Preterm-birth"
               ),
    nrow = 2,
    ncol = 2
    )
```

save stratum stratified weight matched
```{r}
ggsave(
  filename = paste0(output_dir, "/conc_25OHD_child_covariate_correlations_entropy_balanced.jpeg"),
  plot = cowplot::plot_grid(
    love.plot(results_list_child[[1]][[2]], title = ""),
    love.plot(results_list_child[[2]][[2]], title = ""),
    love.plot(results_list_child[[3]][[2]], title = ""),
    love.plot(results_list_child[[4]][[2]], title = ""),
    labels = c("A. Season",
               "B. Sex",
               "C. Birth weight category",
               "D. Preterm-birth"
               )
    ),
  dpi = 300,
  units = "in",
  type = "cairo",
  width = 10.3,
  height = 14.7
)
```


```{r}
for(ii in seq_along(names(results_list_child))) {
  if (ii == 1) {
    child_25OHD_diagnosis_results <- results_list_child[[ii]][[1]]
  } else {
    child_25OHD_diagnosis_results <- rbind(child_25OHD_diagnosis_results, results_list_child[[ii]][[1]])
  }
}
```


```{r}
# child_25OHD_diagnosis_results <- child_25OHD_diagnosis_results %>%
#   dplyr::mutate(
#     strata = ifelse(row_number()> 1 & strata == lag(strata),
#                     NA_character_,
#                     gsub("_"," ",gsub("_cat","",gsub("_dichotomous","",gsub("Birth_weight_cat","Birth weight category",strata))))
#                     )
#     ) %>%
#   dplyr::mutate_if(is.numeric, function(x) ifelse(x<0.001, formatC(x, format = "e", 3), round(x,3))) %>%
#   dplyr::rename(
#     Strata = strata,
#     Level = level,
#     OR = estimate,
#     `P-value` = p.value
#     ) %>%
#   dplyr::mutate(
#         `OR (95% CI)` = paste0(as.character(OR)," (",as.character(conf.low),"-",as.character(conf.high),")")
#   ) %>%
#   dplyr::select(-all_of(c("OR","conf.low", "conf.high"))) %>%
#   dplyr::select(all_of(c("Strata", "Level","OR (95% CI)", "P-value")))

flex_child_25OHD_diagnosis_results <- child_25OHD_diagnosis_results %>%
  as.data.frame() %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::color(
    i = ~ `P-value` < 0.05,
    j = c("Level", "P-value"),
    color = "red")

flex_child_25OHD_diagnosis_results
```

```{r save_flex_child_25OHD_diagnosis_results}
flextable::save_as_docx(
  `Conditional regression: stratified interaction model with child covariate balancing` = flex_child_25OHD_diagnosis_results,
  path = paste0(output_dir,"/flex_conditional_regression_child_cov_stratified_interaction_model.docx"),
  pr_section = sect_properties
)
```

```{r}
maternal_covariates <- c(
  "First_degree_T1D_heredity_cat",
  "Smoking_during_pregnancy",
  "Fish_any_pregnancy",
  "Parity",
  "Delivery_mode",
  "Vitamin_mineral_supplement_pregnancy"
)
```

```{r}
df_maternal <- vitamin_d_case_control_plus_abis %>%
  dplyr::select(
    all_of(c("ABISnr", "diagnosis", "conc_25OHD", maternal_covariates))) %>%
  tidyr::drop_na()
```

no stratification
```{r echo=FALSE}
w_25OHD_maternal <- weightit(
  conc_25OHD ~  First_degree_T1D_heredity_cat + Smoking_during_pregnancy + 
    Fish_any_pregnancy + Parity + Delivery_mode + Vitamin_mineral_supplement_pregnancy,
  data = df_maternal
  #,
  # method = "ebal"
)


fit_25OHD_maternal <- glm_weightit(
  diagnosis ~ conc_25OHD * (
    First_degree_T1D_heredity_cat + Smoking_during_pregnancy + 
      Fish_any_pregnancy + Parity + Delivery_mode + Vitamin_mineral_supplement_pregnancy
  ),
  data = df_maternal,
  weightit = w_25OHD_maternal,
  family = binomial
)

avg_25OHD_maternal <- avg_comparisons(
  fit_25OHD_maternal,
  type = "link",
  comparison = "lnor",
  transform = "exp"
) %>%
  broom::tidy() %>%
  as.data.frame() %>%
  dplyr::mutate(term = ifelse(
    row_number() > 1 & term == lag(term),
    NA_character_,
    gsub("_", " ", gsub(
      "_cat", "", gsub(
        "_dichotomous",
        "",
        gsub("Birth_weight_cat", "Birth weight category", term)
      )
    ))
  )) %>%
  dplyr::mutate_if(is.numeric, function(x) ifelse(x<0.001, formatC(x, format = "e", 3), round(x,3))) %>%
  dplyr::rename(
    Term = term,
    Contrast = contrast,
    OR = estimate,
    `P-value` = p.value
    ) %>%
  dplyr::mutate(
        `OR (95% CI)` = paste0(as.character(OR)," (",as.character(conf.low),"-",as.character(conf.high),")")
  ) %>%
  dplyr::select(all_of(c("Term", "Contrast","OR (95% CI)", "P-value")))

first_row <- avg_25OHD_maternal %>%
  dplyr::filter(Term == "conc 25OHD")
other_rows <- avg_25OHD_maternal %>%
  dplyr::filter(Term != "conc 25OHD" | is.na(Term))

avg_25OHD_maternal <- rbind(first_row, other_rows)
  
flex_avg_25OHD_maternal_cov_matched_and_with_interaction <- avg_25OHD_maternal %>%
  as.data.frame() %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::color(
    i = ~ `P-value` < 0.05,
    j = c("Contrast", "P-value"),
    color = "red")
```

```{r save_flex_avg_25OHD_maternal_cov_matched_and_with_interaction}
flextable::save_as_docx(
  `Conditional regression: interaction model with maternal covariate balancing` = flex_avg_25OHD_maternal_cov_matched_and_with_interaction,
  path = paste0(output_dir,"/flex_conditional_regression_maternal_cov_interaction_model.docx"),
  pr_section = sect_properties
)
```

diagnosis ~ conc_25OHD with stratified maternal covariates
```{r warning=FALSE}
results_list_maternal <- lapply(maternal_covariates, function(stratum) {

  curr_strata <- gsub("_", " ", 
                      gsub("_cat", "", 
                           gsub( "_during_pregnancy", " (pregnancy) ", 
                                 gsub("_any_pregnancy", " any (pregnancy) ", 
                                      stratum)
                                 )
                           )
                      )

    covariates2match <- paste0(
      setdiff(c("First_degree_T1D_heredity_cat", "Smoking_during_pregnancy",
    "Fish_any_pregnancy", "Parity", "Delivery_mode", "Vitamin_mineral_supplement_pregnancy"), stratum), collapse = " + ")
    formula2use <- paste0("conc_25OHD ~ ", covariates2match)
    
    print(covariates2match)
    print(formula2use)
    
    w <- weightit(
    as.formula(formula2use),
    data = df_maternal,
    by = as.formula(paste0("~", stratum))
    #,
    # method = "ebal"
    )
    
    formula2fit <- paste0("diagnosis ~ conc_25OHD * (",covariates2match,")")
    print(formula2fit)
    print("###")

    fit <- glm_weightit(
    as.formula(formula2fit),
    data = df_maternal,
    weightit = w,
    family = binomial
    )

    avg <- avg_comparisons(
    fit,
    variables = "conc_25OHD",
    by = stratum,
    type = "link",
    comparison = "lnor",
    transform = "exp"
    ) %>%
    broom::tidy() %>%
    dplyr::mutate_if(
      is.numeric, function(x) ifelse(x<0.001, 
                                     formatC(x, format = "e", 2),  
                                     formatC(x, format = "f", 3))) %>%
    dplyr::mutate(
      Term = ifelse(row_number()>1 & term == lag(term), NA_character_, gsub("conc_25OHD","conc 25(OH)D",term)),
       `OR (95% CI)` = paste0(as.character(estimate)," (",as.character(conf.low),"-",as.character(conf.high),")"),
       `P-value` = p.value,
      Stratum = curr_strata,
      Stratum = ifelse(row_number()==1, Stratum,NA_character_),
    ) %>%
    dplyr::rename(
      Contrast = contrast,
      Level = sym(stratum)
     ) %>%
    dplyr::select(Term, Contrast, Stratum, Level,`OR (95% CI)`,  `P-value`)

  return(list(avg,w, fit))
})
names(results_list_maternal) <- maternal_covariates
```


```{r}
for(ii in seq_along(names(results_list_maternal))) {
  if (ii == 1) {
    maternal_25OHD_diagnosis_results <- results_list_maternal[[ii]][[1]]
  } else {
    maternal_25OHD_diagnosis_results <- rbind(maternal_25OHD_diagnosis_results, results_list_maternal[[ii]][[1]])
  }
}

maternal_25OHD_diagnosis_results <- as.data.frame(maternal_25OHD_diagnosis_results)
```

```{r}
ggsave(
  filename = paste0(output_dir, "/conc_25OHD_maternal_covariate_correlations_entropy_balanced.jpeg"),
  plot = cowplot::plot_grid(
    love.plot(results_list_maternal[[1]][[2]], title = ""),
    love.plot(results_list_maternal[[2]][[2]], title = ""),
    love.plot(results_list_maternal[[3]][[2]], title = ""),
    love.plot(results_list_maternal[[4]][[2]], title = ""),
    love.plot(results_list_maternal[[5]][[2]], title = ""),
    love.plot(results_list_maternal[[6]][[2]], title = ""),
    labels = c("A. First-degree T1D heredity",    
               "B. Smoking during pregnancy",
               "C. Fish intake pregnancy",
               "D. Parity",
               "E. Delivery mode",
               "F. Vitamin supplement pregnancy"
               ),
    nrow = 3, ncol = 2
    ),
  dpi = 300,
  units = "in",
  type = "cairo",
  width = 11.67,
  height = 16.27
)
```

```{r}
flex_stratified_maternal_25OHD_diagnosis_results <- maternal_25OHD_diagnosis_results %>%
  as.data.frame() %>%
  flextable::flextable() %>%
  flextable::autofit() %>%
  flextable::color(
    i = ~ `P-value` < 0.05,
    j = c("Contrast", "P-value"),
    color = "red")

flex_stratified_maternal_25OHD_diagnosis_results
```

```{r save_flex_stratified_maternal_25OHD_diagnosis_results}
flextable::save_as_docx(
  `Conditional regression: stratified interaction model with maternal covariate balancing` = flex_stratified_maternal_25OHD_diagnosis_results,
  path = paste0(output_dir,"/flex_conditional_regression_maternal_cov_stratified_interaction_model.docx"),
  pr_section = sect_properties
)
```
